### 2024-12-17 paper/paper.qmd
start rough outline of paper

---
title: "BlockTensorDecompositions.jl: A Unified Constrained Tensor Decomposition Julia Package"
subtitle: ""
authors:
  - name: "Nicholas J. E. Richardson"
    affiliations:
      - ref: math
  - name: "Noah Marusenko"
    affiliations:
      - ref: cs
  - name: "Michael P. Friedlander"
    affiliations:
      - ref: mathcs
affiliations:
  - id: math
    name: "Department of Mathematics\n"
  - id: cs
    name: "Department of Computer Science\n"
  - id: mathcs
    name: "Departments of Mathematics and Computer Science\n"
date: ""
date-format: "D MMMM YYYY"
slide-number: true
categories: [tensors, signal processing, julia]
execute:
  freeze: true
bibliography: references.bib
csl: citationstyles/ieee-compressed-in-text-citations.csl # Citation Style
callout-appearance: minimal
callout-icon: false
format:
  typst:
    toc: true
    section-numbering: 1.1.a.i
    keep-typ: true
    mainfont: "Libertinus Serif"
editor:
  render-on-save: true
#appendix-style: plain
# nocite: |
#   @firstCitation
#embed-resources: true ## uncomment when sharing HTML file
---

# Introduction

### 2024-12-17 paper/fixtypst.jl

@@ -1,0 +1,71 @@
+# fixtypst.jl
+#
+# Script to fix issues when rendering Quarto markdown
+# .qmd files to Typst .typ files. These issues are
+# because of Pandoc's conversion issues.
+
+filename = "paper"
+extention = ".typ"
+
+patterns = (
+    # Bad symbol arrangement
+    ";." => ".", # citations with sectioning `[@this, Sec. 1] -> @this[Sec. 1];.`
+    "];@" => "]@", # citations next to each other with sectioning `@this[Sec. 1];@that`
+    " prime" => "'", # plays nicer with variables
+    " ; " => " \\; ", # semi-colon in mathmode should be escaped
+
+    # Fix norm and absolute value delimiters
+    r"∥([\w\W]+?)∥" => s"norm(\1)",
+    r"parallel([\w\W]+?)parallel" => s"norm(\1)",
+    r"\\\|([\w\W]+?)\\\|" => s"abs(\1)",
+    r"lr\(\|([\w\W]+?)\|\)" => s"abs(\1)",
+    r"abs\(\s([\w\W]+?)\s\)" => s"abs(\1)", # remove spaces around absolute value signs
+
+    # Fix inner product and double bracket left-right delimiters
+    r"angle.l([\w\W]+?)angle.r" => s"lr(angle.l\1angle.r)",
+    r"bracket\.l\.double([\w\W]+?)bracket\.r\.double" => s"lr(bracket.l.double\1bracket.r.double)",
+    " | " => " mid(|) ", # midline as part of sets: \left\{x \middle| x > 0 \right\}
+
+    # Figure grids with 4 columns should have a smaller gutter
+    "#grid(columns: 4, gutter: 2em" => "#grid(columns: 4, gutter: 1em",
+
+    # Missing callout box icons
+    "fa-lightbulb()" => "\"❔\"",
+    "fa-info()" => "\"❕\"",
+
+    # Custom citation style in bibliography
+    ".bib\")" => ".bib\", style: \"citationstyles/ieee-compressed-in-text-citations.csl\")",
+
+    # Theorem numbering base_level (only show Definition 4.1, not 4.1.1.1 for sub-subsections)
+    r"thmbox\(([\w\W]+?)\)" => s"thmbox(\1, base_level: 1)"
+    )
+
+"""
+    fixit(filename, patterns; extention=".typ", append_filename="_fixed")
+
+Uses patterns to perform find-and-replace on `filename`.
+
+
+Filename should be given without the file extension. Pattern can be any iterable of Pairs
[…]

### 2024-12-18 paper/paper.qmd
more paper outline work

@@
   - id: math
-    name: "Department of Mathematics\n"
+    name: "Department of Mathematics"
   - id: cs
-    name: "Department of Computer Science\n"
+    name: "Department of Computer Science"
   - id: mathcs
@@
 # Introduction
+
+- Tenors are useful in many applications
+- Need tools for fast and efficient decompositions
+
+For the scientific user, it would be most useful for there to be a single piece of software that can take as input, any reasonable type of factorization model, and constraints on the individual factors, and produce a factorization without worrying the user about the details of what rank to select, how the constraints should be enforced, and how to optimize for performance.
+Of course a knowledgeable user may still want the ability to tweak the convergence criteria used, the loss function optimized, or what statistics to record each iteration.
+These are the core specification for BlockTensorDecompositions.jl.
+
+## Related tools
+
+- Packages within Julia
+- Other languages
+- Hint at why I developed this
+
+Beyond the external usefulness already mentioned, this package offers a playground for fair comparisons of different parameters and options for performing tensor factorizations across various decomposition models.
+There exist packages for working with tensors in languages like Python (TensorFlow [@martin_abadi_tensorflow_2015], PyTorch [@ansel_pytorch_2024], and TensorLy [@kossaifi_tensorly_2019]), MATLAB (Tensor Toolbox [@bader_tensor_2023]), R (rTensor [@li_rtensor_2018]), and Julia (TensorKit.jl [@jutho_juthotensorkitjl_2024], Tullio.jl [@abbott_mcabbotttulliojl_2023], OMEinsum.jl [@peter_under-peteromeinsumjl_2024], and TensorDecompositions.jl [@wu_yunjhongwutensordecompositionsjl_2024]).
+But they only provide a groundwork for basic manipulation of tensors and the most common tensor decomposition models and algorithms, and are not equipped to handle arbitrary user defined constraints and factorization models.
+
+Some progress towards building a unified framework has been made [@xu_BlockCoordinateDescent_2013; @kim_algorithms_2014; @yang_unified_2011].
+But these approaches don't operate on the high dimensional tensor data natively and rely on matricizations of the problem, or only consider nonnegative constraints.
+They also don't provide an all-in-one package for executing their frameworks.
+
+## Contributions
+
+- Fast and flexible tensor decomposition package
+- Framework for creating and performing custom
+  - tensor decompositions
+  - constrained factorization (the what)
+  - iterative updates (the how)
+- Implement new "tricks"
+  - a (Lipschitz) matrix step size for efficient sub-block updates
+  - multi-scaled factorization when tensor entries are discretizations of a continuous function
+  - partial projection and rescaling to enforce linear constraints (rather than Euclidean projection)
+  - ?? rank detection ??
+
+# Tensor Decompositions
+- the math section of the paper
+
+## Notation
+- tensor notation, use MATLAB notation for indexing so subscripts can be used for a sequence of tensors
+
+## Common Decompositions
+- Extensions of PCA/ICA/NMF to higher dimensions
+- talk about the most popular Tucker, Tucker-n, CP
+- other decompositions
+  - high order SVD (see Kolda and Bader)
+  - HOSVD (see Kolda, Shifted power method for computing tensor eigenpairs)
+
+## Tensor rank
+- tensor rank
+- constrained rank (nonnegative etc.)
+
+# Computing Decompositions
+- Given a data tensor and a model, how do we fit the model?
+
+## Optimization Problem
+- Least squares (can use KL, 1 norm, etc.)
+
+## Base algorithm
+- Use Block Coordinate Descent / Alternating Proximal Descent
+  - do *not* use alternating least squares (slower for unconstrained problems, no closed form update for general constrained problems)
+
+# Techniques for speeding up convergences
+- As stated, algorithm works
+- But can be slow, especially for constrained or large problems
+
+## Sub-block Descent
+- Use smaller blocks, but descent in parallel (sub-blocks don't wait for other sub-blocks)
+- Can perform this efficiently with a "matrix step-size"
+
+## Momentum
+- This one is standard
+- Use something similar to [@xu_BlockCoordinateDescent_2013]
+- This is compatible with sub-block descent with appropriately defined matrix operations
+
+## Partial Projection and Rescaling
+- for bounded linear constraints
+  - first project
+  - then rescale to enforce linear constraints
+- faster to execute then a projection
+- often does not loose progress because of the rescaling (decomposition dependent)
+
+## Multi-scale
+- use a coarse discretization along continuous dimensions
+- factorize
+- linearly interpolate decomposition to warm start larger decompositions
+
+# Conclusion
+- all-in-one package
+- provide a playground to invent new decompositions
+- like auto-diff for factorizations

### 2025-01-06 paper/paper.qmd
Intro to notation sections

@@
 
-For the scientific user, it would be most useful for there to be a single piece of software that can take as input, any reasonable type of factorization model, and constraints on the individual factors, and produce a factorization without worrying the user about the details of what rank to select, how the constraints should be enforced, and how to optimize for performance.
-Of course a knowledgeable user may still want the ability to tweak the convergence criteria used, the loss function optimized, or what statistics to record each iteration.
+For the scientific user, it would be most useful for there to be a single piece of software that can take as input 1) any reasonable type of factorization model and 2) constraints on the individual factors, and produce a factorization.
+Details like what rank to select, how the constraints should be enforced, and convergence criteria should be handled automatically, but customizable to the knowledgable user.
 These are the core specification for BlockTensorDecompositions.jl.
@@
 
+The main contribution is a description of a fast and flexible tensor decomposition package, along with a public implementation written in Julia: BlockTensorDecompositions.jl.
+This package provides a framework for creating and performing custom tensor decompositions.
+To the author's knowledge, it is the first package to provide automatic factorization to a large class of constrained tensor decompositions problems, as well as a framework for implementing new constraints and iterative algorithms.
+This paper also describes three new techniques not found in the literature that empirically convergence faster than traditional block-coordinate descent.
+
 # Tensor Decompositions
@@
 
+This section reviews the notation used throughout the paper and commonly used tensor decompositions.
+
 ## Notation
 - tensor notation, use MATLAB notation for indexing so subscripts can be used for a sequence of tensors
+
+We use $[N]=\{1, 2, \dots, N\}=\{n\}_{n=1}^N$ to denote integers from $1$ to $N$.
+
+Usually, lower case symbols will be used for the running index, and the capitalized letter will be the maximum letter it runs to. This leads to the convenient shorthand $i\in[I]$, $j\in[J]$, etc.
+
+The set of nonnegative numbers is denoted as $\mathbb{R}_+=\mathbb{R}_{\geq 0}=\left\{x\in \mathbb{R} \middle| x\geq 0\right\}$.
+
+Vectors are denoted with lowercase letters ($x$, $y$, etc.), and matrices and higher order tensors with uppercase letters (commonly $A$, $B$, $C$ and $X$, $Y$, $Z$).
+The order of a tensor is the number of axes it has.
+We would call vectors "order-1" or "1st order" tensors, and matrices "order-2" or "2nd order" tensors.
+
+To avoid confusion between entries of a vector/matrix/tensor and indexing a list of objects, we use square brackets to denote the former, and subscripts to denote the later.
+For example, the entry in the $i$th row and $j$th column of a matrix $A\in\mathbb{R}$ is $A[i,j]$.
+This follows MATLAB/Julia notation where `A[i,j]` points to the entry $A[i,j]$.
+We contrast this with a list of $I$ objects being denoted as $a_1, \dots, a_I$, or more compactly, $\{a_i\}$ when it is clear the index $i\in[I]$.
+
+The $n$-fibres, $n$th mode fibres, or mode $n$ fibres of an $N$th order tensor $A$ are denoted $A[i_1,~\dots,~i_{n-1},~:,~i_{n+1},~\dots,~i_{N}]$.
+For example, the 1-fibres of a matrix $M$ are the column vectors\
+$M[:,~j]$, and the 2-fibres are the row vectors $M[i,~:]$.
+For order-3 tensors, the $1$st, $2$nd, and $3$rd mode fibres $A[:, j, k]$, $A[i,:,:]$, and $A[i,j,:]$ are called the vertical/column, horizontal/row, and depth/tube fibres respectively and are displayed in @fig-tensor-fibres.
+In Julia, the 1-, 2-, and 3-fibres of a third order array `A` would be `eachslice(A, dims=(2,3))`, `eachslice(A, dims=(1,3))`, and `eachslice(A, dims=(1,2))`.
+
+![Fibres of an order $3$ tensor $A$.](figure/tensor_fibres.png){#fig-tensor-fibres}
+
+The $n$-slices, $n$th mode slices, or mode $n$ slices of an $N$th order tensor $A$ are notated with the slice $A[:,~\dots,~:,~i_{n},~:,~\dots,~:]$.
+For matrices, the 1-fibres are the same as the 2-slices (and vice versa), but for $N$th order tensors in general, fibres are always vectors, whereas $n$-slices are $(N-1)$th order tensors.
+For a $3$rd order tensor $A$, the $1$st, $2$nd, and $3$rd mode slices $A[i,:,:]$, $A[:,j,:]$, and $A[:,:,k]$ have special names and are called the horizontal, lateral, and frontal slices and are displayed in @fig-tensor-slices.
+In Julia, the 1-, 2-, and 3-slices of a third order array `A` would be `eachslice(A, dims=1)`, `eachslice(A, dims=2)`, and `eachslice(A, dims=3)`.
+
+![Slices of an order $3$ tensor $A$.](figure/tensor_slices.png){#fig-tensor-slices}
+
+The Frobenius inner product between two tensors $A,B\in\mathbb{R}^{I_1\times\dots\times I_N}$ is denoted
+
+$$
+\left\langle A, B\right\rangle = A \cdot B = \sum_{i_1=1}^{I_1}\dots\sum_{i_N=1}^{I_N} A[i_1,\dots,i_N] B[i_1,\dots,i_N].
+$$
+
+Since we commonly use $I$ as the size of a tensor's dimension, we use $\mathrm{id}_I$ to denote the identity tensor of size $I$ (of the appropriate order).
+When the order is $2$, $\mathrm{id}_I$ is an $I\times I$ matrix with ones along the main diagonal, and zeros elsewhere.
+For higher orders $N$, this is an $\underbrace{I\times \cdots \times I}_{N \text{times}}$ tensor where $\mathrm{id}_I[i_1,\dots,i_N]=1$ when $i_1=\dots=i_N\in[I]$, and is zero otherwise.
 
@@
   - HOSVD (see Kolda, Shifted power method for computing tensor eigenpairs)
+
+<!-- From my thesis proposal: -->
+
+<!-- This section gives a brief tour of the most common tensor factorizations, heavily relying on the framework and notation given by Kolda and Bader [@kolda_TensorDecompositionsApplications_2009].
+They are, however, presented in my own interpretation as matrix factorization generalizations and as different instances of the Tucker factorization.
+
+One way we can extend matrix factorization to multidimensional signals is with a *Tucker-1* factorization.
+Here, each mixture is an $N$th order tensor $Y_i\in\mathbb{R}^{J_1\times \dots\times J_N}$ stored in the $1$-slices of an order-$(N+1)$ tensor $Y\in\mathbb{R}^{I\times J_1\times \dots\times J_N}$, $Y[i, :, \dots, :]=Y_i$, and similarly for $B$.
+Slice-wise, we have the equation
+
+$$
+Y_i = \sum_{r=1}^R A[i,r] B_r,
+$$ {#eq-tucker-1-slices}
+
+which is entry-wise
+
+$$
+Y[i, j_1, \dots, j_N] = \sum_{r=1}^R A[i,r] B[r,j_1, \dots, j_N].
+$$ {#eq-tucker-1}
+
+![Example of a rank $R=3$ Tucker-1 factorization coloured according to @eq-tucker-1-slices.
+Here we have $Y\in\mathbb{R}^{5\times J \times K}, A\in\mathbb{R}^{5\times 3}$, and $B\in\mathbb{R}^{3\times J \times K}$.
+The first horizontal slice of $Y$, $Y_1$, is equal to $A[1,1]B_1+A[1,2]B_2+A[1,3]B_3$.](figure/tucker_1_example.png){#fig-tucker-1-example}
+
+Seeing @eq-tucker-1-slices alongside @fig-tucker-1-example makes it clear that the decomposition is performing a compression along the first axis of $Y$.
+This makes sense in the model where the mixtures are just linear combinations of sources with the same size and dimensions.
+
+In the case where the sources can be further factored along other axes, specifically the first $n$ dimensions, we can use the more general *Tucker-$n$* factorization of an order $N$ tensor $Y$,
+
+$$
+\begin{split}
+Y[i_1,\dots,i_n, & j_1, \dots, j_{N-n+1}] = \\ &\sum_{r_1=1}^{R_1} \dots \sum_{r_n=1}^{R_n} A_1[i_1,r_1]\cdots A_r[i_n,r_n] B[r_1,\dots,r_n, j_1, \dots, j_{N-n+1}].
+\end{split}
+$$ {#eq-tucker-n}
+
+In the case where $n=N$ is the order of the tensor $Y$, we omit the $n$ and call this a *Tucker* factorization. So a Tucker factorization of an order $N$ tensor $Y$ is
+
+$$
+Y[i_1,\dots,i_N] = \sum_{r_1=1}^{R_1} \dots \sum_{r_n=1}^{R_n} A_1[i_1,r_1]\cdots A_r[i_n,r_N] B[r_1,\dots,r_N].
+$$ {#eq-tucker}
+
+All the Tucker decompositions can be written using the $n$-mode product $\times_n$ between a tensor $A$ and matrix $B$, defined as
+
+$$
+(A \times_n B)[i_1, \dots, i_{n-1},j,i_{n+1},\dots,N] = \sum_{i_{n}=1}^{I_n} A[i_1, \dots, i_{n-1},i_n,i-{n+1},\dots,N]B[i_n,j].
+$$
+
+They can also be written with an $n$ product shorthand, or with bracket notation. For example, the full tucker decomposition shown in @eq-tucker can be written as
+
+$$
+\begin{align*}
+Y &= B\times_1 A_1 \times_2 \dots\times_N A_N = B \bigtimes_n A_n = \llbracket B; A_1, \dots, A_N \rrbracket \\
+\text{or}\quad Y &= A_0\times_1 \dots\times_N A_N = A_0 \bigtimes_n A_n = \llbracket A_0; A_1, \dots, A_N \rrbracket.
+\end{align*}
+$$
+
+To ease notation further, we sometimes consider the *core* tensor $B$ to be the zeroth factor $A_0=B$ so we can say $\{A_n\}_{n=0}^N$ are the factors of $Y$.
+Since it can be hard to visualize the product, consider the case of a Tucker factorization of a third order tensor $Y$:
+
+$$
+Y[i,j,k] = \sum_{r=1}^R\sum_{s=1}^S\sum_{t=1}^T A_1[i,r] A_2[j,s] A_3[k,t] B[r,s,t].
+$$
+
+We can picture this as multiplying the core factor $B$ with matrix factors $A_1, A_2, A_3$ on all sizes to construct $Y$ as shown in @fig-tucker.
+
+![Tucker factorization of a $3$rd order tensor $Y$.](figure/tucker_decomposition_order_3.png){#fig-tucker}
+
+The Tucker-$n$ factorization of an order $N$ tensor can be thought of as a special case of the (full) Tucker factorization, where the matrix factors $A_{n+1}, \dots, A_{N}$ are identity matrices.
+
+In the case where the core tensor $B\in\mathbb{R}^{R_1\times \dots\times R_N}$ of the Tucker factorization is cubic ($R_1=\dots=R_N$) and diagonal ($B[r_1,\dots,r_N]=\lambda_r$ if $r_1=\dots=r_N=r$ and $0$ otherwise), we have the **can**onical **decomp**osition/**para**llel **fac**tors model (CANDECOMP/PARAFAC or CP for short):
+
+$$
+Y[i_1,\dots,i_N] = \sum_{r=1}^{R} \lambda_r A_1[i_1,r]\cdots A_N[i_N,r].
+$$
+
+This can be written as
+
+$$
+Y = \llbracket \lambda; A_1, \dots, A_N\rrbracket.
+$$
+
+Sometimes the factor $\lambda_r$ is scaled so that the columns $A[:,r]$ are normalized $\lVert A[:,r]\rVert_2=1$ for each $r\in[R]$, or it is absorbed into the matrix factors and omitted from the decomposition.
+For the remainder of this text, I'll use CP decomposition to mean the latter case where $\lambda_r=1$ for all $r\in[R]$, as showing in @eq-cp:
+
+$$
+Y[i_1,\dots,i_N] = \sum_{r=1}^{R} A_1[i_1,r]\cdots A_N[i_N,r].
+$$ {#eq-cp}
+
+Other factorization models are used that combine aspects of CP and Tucker decomposition [@kolda_TensorDecompositionsApplications_2009], are specialized for order $3$ tensors [@qi_TripleDecompositionTensor_2020; @wu_manifold_2022], or provide alternate decomposition models entirely like tensor-trains [@oseledets_tensor-train_2011].
+But the (full) Tucker, and its special cases Tucker-$n$, and CP decomposition are most commonly used extensions of the low-rank matrix factorization.
+These factorizations are summarized in @tbl-tensor-factorizations.
+
+| Name       | Bracket Notation                             | $n$-mode Product                                         | Entry-wise   |
+|------------|----------------------------------------------|----------------------------------------------------------|--------------|
+| Tucker-$1$ | $\llbracket A_0; A_1 \rrbracket$             | $A_0\times_1 A_1$                                        | @eq-tucker-1 |
+| Tucker-$n$ | $\llbracket A_0; A_1, \dots, A_n \rrbracket$ | $A_0\times_1 A_1 \times_2 \dots\times_n A_n$             | @eq-tucker-n |
+| Tucker     | $\llbracket A_0; A_1, \dots, A_N \rrbracket$ | $A_0\times_1 A_1 \times_2 \dots\times_N A_N$             | @eq-tucker   |
+| CP         | $\llbracket A_1, \dots, A_N \rrbracket$      | $\mathrm{id}_{R}\times_1 A_1 \times_2 \dots\times_N A_N$ | @eq-cp       |
+
+: Summary of common tensor factorizations. Here, $N$ is the order of the factorized tensor. {#tbl-tensor-factorizations tbl-colwidths="[15,25,35,25]"}
+ -->
 

### 2025-01-06 paper/fixtypst.jl

@@ -34,7 +34,7 @@ patterns = (
     "fa-info()" => "\"❕\"",
 
     # Custom citation style in bibliography
-    ".bib\")" => ".bib\", style: \"citationstyles/ieee-compressed-in-text-citations.csl\")",
+    #".bib\")" => ".bib\", style: \"citationstyles/ieee-compressed-in-text-citations.csl\")",
 
     # Theorem numbering base_level (only show Definition 4.1, not 4.1.1.1 for sub-subsections)
     r"thmbox\(([\w\W]+?)\)" => s"thmbox(\1, base_level: 1)"

### 2025-01-06 src/Core/utils.jl
add DiagonalTuple todo

@@ -57,6 +57,8 @@ end
 
 const DiagonalTuple{N, T} = NTuple{N, Diagonal{T}} # Tuple type where you have a tuple of Diagonal matrices, possibly of varying sizes
 
+# TODO: Use meta programming to define these DiagonalTuple methods
+
 Base.:+(X, DT::DiagonalTuple) = map(D -> X + D, DT)
 Base.:+(DT::DiagonalTuple, X) = map(D -> D + X, DT)
 Base.:+(DT1::DiagonalTuple{N}, DT2::DiagonalTuple{N}) where {N} = Tuple(C + D for (C, D) in zip(DT1, DT2))

### 2025-01-06 src/BlockTensorDecomposition.jl
Add eachfibre function

@@ -6,7 +6,7 @@ using .Core
 
 # Basic functionality
 #include("./utils.jl")
-export SuperDiagonal, abs_randn, all_recursive, getnotindex, geomean, identityslice, interlace, norm2, proj_one_hot, projsplx, proj_one_hot!, projsplx!
+export SuperDiagonal, abs_randn, all_recursive, eachfibre, getnotindex, geomean, identityslice, interlace, norm2, proj_one_hot, projsplx, proj_one_hot!, projsplx!
 #include("./tensorproducts.jl")
 export ×₁, nmp, nmode_product, mtt, slicewise_dot, tuckerproduct, cpproduct
 

### 2025-01-06 src/Core/Core.jl

@@ -16,7 +16,7 @@ using LinearAlgebra: LinearAlgebra, diag
 
 # Basic functionality
 include("./utils.jl")
-export SuperDiagonal, abs_randn, all_recursive, getnotindex, geomean, identityslice, interlace, norm2, proj_one_hot, projsplx, proj_one_hot!, projsplx!
+export SuperDiagonal, abs_randn, all_recursive, eachfibre, getnotindex, geomean, identityslice, interlace, norm2, proj_one_hot, projsplx, proj_one_hot!, projsplx!
 include("./tensorproducts.jl")
 export ×₁, nmp, nmode_product, mtt, slicewise_dot, tuckerproduct, cpproduct
 

### 2025-01-06 src/Core/utils.jl

@@ -86,6 +86,20 @@ getnotindex(A, i::Int) = A[eachindex(A) .!= i]
 getnotindex(A, I) = A[eachindex(A) .∉ (I,)]
 
 """
+    eachfibre(A::AbstractArray; n::Integer, kwargs...)
+
+Creates views of `A` that are that `n`-fibres of `A`.
+
+Shorthand for `eachslice(A; dims=(1,...,n-1,n+1,...,ndims(A)))`.
+
+See `eachslice`.
+"""
+function eachfibre(A::AbstractArray; n::Integer, kwargs...)
+    dims = Tuple([1:n-1; n+1:ndims(A)]) # each integer from 1 to N, excluding n
+    return eachslice(A; dims, kwargs...)
+end
+
+"""
     swapdims(A::AbstractArray, a::Integer, b::Integer=1)
 
 Swap dimensions `a` and `b`.

### 2025-01-06 test/runtests.jl

@@ -60,6 +60,13 @@ const VERBOSE = true
             @test all_recursive(x -> x > 0, collect.([1:10, 2:11, 3:5, [1:3, 2]])) == true
             @test all_recursive(x -> x < 11, [1:10, 2:11, 3:5, [1:3, 2]]) == false
         end
+
+        @testset "eachfibre" begin
+            A = reshape(collect(1:60), 3,4,5)
+            @test eachfibre(A; n=1) == eachslice(A; dims=(2,3))
+            @test eachfibre(A; n=2) == eachslice(A; dims=(1,3))
+            @test eachfibre(A; n=3) == eachslice(A; dims=(1,2))
+        end
     end
 
     @testset "Products" begin

### 2025-01-07 src/Core/utils.jl
make identity_tensor

@@ -53,6 +53,14 @@ function Base.getindex(S::SuperDiagonal, I::Vararg{Int})
     end
 end
 
+"""
+    identity_tensor(N, ndims)
+
+Creates a SuperDiagonal array of ones with size N × ... × N of order `ndims`.
+"""
+identity_tensor(N, ndims) = SuperDiagonal(ones(Int, N), ndims)
+identity_tensor(T, N, ndims) = SuperDiagonal(ones(T, N), ndims)
+
 #######################################################
 
 const DiagonalTuple{N, T} = NTuple{N, Diagonal{T}} # Tuple type where you have a tuple of Diagonal matrices, possibly of varying sizes

### 2025-01-07 paper/paper.qmd
Notation section organize

@@
 
+### Sets
+
+The set of nonnegative numbers is denoted as $\mathbb{R}_+=\mathbb{R}_{\geq 0}=\left\{x\in \mathbb{R} \middle| x\geq 0\right\}$.
+
 We use $[N]=\{1, 2, \dots, N\}=\{n\}_{n=1}^N$ to denote integers from $1$ to $N$.
@@
 
-The set of nonnegative numbers is denoted as $\mathbb{R}_+=\mathbb{R}_{\geq 0}=\left\{x\in \mathbb{R} \middle| x\geq 0\right\}$.
+We use a capital delta $\Delta$ to denote sets of vectors or higher order tensors where the slices or fibres along a specified dimension sum to $1$ i.e. generalized simplexes.
+
+Usually, we use script letters ($\mathcal{A}, \mathcal{B}, \mathcal{C}, etc.$) for other sets.
+
+### Vectors, Matrices, and Tensors
 
@@
 
+The $n$-slices, $n$th mode slices, or mode $n$ slices of an $N$th order tensor $A$ are notated with the slice $A[:,~\dots,~:,~i_{n},~:,~\dots,~:]$.
+For a $3$rd order tensor $A$, the $1$st, $2$nd, and $3$rd mode slices $A[i,:,:]$, $A[:,j,:]$, and $A[:,:,k]$ have special names and are called the horizontal, lateral, and frontal slices and are displayed in @fig-tensor-slices.
+In Julia, the 1-, 2-, and 3-slices of a third order array `A` would be `eachslice(A, dims=1)`, `eachslice(A, dims=2)`, and `eachslice(A, dims=3)`.
+
+![Slices of an order $3$ tensor $A$.](figure/tensor_slices.png){#fig-tensor-slices}
+
 The $n$-fibres, $n$th mode fibres, or mode $n$ fibres of an $N$th order tensor $A$ are denoted $A[i_1,~\dots,~i_{n-1},~:,~i_{n+1},~\dots,~i_{N}]$.
@@
 For order-3 tensors, the $1$st, $2$nd, and $3$rd mode fibres $A[:, j, k]$, $A[i,:,:]$, and $A[i,j,:]$ are called the vertical/column, horizontal/row, and depth/tube fibres respectively and are displayed in @fig-tensor-fibres.
-In Julia, the 1-, 2-, and 3-fibres of a third order array `A` would be `eachslice(A, dims=(2,3))`, `eachslice(A, dims=(1,3))`, and `eachslice(A, dims=(1,2))`.
+Natively in Julia, the 1-, 2-, and 3-fibres of a third order array `A` would be `eachslice(A, dims=(2,3))`, `eachslice(A, dims=(1,3))`, and `eachslice(A, dims=(1,2))`.
+BlockTensorDecomposition.jl defines the function `eachfibre(A; n)` to do exactly this.
+For example, the 1-fibres of an array `A` would be `eachfibre(A, n=1)`.
+
+For matrices, the 1-fibres are the same as the 2-slices (and vice versa), but for $N$th order tensors in general, fibres are always vectors, whereas $n$-slices are $(N-1)$th order tensors.
 
@@
 
-The $n$-slices, $n$th mode slices, or mode $n$ slices of an $N$th order tensor $A$ are notated with the slice $A[:,~\dots,~:,~i_{n},~:,~\dots,~:]$.
-For matrices, the 1-fibres are the same as the 2-slices (and vice versa), but for $N$th order tensors in general, fibres are always vectors, whereas $n$-slices are $(N-1)$th order tensors.
-For a $3$rd order tensor $A$, the $1$st, $2$nd, and $3$rd mode slices $A[i,:,:]$, $A[:,j,:]$, and $A[:,:,k]$ have special names and are called the horizontal, lateral, and frontal slices and are displayed in @fig-tensor-slices.
-In Julia, the 1-, 2-, and 3-slices of a third order array `A` would be `eachslice(A, dims=1)`, `eachslice(A, dims=2)`, and `eachslice(A, dims=3)`.
+Since we commonly use $I$ as the size of a tensor's dimension, we use $\mathrm{id}_I$ to denote the identity tensor of size $I$ (of the appropriate order).
+When the order is $2$, $\mathrm{id}_I$ is an $I\times I$ matrix with ones along the main diagonal, and zeros elsewhere.
+For higher orders $N$, this is an $\underbrace{I\times \cdots \times I}_{N \text{times}}$ tensor where $\mathrm{id}_I[i_1,\dots,i_N]=1$ when $i_1=\dots=i_N\in[I]$, and is zero otherwise.
 
-![Slices of an order $3$ tensor $A$.](figure/tensor_slices.png){#fig-tensor-slices}
+BlockTensorDecomposition.jl defines `identity_tensor`
+### Operations
 
@@
 
-Since we commonly use $I$ as the size of a tensor's dimension, we use $\mathrm{id}_I$ to denote the identity tensor of size $I$ (of the appropriate order).
-When the order is $2$, $\mathrm{id}_I$ is an $I\times I$ matrix with ones along the main diagonal, and zeros elsewhere.
-For higher orders $N$, this is an $\underbrace{I\times \cdots \times I}_{N \text{times}}$ tensor where $\mathrm{id}_I[i_1,\dots,i_N]=1$ when $i_1=\dots=i_N\in[I]$, and is zero otherwise.
+Julia's standard library package LinearAlgebra implements the Frobenius inner product with `dot(A, B)` or `A ⋅ B`.
 

### 2025-01-08 paper/paper.qmd
First pass of section Tensor Decompositions

@@
 
-The set of nonnegative numbers is denoted as $\mathbb{R}_+=\mathbb{R}_{\geq 0}=\left\{x\in \mathbb{R} \middle| x\geq 0\right\}$.
+The set of real number is denoted as $\mathbb{R}$ and its restrictions to nonnegative numbers is denoted as $\mathbb{R}_+=\mathbb{R}_{\geq 0}=\left\{x\in \mathbb{R} \middle| x\geq 0\right\}$.
 
@@
 
-We use a capital delta $\Delta$ to denote sets of vectors or higher order tensors where the slices or fibres along a specified dimension sum to $1$ i.e. generalized simplexes.
+We use a capital delta $\Delta$ to denote sets of vectors or higher order tensors where the slices or fibres along a specified dimension sum to $1$, i.e. generalized simplexes.
 
-Usually, we use script letters ($\mathcal{A}, \mathcal{B}, \mathcal{C}, etc.$) for other sets.
+Usually, we use script letters ($\mathcal{A}, \mathcal{B}, \mathcal{C},$ etc.) for other sets.
 
@@
 
-BlockTensorDecomposition.jl defines `identity_tensor`
+BlockTensorDecomposition.jl defines `identity_tensor(I, ndims)` to construct $\mathrm{id}_I$.
+
 ### Operations
@@
 
-## Common Decompositions
+The $n$-slice dot product $\cdot_n$ between two tensors $A\in\mathbb{R}^{I_{1},\dots, I_{n-1},J,I_{n+1},\dots, I_{N}}$ and $B\in\mathbb{R}^{I_{1},\dots, I_{n-1},K,I_{n+1},\dots, I_{N}}$ returns a matrix $(A \cdot_n B)\in\mathbb{R}^{J\times K}$ with entries
+
+$$
+(A \cdot_n B)[j, k] = {\sum_{i_{1}\dots i_{n-1}i_{n+1}\dots i_{N}}} A[i_{1},\dots, i_{n-1},j,i_{n+1},\dots, i_{N}]B[i_{1},\dots, i_{n-1},k,i_{n+1},\dots, i_{N}].
+$$
+
+This product can also be thought of as taking the dot product between all pairs of $n$th order slices of $A$ and $B$: $(A \cdot_n B)[j, k] = A_j \cdot B_k$.
+
+BlockTensorDecomposition.jl defines this operation with `slicewise_dot(A, B, n)`.
+In the special case where $A=B$, a more efficient method that only computes entries where $i \leq j$ is defined since $A \cdot_n A$ is a symmetric matrix.
+
+The $n$-mode product $\times_n$ between a tensor $A\in\mathbb{R}^{I_1 \times \dots \times I_N}$ and matrix $B\in\mathbb{R}^{I_n\times J}$, returns a tensor $(A \times_n B)\in\mathbb{R}^{I_1 \times \dots \times I_{n-1} \times J \times I_{n+1} \times \dots \times I_N}$ with entries
+
+$$
+(A \times_n B)[i_1, \dots, i_{n-1},j,i_{n+1},\dots,i_N] = \sum_{i_{n}=1}^{I_n} A[i_1, \dots, i_{n-1},i_n,i_{n+1},\dots,i_N]B[i_n,j].
+$$
+
+BlockTensorDecomposition.jl defines this operation with `nmode_product(A, B, n)`.
+
+```julia
+function nmode_product(A::AbstractArray, B::AbstractMatrix, n::Integer)
+    # convert the problem to the mode-1 product
+    Aperm = swapdims(A, n)
+    Cperm = Aperm ×₁ B
+    return swapdims(Cperm, n) # swap back
+end
+
+function ×₁(A::AbstractArray, B::AbstractMatrix)
+    # Turn the 1-mode product into matrix-matrix multiplication
+    sizeA = size(A)
+    Amat = reshape(A, sizeA[1], :)
+
+    # Initialize the output tensor
+    C = zeros(size(B)[1], sizeA[2:end]...)
+    Cmat = reshape(C, size(B)[1], prod(sizeA[2:end]))
+
+    # Perform matrix-matrix multiplication Cmat = B*Amat
+    mul!(Cmat, B, Amat)
+
+    return C # Output entries of Cmat in tensor form
+end
+
+function swapdims(A::AbstractArray, a::Integer, b::Integer=1)
+    # Construct a permutation where a and b are swapped
+    # e.g. [4, 2, 3, 1, 5, 6] when a=4 and b=1
+    dims = collect(1:ndims(A))
+    dims[a] = b; dims[b] = a
+    return permutedims(A, dims)
+end
+```
+
+The Frobenius norm of a tensor $A$ is the square root of its dot product with itself
+
+$$
+\left\lVert A \right\rVert_F = \sqrt{\left\langle A, A\right\rangle}.
+$$
+
+For vectors $v$, this is equivalent to the (Euclidean) 2-norm
+
+$$
+\left\lVert v \right\rVert_F = \left\lVert v \right\rVert_2 = \sqrt{\left\langle v, v\right\rangle}.
+$$
+
+For matrices $M$, the (Operator) 2-norm is defined as
+
+$$
+\left\lVert M \right\rVert_2 = \arg\,\max_{\left\lVert v \right\rVert_2 = 1} \left\lVert Mv \right\rVert_2 = \sigma_1(M)
+$$
+
+where $\sigma_1(M)$ is the largest singular value of $M$.
+
+For tensors $T$, the (Operator) 2-norm needs to be defined in terms of how we treat them as function on other tensors.
+There is a canonical way to do this for vectors $x \to v^{\top}x$ and matrices $x \to Mx$, but not tensors.
+This is relevant to @sec-lipschitz-computation where the Lipschitz step-size is computed in terms of the Operator norm of the Hessian of our objective function.
+
+## Common Decompositions {#sec-common-decompositions}
 - Extensions of PCA/ICA/NMF to higher dimensions
@@
 
-<!-- From my thesis proposal: -->
+A tensor decomposition is a factorization of a tensor into multiple (usually smaller) tensors, that can be recombined into the original tensor.
+Computationally, we can think of a generic decomposition as storing factors $(A,B,C,...)$ and operations $(\times_a,\times_b,...)$ for combining them.
+This is what we do in BlockTensorDecomposition.jl.
 
-<!-- This section gives a brief tour of the most common tensor factorizations, heavily relying on the framework and notation given by Kolda and Bader [@kolda_TensorDecompositionsApplications_2009].
-They are, however, presented in my own interpretation as matrix factorization generalizations and as different instances of the Tucker factorization.
+```julia
+struct GenericDecomposition{T, N} <: AbstractArray{T, N}
+    factors::Tuple{Vararg{AbstractArray{T}}} # e.g. (A, B, C)
+    contractions::Tuple{Vararg{Function}} # e.g. (×₁, ×₂)
+end
+# Y = A ×₁ B ×₂ C
+array(G::GenericDecomposition) = multifoldl(contractions(G), factors(G))
+```
 
-One way we can extend matrix factorization to multidimensional signals is with a *Tucker-1* factorization.
-Here, each mixture is an $N$th order tensor $Y_i\in\mathbb{R}^{J_1\times \dots\times J_N}$ stored in the $1$-slices of an order-$(N+1)$ tensor $Y\in\mathbb{R}^{I\times J_1\times \dots\times J_N}$, $Y[i, :, \dots, :]=Y_i$, and similarly for $B$.
-Slice-wise, we have the equation
+The function `multifoldl` applies the given operations between each factor, from left to right.
+
+```julia
+function multifoldl(ops, args)
+    @assert (length(ops) + 1) == length(args)
+    x = args[begin]
+    for (op, arg) in zip(ops, args[begin+1:end])
+        x = op(x, arg)
+    end
+    return x
+end
+```
+
+Different types of decompositions define different operations, and different "ranks" of the same decomposition specific the sizes of the factors used.
+
+A commonly used family of decompositions can be derived from the Tucker decomposition.
+
+::: {#def-tucker-decomposition}
+A rank-$(R_1,\dots,R_N)$ Tucker decomposition of a tensor $Y\in\mathbb{R}^{I_1\times\dots\times I_N}$ produces $N$ matrices $A_n\in\mathbb{R}^{I_n\times R_n}$, $n\in[N]$, and core tensor $B\in\mathbb{R}^{R_1\times\dots\times R_N}$ such that
 
 $$
-Y_i = \sum_{r=1}^R A[i,r] B_r,
-$$ {#eq-tucker-1-slices}
+Y[i_1,\dots,i_N] = \sum_{r_1=1}^{R_1} \dots \sum_{r_N=1}^{R_N} A_1[i_1,r_1]\cdots A_r[i_N,r_N] B[r_1,\dots,r_N]
+$$ {#eq-tucker}
 
-which is entry-wise
+entry-wise. More compactly, this decomposition can be written using the $n$-mode product, or with double brackets
 
 $$
-Y[i, j_1, \dots, j_N] = \sum_{r=1}^R A[i,r] B[r,j_1, \dots, j_N].
-$$ {#eq-tucker-1}
-
-![Example of a rank $R=3$ Tucker-1 factorization coloured according to @eq-tucker-1-slices.
-Here we have $Y\in\mathbb{R}^{5\times J \times K}, A\in\mathbb{R}^{5\times 3}$, and $B\in\mathbb{R}^{3\times J \times K}$.
-The first horizontal slice of $Y$, $Y_1$, is equal to $A[1,1]B_1+A[1,2]B_2+A[1,3]B_3$.](figure/tucker_1_example.png){#fig-tucker-1-example}
-
-Seeing @eq-tucker-1-slices alongside @fig-tucker-1-example makes it clear that the decomposition is performing a compression along the first axis of $Y$.
-This makes sense in the model where the mixtures are just linear combinations of sources with the same size and dimensions.
-
-In the case where the sources can be further factored along other axes, specifically the first $n$ dimensions, we can use the more general *Tucker-$n$* factorization of an order $N$ tensor $Y$,
-
-$$
-\begin{split}
-Y[i_1,\dots,i_n, & j_1, \dots, j_{N-n+1}] = \\ &\sum_{r_1=1}^{R_1} \dots \sum_{r_n=1}^{R_n} A_1[i_1,r_1]\cdots A_r[i_n,r_n] B[r_1,\dots,r_n, j_1, \dots, j_{N-n+1}].
-\end{split}
-$$ {#eq-tucker-n}
-
-In the case where $n=N$ is the order of the tensor $Y$, we omit the $n$ and call this a *Tucker* factorization. So a Tucker factorization of an order $N$ tensor $Y$ is
-
-$$
-Y[i_1,\dots,i_N] = \sum_{r_1=1}^{R_1} \dots \sum_{r_n=1}^{R_n} A_1[i_1,r_1]\cdots A_r[i_n,r_N] B[r_1,\dots,r_N].
-$$ {#eq-tucker}
-
-All the Tucker decompositions can be written using the $n$-mode product $\times_n$ between a tensor $A$ and matrix $B$, defined as
-
-$$
-(A \times_n B)[i_1, \dots, i_{n-1},j,i_{n+1},\dots,N] = \sum_{i_{n}=1}^{I_n} A[i_1, \dots, i_{n-1},i_n,i-{n+1},\dots,N]B[i_n,j].
+Y = B\times_1 A_1 \times_2 \dots\times_N A_N = B \bigtimes_{n} A_n = \llbracket B; A_1, \dots, A_N \rrbracket.
 $$
 
-They can also be written with an $n$ product shorthand, or with bracket notation. For example, the full tucker decomposition shown in @eq-tucker can be written as
+:::
 
-$$
-\begin{align*}
-Y &= B\times_1 A_1 \times_2 \dots\times_N A_N = B \bigtimes_n A_n = \llbracket B; A_1, \dots, A_N \rrbracket \\
-\text{or}\quad Y &= A_0\times_1 \dots\times_N A_N = A_0 \bigtimes_n A_n = \llbracket A_0; A_1, \dots, A_N \rrbracket.
-\end{align*}
-$$
-
-To ease notation further, we sometimes consider the *core* tensor $B$ to be the zeroth factor $A_0=B$ so we can say $\{A_n\}_{n=0}^N$ are the factors of $Y$.
-Since it can be hard to visualize the product, consider the case of a Tucker factorization of a third order tensor $Y$:
-
-$$
-Y[i,j,k] = \sum_{r=1}^R\sum_{s=1}^S\sum_{t=1}^T A_1[i,r] A_2[j,s] A_3[k,t] B[r,s,t].
-$$
-
-We can picture this as multiplying the core factor $B$ with matrix factors $A_1, A_2, A_3$ on all sizes to construct $Y$ as shown in @fig-tucker.
+Sometimes we write $A_0=B$ to ease notation, and suggest the "zeroth" factor of the tucker decomposition is the core tensor $B$.
+In the special case when $N=3$, we can visualize Tucker decomposition as multiplying the core tensor by matrices on all three sides as shown in @fig-tucker.
 
@@
 
-The Tucker-$n$ factorization of an order $N$ tensor can be thought of as a special case of the (full) Tucker factorization, where the matrix factors $A_{n+1}, \dots, A_{N}$ are identity matrices.
+Setting all the matrices of a Tucker decomposition to the identity matrix but the first gives the Tucker-$1$ decomposition.
 
-In the case where the core tensor $B\in\mathbb{R}^{R_1\times \dots\times R_N}$ of the Tucker factorization is cubic ($R_1=\dots=R_N$) and diagonal ($B[r_1,\dots,r_N]=\lambda_r$ if $r_1=\dots=r_N=r$ and $0$ otherwise), we have the **can**onical **decomp**osition/**para**llel **fac**tors model (CANDECOMP/PARAFAC or CP for short):
+::: {#def-tucker-1-decomposition}
+A rank-$R$ Tucker-$1$ decomposition of a tensor $Y\in\mathbb{R}^{I_1\times\dots\times I_N}$ produces a matrix $A\in\mathbb{R}^{I_1\times R}$, and core tensor $B\in\mathbb{R}^{R\times I_2\times\dots\times I_N}$ such that
 
 $$
-Y[i_1,\dots,i_N] = \sum_{r=1}^{R} \lambda_r A_1[i_1,r]\cdots A_N[i_N,r].
+Y[i_1,\dots,i_N] = \sum_{r=1}^{R} A[i_1,r] B[r, i_2,\dots,i_N]
+$$ {#eq-tucker-1}
+
+entry-wise or more compactly,
+
+$$
+Y = AB = B\times_1 A = \llbracket B; A \rrbracket.
 $$
 
-This can be written as
+:::
+
+Note we extend the usual definition of matrix-matrix multiplication
 
 $$
-Y = \llbracket \lambda; A_1, \dots, A_N\rrbracket.
+(AB)[i,j] = \sum_{r=1}^R A[i,r] B[r, j]
 $$
 
-Sometimes the factor $\lambda_r$ is scaled so that the columns $A[:,r]$ are normalized $\lVert A[:,r]\rVert_2=1$ for each $r\in[R]$, or it is absorbed into the matrix factors and omitted from the decomposition.
-For the remainder of this text, I'll use CP decomposition to mean the latter case where $\lambda_r=1$ for all $r\in[R]$, as showing in @eq-cp:
+to tensors $B$ in the compact notation for Tucker-1 decomposition $Y=AB$.
+
+More generally, any number of matrices can be set to the identity matrix giving the Tucker-$n$ decomposition.
+
+::: {#def-tucker-n-decomposition}
+A rank-$(R_1,\dots,R_n)$ Tucker-$n$ decomposition of a tensor $Y\in\mathbb{R}^{I_1\times\dots\times I_N}$ produces $n$ matrices $A_1,\dots,A_n$, and core tensor $B\in\mathbb{R}^{R_1\times\dots\times R_n \times I_{n+1}\times\dots\times I_N}$ such that
 
 $$
-Y[i_1,\dots,i_N] = \sum_{r=1}^{R} A_1[i_1,r]\cdots A_N[i_N,r].
+Y[i_1,\dots,i_N] = \sum_{r_1=1}^{R_1} \dots \sum_{r_N=1}^{R_n} A_1[i_1,r_1]\cdots A_n[i_N,r_n] B[r_1,\dots,r_n,i_{n+1},\dots,i_{N}]
+$$ {#eq-tucker-n}
+
+entry-wise, or compactly written in the following three ways,
+$$
+\begin{align}
+Y &= B\times_1 A_1 \times_2 \dots\times_n A_n \times_{n+1} \mathrm{id}_{I_{n+1}} \times_{n+2} \dots \times_{N} \mathrm{id}_{I_{N}}\\
+Y &= B\times_1 A_1 \times_2 \dots\times_n A_n\\
+Y &= \llbracket B; A_1, \dots, A_n \rrbracket.
+\end{align}
+$$
+
+:::
+
+Lastly, if we set the core tensor $B$ to the identity tensor $\mathrm{id}_{R}$, we obtain the **can**onical **decomp**osition/**para**llel **fac**tors model (CANDECOMP/PARAFAC or CP for short).
+
+::: {#def-cp-decomposition}
+A rank-$R$ CP decomposition of a tensor $Y\in\mathbb{R}^{I_1\times\dots\times I_N}$ produces $N$ matrices $A_n\in\mathbb{R}^{I_n\times R}$, such that
+
+$$
+Y[i_1,\dots,i_N] = \sum_{r=1}^{R} A_1[i_1,r]\cdots A_r[i_N,r]
 $$ {#eq-cp}
 
+entry-wise. More compactly, this decomposition can be written using the $n$-mode product, or with double brackets
+
+$$
+Y = \mathrm{id}_{R} \times_1 A_1 \times_2 \dots\times_N A_N = \mathrm{id}_{R} \bigtimes_{n} A_n = \llbracket A_1, \dots, A_N \rrbracket.
+$$
+
+:::
+
+Note CP decomposition is sometimes referred to as Kruskal decomposition, and requires the core only be diagonal (and not necessarily identity) and the factors $A_n$ have normalized columns $\left\lVert A_n[:,r]\right\rVert_2 =1$.
+
 Other factorization models are used that combine aspects of CP and Tucker decomposition [@kolda_TensorDecompositionsApplications_2009], are specialized for order $3$ tensors [@qi_TripleDecompositionTensor_2020; @wu_manifold_2022], or provide alternate decomposition models entirely like tensor-trains [@oseledets_tensor-train_2011].
-But the (full) Tucker, and its special cases Tucker-$n$, and CP decomposition are most commonly used extensions of the low-rank matrix factorization.
+But the (full) Tucker, and its special cases Tucker-$n$, and CP decomposition are most commonly used extensions of the low-rank matrix factorization to tensors.
 These factorizations are summarized in @tbl-tensor-factorizations.
@@
 |------------|----------------------------------------------|----------------------------------------------------------|--------------|
+| Tucker     | $\llbracket A_0; A_1, \dots, A_N \rrbracket$ | $A_0\times_1 A_1 \times_2 \dots\times_N A_N$             | @eq-tucker   |
 | Tucker-$1$ | $\llbracket A_0; A_1 \rrbracket$             | $A_0\times_1 A_1$                                        | @eq-tucker-1 |
 | Tucker-$n$ | $\llbracket A_0; A_1, \dots, A_n \rrbracket$ | $A_0\times_1 A_1 \times_2 \dots\times_n A_n$             | @eq-tucker-n |
-| Tucker     | $\llbracket A_0; A_1, \dots, A_N \rrbracket$ | $A_0\times_1 A_1 \times_2 \dots\times_N A_N$             | @eq-tucker   |
 | CP         | $\llbracket A_1, \dots, A_N \rrbracket$      | $\mathrm{id}_{R}\times_1 A_1 \times_2 \dots\times_N A_N$ | @eq-cp       |
@@
 : Summary of common tensor factorizations. Here, $N$ is the order of the factorized tensor. {#tbl-tensor-factorizations tbl-colwidths="[15,25,35,25]"}
- -->
+
+TODO add discussion on other decompositions
+- high order SVD (see Kolda and Bader)
+- HOSVD (see Kolda, Shifted power method for computing tensor eigenpairs)
+
+Tensor decompositions are not nessisarily unique.
+It should be clear that scaling one factor by $x\neq 0$ and dividing another by $x$ yields the same original tensor.
+Furthermore, fibres and slices can be permuted without affecting the the original tensor.
+Up to these manipulations, for a fixed rank, there exist criteria that ensures their decompositions are unique [@kolda_TensorDecompositionsApplications_2009;@kruskal_three-way_1977;@bhaskara_uniqueness_2014].
+
+### Representing Tucker Decompositions
+
+There are implemented in BlockTensorDecomposition.jl and can be called, for a third order tensor, with `Tucker((B, A₁, A₂, A₃))`, `Tucker1((B, A₁))`, and `CPDecomposition((A₁, A₂, A₃))`.
+These Julia `structs` store the tensor in its factored form.
+If the recombined tensor or particular entries are requested, Julia dispatches on the type of decomposition and calls a particular method of `array` or `getindex`.
+The implementations for efficient array construction and index access are provided below.
+
+```julia
+array(T::Tucker) = multifoldl(tucker_contractions(ndims(T)), factors(T))
+tucker_contractions(N) = Tuple((G, A) -> nmode_product(G, A, n) for n in 1:N)
+```
+
+TODO add getindex method for Tucker type
+
+```julia
+function array(T::Tucker1)
+    B, A = factors(T)
+    return B ×₁ A
+end
+
+function getindex(T::Tucker1, I::Vararg{Int})
+    B, A = factors(T)
+    i = I[1]
+    J = I[begin+1:end] # J = (I₂, I₃, ..., I_N)
+    return (@view A[i, :]) ⋅ view(B, :, J...)
+end
+```
+
+```julia
+array(CPD::CPDecomposition) =
+  mapreduce(vector_outer, +, zip((eachcol.(factors(CPD)))...))
+
+vector_outer(v) = reshape(kron(reverse(v)...),length.(v))
+
+getindex(CPD::CPDecomposition, I::Vararg{Int}) =
+  sum(reduce(.*, (@view f[i,:]) for (f,i) in zip(factors(CPD), I)))
+```
 
@@
 - constrained rank (nonnegative etc.)
+
+The rank of a matrix $Y\in\mathbb{R}^{I\times J}$ can be defined as the smallest $R\in\mathbb{Z}_+$ such that there exists an exact factorization $Y=AB$ for some $A\in\mathbb{R}^{I\times R}$ and $B\in\mathbb{R}^{R\times J}$.
+
+Although this can be extended to higher order tensors, we must specify under which factorization model we are using.
+For example, the *CP-rank* $R$ of a tensor $Y$ is the smallest such $R$ that omits an exact CP decomposition of $Y$.
+
+::: {#def-cp-rank}
+The CP rank of a tensor $Y\in\mathbb{R}^{I_1\times \dots \times I_N}$ is the smallest $R$ such that there exist factors $A_n\in\mathbb{R}^{I_n \times R}$ and $Y = \llbracket A_1, \dots, A_N\rrbracket$,
+$$
+\text{rank}_\text{CP} (Y) = \min \left\{R \middle| \exists A_n\in\mathbb{R}^{I_n \times R},\, n\in[N] \quad \text{s.t.}\quad Y = \llbracket A_1, \dots, A_N\rrbracket \right\}
+$$
+
+:::
+
+In a similar way, we can define the *Tucker-1-rank* $R$.
+
+::: {#def-tucker-1-rank}
+The Tucker-1 rank of a tensor $Y\in\mathbb{R}^{I_1\times \dots \times I_N}$ is the smallest $R$ such that there exist factors $A\in\mathbb{R}^{I_1 \times R}$ and $B\in\mathbb{R}^{R \times I_2 \times \dots \times I_N}$ where $Y = AB$
+
+$$
+\text{rank}_\text{Tucker-1} (Y) = \min \left\{R \middle| \exists A_n\in\mathbb{R}^{I_n \times R},B\in\mathbb{R}^{R \times I_2 \times \dots \times I_N}\, \quad \text{s.t.}\quad Y = AB\right\}
+$$
+
+:::
+
+For the Tucker and Tucker-$n$ decompositions, we instead call a particular factorization **a** rank-$(R_1,\dots,R_N)$ Tucker factorization or **a** rank-$(R_1,\dots,R_n)$ Tucker-$n$ factorization, rather than **the** CP- or Tucker-$1$-rank of a tensor or **the** rank of a matrix.
+
+One reason CP and Tucker-$1$ only need a single rank $R$ can be explained by considering the case when the order of the tensor $N=2$ (matrices).
+The two factorizations become equivalent and are equal to low-rank $R$ matrix factorization $Y=AB$.
+In fact, Tucker-$1$ is always equivalent to a low-rank matrix factorization, if you consider a flattening of the tensor to arrange the entries as a matrix.
+
+The idea of tensor rank can be generalized further to constrained rank. These are the smallest rank $R$ such that the factors in the decomposition obey the given set of constraints.
+
+For example, the nonnegative Tucker-1 rank is defined as
+$$
+\text{rank}_{\text{Tucker-1}}^+ (Y) = \min \left\{R \middle| \exists A_n\in\mathbb{R}_+^{I_n \times R},B\in\mathbb{R}_+^{R \times I_2 \times \dots \times I_N}\, \quad \text{s.t.}\quad Y = AB\right\}.
+$$
+
+More restrictive constraints increase the rank of the tensor since there is less freedom in selecting the factors.
 
@@
   - do *not* use alternating least squares (slower for unconstrained problems, no closed form update for general constrained problems)
+
+### Computing Gradients
+- Use Auto diff generally
+- But hand-crafted gradients and Lipschitz calculations *can* be faster (e.g. symmetrized slicewise dot product)
+
+### Computing Lipschitz Step-sizes {#sec-lipschitz-computation}
 

### 2025-01-09 paper/paper.qmd
start on computing decomposition section

@@
 $$
-\text{rank}_\text{CP} (Y) = \min \left\{R \middle| \exists A_n\in\mathbb{R}^{I_n \times R},\, n\in[N] \quad \text{s.t.}\quad Y = \llbracket A_1, \dots, A_N\rrbracket \right\}
+\text{rank}_\text{CP} (Y) = \min \left\{R \middle| \exists A_n\in\mathbb{R}^{I_n \times R},\, n\in[N] \quad \text{s.t.}\quad Y = \llbracket A_1, \dots, A_N\rrbracket \right\}.
 $$
@@
 
+Most tensor decomposition algorithms require the rank as input [CITE] since calculating the rank of the tensor can be NP-hard in general [@vavasis_complexity_2010].
+For applications where the rank is not known a priori, a common strategy is to attempt a decomposition for a variety of ranks, and select the model with smallest rank that still achieves good fit between the factorization and the original tensor.
+
 # Computing Decompositions
 - Given a data tensor and a model, how do we fit the model?
+
+Many tensor decompositions algorithms exist in the literature.
+Usually, they cyclically (or in a random order) update factors until their reconstruction satisfies some convergence criterion.
+The base algorithm described in @sec-base-algorithm provides flexible framework for wide class of constrained tensor factorization problems.
+This framework was selected based on empirical observations where it outperforms other similar algorithms, and has also been observed in the literature  [@xu_BlockCoordinateDescent_2013].
 
@@
 
-## Base algorithm
+## Base algorithm {#sec-base-algorithm}
 - Use Block Coordinate Descent / Alternating Proximal Descent
@@
 
-## Partial Projection and Rescaling
+# Partial Projection and Rescaling
 - for bounded linear constraints
@@
 
-## Multi-scale
+# Multi-scale
 - use a coarse discretization along continuous dimensions

### 2025-01-10 src/Core/factorize.jl
spelling error

@@ -9,7 +9,7 @@ factorize(Y; kwargs...) =
 	_factorize(Y; (default_kwargs(Y; kwargs...))...)
 
 """
-Inner level function once keyword agruments are set
+Inner level function once keyword arguments are set
 """
 function _factorize(Y; kwargs...)
 	decomposition, kwargs = initialize_decomposition(Y; kwargs...)

### 2025-01-10 paper/paper.qmd
Start Optimization Problem section

@@
 
+Ideally, we would be given a data tensor $Y$ and decomposition model, and compute an exact factorization of $Y$ into its factors.
+Because there is often measurement, numerical, or modeling error, an exact factorization of $Y$ for a particular rank may not exist.
+To over come this, we instead try to fit the model to the data.
+Let $X$ be the reconstruction of factors $A_1,\dots,A_N$ according to some decomposition for a fixed rank.
+We assume we know the size of the factors $A_1,\dots,A_N$ and how they are combined to produce a tensor the same size of $Y$, i.e.
+the map $g: (A_1,\dots,A_N) \mapsto X$.
+
+There are many loss functions that can be used to determine how close the model $X$ is to the data $Y$.
+In principle, any distance or divergence $d(Y,X)$ could be used.
+We use the $L_2$ loss or least-squares distance between the tensors $\left\lVert X - Y\right\rVert _F^2$, but other losses are used for tensor decomposition in practice such as the KL divergence [CITE].
+
+The main optimization we must solve is now given.
+
+::: {#def-constrained-least-squares}
+
+The constrained least-squares tensor factorization problem is to solve
+
+$$
+\min_{A_1, \dots, A_N} \left\lVert g(A_1,\dots,A_N) - Y\right\rVert _F^2 \quad \text{s.t.} \quad (A_1,\dots,A_N) \in\mathcal{C}_1 \times \dots \times \mathcal{C}_N
+$$ {#eq-constrained-least-squares}
+
+for a given data tensor $Y$, constraints $\mathcal{C}_1, \dots, \mathcal{C}_N$, and decomposition model $g$ with fixed rank.
+
+:::
+
 ## Base algorithm {#sec-base-algorithm}
@@
   - do *not* use alternating least squares (slower for unconstrained problems, no closed form update for general constrained problems)
+
+Let $f(A_1, \dots, A_N) := \left\lVert g(A_1,\dots,A_N) - Y\right\rVert _F^2$ be the objective function we wish to minimize in @eq-constrained-least-squares.
+Following Xu and Yin [@xu_BlockCoordinateDescent_2013], the general approach we take to minimize $f$ is to apply block coordinate descent using each factor as a different block.
+Let $A_n^t$ be the $t$th iteration of the $n$th factor, and let
+
+$$
+f_n^t(A_n) := \left\lVert g(A_{1}^{t+1}, \dots, A_{n-1}^{t+1}, A_n, A_{n+1}^t, \dots , A_{N}^t)- Y \right\rVert _F^2
+$$
+
+be the (partially updated) objective function at iteration $t$ for factor $n$.
+
+Given initial factors $A_1^0,\dots,A_N^0$, we cycle through the factors $n\in[N]$ and perform the update
+
+$$
+A_n^{t+1} \leftarrow \arg\min_{A_n\in\mathcal{C}_n} \left\langle \nabla f_n^t(A_n^{t}), A_n - A_n^{t} \right\rangle + \frac{L_n^t}{2}\left\lVert A_n - A_n^{t} \right\rVert_F^2,
+$$
+
+for $t=1,2,\dots$ until some convergence criterion is satisfied (see @sec-convergence-criteria).
+
+This implicit update has the *projected gradient descent* closed form solution for convex constraints $\mathcal{C}_n$,
+
+$$
+A_n^{t+1} \leftarrow P_{\mathcal{C}}_n\left( A_n^{t} - \frac{1}{L_n^t}\nabla f_n^t(A_n^{t})\right).
+$$ {#eq-proximal-explicit}
+
+We typically choose $L_n^t$ to be the Lipschitz constant of $\nabla f_n^t$, since it is a sufficient condition to guarantee $f_n^t(A_n^{t+1}) \leq f_n^t(A_n^{t})$, but other step sizes can be used in theory [@nesterov_NonlinearOptimization_2018, Sec.
+1.2.3].
+
+?ASIDE?
+To write $\nabla f_n^t$, we have assumed (block) differentiability of the decomposition model $g$.
+In practice, most decompositions are "block-linear" (freeze all factors but one and you have a linear function) and in rare cases are "block-affine".
+"block-affine" is enough to ensure $f_n^t$ is convex (i.e.
+$f$ is "block-convex") so the updates @eq-proximal-explicit converge to a Nash equilibrium (block minimizer).
+
+### High level code
+
+To ensure the code stays flexible, the main algorithm of BlockTensorDecomposition.jl, `factorize` is defined at a very high level.
+
+<!--
+```julia
+factorize(Y; kwargs...) =
+    _factorize(Y; (default_kwargs(Y; kwargs...))...)
+
+"""
+Inner level function once keyword arguments are set
+"""
+function _factorize(Y; kwargs...)
+    decomposition, kwargs = initialize_decomposition(Y; kwargs...)
+    previous, updateprevious! = initialize_previous(decomposition, Y; kwargs...)
+    parameters, updateparameters! =
+        initialize_parameters(decomposition, Y, previous; kwargs...)
+
+    # more kwargs get added here, so need to return kwargs
+    update!, kwargs = make_update!(decomposition, Y; kwargs...)
+
+    stats_data, getstats =
+        initialize_stats(decomposition, Y, previous, parameters; kwargs...)
+
+    converged = make_converged(; kwargs...)
+
+    # freeze the kwargs from a Dictionary to a NamedTuple for type stability
+    kwargs = NamedTuple(kwargs)
+
+    while !converged(stats_data; kwargs...)
+        # Update the decomposition
+        # This is one cycle of updates on each factor in the decomposition
+        update!(decomposition; parameters...)
+
+        # Update parameters used for the next update of the decomposition
+        # this could be the next stepsize or other info used by update!
+        updateparameters!(parameters, decomposition, previous)
+
+        # Save stats
+        push!(stats_data,
+            getstats(decomposition, Y, previous, parameters, stats_data))
+
+        # Update the one or two previous iterates
+        # We do not save every iterate since that is too much memory
+        updateprevious!(previous, parameters, decomposition)
+    end
+
+    # Perform one final constrain if requested
+    if kwargs[:constrain_output]
+        kwargs = finalconstrain!(decomposition; kwargs...)
+        updateparameters!(parameters, decomposition, previous)
+        push!(stats_data,
+            getstats(decomposition, Y, previous, parameters, stats_data))
+    end
+
+    return decomposition, stats_data, kwargs
+end
+```
+
+Abstracting further, we can see the general structure. -->
+
+```julia
+function factorize(Y; kwargs...)
+    decomposition, converged, update!, updateprevious!, updateparameters!, stats_data, getstats, kwargs = initialize(Y; kwargs...)
+
+    while !converged(stats_data; kwargs...)
+        update!(decomposition; parameters...)
+
+        updateparameters!(parameters, decomposition, previous)
+
+        push!(stats_data,
+            getstats(decomposition, Y, previous, parameters, stats_data))
+
+        updateprevious!(previous, parameters, decomposition)
+    end
+
+    postprocess!(decomposition, parameters, stats_data, Y; kwargs...)
+
+    return decomposition, stats_data, kwargs
+end
+```
+
+The magic of the code is in defining the functions at runtime for a particular decomposition requested, from a reasonable set of default keyword arguments.
+This is discussed further in @sec-flexibility.
 
@@
 
-# Techniques for speeding up convergences
+# Computational Techniques
+
+## For Improving Convergence Speed
 - As stated, algorithm works
@@
 
-## Sub-block Descent
+### Sub-block Descent
 - Use smaller blocks, but descent in parallel (sub-blocks don't wait for other sub-blocks)
@@
 
-## Momentum
+### Momentum
 - This one is standard
@@
 - This is compatible with sub-block descent with appropriately defined matrix operations
+
+## For Flexibility {#sec-flexibility}
+- there are a number of software engineering techniques used
+- these help flexibility for hot swapping and a language for making custom...
+  - convergence criterion (and having multiple stopping conditions)
+  - probing info during the iterations (stats collected at the end)
+  - having multiple constraints and ways to enforce them
+  - cyclically or partially randomly or fully randomly update factors
+- smart enough to apply these in a reasonable order
+
+### Convergence Criteria and Stats {#sec-convergence-criteria}
+- Can request info about any factor at each outer iteration
+- any subset of stats can be the convergence criteria
+
+### `BlockUpdate` Language
+- construct the updates as a list of updates
+- very functional programming
+- can apply them in sequence or in a random order (or partially random)
+
+### Constraints
+- one type of update (other than the typical GD update)
+- can combine them with composition
+  - which is different than projecting onto their intersection!
+- Constraint updates combine the constraint with how they are enforced
+  - need to go together since there are multiple ways to enforce them e.g. simplex (see next section)
 

### 2025-01-10 src/Core/factorize.jl
attempt to abstract _factorize further

@@ -8,10 +8,81 @@ See [`default_kwargs`](@ref) for the default keywords.
 factorize(Y; kwargs...) =
 	_factorize(Y; (default_kwargs(Y; kwargs...))...)
 
+# """
+# Inner level function once keyword arguments are set
+# """
+# function _factorize(Y; kwargs...)
+# 	decomposition, kwargs = initialize_decomposition(Y; kwargs...)
+# 	previous, updateprevious! = initialize_previous(decomposition, Y; kwargs...)
+# 	parameters, updateparameters! = initialize_parameters(decomposition, Y, previous; kwargs...)
+
+# 	# one pass of the constraints is possibly applied so note decomposition could be mutated
+# 	# and more kwargs get added here, so need to return kwargs
+# 	update!, kwargs = make_update!(decomposition, Y; kwargs...)
+
+# 	stats_data, getstats = initialize_stats(decomposition, Y, previous, parameters; kwargs...)
+# 	#@show stats_data
+# 	converged = make_converged(; kwargs...)
+
+# 	kwargs = NamedTuple(kwargs) # freeze the kwargs from a Dictionary to a NamedTuple for type stability
+
+# 	while !converged(stats_data; kwargs...)
+# 		# Update the decomposition
+# 		# This is possibly one cycle of updates on each factor in the decomposition
+# 		update!(decomposition; parameters...)
+
+# 		# Update parameters used for the next update of the decomposition
+# 		# this could be the next stepsize or other info used by update!
+# 		updateparameters!(parameters, decomposition, previous)
+
+# 		# Save stats
+# 		push!(stats_data, getstats(decomposition, Y, previous, parameters, stats_data))
+
+# 		# Update the one or two previous iterates, used for momentum, spg stepsizes
+# 		# We do not save every iterate since that is too much memory
+# 		updateprevious!(previous, parameters, decomposition)
+# 	end
+
+# 	# Perform one final constrain if requested
+# 	if kwargs[:constrain_output]
+# 		kwargs = finalconstrain!(decomposition; kwargs...)
+# 		updateparameters!(parameters, decomposition, previous)
+# 		push!(stats_data, getstats(decomposition, Y, previous, parameters, stats_data))
+# 	end
+
+# 	return decomposition, stats_data, kwargs
+# end
+
 """
 Inner level function once keyword arguments are set
 """
 function _factorize(Y; kwargs...)
+    decomposition, previous, updateprevious!, parameters, updateparameters!,
[…]

### 2025-01-10 test/runtests.jl

@@ -488,7 +488,10 @@ end
     A2 = matrix_factor(X, 1)
     B2 = core(X)
 
-    @test norm(X-Y1)/norm(Y1) <= 0.001 # less than 0.1% relative error
+    @test_broken norm(X-Y1)/norm(Y1) <= 0.001 # should be a bit worse even though tolerence=0.001
+        # since we auto-apply constraints at the last step
+
+    @test norm(X-Y1)/norm(Y1) <= 0.005 # should still be reasonably good
 
     A1, B1, rel_errors, norm_grad, dist_Ncone = nnmtf(Y1, R)
 

### 2025-01-14 src/Core/blockupdates.jl
change default of tuckerproduct to take full list of matrices, and ignore excluded matrices

@@ -83,7 +83,7 @@ function make_lipschitz(T::Tucker, n::Integer, Y::AbstractArray; objective::L2, 
     elseif n in 1:N # the matrix is the zeroth factor
         function lipschitz_matrix(T::AbstractTucker; kwargs...)
             matrices = matrix_factors(T)
-            TExcludeAn = tuckerproduct(core(T), getnotindex(matrices, n); exclude=n)
+            TExcludeAn = tuckerproduct(core(T), matrices; exclude=n)
             return opnorm(slicewise_dot(TExcludeAn, TExcludeAn; dims=n))
         end
         return lipschitz_matrix
@@ -98,7 +98,7 @@ function make_lipschitz(T::CPDecomposition, n::Integer, Y::AbstractArray; object
     if n in 1:N # the matrix is the zeroth factor
         function lipschitz_matrix(T::AbstractTucker; kwargs...)
             matrices = matrix_factors(T)
-            TExcludeAn = tuckerproduct(core(T), getnotindex(matrices, n); exclude=n) # TODO optimize this to avoid making the super diagonal core
+            TExcludeAn = tuckerproduct(core(T), matrices; exclude=n) # TODO optimize this to avoid making the super diagonal core
             return opnorm(slicewise_dot(TExcludeAn, TExcludeAn; dims=n))
         end
         return lipschitz_matrix
@@ -133,7 +133,7 @@ function make_block_lipschitz(T::CPDecomposition, n::Integer, Y::AbstractArray; 
     if n in 1:N # the matrix is the zeroth factor
         function lipschitz_matrix(T::AbstractTucker; kwargs...)
             matrices = matrix_factors(T)
-            TExcludeAn = tuckerproduct(core(T), getnotindex(matrices, n); exclude=n) # TODO optimize this to avoid making the super diagonal core
+            TExcludeAn = tuckerproduct(core(T), matrices; exclude=n) # TODO optimize this to avoid making the super diagonal core
             return Diagonal_col_norm(slicewise_dot(TExcludeAn, TExcludeAn; dims=n))
         end
         return lipschitz_matrix
@@ -154,7 +154,7 @@ function make_block_lipschitz(T::Tucker, n::Integer, Y::AbstractArray; objective
     elseif n in 1:N # the matrix is the zeroth factor
         function lipschitz_matrix(T::AbstractTucker; kwargs...)
             matrices = matrix_factors(T)
[…]

### 2025-01-14 src/Core/tensorproducts.jl

@@ -108,19 +108,25 @@ arguments as matricies in a Tucker decomposition.
 
 Example
 -------
-tuckerproduct(G, A, B, C) == G ×₁ A ×₂ B ×₃ C
-tuckerproduct(G, A, C; exclude=2) == G ×₁ A ×₃ C
+tuckerproduct(G, (A, B, C)) == G ×₁ A ×₂ B ×₃ C
+tuckerproduct(G, (A, B, C); exclude=2) == G ×₁ A ×₃ C
+tuckerproduct(G, (A, B, C); exclude=2, excludes_missing=false) == G ×₁ A ×₃ C
+tuckerproduct(G, (A, C); exclude=2, excludes_missing=true) == G ×₁ A ×₃ C
 """
-function tuckerproduct(core, matricies; exclude=nothing)
+function tuckerproduct(core, matricies; exclude=nothing, excludes_missing=false)
     N = ndims(core)
     if isnothing(exclude)
         N == length(matricies) ||
             throw(ArgumentError("expected $N number of matricies, got $(length(matricies))"))
         return multifoldl(tucker_contractions(N), (core, matricies...))
-    else
+    elseif excludes_missing
         N == length(matricies) + length(exclude) ||
         throw(ArgumentError("expected $N number of matricies, got $(length(matricies))"))
         return multifoldl(getnotindex(tucker_contractions(N), exclude), (core, matricies...))
+    else
+        N == length(matricies) ||
+            throw(ArgumentError("expected $N number of matricies, got $(length(matricies))"))
+        return multifoldl(getnotindex(tucker_contractions(N), exclude), (core, getnotindex(matricies, exclude)...))
     end
 end
 tuckerproduct(core, matricies...; kwargs...) = tuckerproduct(core, matricies; kwargs...)

### 2025-01-14 test/runtests.jl

@@ -307,7 +307,7 @@ end
 
         for n in 1:3
             An = factor(T, n)
-            CM = tuckerproduct(C, getnotindex(matricies, n); exclude=n)
+            CM = tuckerproduct(C, matricies; exclude=n)
 
             # two ways of calculating the block gradient of ||T - Y||_F^2 w.r.t. An
             grad1 = slicewise_dot(T - Y, CM; dims=n)

### 2025-01-14 paper/paper.qmd
continue on Computing Gradients section

@@
   render-on-save: true
+#engine: julia
 #appendix-style: plain
@@
 
+The transpose $A^{\top}\in\mathbb{R}^{J\times I}$ of a matrix $A\in\mathbb{R}^{I\times J}$ flips entries along the main diagonal: $A^{\top}[j,i] = A[i,j]$.
+In Julia, the transpose of a matrix is typed with a single apostrophe `A'`.
+
 The $n$-slices, $n$th mode slices, or mode $n$ slices of an $N$th order tensor $A$ are notated with the slice $A[:,~\dots,~:,~i_{n},~:,~\dots,~:]$.
@@
 
-The $n$-slice dot product $\cdot_n$ between two tensors $A\in\mathbb{R}^{I_{1},\dots, I_{n-1},J,I_{n+1},\dots, I_{N}}$ and $B\in\mathbb{R}^{I_{1},\dots, I_{n-1},K,I_{n+1},\dots, I_{N}}$ returns a matrix $(A \cdot_n B)\in\mathbb{R}^{J\times K}$ with entries
+The $n$-slice dot product $\cdot_n$ between two tensors $A\in\mathbb{R}^{K_{1},\dots, K_{n-1},I,K_{n+1},\dots, K_{N}}$ and $B\in\mathbb{R}^{K_{1},\dots, K_{n-1},J,K_{n+1},\dots, K_{N}}$ returns a matrix $(A \cdot_n B)\in\mathbb{R}^{I\times J}$ with entries
 
 $$
-(A \cdot_n B)[j, k] = {\sum_{i_{1}\dots i_{n-1}i_{n+1}\dots i_{N}}} A[i_{1},\dots, i_{n-1},j,i_{n+1},\dots, i_{N}]B[i_{1},\dots, i_{n-1},k,i_{n+1},\dots, i_{N}].
+(A \cdot_n B)[i, j] = {\sum_{k_{1}\dots k_{n-1}k_{n+1}\dots k_{N}}} A[k_{1},\dots, k_{n-1},i,k_{n+1},\dots, k_{N}]B[k_{1},\dots, k_{n-1},j,k_{n+1},\dots, k_{N}].
 $$
 
-This product can also be thought of as taking the dot product between all pairs of $n$th order slices of $A$ and $B$: $(A \cdot_n B)[j, k] = A_j \cdot B_k$.
+This product can also be thought of as taking the dot product $(A \cdot_n B)[i, j] = A_i \cdot B_j$ between all pairs of $n$th order slices of $A$ and $B$, which exactly how BlockTensorDecomposition.jl defines the operation.
+
+```julia
+function slicewise_dot(A::AbstractArray, B::AbstractArray; dims=1)
+    C = zeros(size(A, dims), size(B, dims))
+    if A === B # use faster routine if they are the same
+        return _slicewise_self_dot!(C, A; dims)
+    end
+
+    for (i, A_slice) in enumerate(eachslice(A; dims))
+        for (j, B_slice) in enumerate(eachslice(B; dims))
+            C[i, j] = A_slice ⋅ B_slice
+        end
+    end
+    return C
+end
+
+function _slicewise_self_dot!(C, A; dims=1)
+    enumerated_A_slices = enumerate(eachslice(A; dims))
+    for (i, Ai_slice) in enumerated_A_slices
+        for (j, Aj_slice) in enumerated_A_slices
+            if i > j
+                continue
+            else # only compute the upper triangle entries of C
+                C[i, j] = Ai_slice ⋅ Aj_slice
+            end
+        end
+    end
+    return Symmetric(C) # indexing C[2,1] points to the entry in C[1,2]
+end
+```
 
@@
     # Initialize the output tensor
-    C = zeros(size(B)[1], sizeA[2:end]...)
-    Cmat = reshape(C, size(B)[1], prod(sizeA[2:end]))
+    C = zeros(size(B, 1), sizeA[2:end]...)
+    Cmat = reshape(C, size(B, 1), prod(sizeA[2:end]))
 
@@
 ```
+
+::: {.callout-note}
+
+If we were only working with a fixed order of tensors, we could have defined `×₁` entry-wise with `Tullio.jl`. The function definition `tullio×₁` below gives an example for order three tensors.
+
+```julia
+function tullio×₁(A::AbstractArray{_,3}, B::AbstractMatrix)
+  @tullio C[i, j, k] := A[r, j, k] * B[i, r]
+  return C
+end
+```
+
+But we would need a new definition for each ordered tensor, or use Julia's meta programming to write a method for each order at runtime.
+
+:::
 
@@
 A tensor decomposition is a factorization of a tensor into multiple (usually smaller) tensors, that can be recombined into the original tensor.
+To make a common interface for decompositions, we make an abstract subtype of Julia's `AbstractArray`, and subtype `AbstractDecomposition` for our concrete tensor decompositions.
+
+```julia
+abstract type AbstractDecomposition{T, N} <: AbstractArray{T, N} end
+```
+
 Computationally, we can think of a generic decomposition as storing factors $(A,B,C,...)$ and operations $(\times_a,\times_b,...)$ for combining them.
@@
 ```julia
-struct GenericDecomposition{T, N} <: AbstractArray{T, N}
+struct GenericDecomposition{T, N} <: AbstractDecomposition{T, N}
     factors::Tuple{Vararg{AbstractArray{T}}} # e.g. (A, B, C)
@@
     @assert (length(ops) + 1) == length(args)
-    x = args[begin]
-    for (op, arg) in zip(ops, args[begin+1:end])
+    x, xs... = args
+    for (op, arg) in zip(ops, xs)
         x = op(x, arg)
@@
 These Julia `structs` store the tensor in its factored form.
+We could define the contractions for these types and use the common interface provided by `array`, but it turns out we can reconstruct the whole tensor more efficiently.
 If the recombined tensor or particular entries are requested, Julia dispatches on the type of decomposition and calls a particular method of `array` or `getindex`.
@@
     B, A = factors(T)
-    i = I[1]
-    J = I[begin+1:end] # J = (I₂, I₃, ..., I_N)
+    i, J... = I # (i, J) = (I[1], I[begin+1:end])
     return (@view A[i, :]) ⋅ view(B, :, J...)
@@
 $$
-\min_{A_1, \dots, A_N} \left\lVert g(A_1,\dots,A_N) - Y\right\rVert _F^2 \quad \text{s.t.} \quad (A_1,\dots,A_N) \in\mathcal{C}_1 \times \dots \times \mathcal{C}_N
+\min_{A_1, \dots, A_N} \frac{1}{2} \left\lVert g(A_1,\dots,A_N) - Y\right\rVert _F^2 \quad \text{s.t.} \quad (A_1,\dots,A_N) \in\mathcal{C}_1 \times \dots \times \mathcal{C}_N
 $$ {#eq-constrained-least-squares}
@@
 
+Note the problem would have the same solutions as simply using the objective $\left\lVert g(A_1,\dots,A_N) - Y\right\rVert$ without squaring and dividing by $2$.
+We define the objective in @eq-constrained-least-squares to make computing the function value and gradients faster.
+
 ## Base algorithm {#sec-base-algorithm}
@@
 
-Let $f(A_1, \dots, A_N) := \left\lVert g(A_1,\dots,A_N) - Y\right\rVert _F^2$ be the objective function we wish to minimize in @eq-constrained-least-squares.
+Let $f(A_1, \dots, A_N) := \frac{1}{2}\left\lVert g(A_1,\dots,A_N) - Y\right\rVert _F^2$ be the objective function we wish to minimize in @eq-constrained-least-squares.
 Following Xu and Yin [@xu_BlockCoordinateDescent_2013], the general approach we take to minimize $f$ is to apply block coordinate descent using each factor as a different block.
@@
 $$
-f_n^t(A_n) := \left\lVert g(A_{1}^{t+1}, \dots, A_{n-1}^{t+1}, A_n, A_{n+1}^t, \dots , A_{N}^t)- Y \right\rVert _F^2
+f_n^t(A_n) := \frac{1}{2}\left\lVert g(A_{1}^{t+1}, \dots, A_{n-1}^{t+1}, A_n, A_{n+1}^t, \dots , A_{N}^t)- Y \right\rVert _F^2
 $$
@@
 
-To ensure the code stays flexible, the main algorithm of BlockTensorDecomposition.jl, `factorize` is defined at a very high level.
+To ensure the code stays flexible, the main algorithm of BlockTensorDecomposition.jl, `factorize`, is defined at a very high level.
 
-<!--
 ```julia
@@
 function _factorize(Y; kwargs...)
-    decomposition, kwargs = initialize_decomposition(Y; kwargs...)
-    previous, updateprevious! = initialize_previous(decomposition, Y; kwargs...)
-    parameters, updateparameters! =
-        initialize_parameters(decomposition, Y, previous; kwargs...)
-
-    # more kwargs get added here, so need to return kwargs
-    update!, kwargs = make_update!(decomposition, Y; kwargs...)
-
-    stats_data, getstats =
-        initialize_stats(decomposition, Y, previous, parameters; kwargs...)
-
-    converged = make_converged(; kwargs...)
-
-    # freeze the kwargs from a Dictionary to a NamedTuple for type stability
-    kwargs = NamedTuple(kwargs)
+    decomposition, previous, updateprevious!, parameters, updateparameters!,
+	update!, stats_data, getstats, converged, kwargs = initialize(Y, kwargs)
 
     while !converged(stats_data; kwargs...)
-        # Update the decomposition
-        # This is one cycle of updates on each factor in the decomposition
+		# Usually one cycle of updates through each factor in the decomposition
         update!(decomposition; parameters...)
 
-        # Update parameters used for the next update of the decomposition
-        # this could be the next stepsize or other info used by update!
-        updateparameters!(parameters, decomposition, previous)
-
-        # Save stats
-        push!(stats_data,
-            getstats(decomposition, Y, previous, parameters, stats_data))
-
-        # Update the one or two previous iterates
-        # We do not save every iterate since that is too much memory
-        updateprevious!(previous, parameters, decomposition)
-    end
-
-    # Perform one final constrain if requested
-    if kwargs[:constrain_output]
-        kwargs = finalconstrain!(decomposition; kwargs...)
-        updateparameters!(parameters, decomposition, previous)
-        push!(stats_data,
-            getstats(decomposition, Y, previous, parameters, stats_data))
-    end
-
-    return decomposition, stats_data, kwargs
-end
-```
-
-Abstracting further, we can see the general structure. -->
-
-```julia
-function factorize(Y; kwargs...)
-    decomposition, converged, update!, updateprevious!, updateparameters!, stats_data, getstats, kwargs = initialize(Y; kwargs...)
-
-    while !converged(stats_data; kwargs...)
-        update!(decomposition; parameters...)
-
+		# This could be the next stepsize or other info used by update!
         updateparameters!(parameters, decomposition, previous)
@@
 
+		# Update one or two previous iterates. For example, used for momentum
         updateprevious!(previous, parameters, decomposition)
@@
 
-    postprocess!(decomposition, parameters, stats_data, Y; kwargs...)
+    kwargs = postprocess!(decomposition, Y, previous, parameters, stats_data, updateparameters!, getstats, kwargs)
 
@@
 
+Generally, we can use automatic differentiation on $f$ to compute gradients.
+Some care needs to be taken otherwise the forward or backwards pass will have to be recompiled every iteration since the factors are updated every iteration.
+
+But for Tucker decompositions, we can compute gradients faster than what an automatic differentiation scheme would give, by taking advantage of symmetry and other computational shortcuts.
+
+Starting with the Tucker-1 decomposition (@def-tucker-1-decomposition), we would like to compute $\nabla_B f(B, A)$ and $\nabla_A f(B, A)$ for $f(B, A) = \frac{1}{2} \left\lVert AB - Y\right\rVert_F^2$ for a given input $Y$.
+We have the gradient
+
+$$
+\nabla_B f(B, A) = A^{\top}(AB - Y) = (B \times_{1} A - Y) \times_1 A^\top
+$$ {#eq-tucker-1-gradient-1}
+
+by chain rule, but it is more efficient to calculate the gradient as
+
+$$
+\nabla_B f(B, A) = (A^{\top} A)B - A^{\top}Y = B \times_1 (A^{\top} A) - Y\times_1 A^{\top}.
+$$ {#eq-tucker-1-gradient-2}
+
+^[Seeing @eq-tucker-1-gradient-1 and @eq-tucker-1-gradient-2 written using the $1$-mode product shows how it is "backwards" to normal matrix-matrix multiplication.]For $A\in\mathbb{R}^{I\times R}$, $B\in\mathbb{R}^{R\times J\times K}$, and $Y\in\mathbb{R}^{I\times J \times K}$, @eq-tucker-1-gradient-1 requires
+
+$$
+\underbrace{ 2IJKR }_{ AB-Y }+\underbrace{ IJK(2I-1) }_{ A^\top (AB-Y) } \sim 2IJKR+2I^2JK
+$$
+floating point operations (FLOPS) whereas @eq-tucker-1-gradient-2 only uses
+
+$$
+\underbrace{ \frac{R(R+1)}{2}(2I-1) }_{ A^\top A } + \underbrace{ RJK(2I-1) }_{ A^\top Y } + \underbrace{ 2R^2JK }_{ (A^{\top} A)B - (A^{\top}Y) } \sim 2IJKR+2R^2JK + IR^2
+$$
+
+FLOPS^[Note we have the smaller factor ${R(R+1)}/{2}$ and not the expected $R^2$ number of entries needed to compute $A^\top A$.
+The product is a symmetric matrix so only the upper or lower triangle of entries needs to be computed.].
+So for small ranks $R \ll I$, @eq-tucker-1-gradient-2 is cheaper.
+
+A similar story can be said about $\nabla_A f(B, A)$ which is most efficiently computed as
+
+$$
+\nabla_A f(B, A) = A(B\cdot _1 B) - Y\cdot_1 B.
+$$
+
+The associated implementation with BlockTensorDecomposition.jl is shown below.
+We define a `make_gradient` which takes the decomposition, factor index `n`, and data tensor `Y`, and creates a function that computes the gradient for the same type of decomposition.
+This lets us manipulate the function that computes the gradient, rather than just the computed gradient.
+
+```julia
+function make_gradient(T::Tucker1, n::Integer, Y::AbstractArray; objective::L2, kwargs...)
+    if n==0 # the core is the zeroth factor
+        function gradient0(T::Tucker1; kwargs...)
+            (B, A) = factors(T)
+            AA = A'A
+            YA = Y×₁A'
+            grad = B×₁AA - YA
+            return grad
+        end
+        return gradient0
+    elseif n==1 # the matrix is the first factor
+        function gradient1(T::Tucker1; kwargs...)
+            (B, A) = factors(T)
+            BB = slicewise_dot(B, B)
+            YB = slicewise_dot(Y, B)
+            grad = A*BB - YB
+            return grad
+        end
+        return gradient1
+    else
+        error("No $(n)th factor in Tucker1")
+    end
+end
+```
+
+Similarly, we also have special methods for the Tucker and CP Decomposition.
+
+The gradient with respect to the core for a full Tucker factorization is
+
+$$
+\nabla_B f(B, A_1, \dots, A_N) = B \bigtimes_n A_n^\top A_n - Y \bigtimes_{n} A_n^\top,
+$$
+
+and the gradient with respect to the matrix factor $A_m$ is
+
+$$
+\nabla_{A_m} f(B, A_1, \dots, A_N) = A_m (B \bigtimes_{n\neq m} A_n) \cdot_m (B \bigtimes_{n\neq m} A_n) - Y \cdot_m (B \bigtimes_{n\neq m} A_n).
+$$
+
+```julia
+function make_gradient(T::Tucker, n::Integer, Y::AbstractArray; objective::L2, kwargs...)
+    N = ndims(T)
+    if n==0 # the core is the zeroth factor
+        function gradient_core(T::AbstractTucker; kwargs...)
+            C = core(T)
+            matrices = matrix_factors(T)
+            gram_matrices = map(A -> A'A, matrices) # gram matrices AA = A'A, BB = B'B...
+            YAB = tuckerproduct(Y, adjoint.(matrices)) # Y ×₁ A' ×₂ B' ...
+            grad = tuckerproduct(C, gram_matrices) - YAB
+            return grad
+        end
+        return gradient_core
+
+    elseif n in 1:N # the matrix factors start at m=1
+        function gradient_matrix(T::AbstractTucker; kwargs...)
+            matrices = matrix_factors(T)
+            TExcludeAn = tuckerproduct(core(T), matrices; exclude=n)
+            An = factor(T, n)
+            grad = An*slicewise_dot(TExcludeAn, TExcludeAn; dims=n) - slicewise_dot(Y, TExcludeAn; dims=n)
+            return grad
+        end
+        return gradient_matrix
+
+    else
+        error("No $(n)th factor in Tucker")
+    end
+end
+```
+
+```julia
+function make_gradient(T::CPDecomposition, n::Integer, Y::AbstractArray; objective::L2, kwargs...)
+    N = ndims(T)
+    if n in 1:N # the matrix factors start at m=1
+        function gradient_matrix(T::AbstractTucker; kwargs...)
+            matrices = matrix_factors(T)
+            TExcludeAn = tuckerproduct(core(T), matrices; exclude=n)
+            An = factor(T, n)
+            grad = An*slicewise_dot(TExcludeAn, TExcludeAn; dims=n) - slicewise_dot(Y, TExcludeAn; dims=n)
+            return grad
+        end
+        return gradient_matrix
+
+    else
+        error("No $(n)th factor in Tucker")
+    end
+end
+```
+
+The function `tuckerproduct(B, (A₁, ..., Aₙ))` computes
+$$
+B \bigtimes_{n} A_n = \llbracket B; A_1,\dots,A_N \rrbracket,
+$$
+
+and can optionally "exclude" one of the matrix factors `tuckerproduct(B, (A₁, ..., Aₙ); exclude=m)` to compute
+
+$$
+B \bigtimes_{n\neq m} A_n = \llbracket B; A_1,\dots,A_{n-1}, \text{id}_{I_n}, A_{n+1},\dots,A_{N} \rrbracket
+$$
+where $B\in\mathbb{R}^{I_1\times \dots \times I_N}$.
+
 ### Computing Lipschitz Step-sizes {#sec-lipschitz-computation}
+
+Similar to automatic differentiation, there exist "automatic Lipschitz" calculations to upper bound the Lipschitz constant of a function [@virmaux_lipschitz_2018].
 

### 2025-01-14 example/constrainedCP.jl
fix spelling of matrices

@@ -7,9 +7,9 @@ R = 5
 D = 2
 
 fact = BlockTensorDecomposition.factorize
-matricies = [abs_randn(N, R) for _ in 1:D]
-l1scale_cols!.(matricies)
-Ydecomp = CPDecomposition(Tuple(matricies))#abs_randn
+matrices = [abs_randn(N, R) for _ in 1:D]
+l1scale_cols!.(matrices)
+Ydecomp = CPDecomposition(Tuple(matrices))#abs_randn
 @assert all(check.(l1scale_cols!, factors(Ydecomp)))
 Y = array(Ydecomp)
 

### 2025-01-14 src/Core/decomposition.jl

@@ -307,7 +307,7 @@ end=#
 # TODO add optimized indexing for full tucker decomposition
 
 """
-CP decomposition. Takes the form of an outerproduct of multiple matricies.
+CP decomposition. Takes the form of an outerproduct of multiple matrices.
 
 For example, a rank r CP decomposition of an order three tensor D would be, entry-wise,
 D[i, j, k] = sum_r A[i, r] * B[j, r] * C[k, r]).

### 2025-01-14 src/Core/tensorproducts.jl

@@ -60,7 +60,7 @@ Constracts all but the dimentions `dimsA` and `dimsB` of A and B by performing a
 
 Generalizes `@einsum C[s,r] := A[s,j,k]*B[r,j,k]` to arbitrary dimentions.
 
-For example, if A and B are both matricies, slicewise_dot(A, B) == A*B'
+For example, if A and B are both matrices, slicewise_dot(A, B) == A*B'
 """
 function slicewise_dot(A::AbstractArray, B::AbstractArray; dims=1, dimsA=dims, dimsB=dims)
     C = zeros(size(A, dimsA), size(B, dimsB)) # Array{promote_type(T, U), 2}(undef, size(A, 1), size(B, 1)) doesn't seem to be faster
@@ -104,7 +104,7 @@ tucker_contractions(N) = Tuple((G, A) -> nmp(G, A, n) for n in 1:N)
     tuckerproduct(G, A, B, ...)
 
 Multiplies the inputs by treating the first argument as the core and the rest of the
-arguments as matricies in a Tucker decomposition.
+arguments as matrices in a Tucker decomposition.
 
 Example
 -------
@@ -113,42 +113,42 @@ tuckerproduct(G, (A, B, C); exclude=2) == G ×₁ A ×₃ C
 tuckerproduct(G, (A, B, C); exclude=2, excludes_missing=false) == G ×₁ A ×₃ C
 tuckerproduct(G, (A, C); exclude=2, excludes_missing=true) == G ×₁ A ×₃ C
 """
-function tuckerproduct(core, matricies; exclude=nothing, excludes_missing=false)
+function tuckerproduct(core, matrices; exclude=nothing, excludes_missing=false)
     N = ndims(core)
     if isnothing(exclude)
-        N == length(matricies) ||
-            throw(ArgumentError("expected $N number of matricies, got $(length(matricies))"))
-        return multifoldl(tucker_contractions(N), (core, matricies...))
+        N == length(matrices) ||
+            throw(ArgumentError("expected $N number of matrices, got $(length(matrices))"))
+        return multifoldl(tucker_contractions(N), (core, matrices...))
     elseif excludes_missing
-        N == length(matricies) + length(exclude) ||
-        throw(ArgumentError("expected $N number of matricies, got $(length(matricies))"))
-        return multifoldl(getnotindex(tucker_contractions(N), exclude), (core, matricies...))
+        N == length(matrices) + length(exclude) ||
[…]

### 2025-01-14 test/factorize_initialization.jl

@@ -6,9 +6,9 @@ R = 5
 D = 2
 
 fact = BlockTensorDecomposition.factorize
-matricies = [abs_randn(N, R) for _ in 1:D]
-l1scale_cols!.(matricies)
-Ydecomp = CPDecomposition(Tuple(matricies))#abs_randn
+matrices = [abs_randn(N, R) for _ in 1:D]
+l1scale_cols!.(matrices)
+Ydecomp = CPDecomposition(Tuple(matrices))#abs_randn
 @assert all(check.(l1scale_cols!, factors(Ydecomp)))
 Y = array(Ydecomp)
 

### 2025-01-14 test/runtests.jl

@@ -303,11 +303,11 @@ end
         T = Tucker((10,11,12), (3,4,5))
         Y = randn(10,11,12)
         C = core(T)
-        matricies = matrix_factors(T)
+        matrices = matrix_factors(T)
 
         for n in 1:3
             An = factor(T, n)
-            CM = tuckerproduct(C, matricies; exclude=n)
+            CM = tuckerproduct(C, matrices; exclude=n)
 
             # two ways of calculating the block gradient of ||T - Y||_F^2 w.r.t. An
             grad1 = slicewise_dot(T - Y, CM; dims=n)

### 2025-01-27 paper/paper.qmd
Redefine tensor operations and add to Lipschitz constant section

@@
 
-### Operations
+### Products of Tensors
 
-The Frobenius inner product between two tensors $A,B\in\mathbb{R}^{I_1\times\dots\times I_N}$ is denoted
+The outer product $\otimes$ between two tensors $A\in\mathbb{R}^{I_1\times\dots\times I_M}$ and $B\in\mathbb{R}^{J_1\times\dots\times J_N}$ yields an order $M+N$ tensor $A\otimes B \in \mathbb{R}^{I_1\times\dots\times I_M\times J_1\times\dots\times J_N}$ that is entry-wise
+
+$$
+(A\otimes B)[i_1,\dots,i_M,j_1,\dots,j_N] = A[i_1,\dots,i_M] B[j_1,\dots,j_N].
+$$
+
+TODO Define in BlockTensorDecomposition.jl
+
+The Frobenius inner product between two tensors $A,B\in\mathbb{R}^{I_1\times\dots\times I_N}$ yields a real number $A \cdot B\in\mathbb{R}$ and is defined as
 
@@
 
+The $n$-mode product and $n$-slice product can be thought of as opposites of each other.
+The $n$-mode product sums over just the $n$th dimension of the first tensor, whereas the $n$-slice product sums over all but the $n$th dimension.
+
+We can extend the $n$-mode product to sum over multiple indices between two tensors.
+
+The (complete) multi-mode product $\times_{1,\dots,n}=\times_{[n]}$ between a tensor $A\in\mathbb{R}^{I_1 \times \dots \times I_N}$ and tensor $B\in\mathbb{R}^{I_1\times\dots\times I_n}$, returns a tensor $(A \times_{[n]} B)\in\mathbb{R}^{I_{n+1} \times \dots \times I_N}$ with entries
+
+$$
+(A \times_{[n]} B)[i_{n+1},\dots,i_N] = \sum_{i_1, \dots ,i_n} A[i_1, \dots,i_n,i_{n+1},\dots,i_N]B[i_1, \dots ,i_n].
+$$
+
+### Gradients, Norms, and Lipschitz Constants
+
+The gradient $\nabla f:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}^{I_1\times \dots \times I_N}$ of a (differentiable) function $f:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}$ is defined entry-wise for a tensor $A\in \mathbb{R}^{I_1\times \dots \times I_N}$ by
+
+$$
+\nabla f(A) [i_1, \dots, i_N] = \frac{\partial f}{\partial A[i_1, \dots, i_N]} (A).
+$$
+
+The Hessian $\nabla^2 f:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}^{(I_1\times \dots \times I_N)^2}$ of a (second-differentiable) function $f:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}$ is the gradient of the gradient and is defined for a tensor $A\in\mathbb{R}^{I_1\times \dots \times I_N}$ entry-wise by
+
+$$
+\nabla^2 f(A) [i_1, \dots, i_N, j_1, \dots , j_N] = \frac{\partial^2 f}{\partial A[i_1, \dots , i_N]\partial A[j_1, \dots , j_N]} (A).
+$$
+
+For a tensor input $A$ of order $N$, the Hessian tensor $\nabla^2 f(A)$ is of order $2N$.
+
 The Frobenius norm of a tensor $A$ is the square root of its dot product with itself
@@
 
-For matrices $M$, the (Operator) 2-norm is defined as
+For matrices $M$, the (operator) 2-norm is defined as
 
 $$
-\left\lVert M \right\rVert_2 = \arg\,\max_{\left\lVert v \right\rVert_2 = 1} \left\lVert Mv \right\rVert_2 = \sigma_1(M)
+\left\lVert M \right\rVert_{\text{op}} = \left\lVert M \right\rVert_2 = \sup_{\lVert v \rVert_2 = 1} \left\lVert Mv \right\rVert_2 = \sigma_1(M)
 $$
@@
 
-For tensors $T$, the (Operator) 2-norm needs to be defined in terms of how we treat them as function on other tensors.
+For tensors $T$, the (operator) 2-norm needs to be defined in terms of how we treat them as function on other tensors.
 There is a canonical way to do this for vectors $x \to v^{\top}x$ and matrices $x \to Mx$, but not tensors.
 This is relevant to @sec-lipschitz-computation where the Lipschitz step-size is computed in terms of the Operator norm of the Hessian of our objective function.
+In the case of the Hessian tensor $\nabla^2 f(A)$ evaluated at $A$, we consider the function $X \to \nabla^2 f(A) \times_{[N]} X$.
+This gives us
+
+$$
+\left\lVert \nabla^2 f(A) \right\rVert_{\text{op}} = \sup_{\left\lVert X \right\rVert_F = 1} \left\lVert \nabla^2 f(A) \times_{[N]} X \right\rVert_F = \sigma_1(\text{mat}(\nabla^2 f(A)))
+$$
+
+where $\text{mat}:\mathbb{R}^{(I_1\times \dots \times I_N)^2}\to\mathbb{R}^{(I_1 \dots I_N)\times (I_1 \dots I_N)}$ takes a $2N$-ordered tensor and flattens it into a square $(I_1 \dots I_N)\times (I_1 \dots I_N)$ matrix.
+
+A function $g:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}^{I_1\times \dots \times I_N}$ is $L$-Lipschitz when
+
+$$
+\left\lVert g(A) - g(B) \right\rVert_F \leq L \left\lVert A - B \right\rVert_F, \quad \forall A,B\in\mathbb{R}^{I_1\times \dots \times I_N}.
+$$
+
+We call the smallest such $L$ *the* Lipschitz constant of $g$.
+
+A differentiable function $f:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}$ is $L$-smooth when its gradient $g=\nabla f$ is $L$-Lipschitz.
+
+::: {#thm-quadratic-smoothness}
+# Quadratic Smoothness
+Let $f:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}$ be a quadratic function
+
+$$
+f(X) = \frac{1}{2}A(X,X) + B(X) + C
+$$
+
+of $X$ with bilinear function $A:(\mathbb{R}^{I_1\times \dots \times I_N})^2\to\mathbb{R}$, linear function $B:\mathbb{R}^{I_1\times \dots \times I_N}\to\mathbb{R}$, and constant $C\in\mathbb{R}^{I_1\times \dots \times I_N}$.
+
+Then
+
+1. the Hessian $\nabla^2 f$ is a constant function that evaluates to $\nabla^2 f(X) = D$ at every point $X$ for some $D \in \mathbb{R}^{(I_1\times \dots \times I_N)^2}$,
+2. the tensor $D$ only depends on the bilinear function $A$ (and not on $B$ and $C$), and
+2. the quadratic function $f$ is $L$-smooth with constant
+
+$$
+L = \left\lVert D \right\rVert _{\text{op}} = \sigma_1(\operatorname{mat}D).
+$$
+
+:::
 
@@
 Y = B\times_1 A_1 \times_2 \dots\times_N A_N = B \bigtimes_{n} A_n = \llbracket B; A_1, \dots, A_N \rrbracket.
+$$ {#eq-tucker-product}
+
+:::
+
+The *Tucker Product* defined by @eq-tucker-product is implemented in BlockTensorDecomposition.jl with `tuckerproduct(B, (A1, ..., AN))` and computes
+$$
+B \bigtimes_{n} A_n = \llbracket B; A_1,\dots,A_N \rrbracket.
 $$
 
-:::
+It can also optionally "exclude" one of the matrix factors with the call `tuckerproduct(B, (A1, ..., AN); exclude=n)` to compute
+
+$$
+B \bigtimes_{m\neq n} A_m = \llbracket B; A_1,\dots,A_{n-1}, \text{id}_{R_n}, A_{n+1},\dots,A_{N} \rrbracket.
+$$
+
+```julia
+function tuckerproduct(core, matrices; exclude=nothing)
+    N = ndims(core)
+    if isnothing(exclude)
+        return multifoldl(tucker_contractions(N), (core, matrices...))
+    else
+        return multifoldl(getnotindex(tucker_contractions(N), exclude), (core, getnotindex(matrices, exclude)...))
+    end
+end
+
+tucker_contractions(N) = Tuple((B, A) -> nmode_product(B, A, n) for n in 1:N)
+```
 
@@
 
-### Computing Gradients
+### Computing Gradients {#sec-gradient-computation}
 - Use Auto diff generally
@@
 
+::: callout-note
+For the family of Tucker decompositions, the objective function $f$ is "block-quadratic" with respect to the factors.
+This means the gradient with respect to a factor is an affine function of that factor.
+This is exactly what we see in @eq-tucker-1-gradient-2 where $B$ is multiplied by the "slope" $A^\top A$ plus a shift of $-Y\times_1 A^\top$.
+:::
+
 The associated implementation with BlockTensorDecomposition.jl is shown below.
@@
     if n==0 # the core is the zeroth factor
-        function gradient0(T::Tucker1; kwargs...)
-            (B, A) = factors(T)
+        function gradient0(X::Tucker1; kwargs...)
+            (B, A) = factors(X)
             AA = A'A
@@
     elseif n==1 # the matrix is the first factor
-        function gradient1(T::Tucker1; kwargs...)
-            (B, A) = factors(T)
+        function gradient1(X::Tucker1; kwargs...)
+            (B, A) = factors(X)
             BB = slicewise_dot(B, B)
@@
 
-and the gradient with respect to the matrix factor $A_m$ is
+and the gradient with respect to the matrix factor $A_n$ is
 
 $$
-\nabla_{A_m} f(B, A_1, \dots, A_N) = A_m (B \bigtimes_{n\neq m} A_n) \cdot_m (B \bigtimes_{n\neq m} A_n) - Y \cdot_m (B \bigtimes_{n\neq m} A_n).
+\nabla_{A_n} f(B, A_1, \dots, A_N) = A_n ( \tilde{X}_n \cdot_n \tilde{X}_n) - Y \cdot_n \tilde{X}_n
+$$
+
+where
+
+$$
+\tilde{X}_n = (B \bigtimes_{m\neq n} A_m) = \llbracket B; A_1,\dots,A_{n-1}, \text{id}_{R_n}, A_{n+1},\dots,A_{N} \rrbracket.
 $$
@@
     if n==0 # the core is the zeroth factor
-        function gradient_core(T::AbstractTucker; kwargs...)
-            C = core(T)
-            matrices = matrix_factors(T)
-            gram_matrices = map(A -> A'A, matrices) # gram matrices AA = A'A, BB = B'B...
-            YAB = tuckerproduct(Y, adjoint.(matrices)) # Y ×₁ A' ×₂ B' ...
-            grad = tuckerproduct(C, gram_matrices) - YAB
+        function gradient_core(X::AbstractTucker; kwargs...)
+            B = core(X)
+            matrices = matrix_factors(X)
+            gram_matrices = map(A -> A'A, matrices) # gram matrices AA = A'A,
+                                                    # BB = B'B, ...
+            grad = tuckerproduct(B, gram_matrices)
+                 - tuckerproduct(Y, adjoint.(matrices))
             return grad
@@
     elseif n in 1:N # the matrix factors start at m=1
-        function gradient_matrix(T::AbstractTucker; kwargs...)
-            matrices = matrix_factors(T)
-            TExcludeAn = tuckerproduct(core(T), matrices; exclude=n)
-            An = factor(T, n)
-            grad = An*slicewise_dot(TExcludeAn, TExcludeAn; dims=n) - slicewise_dot(Y, TExcludeAn; dims=n)
+        function gradient_matrix(X::AbstractTucker; kwargs...)
+            B = core(X)
+            matrices = matrix_factors(X)
+            Aₙ = factor(X, n)
+            X̃ₙ = tuckerproduct(B, matrices; exclude=n)
+            grad = Aₙ * slicewise_dot(X̃ₙ, X̃ₙ; dims=n)
+                   - slicewise_dot(Y, X̃ₙ; dims=n)
             return grad
@@
 
+For the CP Decomposition, we can simply treat the core as $B=\text{id}_R$ and compute the gradient with respect to the matrix factors similarly to the Tucker decomposition:
+
+$$
+\nabla_{A_n} f(A_1, \dots, A_N) = A_n (\tilde{X}_n \cdot_n \tilde{X}_n) - Y \cdot_n \tilde{X}_n
+$$
+
+where
+
+$$
+\tilde{X}_n = (\text{id}_R \bigtimes_{m\neq n} A_m) = \llbracket \text{id}_R; A_1,\dots,A_{n-1}, \text{id}_{R}, A_{n+1},\dots,A_{N} \rrbracket.
+$$
+
 ```julia
@@
     if n in 1:N # the matrix factors start at m=1
-        function gradient_matrix(T::AbstractTucker; kwargs...)
-            matrices = matrix_factors(T)
-            TExcludeAn = tuckerproduct(core(T), matrices; exclude=n)
-            An = factor(T, n)
-            grad = An*slicewise_dot(TExcludeAn, TExcludeAn; dims=n) - slicewise_dot(Y, TExcludeAn; dims=n)
+        function gradient_matrix(X::AbstractTucker; kwargs...)
+            B = core(X)
+            matrices = matrix_factors(X)
+            Aₙ = factor(X, n)
+            X̃ₙ = tuckerproduct(B, matrices; exclude=n)
+            grad = Aₙ * slicewise_dot(X̃ₙ, X̃ₙ; dims=n)
+                   - slicewise_dot(Y, X̃ₙ; dims=n)
             return grad
@@
 
-The function `tuckerproduct(B, (A₁, ..., Aₙ))` computes
-$$
-B \bigtimes_{n} A_n = \llbracket B; A_1,\dots,A_N \rrbracket,
-$$
-
-and can optionally "exclude" one of the matrix factors `tuckerproduct(B, (A₁, ..., Aₙ); exclude=m)` to compute
-
-$$
-B \bigtimes_{n\neq m} A_n = \llbracket B; A_1,\dots,A_{n-1}, \text{id}_{I_n}, A_{n+1},\dots,A_{N} \rrbracket
-$$
-where $B\in\mathbb{R}^{I_1\times \dots \times I_N}$.
-
 ### Computing Lipschitz Step-sizes {#sec-lipschitz-computation}
@@
 
+For the family of Tucker decompositions, we can compute the Lipschitz constants efficiently similar to how we compute the gradient in @sec-gradient-computation.
+
+::: callout-note
+For the family of Tucker decompositions, the objective function $f$ is "block-quadratic" with respect to the factors.
+We saw in @eq-tucker-1-gradient-2 that the gradient is an affine map where $B$ is multiplied by the "slope" $A^\top A$ plus a shift of $-Y\times_1 A^\top$.
+
+This means the Lipschitz constant of the gradient for that block is the operator norm of the "slope".
+:::
+
+Since our least-squares objective $f$ is a block-quadratic function in each factor, we can use @thm-quadratic-smoothness to find the smoothness $L_{A_n}$ with respect to each factor $A_n$ by calculating the largest eigenvalue of the partial Hessian $\nabla^2_{A_n}f(A_1,\dots,A_N)$.
+
+Starting with the Tucker-1 decomposition (@def-tucker-1-decomposition) objective $f(B,A)=\frac{1}{2}\left\lVert AB-Y\right\rVert_F^2$, we have
+
+$$
+\nabla^2_{B}f(B,A) = A
+$$
+
+TODO Explain how we get the following.
+General idea is to take partial Hessian of our quadratic function with respect to one block, and get that the operator norm of the hessian (as a map from $\mathbb{R}^{I_1\times \dots \times I_N}$ to itself) is the Lipschitz constant of the gradient.
+See Obsidian note "Gradients Hessians and Lipschitz Constants of a Tucker Decomposition".
+
+```julia
+function make_lipschitz(T::Tucker1, n::Integer, Y::AbstractArray; objective::L2, kwargs...)
+    if n==0 # the core is the zeroth factor
+        function lipschitz0(T::Tucker1; kwargs...)
+            A = matrix_factor(T, 1)
+            return opnorm(A'A)
+        end
+        return lipschitz0
+
+    elseif n==1 # the matrix is the zeroth factor
+        function lipschitz1(T::Tucker1; kwargs...)
+            C = core(T)
+            return opnorm(slicewise_dot(C, C))
+        end
+        return lipschitz1
+
+    else
+        error("No $(n)th factor in Tucker1")
+    end
+end
+
+function make_lipschitz(T::Tucker, n::Integer, Y::AbstractArray; objective::L2, kwargs...)
+    N = ndims(T)
+    if n==0 # the core is the zeroth factor
+        function lipschitz_core(T::AbstractTucker; kwargs...)
+            #matrices = matrix_factors(T)
+            #gram_matrices = map(A -> A'A, matrices)
+            #return prod(opnorm.(gram_matrices))
+            return prod(A -> opnorm(A'A), matrix_factors(T))
+        end
+        return lipschitz_core
+
+    elseif n in 1:N # the matrix is the zeroth factor
+        function lipschitz_matrix(T::AbstractTucker; kwargs...)
+            matrices = matrix_factors(T)
+            TExcludeAn = tuckerproduct(core(T), matrices; exclude=n)
+            return opnorm(slicewise_dot(TExcludeAn, TExcludeAn; dims=n))
+        end
+        return lipschitz_matrix
+
+    else
+        error("No $(n)th factor in Tucker")
+    end
+end
+
+function make_lipschitz(T::CPDecomposition, n::Integer, Y::AbstractArray; objective::L2, kwargs...)
+    N = ndims(T)
+    if n in 1:N # the matrix is the zeroth factor
+        function lipschitz_matrix(T::AbstractTucker; kwargs...)
+            matrices = matrix_factors(T)
+            TExcludeAn = tuckerproduct(core(T), matrices; exclude=n) # TODO optimize this to avoid making the super diagonal core
+            return opnorm(slicewise_dot(TExcludeAn, TExcludeAn; dims=n))
+        end
+        return lipschitz_matrix
+
+    else
+        error("No $(n)th factor in Tucker")
+    end
+end
+```
+
 # Computational Techniques
+- As stated, algorithm works
+- But can be slow, especially for constrained or large problems
+
+As stated, the algorithm described in @sec-base-algorithm works.
+It will converge to a solution to our optimization problem and factorize the input tensor.
+It is worth discussing how the algorithm can be modified to improve convergence to maintain quick convergence for large problems, and what sort of architectural methods are used to allow for maximum flexibility, without over engineering the package.
 
 ## For Improving Convergence Speed
-- As stated, algorithm works
-- But can be slow, especially for constrained or large problems
+
+There are a few techniques used to assist convergence.
+Two ideas that are well studied are discussed in this section.
+They are 1) breaking up the updates into smaller blocks, and 2) using momentum or acceleration.
+What is perhaps novel is considering the synergy between these two ideas.
+
+Two more techniques are implemented in BlockTensorDecomposition.jl to improve convergence.
+To the authors knowledge, these are new to tensor factorization, but may or may not be applicable depending on the exact factorization problem or data being studied.
+For these reasons, these other techniques are discussed separately in @sec-ppr and @sec-multi-scale.
 
@@
 - Can perform this efficiently with a "matrix step-size"
+
+When using block coordinate descent as in @sec-base-algorithm, it is natural to treat each factor as its own block.
+This requires the fewest blocks while ensuring the objective is still convex with respect to each block.
+We could just as easily use smaller blocks.
+
+In [@lee_algorithms_2000], the multiplicative update algorithm used to solve nonnegative matrix factorization problems is written in a way so that each row of the first factor
 
@@
 
-# Partial Projection and Rescaling
+# Partial Projection and Rescaling {#sec-ppr}
 - for bounded linear constraints
@@
 
-# Multi-scale
+# Multi-scale {#sec-multi-scale}
 - use a coarse discretization along continuous dimensions

### 2025-01-27 test/opnorm/faster_opnorm.qmd
make opnorm test

---
title: Calculating the Operator Norm in Julia
date: 27 January 2025
format: html
engine: julia
editor:
  render-on-save: true
---

# Context
This document tests different ways to calculate the operator norm of a matrix in Julia.

The operator norm is needed to calculate the step-size used in each gradient descent update, so it is important to ensure this operation fast.

# Background

We define the operator norm of a matrix $A\in\mathbb{R}^{I\times J}$ to be
$$
\left\lVert A \right\rVert_{op} = \sup_{\lVert v \rVert_{2}} \left\lVert A v\right\rVert_{2}.
$$

It is true that
$$
\left\lVert A^\top \right\rVert_{op} = \left\lVert A \right\rVert_{op}
$$

and

$$
\left\lVert A^\top A \right\rVert_{op} = \left\lVert A A^\top  \right\rVert_{op} = \left\lVert A \right\rVert_{op}^2.
$$

# Tests

Let's look at a few ways we can find the norm of a matrix in Julia. To ensure a fair comparison, we will wrap all of these in a function, and test them out on three types of dense matrices: tall, wide, and square.

## Possible routines
```{julia}
using Pkg; Pkg.activate(".")
Pkg.add(["LinearAlgebra","Random","Arpack","BenchmarkTools"])
#Pkg.add(path="C:\\Users\\Nicholas\\OneDrive - UBC\\Research\\general-block-decomposition\\BlockTensorDecomposition.jl")
```
```{julia}
using Random: rand, randn
using LinearAlgebra: svdvals, opnorm, Symmetric
using Arpack: svds
using BlockTensorDecomposition: slicewise_dot
using BenchmarkTools
```

Now for the set of function
```{julia}
fopnorm1(X) = opnorm(X)
fopnorm2(X) = sqrt(opnorm(X'X))
fopnorm3(X) = sqrt(opnorm(X*X'))
fopnorm4(X) = sqrt(opnorm(Symmetric(X'X)))
fopnorm5(X) = sqrt(opnorm(Symmetric(X*X')))

fopnorms = [fopnorm1,fopnorm2,fopnorm3,fopnorm4,fopnorm5]

svdvals_top(X) = svdvals(X)
fsvdvals1(X) = svdvals_top(X)
fsvdvals2(X) = sqrt(svdvals_top(X'X))
fsvdvals3(X) = sqrt(svdvals_top(X*X'))
fsvdvals4(X) = sqrt(svdvals_top(Symmetric(X'X)))
fsvdvals5(X) = sqrt(svdvals_top(Symmetric(X*X')))

fopnorms = [fsvdvals1,fsvdvals2,fsvdvals3,fsvdvals4,fsvdvals5]

svds_top(X) = svds(X;nsv=1)[1].S[1]
fsvds1(X) = svds_top(X)
fsvds2(X) = sqrt(svds_top(X'X))
fsvds3(X) = sqrt(svds_top(X*X'))
fsvds4(X) = sqrt(svds_top(Symmetric(X'X)))
fsvds5(X) = sqrt(svds_top(Symmetric(X*X')))

fsvdss = [fsvds1,fsvds2,fsvds3,fsvds4,fsvds5]
```
```{julia}
b = @benchmark fopnorm1(A) setup=(A=randn(10,100))
```

```julia
dims = [
    (100, 10),
    (10, 100),
    (100, 100),
    (10, 10)
]

for (N, M) in dims
    A = randn(N, M)
    b = @benchmark opnorm(A)
    display(b)
end
```

### 2025-01-27 test/opnorm/faster_opnorm.jl

@@ -1,0 +1,132 @@
+# Context
+#This document tests different ways to calculate the operator norm of a matrix in Julia.
+
+#The operator norm is needed to calculate the step-size used in each gradient descent update, so it is important to ensure this operation fast.
+
+# Background
+
+# We define the operator norm of a matrix $A\in\mathbb{R}^{I\times J}$ to be
+# $$
+# \left\lVert A \right\rVert_{op} = \sup_{\lVert v \rVert_{2}} \left\lVert A v\right\rVert_{2}.
+# $$
+
+# It is true that
+# $$
+# \left\lVert A^\top \right\rVert_{op} = \left\lVert A \right\rVert_{op}
+# $$
+
+# and
+
+# $$
+# \left\lVert A^\top A \right\rVert_{op} = \left\lVert A A^\top  \right\rVert_{op} = \left\lVert A \right\rVert_{op}^2.
+# $$
+
+# # Tests
+
+# Let's look at a few ways we can find the norm of a matrix in Julia. To ensure a fair comparison, we will wrap all of these in a function, and test them out on three types of dense matrices: tall, wide, and square.
+
+## Possible routines
+# ```{julia}
+# using Pkg; Pkg.activate(".")
+# Pkg.add(["LinearAlgebra","Random","Arpack","BenchmarkTools"])
+# #Pkg.add(path="C:\\Users\\Nicholas\\OneDrive - UBC\\Research\\general-block-decomposition\\BlockTensorDecomposition.jl")
+# ```
+
+using Random: rand, randn
+using LinearAlgebra: svdvals, opnorm, Symmetric
+using Arpack: svds
+using BlockTensorDecomposition: slicewise_dot
+using BenchmarkTools
+
+
+# Now for the test set of functions
+
+fopnorm1(X) = opnorm(X)
+fopnorm2(X) = sqrt(opnorm(X'X))
+fopnorm3(X) = sqrt(opnorm(X*X'))
+fopnorm4(X) = sqrt(opnorm(Symmetric(X'X)))
+fopnorm5(X) = sqrt(opnorm(Symmetric(X*X')))
+
+fopnorms = [fopnorm1,fopnorm2,fopnorm3,fopnorm4,fopnorm5]
+
+svdvals_top(X) = svdvals(X)[1]
+fsvdvals1(X) = svdvals_top(X)
[…]

### 2025-02-04 src/Core/blockupdates.jl
typo in lipschitz error message

@@ -104,7 +104,7 @@ function make_lipschitz(T::CPDecomposition, n::Integer, Y::AbstractArray; object
         return lipschitz_matrix
 
     else
-        error("No $(n)th factor in Tucker")
+        error("No $(n)th factor in CPDecomposition")
     end
 end
 

### 2025-02-04 paper/fixtypst.jl
fix centered dots and add symbol for warning callouts

@@ -32,6 +32,11 @@ patterns = (
     # Missing callout box icons
     "fa-lightbulb()" => "\"❔\"",
     "fa-info()" => "\"❕\"",
+    "fa-exclamation-triangle()" => "\"⚠\"",
+
+    # \dots always gets converted to dots.h, sometimes it should be dots.h.c (TeX's \cdots)
+    "times dots.h times" => "times dots.h.c times",
+    "times.circle dots.h times.circle" => "times.circle dots.h.c times.circle",
 
     # Custom citation style in bibliography
     #".bib\")" => ".bib\", style: \"citationstyles/ieee-compressed-in-text-citations.csl\")",

### 2025-02-04 paper/paper.qmd
first pass of sub-block descent section

@@
 
+::: {#def-outer-product}
+
 The outer product $\otimes$ between two tensors $A\in\mathbb{R}^{I_1\times\dots\times I_M}$ and $B\in\mathbb{R}^{J_1\times\dots\times J_N}$ yields an order $M+N$ tensor $A\otimes B \in \mathbb{R}^{I_1\times\dots\times I_M\times J_1\times\dots\times J_N}$ that is entry-wise
@@
 $$
+
+:::
 
@@
 
+The $n$-slice product of a tensor with itself $X \cdot_n X$ should be thought of as a generalization of the Gram matrix $X^\top X$ since it considers the matrix generated by taking the dot product between every $n$th mode slice, just like how the Gram matrix considers the dot product between every pair of columns.
+
 The $n$-mode product $\times_n$ between a tensor $A\in\mathbb{R}^{I_1 \times \dots \times I_N}$ and matrix $B\in\mathbb{R}^{I_n\times J}$, returns a tensor $(A \times_n B)\in\mathbb{R}^{I_1 \times \dots \times I_{n-1} \times J \times I_{n+1} \times \dots \times I_N}$ with entries
@@
 
-The (complete) multi-mode product $\times_{1,\dots,n}=\times_{[n]}$ between a tensor $A\in\mathbb{R}^{I_1 \times \dots \times I_N}$ and tensor $B\in\mathbb{R}^{I_1\times\dots\times I_n}$, returns a tensor $(A \times_{[n]} B)\in\mathbb{R}^{I_{n+1} \times \dots \times I_N}$ with entries
-
-$$
-(A \times_{[n]} B)[i_{n+1},\dots,i_N] = \sum_{i_1, \dots ,i_n} A[i_1, \dots,i_n,i_{n+1},\dots,i_N]B[i_1, \dots ,i_n].
-$$
+The multi-mode product $\times_{1,\dots,n}=\times_{1:n}= \times_{[n]}$ between a tensor $A\in\mathbb{R}^{I_1 \times \dots \times I_{N}}$ and tensor $B\in\mathbb{R}^{I_1\times \dots\times I_{n}}$, returns a tensor $(A \times_{[n]} B)\in\mathbb{R}^{I_{n+1} \times \dots \times I_{N}}$ with entries
+
+$$
+(A \times_{[n]} B)[i_{n+1},\dots,i_{N}] = \sum_{i_1, \dots ,i_n} A[i_1, \dots,i_{n}, i_{n+1}, \dots,i_{N}]B[i_1,\dots,i_{n}].
+$$
+
+This product contracts the first $n$ indexes of $A$ with every index of $B$.
+
+More generally, we can contract any number of indexes such as the last $n$ indexes of $A$ with every index of $B$ with $\times_{N-n+1,\dots,N}=\times_{(N-n+1):N}=\times_{[-n]}$,
+
+$$
+(A \times_{[-n]} B)[i_1, \dots,i_{n}] = \sum_{i_{n+1}, \dots,i_{N}} A[i_1, \dots,i_{N-n}, i_{N-n+1}, \dots,i_{N}]B[i_{N-n+1}, \dots,i_{N}],
+$$
+
+or specific indexes.
+For example, we would define $(A\times_{1,3,5}B)\in\mathbb{R}^{I_2\times I_4\times I_6}$ where $A\in\mathbb{R}^{I_1\times\dots\times I_6}$ and $B\in\mathbb{R}^{I_1\times I_3\times I_5}$ to be
+
+$$
+(A \times_{1,3,5} B)[i_2,i_4,i_6] = \sum_{i_1,i_3,i_5} A[i_1,i_2,i_3,i_4,i_5,i_6]B[i_1,i_3,i_5].
+$$
+
+When $A$ a *half-symmetric*^[For example, the Hessian of a scalar function (see @def-hessian).] tensor of order $2N$
+
+$$
+A[i_1, \dots,i_{N}, i_{N+1}, \dots,i_{2N}] = A[i_{N+1}, \dots,i_{2N},i_1, \dots,i_{N}],
+$$ {#eq-half-symmetric}
+
+we have
+
+$$
+A \times_{[N]} B = A \times_{[-N]} B
+$$
+
+for tensors $B$ of order $N$.
+
+<!-- The half product $\times$ between a tensor $A\in\mathbb{R}^{I_1 \times \dots \times I_{2N}}$ and tensor $B\in\mathbb{R}^{I_2\times I_4\times\dots\times I_{2N}}$, returns a tensor $(A \times B)\in\mathbb{R}^{I_1 \times I_3 \times \dots \times I_{2N-1}}$ with entries
+
+$$
+(A \times B)[i_1,i_3,\dots,i_{2N-1}] = \sum_{i_1, \dots ,i_n} A[i_1, \dots,i_{2N}]B[i_2,i_4,\dots,i_{2N}].
+$$
+
+This product contracts every other index of $A$ with every index of $B$. -->
 
@@
 
+::: {#def-gradient}
+
+# Gradient
+
 The gradient $\nabla f:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}^{I_1\times \dots \times I_N}$ of a (differentiable) function $f:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}$ is defined entry-wise for a tensor $A\in \mathbb{R}^{I_1\times \dots \times I_N}$ by
@@
 
-The Hessian $\nabla^2 f:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}^{(I_1\times \dots \times I_N)^2}$ of a (second-differentiable) function $f:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}$ is the gradient of the gradient and is defined for a tensor $A\in\mathbb{R}^{I_1\times \dots \times I_N}$ entry-wise by
+:::
+
+::: {#def-hessian}
+
+# Hessian
+
+The Hessian $\nabla^2 f:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}^{(I_1\times \dots \times I_N)^2}$ of a second-differentiable function $f:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}$ is the gradient of the gradient and is defined for a tensor $A\in\mathbb{R}^{I_1\times \dots \times I_N}$ entry-wise by
 
@@
 
+:::
+
 For a tensor input $A$ of order $N$, the Hessian tensor $\nabla^2 f(A)$ is of order $2N$.
 
+See @sec-hessian-from-gradient for how this definition can be reproduced by performing two gradients $\nabla^2 f = \nabla(\nabla f)$.
+
+<!-- ::: {.callout-note}
+We could have defined the Hessian of a scalar function as
+
+$$
+\nabla^2 f(A) [i_1,j_1, \dots, i_N, j_N] = \frac{\partial^2 f}{\partial A[i_1, \dots , i_N]\partial A[j_1, \dots , j_N]} (A).
+$$
+
+Notice how the dimensions are reordered to have all $i_n$'s followed by all the $j_n$'s rather than paired together in our definition.
+Pairing them together as shown in @eq-hessian often leads to cleaner results when using the usual outer product $\otimes$.
+::: -->
+
+::: {#def-frobenius-norm}
+
 The Frobenius norm of a tensor $A$ is the square root of its dot product with itself
@@
 
+:::
+
 For vectors $v$, this is equivalent to the (Euclidean) 2-norm
@@
 
-For matrices $M$, the (operator) 2-norm is defined as
-
-$$
-\left\lVert M \right\rVert_{\text{op}} = \left\lVert M \right\rVert_2 = \sup_{\lVert v \rVert_2 = 1} \left\lVert Mv \right\rVert_2 = \sigma_1(M)
+For matrices $M$, the ($2\to 2$) operator norm is defined as
+
+$$
+\left\lVert M \right\rVert_{\text{op}} = \sup_{\lVert v \rVert_2 = 1} \left\lVert Mv \right\rVert_2 = \sigma_1(M)
 $$
@@
 
-For tensors $T$, the (operator) 2-norm needs to be defined in terms of how we treat them as function on other tensors.
-There is a canonical way to do this for vectors $x \to v^{\top}x$ and matrices $x \to Mx$, but not tensors.
-This is relevant to @sec-lipschitz-computation where the Lipschitz step-size is computed in terms of the Operator norm of the Hessian of our objective function.
-In the case of the Hessian tensor $\nabla^2 f(A)$ evaluated at $A$, we consider the function $X \to \nabla^2 f(A) \times_{[N]} X$.
-This gives us
-
-$$
-\left\lVert \nabla^2 f(A) \right\rVert_{\text{op}} = \sup_{\left\lVert X \right\rVert_F = 1} \left\lVert \nabla^2 f(A) \times_{[N]} X \right\rVert_F = \sigma_1(\text{mat}(\nabla^2 f(A)))
-$$
-
+For tensors $T$, the operator-norm is ambiguous since there are multiple ways we can treat tensors as function on other tensors.
+There is a canonical way to do this for vectors $x \mapsto v^{\top}x$ and matrices $x \mapsto Mx$, but not tensors.
+In the case of the Hessian tensor $\nabla^2 f(A)\in\mathbb{R}^{(I_1\times\dots\times I_N)^2}$ evaluated at $A\in\mathbb{R}^{I_1\times\dots\times I_N}$, it is natural to consider the function $X \mapsto \nabla^2 f(A) \times_{[N]} X$ for $X\in\mathbb{R}^{I_1\times\dots\times I_N}$.
+This gives us our definition of the operator norm on tensors.
+
+::: {#def-operator-norm}
+
+# Operator Norm
+
+The operator norm of a half-symmetric tensor $A\in\mathbb{R}^{(I_1\times\dots\times I_N)^2}$ (@eq-half-symmetric) is defined as
+
+$$
+\left\lVert A \right\rVert_{\text{op}} = \sup_{\left\lVert X \right\rVert_F = 1} \left\lVert A \times_{[N]} X \right\rVert_F.
+$$ {#eq-operator-norm}
+
+In @eq-operator-norm, $X\in\mathbb{R}^{I_1\times\dots\times I_N}$. Note that this definition agrees with the usual operator norm on matrices when $N=1$.
+
+:::
+
+::: {#thm-operator-norm-outer-product}
+
+# Norm of an outer product
+
+Let $T = A_1 \otimes \dots \otimes A_N \in \mathbb{R}^{(I_1\times \dots \times I_N)^2}$ where $A_n\in\mathbb{R}^{I_n\times I_n}$ are symmetric matrices.
+
+Then $T$ is half-symmetric and
+
+$$
+\left\lVert T \right\rVert_{\text{op}} = \prod_{n=1}^N \left\lVert A_n \right\rVert_{\text{op}}.
+$$
+
+:::
+
+::: {.callout-warning}
+
+According to how the outer product $\otimes$ is defined in @def-outer-product, the product $A_1 \otimes \dots \otimes A_N$ shown in @thm-operator-norm-outer-product is really an element of $\mathbb{R}^{I_1\times I_1 \times \dots \times I_N\times I_N}$.
+Note how the indexes are ordered differently than an element of $\mathbb{R}^{(I_1\times \dots \times I_N)^2} = \mathbb{R}^{I_1\times \dots \times I_N \times I_1\times \dots \times I_N}$.
+Correcting for this with explicit notation becomes cumbersome and would require tensor transposes, a new definition of an outer product, or reordering of indexes in the definition of a half-symmetric tensor.
+These can have knock-on effects to the definition of the Hessian, multi-mode product, and the operator norm.
+
+To avoid the headache, the equality
+
+$$
+T=A_1 \otimes \dots \otimes A_N
+$$
+
+should be thought of as the following entry-wise equation
+
+$$
+T[i_1,\dots,i_N,j_1,\dots,j_N] = A_1[i_1,j_1] \cdots A_N[i_N,j_N].
+$$ {#eq-corrected-outer-product}
+
+With the outer product understood as @eq-corrected-outer-product, the results of @thm-operator-norm-outer-product that $T$ is half-symmetric and its operator norm is the product of the operator norms of the constituent matrices is true.
+
+:::
+
+<!-- ^[This is relevant to @thm-quadratic-smoothness and @sec-lipschitz-computation where the Lipschitz step-size is computed in terms of the Operator norm of the Hessian of our objective function.] -->
+
+<!-- = \sigma_1(\text{mat}(\nabla^2 f(A)))
 where $\text{mat}:\mathbb{R}^{(I_1\times \dots \times I_N)^2}\to\mathbb{R}^{(I_1 \dots I_N)\times (I_1 \dots I_N)}$ takes a $2N$-ordered tensor and flattens it into a square $(I_1 \dots I_N)\times (I_1 \dots I_N)$ matrix.
+-->
+
+::: {#def-lipschitz}
+
+# Lipschitz Function
 
@@
 
+:::
+
+::: {#def-smooth}
+
+# Smooth Function
+
 A differentiable function $f:\mathbb{R}^{I_1\times \dots \times I_N}\to \mathbb{R}$ is $L$-smooth when its gradient $g=\nabla f$ is $L$-Lipschitz.
+
+:::
 
@@
 
-Then
+Then:
 
@@
 $$
-L = \left\lVert D \right\rVert _{\text{op}} = \sigma_1(\operatorname{mat}D).
+L = \left\lVert D \right\rVert _{\text{op}}.
 $$
@@
 
-For the family of Tucker decompositions, we can compute the Lipschitz constants efficiently similar to how we compute the gradient in @sec-gradient-computation.
-
-::: callout-note
+For the family of Tucker decompositions, we can compute the Lipschitz constants of the gradient efficiently similar to how we compute the gradient in @sec-gradient-computation with the following corollaries of @thm-quadratic-smoothness.
+
+::: {#cor-least-squares-matrix}
+
+Let $B\in\mathbb{R}^{R_1\times\dots\times R_N}$, $A_m\in\mathbb{R}^{I_m\times R_m}$, and $Y\in\mathbb{R}^{I_1\times\dots\times I_N}$. The function
+
+$$
+f(A)=\frac{1}{2} \left\lVert [B; A_1,\dots,A_{n-1},A,A_{n+1},\dots,A_N] - Y \right\rVert_F^2
+$$
+
+is quadratic, and $L$-smooth with constant
+
+$$
+L_{A_n} = \left\lVert \tilde{X}_n \cdot_n \tilde{X}_n \right\rVert _{\text{op}}
+$$
+
+where
+
+$$
+\tilde{X}_n = (B \bigtimes_{m\neq n} A_m) = \llbracket B; A_1,\dots,A_{n-1}, \text{id}_{R_n}, A_{n+1},\dots,A_{N} \rrbracket.
+$$
+
+::: {.proof}
+The result follows from @thm-operator-norm-outer-product and @thm-quadratic-smoothness.
+:::
+
+:::
+
+::: {#cor-least-squares-core}
+
+Let $A_n\in\mathbb{R}^{I_n\times R_n}$, and $Y\in\mathbb{R}^{I_1\times\dots\times I_N}$. The function
+
+$$
+f(B)=\frac{1}{2} \left\lVert [B; A_1,\dots,A_N] - Y \right\rVert_F^2
+$$
+
+is quadratic, and $L$-smooth with constant
+
+$$
+L_{B} = \prod_{n=1}^N\left\lVert A_n^\top A_n \right\rVert _{\text{op}}.
+$$
+
+::: {.proof}
+The result follows from @thm-operator-norm-outer-product and @thm-quadratic-smoothness.
+:::
+
+:::
+
+This yields the following efficient implementations.
+
+::: {.callout-note}
+
+It is tempting to use the identity $\left\lVert A^\top A \right\rVert_{\text{op}} = \left\lVert A \right\rVert_{\text{op}}^2$ to calculate the Lipschitz constant without forming $A^\top A$.
+For tall dense matrices, using this identity is slower and more memory intensive as of Julia 1.11.2.
+See [LinearAlgebra.jl issue 1185](https://github.com/JuliaLang/LinearAlgebra.jl/issues/1185) on Github.
+
+:::
+
+<!-- ::: callout-note
 For the family of Tucker decompositions, the objective function $f$ is "block-quadratic" with respect to the factors.
@@
 See Obsidian note "Gradients Hessians and Lipschitz Constants of a Tucker Decomposition".
-
-```julia
-function make_lipschitz(T::Tucker1, n::Integer, Y::AbstractArray; objective::L2, kwargs...)
-    if n==0 # the core is the zeroth factor
-        function lipschitz0(T::Tucker1; kwargs...)
-            A = matrix_factor(T, 1)
-            return opnorm(A'A)
-        end
-        return lipschitz0
-
-    elseif n==1 # the matrix is the zeroth factor
-        function lipschitz1(T::Tucker1; kwargs...)
-            C = core(T)
-            return opnorm(slicewise_dot(C, C))
-        end
-        return lipschitz1
-
-    else
-        error("No $(n)th factor in Tucker1")
-    end
-end
-
+-->
+```julia
 function make_lipschitz(T::Tucker, n::Integer, Y::AbstractArray; objective::L2, kwargs...)
@@
     if n==0 # the core is the zeroth factor
-        function lipschitz_core(T::AbstractTucker; kwargs...)
-            #matrices = matrix_factors(T)
-            #gram_matrices = map(A -> A'A, matrices)
-            #return prod(opnorm.(gram_matrices))
-            return prod(A -> opnorm(A'A), matrix_factors(T))
+        function lipschitz_core(X::AbstractTucker; kwargs...)
+            return prod(A -> opnorm(A'A), matrix_factors(X))
         end
@@
     elseif n in 1:N # the matrix is the zeroth factor
-        function lipschitz_matrix(T::AbstractTucker; kwargs...)
-            matrices = matrix_factors(T)
-            TExcludeAn = tuckerproduct(core(T), matrices; exclude=n)
-            return opnorm(slicewise_dot(TExcludeAn, TExcludeAn; dims=n))
+        function lipschitz_matrix(X::AbstractTucker; kwargs...)
+            B = core(X)
+            matrices = matrix_factors(X)
+            X̃ₙ = tuckerproduct(B, matrices; exclude=n)
+            return opnorm(slicewise_dot(X̃ₙ, X̃ₙ; dims=n))
         end
@@
 end
-
+```
+
+In the case of Tucker decomposition, the Lipschitz constants simplify to
+
+$$
+L_{B} = \left\lVert A^{\top}A\right\rVert_{\text{op}}, \quad L_{B} = \left\lVert B \cdot_{1} B\right\rVert_{\text{op}}.
+$$
+
+```julia
+function make_lipschitz(T::Tucker1, n::Integer, Y::AbstractArray; objective::L2, kwargs...)
+    if n==0 # the core is the zeroth factor
+        function lipschitz0(X::Tucker1; kwargs...)
+            A = matrix_factor(X, 1)
+            return opnorm(A'A)
+        end
+        return lipschitz0
+
+    elseif n==1 # the matrix is the zeroth factor
+        function lipschitz1(X::Tucker1; kwargs...)
+            B = core(X)
+            return opnorm(slicewise_dot(B, B))
+        end
+        return lipschitz1
+
+    else
+        error("No $(n)th factor in Tucker1")
+    end
+end
+```
+
+Lastly, for CP decomposition, the Lipschitz constants for the matrices can be calculated similarly to the Tucker decomposition.
+
+```julia
 function make_lipschitz(T::CPDecomposition, n::Integer, Y::AbstractArray; objective::L2, kwargs...)
     N = ndims(T)
-    if n in 1:N # the matrix is the zeroth factor
-        function lipschitz_matrix(T::AbstractTucker; kwargs...)
-            matrices = matrix_factors(T)
-            TExcludeAn = tuckerproduct(core(T), matrices; exclude=n) # TODO optimize this to avoid making the super diagonal core
-            return opnorm(slicewise_dot(TExcludeAn, TExcludeAn; dims=n))
+    if n in 1:N
+        function lipschitz_matrix(X::CPDecomposition; kwargs...)
+            id = core(X)
+            matrices = matrix_factors(X)
+            X̃ₙ = tuckerproduct(id, matrices; exclude=n)
+            return opnorm(slicewise_dot(X̃ₙ, X̃ₙ; dims=n))
         end
@@
     else
-        error("No $(n)th factor in Tucker")
+        error("No $(n)th factor in CPDecomposition")
     end
@@
 
-In [@lee_algorithms_2000], the multiplicative update algorithm used to solve nonnegative matrix factorization problems is written in a way so that each row of the first factor
+In the case of Tucker decomposition, one modification of the update shown in @eq-proximal-explicit would be to update each column $a_{n,r}$, $r=1,\dots,R_n$ of the matrix $A_n$ separately.
+This would be suitable if the constraint that $A_N\in\mathcal{C}_n$ can be broken up further into the constraints $a_{n,r}\in\mathcal{C}_{n,r}$.
+This is shown in the following update scheme:
+
+$$
+a_{n,r}^{t+1} \leftarrow P_{\mathcal{C}}_{n,r}\left( a_{n,r}^{t} - \frac{1}{L_{n,r}^t}\nabla f_{n,r}^t(a_{n,r}^{t})\right),
+$$ {#eq-sub-block-update-proper}
+
+where $f_{n,r}^t(a)=\frac{1}{2} \left\lVert [B; A_1^{t+1},\dots,A_{n-1}^{t+1},A_{n,r}^t(a),A_{n+1}^{t},\dots,A_N^{t}] - Y \right\rVert_F^2$ and
+
+$$
+A_{n,r}^t(a) = \begin{bmatrix}
+\uparrow &  & \uparrow & \uparrow & \uparrow &  & \uparrow \\
+a_{n,1}^{t+1} & \cdots & a_{n,r-1}^{t+1}  & a  & a_{n,r+1}^{t}  & \cdots & a_{n,R_{n}}^{t}  \\
+\downarrow &  & \downarrow & \downarrow & \downarrow &  & \downarrow \\
+\end{bmatrix}.
+$$ {#eq-proper-column-update}
+
+In theory, the block update shown in @eq-sub-block-update-proper should be a bit more expensive than using the larger blocks on the matrices $A$ shown in @eq-proximal-explicit, since the gradient needs to be recomputed $R_n$ times for each matrix block $n$, rather than only computing the gradient once per block $n$.
+To get around this, we use the fact that $\nabla f_{n,r}^t(a)$ is the $r$th column from the gradient $\nabla f_{n}^t(A)$ where $f_{n}^t(A)=\frac{1}{2} \left\lVert [B; A_1^{t+1},\dots,A_{n-1}^{t+1},A,A_{n+1}^{t},\dots,A_N^{t}] - Y \right\rVert_F^2$.
+So we can approximate @eq-sub-block-update-proper by first calculating the gradient $\nabla f_{n}^t$ at
+
+$$
+\hat{A}_{n}^t = \begin{bmatrix}
+\uparrow &  & \uparrow & \uparrow & \uparrow &  & \uparrow \\
+a_{n,1}^{t} & \cdots & a_{n,r-1}^{t}  & a_{n,r}^{t}  & a_{n,r+1}^{t}  & \cdots & a_{n,R_{n}}^{t}  \\
+\downarrow &  & \downarrow & \downarrow & \downarrow &  & \downarrow \\
+\end{bmatrix},
+$$ {#eq-merged-column-update}
+
+and then updating each sub-block $r$ according to
+
+$$
+a_{n,r}^{t+1} \leftarrow P_{\mathcal{C}}_{n,r}\left( a_{n,r}^{t} - \frac{1}{L_{n,r}^t}\nabla f_{n}^t(\hat{A}_{n}^t)\right).
+$$ {#eq-sub-block-update-half-merged}
+
+Note the difference between @eq-proper-column-update and @eq-merged-column-update is that we don't use the most recent columns $a_{n,j}$ for $j<r$ in @eq-merged-column-update.
+
+The update given in @eq-sub-block-update-half-merged can be merged back to an update on the whole block $A_n$
+
+$$
+A_{n}^{t+1} \leftarrow P_{\mathcal{C}}_{n}\left( A_{n}^{t} -\nabla f_{n}^t(A_{n}^t)(\hat{L}_{n}^t)^{-1} \right)
+$$ {#eq-sub-block-update}
+
+where we have the $R_n\times R_n$ diagonal "Lipschitz Matrix"
+
+$$
+\hat{L}_{n}^t = \begin{bmatrix}
+{L}_{n,1}^t & 0 &  & 0 \\
+0 & {L}_{n,2}^t &  & 0 \\
+ &  & \ddots & \vdots \\
+0 & 0 & \cdots & {L}_{n,R_{n}}^t
+\end{bmatrix}.
+$$
+
+It is not too hard to show that ${L}_{n,r}^t$ is the Euclidean norm of the $r$th column^[We could have used the $r$th row of $\tilde{X}_n \cdot_n \tilde{X}_n$ since this matrix is symmetric.
+Since Julia store matrices in column-major order, many operations that perform column-wise are more efficient than their equivalent row-wise operation.] of the matrix $\tilde{X}_n \cdot_n \tilde{X}_n$ from @cor-least-squares-matrix,
+
+$$
+{L}_{n,r}^t = \left\lVert \left(\tilde{X}_n \cdot_n \tilde{X}_n \right)[:, r] \right\rVert _{2}.
+$$
+
+This leads to the following efficient calculation of the Lipschitz matrix in Julia.
+
+```julia
+function diagonal_lipschitz_matrix(T::Tucker, n::Int; kwargs...)
+    B = core(T)
+    matrices = matrix_factors(T)
+    X̃ₙ = tuckerproduct(B, matrices; exclude=n)
+    return Diagonal_col_norm(slicewise_dot(X̃ₙ, X̃ₙ; dims=n))
+end
+
+Diagonal_col_norm(X) = Diagonal(norm.(eachcol(X)))
+```
+
+We can now compare the merged sub-block update @eq-sub-block-update to the standard projected gradient descent update shown in @eq-proximal-explicit.
+The difference is that we calculate a "matrix step-size" $\hat{L}_{n}^t \in\mathbb{R}^{R_n \times R_n}$ rather than a scalar ${L}_{n}^t\in\mathbb{R}$.
+In practice, this leads to an improvement in convergence speed for two reasons.
+
+First, computing the matrix $\hat{L}_{n}^t$ often faster than the scalar ${L}_{n}^t=\left\lVert \tilde{X}_n \cdot_n \tilde{X}_n \right\rVert _{\text{op}}$.
+The former only requires calculating the Euclidean norm of $R$ vectors for a total cost of $2R_{n}^2$ floating point operations (FLOPs), whereas the latter requires the top eigenvalue of $\tilde{X}_n \cdot_n \tilde{X}_n$.
+This is usually done with a power method or truncated SVD which can be costlier than the flat rate of $2R_{n}^2$ FLOPs.
+
+Secondly, using the matrix $\hat{L}_{n}^t$ means columns where ${L}_{n,r}^t$ is small can take larger descent steps.
+This is because the largest singular value of $\tilde{X}_n \cdot_n \tilde{X}_n$ is an upper bound on the Euclidean norm of each column: ${L}_{n,r}^t \leq {L}_{n}^t$.
+Using the scaler Lipschitz ${L}_{n}^t$ is equivalent to the diagonal matrix
+
+$$
+D = \begin{bmatrix}
+{L}_{n}^t & 0 &  & 0 \\
+0 & {L}_{n}^t &  & 0 \\
+ &  & \ddots & \vdots \\
+0 & 0 & \cdots & {L}_{n}^t
+\end{bmatrix}
+$$
+
+in the merged sub-block update shown in @eq-sub-block-update.
+So each column of $A_n$ is forced to use the worst case (largest) singular value of $\tilde{X}_n \cdot_n \tilde{X}_n$.
+In this way, the matrix $\hat{L}_{n}^t$ acts like a cheap approximate Hessian as if we were doing a quasi-Newton update with step-size $1$.
+
+For completeness, we can perform the same merged sub-block update to update the core $B$.
+In this case, we obtain the more complicated "Lipschitz tensor" $\hat{L}_{B}^t \in\mathbb{R}^{(R_1\times\dots\times R_N)^2}$ defined by
+
+$$
+\hat{L}_{B}^t = \hat{L}_{B,1}^t \otimes \dots \otimes \hat{L}_{B,N}^t
+$$
+
+where each matrix $\hat{L}_{B,n}^t \in\mathbb{R}^{R_n\times R_n}$ is diagonal with non-zero entries
+
+$$
+{L}_{B,n}^t [r,r] = \left\lVert \left((A_n^t)^\top A_n^t\right)[:, r] \right\rVert _{2}.
+$$
+
+The merged sub-block update for the core becomes
+
+$$
+B^{t+1} \leftarrow P_{\mathcal{C}}_{B}\left( B^{t} - \nabla f_{0}^t(B^t) \times_B (\hat{L}_{B}^t)^{-1} \right)
+$$ {#eq-sub-block-update-core}
+
+with the multiplication
+
+$$
+\begin{align}
+\nabla f_{0}^t(B^t) \times_B (\hat{L}_{B}^t)^{-1} &= \nabla f_{0}^t(B^t)  \bigtimes_{ n} (\hat{L}_{B,n}^{t})^{-1} = \llbracket \nabla f_{0}^t(B^t) ; (\hat{L}_{B,1}^{t})^{-1},\dots,(\hat{L}_{B,n}^{t})^{-1} \rrbracket.
+\end{align}
+$$
+
+This should be thought of as normalizing each dimension of the tensor $\nabla f_{0}^t(B^t)$ so that we can take a unit step-size.
+
+Putting the core and matrices Lipschitz calculations together gives us the following Julia code.
+Note we store $\hat{L}_{B}^t$ in factored form as a tuple of diagonal matrices to save space and computation.
+
+```julia
+function make_block_lipschitz(T::AbstractTucker, n::Integer, Y::AbstractArray; objective::L2, kwargs...)
+    N = ndims(T)
+    if n==0 # the core is the zeroth factor
+        function lipschitz_core(X::AbstractTucker; kwargs...)
+            return map(A -> Diagonal_col_norm(A'A), matrix_factors(X))
+        end # Returns a tuple of diagonal matrices
+        return lipschitz_core
+
+    elseif n in 1:N
+        function lipschitz_matrix(X::AbstractTucker; kwargs...)
+            matrices = matrix_factors(X)
+            X̃ₙ = tuckerproduct(core(X), matrices; exclude=n)
+            return Diagonal_col_norm(slicewise_dot(X̃ₙ, X̃ₙ; dims=n))
+        end
+        return lipschitz_matrix
+
+    else
+        error("No $(n)th factor in Tucker")
+    end
+end
+```
 
@@
 - This is compatible with sub-block descent with appropriately defined matrix operations
+
+In practice, we find that extrapolating the iterate based on the prior iterate
+
+$$
+\hat{A}_n^t \leftarrow A_n^t + \omega_n^t (A_n^{t} - A_n^{t-1})
+$$
+
+for some amount of extrapolation $\omega_n^t \geq 0$ before applying the update @eq-sub-block-update greatly improves the speed of descent.
+This can be thought of as a type of momentum where we continue to move in directions that showed a lot of improvement during the last iteration.
+
+Our selection for $\omega_n^t$ follows Xu and Yin's method for block coordinate descent @xu_BlockCoordinateDescent_2013, which is itself inspired by Tseng and Yun's coordinate gradient descent method @tseng_coordinate_2009.
+
+TODO CONTINUE FROM HERE
+
+```julia
+function (U::MomentumUpdate)(x::T; x_last::T, ω, δ, kwargs...) where T
+    n = U.n
+    if checkfrozen(x, n)
+        return x
+    end
+    # TODO avoid redoing this lipschitz calculation and instead store the previous L
+    # TODO generalize this momentum update to allow for other decaying momentums ω
+    L = U.lipschitz(x; kwargs...)
+    L_last = U.lipschitz(x_last; kwargs...)
+    ω = min.(ω, δ .* .√(L_last/L))
+
+    a, a_last = factor(x, n), factor(x_last, n)
+
+    # Equivalent mathematically, but slightly less efficient ways of applying momentum
+    # @. a = a + ω * (a - a_last)
+
+    # @. a += ω * (a - a_last)
+
+    # a .*= 1 + a
+    # a .-= ω .* a_last
+
+    a = U.combine(a, id + ω) # handle diagonal Lipschitz constants
+    a .-= U.combine(a_last, ω)
+end
+```
+
+```julia
+"""update parameters needed for the update"""
+function initialize_parameters(decomposition, Y, previous; momentum::Bool, random_order, recursive_random_order, kwargs...)
+	# parameters for the update step are symbol => value pairs
+	# they are held in a dictionary since we may mutate these for ex. the stepsize
+	parameters = Dict{Symbol, Any}()
+
+	# General Looping
+	parameters[:iteration] = 0
+	parameters[:x_last] = previous[begin] # Last iterate
+	parameters[:random_order] = random_order
+	parameters[:recursive_random_order] = recursive_random_order
+
+	# Momentum
+	if momentum
+		parameters[:t_last] = float(1) # need this field to hold Floats, not Ints
+		parameters[:t] = update_t(float(1))
+		parameters[:ω] = (parameters[:t_last] - 1) / parameters[:t]
+		parameters[:δ] = kwargs[:δ]
+	end
+
+	function updateparameters!(parameters, decomposition, previous)
+		parameters[:iteration] += 1
+		#parameters[:x_last] = previous[begin] # Last iterate updated with updateprevious!
+											   # this ensures we only copy over the decomposition
+											   # once since updateparameters! happens before updateprevious!
+
+											   # This is commented since parameters[:x_last] already points to previous[begin]
+		# Momentum
+		if momentum
+			parameters[:t_last] = parameters[:t]
+			parameters[:t] = update_t(parameters[:t_last])
+			parameters[:ω] = (parameters[:t_last] - 1) / parameters[:t]
+		end
+	end
+
+	return parameters, updateparameters!
+end
+```
 
@@
 - like auto-diff for factorizations
+
+# Appendix {.appendix}
+
+## Building the Hessian from two gradients {#sec-hessian-from-gradient}
+
+To build the Hessian from the definition of the gradient, we first extend the gradient to tensor-valued functions.
+For a function $F:\mathbb{R}^{J_{1}\times \dots \times J_{N}}\to\mathbb{R}^{I_{1}\times\dots \times I_{M}}$ where
+
+$$
+F(X)=[f_{i_{1}\dots i_{M}}(X)]
+$$
+
+is a tensor of scalar functions $f_{i_{1}\dots i_{M}}:T\to \mathbb{R}$, the gradient of $F$ at $X$ is defined entry-wise as
+
+$$
+\begin{align}
+\nabla F(X)[i_{1},\dots,i_{M}][j_{1},\dots,j_{N}]&=\nabla f_{i_{1}\dots i_{M}}(X)[j_{1},\dots,j_{N}] \\
+&=\frac{{\partial f_{i_{1}\dots i_{M}}}}{\partial X[j_{1},\dots,j_{N}]} (X).
+\end{align}
+$$
+
+This treats the gradient at $X$ as a tensor of tensors $\nabla F(X)\in (\mathbb{R}^{I_{1}\times\dots \times I_{M}})^{J_{1}\times \dots \times J_{N}}$.
+This is naturally isomorphic to a tensor of order $M+N$ with entries
+
+$$
+\nabla F(X)[i_{1},\dots,i_{M},j_{1},\dots,j_{N}]=\frac{{\partial f_{i_{1}\dots i_{M}}}}{\partial X[j_{1},\dots,j_{N}]} (X).
+$$ {#eq-tensor-gradient-entries}
+
+So we conclude that the gradient at $X$ of a tensor-valued function $F$ is $\nabla F: \mathbb{R}^{J_{1}\times \dots \times J_{N}}\to \mathbb{R}^{I_{1}\times\dots \times I_{M}\times J_{1}\times \dots \times J_{N}}$ is given by @eq-tensor-gradient-entries.
+
+We can define the Hessian of a scalar function $f:\mathbb{R}^{I_{1}\times\dots \times I_{N}}\to \mathbb{R}$ at $X$ as $\nabla^2 f(X) = \nabla (\nabla f) (X):\mathbb{R}^{I_{1}\times\dots \times I_{N}}\to \mathbb{R}^{(I_{1}\times\dots \times I_{N})^2}$.
+The inner nabla $\nabla$ is the gradient of the scalar function $f$, and the outer nabla $\nabla$ is the gradient of the tensor-valued function $\nabla f$.
+
+This means
+
+$$
+\nabla^2 f(A) [i_1, \dots, i_N, j_1, \dots , j_N] = \frac{\partial^2 f}{\partial A[j_1, \dots , j_N]\partial A[i_1, \dots , i_N]} (A),
+$$
+
+but if the function has continuous second derivatives, we can perform the partial derivatives in either order
+
+$$
+\frac{\partial^2 f}{\partial A[j_1, \dots , j_N]\partial A[i_1, \dots , i_N]} (A) = \frac{\partial^2 f}{\partial A[i_1, \dots , i_N]\partial A[j_1, \dots , j_N]} (A).
+$$

### 2025-02-04 paper/paper.qmd
First pass of momentum section

@@
 
-should be thought of as the following entry-wise equation
+in @thm-operator-norm-outer-product should be thought of as the following entry-wise equation
 
@@
 - Use Auto diff generally
-- But hand-crafted gradients and Lipschitz calculations *can* be faster (e.g. symmetrized slicewise dot product)
+- But hand-crafted gradients and Lipschitz calculations *can* be faster (e.g. symmetrized slice-wise dot product)
 
@@
 
-It is not too hard to show that ${L}_{n,r}^t$ is the Euclidean norm of the $r$th column^[We could have used the $r$th row of $\tilde{X}_n \cdot_n \tilde{X}_n$ since this matrix is symmetric.
+It is not too hard to show that the Lipschitz ${L}_{n,r}^t$ for $\nabla f_{n,r}^t$ is the Euclidean norm of the $r$th column^[We could have used the $r$th row of $\tilde{X}_n \cdot_n \tilde{X}_n$ since this matrix is symmetric.
 Since Julia store matrices in column-major order, many operations that perform column-wise are more efficient than their equivalent row-wise operation.] of the matrix $\tilde{X}_n \cdot_n \tilde{X}_n$ from @cor-least-squares-matrix,
@@
 \begin{align}
-\nabla f_{0}^t(B^t) \times_B (\hat{L}_{B}^t)^{-1} &= \nabla f_{0}^t(B^t)  \bigtimes_{ n} (\hat{L}_{B,n}^{t})^{-1} = \llbracket \nabla f_{0}^t(B^t) ; (\hat{L}_{B,1}^{t})^{-1},\dots,(\hat{L}_{B,n}^{t})^{-1} \rrbracket.
+\nabla f_{0}^t(B^t) \times_B (\hat{L}_{B}^t)^{-1} &= \nabla f_{0}^t(B^t)  \bigtimes_{ n} (\hat{L}_{B,n}^{t})^{-1} \\
+&= \llbracket \nabla f_{0}^t(B^t) ; (\hat{L}_{B,1}^{t})^{-1},\dots,(\hat{L}_{B,n}^{t})^{-1} \rrbracket.
 \end{align}
-$$
+$$ {#eq-tensor-matrices-product}
 
 This should be thought of as normalizing each dimension of the tensor $\nabla f_{0}^t(B^t)$ so that we can take a unit step-size.
+
+TODO use the multi-mode product in stead of defining a new multiplication $\times_B$ here. I think I'll have the same tensor product issue as before where the indices.
 
@@
 $$
-\hat{A}_n^t \leftarrow A_n^t + \omega_n^t (A_n^{t} - A_n^{t-1})
-$$
+\hat{A}_n^t \leftarrow A_n^t + \omega_n^t(A_n^{t} - A_n^{t-1})
+$$ {#eq-extrapolation-ideal}
 
@@
 
-Our selection for $\omega_n^t$ follows Xu and Yin's method for block coordinate descent @xu_BlockCoordinateDescent_2013, which is itself inspired by Tseng and Yun's coordinate gradient descent method @tseng_coordinate_2009.
-
-TODO CONTINUE FROM HERE
-
-```julia
-function (U::MomentumUpdate)(x::T; x_last::T, ω, δ, kwargs...) where T
+Our selection for $\omega_n^t$ follows Xu and Yin's method for block coordinate descent [@xu_BlockCoordinateDescent_2013], which is itself inspired by Tseng and Yun's coordinate gradient descent method [@tseng_coordinate_2009].
+
+Given a parameter^[Usually we pick a number close to $1$.
+For example, we use the default $\delta=0.9999$.] $\delta \in [0, 1)$, we define the momentum parameters and $τ^t$ and $\omega_n^t$ according to the following updates
+
+$$
+\begin{align*}
+\tau^0 &= 1 \\
+\tau^{t+1} &\leftarrow \frac{1}{2} \left(1+\sqrt{1+4\left(\tau^{t}\right)^2}\right) \\
+\hat{\omega}^{t} &\leftarrow \frac{\tau^{t} - 1}{\tau^{t+1}} \\
+\omega_n^{t} &\leftarrow \min\left(\hat{\omega}^{t}, \delta \sqrt{\hat{L}_n^{t-1} (\hat{L}_n^{t})^{-1}} \right).
+\end{align*}
+$$ {#eq-momentum-parameters}
+
+TODO notation is going to get confusing. We use hat/not hat $L$ for the scaler vs matrix/tensor version. But we use hat/not hat $\omega$ for the ideal vs clamped momentum.
+
+What is novel with our approach is that we perform this momentum on the Lipschitz matrices and tensors $\hat{L}_n^{t}$ rather than scaler Lipschitz constant ${L}_n^{t}$.
+In this way, we should interpret the operations shown in @eq-momentum-parameters as operating element-wise.
+This also means the momentum parameter $\omega_n^{t}$ is a matrix or tensor and takes the same shape as $\hat{L}_n^{t}$.
+
+In order to perform @eq-extrapolation-ideal, we use the equivalent but more efficient formulation
+
+$$
+\begin{align}
+\hat{A}_n^{t} &\leftarrow A_n^t (\text{id}_{R_n} + \omega_n^t) - {A}_n^{t-1} \omega_n^t.
+\end{align}
+$$
+
+```julia
+function (U::MomentumUpdate)(X::T; X_last::T, ω, δ, kwargs...) where T
     n = U.n
-    if checkfrozen(x, n)
-        return x
-    end
-    # TODO avoid redoing this lipschitz calculation and instead store the previous L
-    # TODO generalize this momentum update to allow for other decaying momentums ω
-    L = U.lipschitz(x; kwargs...)
-    L_last = U.lipschitz(x_last; kwargs...)
+
+    L = U.lipschitz(X; kwargs...)
+    L_last = U.lipschitz(X_last; kwargs...)
     ω = min.(ω, δ .* .√(L_last/L))
 
-    a, a_last = factor(x, n), factor(x_last, n)
-
-    # Equivalent mathematically, but slightly less efficient ways of applying momentum
-    # @. a = a + ω * (a - a_last)
-
-    # @. a += ω * (a - a_last)
-
-    # a .*= 1 + a
-    # a .-= ω .* a_last
-
-    a = U.combine(a, id + ω) # handle diagonal Lipschitz constants
-    a .-= U.combine(a_last, ω)
-end
-```
-
-```julia
-"""update parameters needed for the update"""
+    A, A_last = factor(X, n), factor(X_last, n)
+
+    A .= U.combine(A, id + ω)
+    A .-= U.combine(A_last, ω)
+end
+```
+
+In the code above, the momentum stores the factor $n$ in acts on, how to compute the Lipschitz constant, matrix, or tensor, and how to combine (multiply) the constant with the factor.
+In the case of matrix factors $A_n$ in a Tucker decomposition, this is simply right matrix-matrix multiplication.
+The core factor $B$ uses $\times_B$ as described in @eq-tensor-matrices-product.
+
+The parameters of the momentum update are handled separately.
+This is to treat the momentum update as "apply this update with these parameters".
+The parameters $\tau^t$ and $\hat{\omega}^t$ are updated by the following function that keeps track of all parameters needed to perform the iteration.
+In this case, we keep track of what iteration $t$ we are at, the previous iterate, and a few options for the order in which to cycle through and update the blocks.
+
+```julia
+update_τ(τ) = 0.5*(1 + sqrt(1 + 4*τ^2))
+
 function initialize_parameters(decomposition, Y, previous; momentum::Bool, random_order, recursive_random_order, kwargs...)
 	# parameters for the update step are symbol => value pairs
-	# they are held in a dictionary since we may mutate these for ex. the stepsize
+	# held in a dictionary since we may mutate these, e.g. the step-size
 	parameters = Dict{Symbol, Any}()
@@
 	parameters[:iteration] = 0
-	parameters[:x_last] = previous[begin] # Last iterate
+	parameters[:X_last] = previous[begin] # Last iterate
 	parameters[:random_order] = random_order
@@
 	if momentum
-		parameters[:t_last] = float(1) # need this field to hold Floats, not Ints
-		parameters[:t] = update_t(float(1))
-		parameters[:ω] = (parameters[:t_last] - 1) / parameters[:t]
+		parameters[:τ_last] = float(1) # need this field to hold Floats, not Ints
+		parameters[:τ] = update_τ(float(1))
+		parameters[:ω] = (parameters[:τ_last] - 1) / parameters[:τ]
 		parameters[:δ] = kwargs[:δ]
@@
 		parameters[:iteration] += 1
-		#parameters[:x_last] = previous[begin] # Last iterate updated with updateprevious!
-											   # this ensures we only copy over the decomposition
-											   # once since updateparameters! happens before updateprevious!
-
+		# parameters[:x_last] = previous[begin]
 # This is commented since parameters[:x_last] already points to previous[begin]
-		# Momentum
+
 		if momentum
-			parameters[:t_last] = parameters[:t]
-			parameters[:t] = update_t(parameters[:t_last])
-			parameters[:ω] = (parameters[:t_last] - 1) / parameters[:t]
+			parameters[:τ_last] = parameters[:τ]
+			parameters[:τ] = update_τ(parameters[:τ_last])
+			parameters[:ω] = (parameters[:τ_last] - 1) / parameters[:τ]
 		end
@@
 ```
+
+### Empirical evidence for
 

### 2025-02-04 src/Core/blockupdates.jl

@@ -648,7 +648,7 @@ function (U::MomentumUpdate)(x::T; x_last::T, ω, δ, kwargs...) where T
     # a .*= 1 + a
     # a .-= ω .* a_last
 
-    a = U.combine(a, id + ω) # handle diagonal Lipschitz constants
+    a .= U.combine(a, id + ω) # handle diagonal Lipschitz constants
     a .-= U.combine(a_last, ω)
 end
 

### 2025-02-05 paper/paper.qmd
first pass of section 4.1

@@
 
-### Empirical evidence for
+### Empirical Evidence for Sub-Block Descent and Momentum
+
+To showcase that the combination of these two tricks can speed up convergence, we will benchmark them by factorizing a random $10\times 10$ tensor (a matrix) with rank $3$.
+The Julia code is shown below, and the results are shown in @tbl-subblock-momentum-results.
+
+```julia
+using BenchmarkTools
+using BlockTensorDecomposition
+
+fact = BlockTensorDecomposition.factorize
+
+options = (
+    :rank => 3,
+    :tolerence => (1, 0.03),
+    :converged => (GradientNNCone, RelativeError),
+    :δ => 0.9,
+)
+
+n_subblock_n_momentum(Y) = fact(Y;
+    do_subblock_updates=false,
+    momentum=false,
+    options...
+)
+
+y_subblock_n_momentum(Y) = fact(Y;
+    do_subblock_updates=true,
+    momentum=false,
+    options...
+)
+
+n_subblock_y_momentum(Y) = fact(Y;
+    do_subblock_updates=false,
+    momentum=true,
+    options...
+)
+
+y_subblock_y_momentum(Y) = fact(Y;
+    do_subblock_updates=true,
+    momentum=true,
+    options...
+)
+
+I, J = 10, 10
+R = 3
+
+@benchmark n_subblock_n_momentum(Y) setup=(Y=Tucker1((I, J), R))
+@benchmark n_subblock_y_momentum(Y) setup=(Y=Tucker1((I, J), R))
+@benchmark y_subblock_n_momentum(Y) setup=(Y=Tucker1((I, J), R))
+@benchmark y_subblock_y_momentum(Y) setup=(Y=Tucker1((I, J), R))
+
+performance_increase(old, new) = (old - new) / new * 100
+```
+
+The code `Tucker1((I, J), R)` produces a random $I\times J$ rank-$R$ matrix by generating two matrices $A\in\mathbb{R}^{I \times R}$ and $B\in\mathbb{R}^{R \times J}$ with standard normal entries, and multiplies them together.
+
+|                   | **No Momentum**            | **Yes Momentum**           |
+|-------------------|----------------------------|----------------------------|
+| **No Sub-Block**  | 48.843 ms                  | 45.738 ms (6.7887% faster) |
+| **Yes Sub-Block** | 27.473 ms (77.785% faster) | **24.350 ms (100.59% faster)** |
+
+: Summary of median times to factorize a random $10\times 10$ rank-3 matrix under different methods.
+The performance increase is given by the formula $(\text{old} - \text{new}) / \text{new}$.
+{#tbl-subblock-momentum-results}
+
+In @tbl-subblock-momentum-results, you can see that having both sub-block descent and momentum yields the fastest factorization.
+Moreover, the performance increase is *more* than simply the performance increases obtained by exclusively sub-block or momentum alone.^[The expected performance increase if sub-block descent and momentum where independence would be $(1+0.067887)(1+0.77785) = 1.89854$ or only 89.854% faster.] This suggests that there is synergy with these two methods and are best used together.
 

### 2025-02-05 src/Core/blockupdates.jl
benchmark subblock and momentum

@@ -648,7 +648,7 @@ function (U::MomentumUpdate)(x::T; x_last::T, ω, δ, kwargs...) where T
     # a .*= 1 + a
     # a .-= ω .* a_last
 
-    a .= U.combine(a, id + ω) # handle diagonal Lipschitz constants
+    a = U.combine(a, id + ω) # handle diagonal Lipschitz constants
     a .-= U.combine(a_last, ω)
 end
 

### 2025-02-05 test/subblock_momentum_benchmark.jl

@@ -1,0 +1,50 @@
+using BenchmarkTools
+using BlockTensorDecomposition
+
+fact = BlockTensorDecomposition.factorize
+
+options = (
+    :rank => 3,
+    :tolerence => (1, 0.03),
+    :converged => (GradientNNCone, RelativeError),
+    :δ => 0.9,
+)
+
+n_subblock_n_momentum(Y) = fact(Y;
+    do_subblock_updates=false,
+    momentum=false,
+    options...
+)
+
+y_subblock_n_momentum(Y) = fact(Y;
+    do_subblock_updates=true,
+    momentum=false,
+    options...
+)
+
+n_subblock_y_momentum(Y) = fact(Y;
+    do_subblock_updates=false,
+    momentum=true,
+    options...
+)
+
+y_subblock_y_momentum(Y) = fact(Y;
+    do_subblock_updates=true,
+    momentum=true,
+    options...
+)
+
+I, J = 10, 10
+R = 3
+
+@benchmark n_subblock_n_momentum(Y) setup=(Y=Tucker1((I, J), R))
+
+@benchmark y_subblock_n_momentum(Y) setup=(Y=Tucker1((I, J), R))
+
+@benchmark n_subblock_y_momentum(Y) setup=(Y=Tucker1((I, J), R))
+
+@benchmark y_subblock_y_momentum(Y) setup=(Y=Tucker1((I, J), R))
+
+performance_increase(old, new) = (old - new) / new * 100
+
+time_decrease(old, new) = (old - new) / old * 100

### 2025-02-07 src/Core/decomposition.jl
improve tucker product calculation and indexing

@@ -19,7 +19,8 @@ abstract type AbstractDecomposition{T, N} <: AbstractArray{T, N} end
 
 # Fallback interface for any AbstractDecomposition to behave like an AbstractArray
 Base.size(D::AbstractDecomposition) = size(array(D))
-Base.getindex(D::AbstractDecomposition, i::Int) = getindex(array(D), i)
+Base.getindex(D::AbstractDecomposition, i::Int) = D[CartesianIndices(D)[i]] # usually more efficient to compute a single entry given the cartesian index rather than linear index
+Base.getindex(D::AbstractDecomposition, I::CartesianIndex) = getindex(D, Tuple(I)...)
 Base.getindex(D::AbstractDecomposition, I::Vararg{Int}) = getindex(array(D), I...)
 
 # copy/deepcopy all factors and any other properties before reconstructing
@@ -266,6 +267,7 @@ isfrozen(T::AbstractTucker, n::Integer) = frozen(T)[n+1]
 
 # AbstractDecomposition Interface
 array(T::AbstractTucker) = multifoldl(contractions(T), factors(T))
+array(T::Tucker) = tuckerproduct(core(T), matrix_factors(T))
 factors(T::AbstractTucker) = T.factors
 contractions(T::Tucker) = tucker_contractions(ndims(T))
 contractions(_::Tucker1) = ((×₁),)
@@ -290,21 +292,17 @@ eachfactorindex(D::AbstractTucker) = 0:(nfactors(D)-1) # 0 based, where core is 
 # Efficient size and indexing for CPDecomposition
 Base.size(T::Tucker) = map(x -> size(x, 1), matrix_factors(T))
 Base.size(T::Tucker1) = (size(matrix_factors(T)[begin], 1), size(core(T))[begin+1:end]...)
-Base.getindex(T::Tucker1, i::Int) = array(T)[i]
+
+Base.getindex(T::Tucker, i::Int) = T[CartesianIndices(T)[i]]
+Base.getindex(T::Tucker, I::Vararg{Int}) = _gettuckerindex(T, I)
+_gettuckerindex(T::Tucker, I) = _gettuckerindex(core(T), matrix_factors(T), I)
+
+Base.getindex(T::Tucker1, i::Int) = T[CartesianIndices(T)[i]]
 function Base.getindex(T::Tucker1, I::Vararg{Int})
     G, A = factors(T)
[…]

### 2025-02-07 src/Core/tensorproducts.jl

@@ -118,7 +118,7 @@ function tuckerproduct(core, matrices; exclude=nothing, excludes_missing=false)
     if isnothing(exclude)
         N == length(matrices) ||
             throw(ArgumentError("expected $N number of matrices, got $(length(matrices))"))
-        return multifoldl(tucker_contractions(N), (core, matrices...))
+        return _fulltuckerproduct(core, matrices)
     elseif excludes_missing
         N == length(matrices) + length(exclude) ||
         throw(ArgumentError("expected $N number of matrices, got $(length(matrices))"))
@@ -130,6 +130,26 @@ function tuckerproduct(core, matrices; exclude=nothing, excludes_missing=false)
     end
 end
 tuckerproduct(core, matrices...; kwargs...) = tuckerproduct(core, matrices; kwargs...)
+
+function _fulltuckerproduct(core, matrices)
+    output_size = first.(size.(matrices))
+    Y = zeros(promote_type(eltype(core), eltype.(matrices)...), output_size)
+    for Is in CartesianIndices(Y)
+        Y[Is] += _gettuckerindex(core, matrices, Is)
+    end
+    return Y
+end
+
+_gettuckerindex(core, matrices, I::CartesianIndex) = _gettuckerindex(core, matrices, Tuple(I))
+
+"""Just computes index I in the tucker product"""
+function _gettuckerindex(core, matrices, I)
+    Y = 0 # zero(promote_type(eltype(core), eltype.(matrices)...))
+    for Rs in CartesianIndices(core) # need to wrap in a Tuple so I can iterate over the index
+        Y += core[Rs] * prod(A[i, r] for (A, i, r) in zip(matrices, I, Tuple(Rs)))
+    end
+    return Y
+end
 
 """
     cpproduct((A, B, C, ...))

### 2025-02-07 src/Core/utils.jl

@@ -93,8 +93,11 @@ Base.min(A::AbstractArray, x::Number) = min.(A, x)
 
 ########################################################
 
-getnotindex(A, i::Int) = A[eachindex(A) .!= i]
-getnotindex(A, I) = A[eachindex(A) .∉ (I,)]
+# ( (x-1) % 4 +1, (x-1) ÷ 4 % 3 + 1, (x-1) ÷ 4 ÷ 3 % 2 + 1)
+
+"""Like `getindex` but returns the compliment to the index or indices requested."""
+getnotindex(A, i::Int; view=false) = view ? (@view A[eachindex(A) .!= i]) : A[eachindex(A) .!= i]
+getnotindex(A, I; view=false) = view ? (@view A[eachindex(A) .∉ (I,)]) : A[eachindex(A) .∉ (I,)]
 
 """
     eachfibre(A::AbstractArray; n::Integer, kwargs...)

### 2025-02-07 test/runtests.jl

@@ -271,6 +271,23 @@ end
     @test G[1] ≈ array(G)[1] # vectorized indexing, check approx because it calculates in a slightly different way
     @test G[1,1,1] ≈ array(G)[1] # regular indexing, check approx because it calculates in a slightly different way
 
+    # Tucker test
+
+    Y = [776 1008 1240 1472; 1072 1392 1712 2032; 1368 1776 2184 2592;;; 1104 1432 1760 2088; 1524 1976 2428 2880; 1944 2520 3096 3672]
+
+    A = reshape(1:6, 3, 2)
+    B = reshape(1:8, 4, 2)
+    C = reshape(1:4, 2, 2)
+    G = reshape(1:8, 2, 2, 2)
+
+    T = Tucker((G, A, B, C))
+
+    @test array(T) == Y
+    @test T[2] == 1072
+    @test typeof(T[2]) == typeof(1072)
+    @test eltype(T) == typeof(1072)
+    @test eltype(T) == typeof(T[2])
+
     # CPDecomposition test
     A = reshape(1:6, 3, 2)
     B = reshape(1:8, 4, 2)

### 2025-02-07 test/subblock_momentum_benchmark.jl
more subblock and momentum benchmarking

@@ -1,13 +1,25 @@
 using BenchmarkTools
+using Logging
 using BlockTensorDecomposition
+
+
+global_logger(SimpleLogger(Warn))
 
 fact = BlockTensorDecomposition.factorize
 
+function nnrankr_matrix((I, J), R)
+    A = randn(I, R) .|> abs
+    B = randn(R, J) .|> abs
+    return A*B
+end
+
 options = (
-    :rank => 3,
-    :tolerence => (1, 0.03),
-    :converged => (GradientNNCone, RelativeError),
-    :δ => 0.9,
+    :rank => (2,3,4),
+    :tolerence => (0.01),
+    :converged => (RelativeError),
+    :δ => 0.9999,
+    :model => Tucker,
+    #:constraints => nonnegative!,
 )
 
 n_subblock_n_momentum(Y) = fact(Y;
@@ -37,14 +49,45 @@ y_subblock_y_momentum(Y) = fact(Y;
 I, J = 10, 10
 R = 3
 
-@benchmark n_subblock_n_momentum(Y) setup=(Y=Tucker1((I, J), R))
+@benchmark n_subblock_n_momentum(Y) setup=(Y=array(Tucker1((I, J), R)))
 
-@benchmark y_subblock_n_momentum(Y) setup=(Y=Tucker1((I, J), R))
+@benchmark y_subblock_n_momentum(Y) setup=(Y=nnrankr_matrix((I, J), R))
 
-@benchmark n_subblock_y_momentum(Y) setup=(Y=Tucker1((I, J), R))
+@benchmark n_subblock_y_momentum(Y) setup=(Y=nnrankr_matrix((I, J), R))
 
-@benchmark y_subblock_y_momentum(Y) setup=(Y=Tucker1((I, J), R))
+@benchmark y_subblock_y_momentum(Y) setup=(Y=nnrankr_matrix((I, J), R))
 
 performance_increase(old, new) = (old - new) / new * 100
 
 time_decrease(old, new) = (old - new) / old * 100
+
+
+###################
+
+
+I, J, K = 10, 10, 10
+R1, R2, R3 = 2, 3, 4
+
+b = @benchmark n_subblock_n_momentum(Y) setup=(Y=Tucker((I, J, K), (R1, R2, R3))|>array)
+display(b)
+b = @benchmark y_subblock_n_momentum(Y) setup=(Y=Tucker((I, J, K), (R1, R2, R3))|>array)
[…]

### 2025-02-07 paper/paper.qmd
Finish first pass of section Convergence Criteria and Stats

@@
 
+TODO Repeat this experiment on a less trivial factorization. What I've done above can be done with an SVD in less time. Ideally use a 10 x 10 x 10 Tucker decomposition with rank 2 x 3 x 4.
+
 ## For Flexibility {#sec-flexibility}
@@
 
+There are a number of software engineering techniques used to ensure BlockTensorDecomposition.jl is flexible and applicable to a wide range of problems.
+These enable key algorithmic choices to be hot-swapped and easily compared with each other.
+
 ### Convergence Criteria and Stats {#sec-convergence-criteria}
@@
 - any subset of stats can be the convergence criteria
+
+Some iterative algorithms produce the exact solution of a problem after a finite number of iterations.
+Generalized minimal residual method (GMRES) is a good example of this [TODO cite!].
+Our algorithm, like many others, only converges to the exact solution in the limit as the number of iterations grow.
+Since we would like a solution in finite time, we must halt the algorithm early.
+
+In finite precision, we can halt the algorithm if we can guarantee the solution is accurate to machine precision.
+This can often be too strict if convergence is not at a fast enough rate.
+Furthermore, depending on *why* we are decomposing a tensor, we may want different stats to be within a given a tolerance.
+BlockTensorDecomposition.jl attempts to solve this issue by defining some standard criteria that can be used to halt the algorithm.
+These are subtypes of the abstract type `AbstractStat` and are listed below.
+
+```julia
+# X is the decomposition model e.g. Tucker((B, A1, A2, A2)))
+# Y is the input tensor we want to decompose
+# norm(X) is the Frobenius norm of X
+
+GradientNorm # norm(∇f(X))
+GradientNNCone # sqrt(sum(norm(∇f(Ai)[Ai .> 0 .| ∇f(Ai) .< 0])^2 for Ai in factors(X)))
+ObjectiveValue # 1/2 norm(X - Y)^2
+ObjectiveRatio # norm(X_last - Y)^2 / norm(X_current - Y)^2
+RelativeError # norm(X - Y) / norm(Y)
+IterateNormDiff # norm(X_current - X_last)
+IterateRelativeDiff # norm(X_current - X_last) / norm(X_last)
+```
+
+Most of these are self explanatory, expect perhaps `GradientNNCone`.
+When performing first order unconstrained optimization, we usually slow down progress when the norm of the gradient is small.
+In the limit, we expect to converge to a stationary point where the gradient is zero.
+When performing optimization under nonnegative constraints, and even further restrictions like simplex constraints, is makes more sense to ignore entries of the gradient where `X` is negative and the gradient is positive since those coordinates of `X` will not change after they are projected back to the nonnegative constraint.
+
+TODO add theory of negative gradient is in the normal cone of the constraint?
+
+As many or as few of these stats can be used in the call to `factorize`.
+Their tolerances can also be set independently of each other.
+The algorithm stops iterating when at least one of the criteria has a value less than its tolerance.
+As a fail safe, we also define one more type, `Iteration`, which is always active and halts the algorithm when that many iterations have past.
+
+The `AbstractStat` type can also be subtyped to create custom stats that may be used to probe the iterates and diagnose issues.
+
+```julia
+EuclidianStepSize
+EuclidianLipschitz
+FactorNorms
+```
+
+These stats are recorded every iteration in a `DataFrame` and are one of the returns of `factorize`.
+
+Finally, there are two auxiliary stats `PrintStats` and `DisplayDecomposition` which can be used to print all stats or the current decomposition each iteration.
 

### 2025-02-28 src/Core/constraint.jl
minor cleanups in constraint.jl

@@ -2,7 +2,7 @@
 Low level code for the types of constraints
 """
 
-"""Abstract parent type for the verious constraints"""
+"""Abstract parent type for the various constraints"""
 abstract type AbstractConstraint <: Function end
 
 function Base.show(io::IO, X::AbstractConstraint)
@@ -23,7 +23,9 @@ end
 
 Returns `true` if `A` satisfies the constraint `C`.
 """
-function check(_::AbstractConstraint, _::AbstractArray); end
+function check(C::AbstractConstraint, _::AbstractArray)
+    throw("checking constraints with type $(typeof(C)) is not implemented")
+end
 
 """
     GenericConstraint <: AbstractConstraint
@@ -155,7 +157,7 @@ function makeNNprojection(norm)
     if norm == l1norm
         return projsplx!
 
-    elseif norm in (l2norm, linftynorm)
+    elseif norm in (l2norm, linftynorm) # TODO should the l2norm have a different projection?
         return proj_one_hot!
 
     else
@@ -249,10 +251,10 @@ const linftynormalize_12slices! = ProjectedNormalization(linftynorm, linftyproje
 
 Main constructor for the constraint where `norm` of `whats_normalized` equals `scale`.
 
-Scale can be a single `Real`, or an `AbstractArray{<:Real}`, but should be brodcast-able
+Scale can be a single `Real`, or an `AbstractArray{<:Real}`, but should be broadcast-able
 with the output of `whats_normalized`.
-Lasly, scale can be a `Function` which will act on an `AbstractArray{<:Real}` and return
-something that is brodcast-able `whats_normalized`.
+Lastly, scale can be a `Function` which will act on an `AbstractArray{<:Real}` and return
+something that is broadcast-able `whats_normalized`.
 """
 struct ScaledNormalization{T<:Union{Real,AbstractArray{<:Real},Function}} <: AbstractNormalization
     norm::Function
@@ -338,7 +340,7 @@ check(C::Entrywise, A::AbstractArray) = all((C.check).(A))
 
 const nonnegative! = Entrywise(ReLU, isnonnegative)
 
-IntervalConstraint(a, b) = Entrywise(x -> clamp(x, a, b), x -> a <= x <= b)
+IntervalConstraint(a, b) = Entrywise(x -> clamp(x, a, b), x -> a ≤ x ≤ b)
 
 function binaryproject(x)
     if x > 0.5

### 2025-02-28 src/Core/utils.jl
optimize geometric mean

@@ -267,9 +267,23 @@ end
     geomean(v)
     geomean(v...)
 
-Geometric mean of a collection of numbers: `prod(v)^(1/length(v))``.
+Geometric mean of a collection: `prod(v)^(1/length(v))`.
+
+If prod(v) is detected to be 0 or Inf, the safer (but slower) implementation `exp(mean(log.(v)))` is used.
 """
-geomean(v) = prod(v)^(1/length(v))
+function geomean(v)
+    p = prod(v)
+    if isinf(p) | iszero(p)
+        # Slower, but more robust to very large or very small values in v
+        # or if length(v) is large
+        return exp(mean(log.(v)))
+        # if there is a zero element in v, this gets propagated correctly. log(0) == -Inf, the mean will keep the mean as -Inf, and exp(-Inf) == 0.0
+    else
+        # Faster, but more sensitive
+        return p^(1/length(v))
+    end
+end
+
 geomean(v...) = geomean(v)
 
 """

### 2025-02-28 test/geomean.jl

@@ -1,0 +1,44 @@
+using BenchmarkTools
+using Random
+
+f1(x) = prod(x)^(1/length(x))
+f2(x) = exp(mean(log.(x)))
+
+absrandn(x) = abs.(randn(x))
+v = absrandn(10)
+
+f1(v) ≈ f2(v)
+
+@benchmark f1(x) setup=(x = absrandn(10000))
+@benchmark f1(x) setup=(x = 100000*absrandn(10000))
+@benchmark f2(x) setup=(x = absrandn(10000))
+@benchmark f2(x) setup=(x = 100000*absrandn(10000))
+
+function geomean1(v)
+    p = prod(v)
+    if isinf(p) | iszero(p)
+        return exp(mean(log.(v)))
+    else
+        # Faster, but more sensitive
+        return p^(1/length(v))
+    end
+end
+
+function geomean2(v)
+    p = prod(v)
+    if iszero(p) | isinf(p)
+        return exp(mean(log.(v)))
+    else
+        # Faster, but more sensitive
+        return p^(1/length(v))
+    end
+end
+
+@benchmark geomean1(x) setup=(x = absrandn(10000))
+@benchmark geomean2(x) setup=(x = absrandn(10000))
+@benchmark geomean1(x) setup=(x = 10*absrandn(10000))
+@benchmark geomean2(x) setup=(x = 10*absrandn(10000))
+
+@benchmark geomean1(x) setup=(x = 10*absrandn(10))
+@benchmark f1(x) setup=(x = 10*absrandn(10))
+@benchmark f2(x) setup=(x = 10*absrandn(10))

### 2025-02-28 paper/paper.qmd
First pass of constraints and start on projection vs rescaling section

@@
 
+For a vector, matrix, or tensor filled with ones, we use $\mathbb{1}\in\mathbb{R}^{I_1 \times \dots \times I_N}$. This can be constructed in Julia with `ones(I₁, ..., Iₙ)`.
+
 ### Products of Tensors
@@
 
-### Sub-block Descent
+### Sub-block Descent {#sec-sub-block-descent}
 - Use smaller blocks, but descent in parallel (sub-blocks don't wait for other sub-blocks)
@@
 Note we store $\hat{L}_{B}^t$ in factored form as a tuple of diagonal matrices to save space and computation.
+
+TODO Compare to preconditioned decent. See [@kunstner_searching_2023;@gao_gradient_2024;@gao_scalable_2024;@qu_optimal_2024]
 
@@
 
-### Momentum
+### Momentum {#sec-momentum}
 - This one is standard
@@
 
-### `BlockUpdate` Language
-- construct the updates as a list of updates
-- very functional programming
-- can apply them in sequence or in a random order (or partially random)
-
-### Constraints
+### Constraints {#sec-constraints}
 - one type of update (other than the typical GD update)
@@
 
-# Partial Projection and Rescaling {#sec-ppr}
+One of the main motivations for developing BlockTensorDecomposition.jl is to solve constrained tensor problems.
+Other code did not have the expressivity to handle constraints beyond the most common constraints on the factors: nonnegativity and Euclidean normalized columns.
+To enable flexibility, BlockTensorDecomposition.jl defines
+
+```julia
+abstract type AbstractConstraint <: Function end
+```
+
+which has two interfaces. The first treats the constraint like a function
+
+```julia
+(C::AbstractConstraint)(A::AbstractArray)
+```
+
+and applies the constraint to an abstract array, and the other interface
+
+```julia
+check(C::AbstractConstraint, A::AbstractArray)
+```
+
+checks if the array `A` satisfies the constraint `C`. A generic constrain only needs to define these two functions.
+
+```julia
+struct GenericConstraint <: AbstractConstraint
+    apply::Function # input a AbstractArray -> mutate it so that `check` would return true
+    check::Function
+end
+
+function (C::GenericConstraint)(D::AbstractArray)
+    (C.apply)(D)
+end
+
+check(C::GenericConstraint, A::AbstractArray) = (C.check)(A)
+```
+
+In general, the function `check` could be defined as the following,
+
+```julia
+function check(C, A)
+    A_copy = copy(A)
+    C(A)
+    return A ≈ A_copy
+end
+```
+
+but there is often an easier way to check if a tensor is constrained than to apply the constraint.
+
+Although our basic block projected gradient descent algorithm @eq-proximal-explicit relies on Euclidean projections to the relevant constraint set, we want to remain flexible and allow for other maps that move an iterate to the constraint set.
+So each constraint needs to store more than just the constraint itself (the *what*), but also the map from an iterate to the constraint set (the *how*).
+When prototyping algorithms, it is worth comparing alternate approaches to see if there is a more efficient method to enforce a given constraint (See @sec-ppr for an example of constraining to a set without Euclidean projections).
+
+The first class of constraints are entry-wise constraints. This is convenient for defining a scalar function that gets applied to all entries of a tensor.
+
+```julia
+struct Entrywise <: AbstractConstraint
+    apply::Function
+    check::Function
+end
+
+function (C::Entrywise)(A::AbstractArray)
+    A .= (C.apply).(A)
+end
+
+check(C::Entrywise, A::AbstractArray) = all((C.check).(A))
+```
+
+Some common examples would be a nonnegative constraint, constraining entries to an interval, or constraining entries to be one or zero.
+
+```julia
+nonnegative! = Entrywise(x -> max(0, x), x ≥ 0)
+
+IntervalConstraint(a, b) = Entrywise(x -> clamp(x, a, b), x -> a ≤ x ≤ b)
+
+binary! = Entrywise(x > 0.5 ? one(x) : zero(x), x -> x in (0, 1))
+```
+
+Another class of constraints are normalizations $\mathcal{C}_{\lVert \cdot \rVert_{a}}=\left\{v \middle| \lVert v \rVert_{a} = 1 \right\}$ for some norm $\lVert \cdot \rVert_{a}$.
+These can be enforced by a (Euclidean) projection
+
+$$
+\hat{v} \in \arg\,\min_{u\in\mathcal{C}_{\lVert \cdot \rVert_a}} \lVert u - v \rVert _2^2,
+$$
+
+or a scaling (assuming $v\neq 0$)
+
+$$
+\hat{v} \leftarrow \frac{v}{\lVert v \rVert_a}.
+$$
+
+These operations agree for the Frobenius norm (entry-wise $2$-norm), but are different operations in general. In BlockTensorDecomposition.jl, we define these classes as the following.
+
+TODO simplify the following code?
+
+```julia
+abstract type AbstractNormalization <: AbstractConstraint end
+```
+
+```julia
+struct ProjectedNormalization <: AbstractNormalization
+    norm::Function # calculate the norm of an AbstractArray
+    projection::Function # mutate an AbstractArray so it has norm == 1
+    whats_normalized::Function # what part of the array is normalized
+                               # e.g. eachrow, or omit for the entire array
+end
+
+ProjectedNormalization(norm, projection; whats_normalized=identityslice) =
+    ProjectedNormalization(norm, projection, whats_normalized)
+
+function (P::ProjectedNormalization)(A::AbstractArray)
+    whats_normalized_A = P.whats_normalized(A)
+    (P.projection).(whats_normalized_A)
+end
+
+check(P::ProjectedNormalization, A::AbstractArray) =
+    all((P.norm).(P.whats_normalized(A)) .≈ 1)
+```
+
+```julia
+struct ScaledNormalization{T<:Union{Real,AbstractArray{<:Real},Function}} <: AbstractNormalization
+    norm::Function # calculate the norm of an AbstractArray
+    whats_normalized::Function # what part of the array is normalized
+    scale::T # the scale that `whats_normalized` should be normalized to
+             # Can be a real, array of reals, or function of the input that
+             # that returns a real or array of reals
+end
+
+ScaledNormalization(norm;whats_normalized=identityslice,scale=1) =
+    ScaledNormalization{typeof(scale)}(norm, whats_normalized, scale)
+
+function (S::ScaledNormalization{T})(A::AbstractArray) where {T<:Union{Real,AbstractArray{<:Real}}}
+    whats_normalized_A = S.whats_normalized(A)
+    A_norm = (S.norm).(whats_normalized_A) ./ S.scale
+    whats_normalized_A ./= A_norm
+    return A_norm
+end
+
+function (S::ScaledNormalization{T})(A::AbstractArray) where {T<:Function}
+    whats_normalized_A = S.whats_normalized(A)
+    A_norm = (S.norm).(whats_normalized_A) ./ S.scale(A)
+    whats_normalized_A ./= A_norm
+    return A_norm
+end
+
+check(S::ScaledNormalization{<:Union{Real,AbstractArray{<:Real}}}, A::AbstractArray) =
+    all((S.norm).(S.whats_normalized(A)) .≈ S.scale)
+
+check(S::ScaledNormalization{<:Function}, A::AbstractArray) =
+    all((S.norm).(S.whats_normalized(A)) .≈ S.scale(A))
+```
+
+We can define some common constraints like $L_1$ normalized through a projection
+
+```julia
+l1normalize! =
+    ProjectedNormalization(l1norm, l1project!)
+l1normalize_rows! =
+    ProjectedNormalization(l1norm, l1project!; whats_normalized=eachrow)
+l1normalize_1slices! =
+    ProjectedNormalization(l1norm, l1project!;
+        whats_normalized=(x -> eachslice(x; dims=1)))
+```
+
+or $L_\infty$ normalized by scaling
+
+```julia
+linftyscale_cols! = ScaledNormalization(linftynorm; whats_normalized=eachcol)
+```
+
+or even normalize the 3-fibres of a third order tensor on average!
+
+```julia
+l2scale_average12slices! = ScaledNormalization(l2norm;
+    whats_normalized=(x -> eachslice(x; dims=1)),
+    scale=(A -> size(A, 2)))
+```
+
+Constraints can also be composed.
+This is *not* the same as the intersection of constraints.
+This is only a handy way to apply multiple constraints in series, but there is no clever logic that interprets the constraints and combines them into a single constraint.
+
+```julia
+struct ComposedConstraint{T<:AbstractConstraint, U<:AbstractConstraint} <: AbstractConstraint
+    outer::T
+    inner::U
+end
+
+function (C::ComposedConstraint)(A::AbstractArray)
+    C.inner(A)
+    C.outer(A)
+end
+
+check(C::ComposedConstraint, A::AbstractArray) =
+    check(C.outer, A) & check(C.inner, A)
+
+Base.:∘(f::AbstractConstraint, g::AbstractConstraint) =
+    ComposedConstraint(f, g)
+```
+
+This means the following three constraints are all different!
+
+```julia
+l1normalize! ∘ nonnegative!
+nonnegative! ∘ l1normalize!
+simplex!
+```
+
+See @sec-ppr for a discussion on why it may be advantages to use one of these constraints over the other.
+
+### `BlockUpdate` Language
+- construct the updates as a list of updates
+- very functional programming
+- can apply them in sequence or in a random order (or partially random)
+
+To put all these pieces together, we define an abstract type that can be subtyped for the various types of update.
+This will include gradient descent steps, momentum updates, and constraint enforcing updates.
+
+```julia
+abstract type AbstractUpdate <: Function end
+```
+
+A generic update is a function that can take some keyword arguments and mutate an iterate.
+
+```julia
+struct GenericUpdate <: AbstractUpdate
+    f::Function
+end
+
+(U::GenericUpdate)(x; kwargs...) = U.f(x; kwargs...)
+```
+
+The first types of updates are gradient descent updates.
+
+```julia
+abstract type AbstractGradientDescent <: AbstractUpdate end
+```
+
+This includes the regular gradient descent update, and also the sub-block descent updates.
+
+```julia
+struct GradientDescent <: AbstractGradientDescent
+    n::Integer
+    gradient::Function
+    step::AbstractStep
+end
+
+function (U::GradientDescent)(x; x_last, kwargs...)
+    n = U.n
+    if checkfrozen(x, n)
+        return x
+    end
+    grad = U.gradient(x; kwargs...)
+    # Note we pass a function for grad_last (lazy) so that we only compute it if needed for the step
+    s = U.step(x; n, x_last, grad, grad_last=(x -> U.gradient(x; kwargs...)), kwargs...)
+    a = factor(x, n)
+    @. a -= s*grad
+end
+```
+
+The only addition to `BlockGradientDescent` is a function that combines the step with the gradient.
+
+```julia
+struct BlockGradientDescent <: AbstractGradientDescent
+    n::Integer
+    gradient::Function
+    step::AbstractStep
+    combine::Function # takes a step (number, matrix, or tensor) and combines it with a gradient
+end
+
+function (U::BlockGradientDescent)(x; x_last, kwargs...)
+    n = U.n
+    if checkfrozen(x, n)
+        return x
+    end
+    grad = U.gradient(x; kwargs...)
+    # Note we pass a function for grad_last (lazy) so that we only compute it if needed for the step
+    s = U.step(x; n, x_last, grad, grad_last=(x -> U.gradient(x; kwargs...)), kwargs...)
+    a = factor(x, n)
+    a .-= U.combine(grad, s)
+end
+```
+
+A separate step type is defined to allow for different types of steps like constant step, Lipschitz, and spectral projected gradient (SPG).
+
+```julia
+abstract type AbstractStep <: Function end
+```
+
+```julia
+struct LipschitzStep <: AbstractStep
+    lipschitz::Function
+end
+
+function (step::LipschitzStep)(x; kwargs...)
+    L = step.lipschitz(x)
+    return L^(-1) # allow for Lipschitz to be a diagonal matrix
+end
+
+function (step::LipschitzStep)(x::Tucker; kwargs...)
+    L = step.lipschitz(x)
+    if typeof(L) <: Tuple # Currently the only case is when we are updating the core of a Tucker factorization
+                          # Using this condition as a way to tell if it is the core we are calculating the constant for
+        return map(X -> X^(-1), L)
+    else
+        return L^(-1) # allow for Lipschitz to be a diagonal matrix
+    end
+end
+```
+
+```julia
+struct ConstantStep <: AbstractStep
+    stepsize::Real
+end
+
+(step::ConstantStep)(x; kwargs...) = step.stepsize
+```
+
+```julia
+struct SPGStep <: AbstractStep
+    min::Real
+    max::Real
+end
+
+SPGStep(;min=1e-10, max=1e10) = SPGStep(min, max)
+
+# Convert an input of the full decomposition, to a calculation on the nth factor
+(step::SPGStep)(x::T; n, x_last::T, grad_last::Function, kwargs...) where {T <: AbstractDecomposition} =
+    step(factor(x,n); x_last=factor(x_last,n), grad_last=grad_last(x_last), kwargs...)
+
+function (step::SPGStep)(x; grad, x_last, grad_last, stepmin=step.min, stepmax=step.max, kwargs...)
+    s = x - x_last
+    y = grad - grad_last
+    sy = (s ⋅ y)
+    if sy <=0
+        return stepmax
+    else
+        suggested_step = (s ⋅ s) / sy
+        return clamp(suggested_step, stepmin, stepmax)
+    end
+end
+```
+
+The gradient and Lipschitz stepsize is calculated by a separate function that gets make on initialization of the `factorization` algorithm (See @sec-gradient-computation and @sec-lipschitz-computation).
+This de-couples applying the gradient descent (the *what*), and the computation of the gradient (the *how*) so that the gradient can be calculated manually (if an efficient method is known), or with automatic differentiation.
+This also allows other updates to use the same computation code.
+For example, momentum updates also use the same Lipschitz calculation function.
+Momentum updates also use the same `combine` function as the sub-block gradient descent updates.
+
+```julia
+struct MomentumUpdate <: AbstractUpdate
+    n::Integer
+    lipschitz::Function
+    combine::Function # How to combine the momentum variable `ω` with a factor `a`
+end
+
+MomentumUpdate(n, lipschitz) = MomentumUpdate(n, lipschitz, (ω, a) -> ω * a)
+
+function MomentumUpdate(GD::AbstractGradientDescent)
+    n, step = GD.n, GD.step
+    @assert typeof(step) <: LipschitzStep
+
+    return MomentumUpdate(n, step.lipschitz)
+end
+
+function MomentumUpdate(GD::BlockGradientDescent)
+    n, step, combine = GD.n, GD.step, GD.combine
+    @assert typeof(step) <: LipschitzStep
+
+    return MomentumUpdate(n, step.lipschitz, combine)
+end
+```
+
+Constraints also get their own abstract subtype of updates.
+
+```julia
+abstract type ConstraintUpdate <: AbstractUpdate end
+```
+
+We define a constructor for applying a constraint to a particular factor $n$. This turns an `AbstractConstraint` into a `ConstraintUpdate`.
+
+```julia
+ConstraintUpdate(n, constraint::GenericConstraint; kwargs...) =
+    GenericConstraintUpdate(n, constraint)
+ConstraintUpdate(n, constraint::ProjectedNormalization; kwargs...) =
+    Projection(n, constraint)
+ConstraintUpdate(n, constraint::Entrywise; kwargs...) =
+    Projection(n, constraint)
+
+function ConstraintUpdate(n, constraint::ScaledNormalization; skip_rescale=false, whats_rescaled=missing, kwargs...)
+    if skip_rescale
+        ismissing(whats_rescaled) ||
+        isnothing(whats_rescaled) ||
+        @warn "skip_rescale=true but whats_rescaled=$whats_rescaled was given. Overriding to whats_rescaled=nothing"
+        return Rescale(n, constraint, nothing)
+    else
+        return Rescale(n, constraint, whats_rescaled)
+    end
+end
+```
+
+A generic constraint update extracts the factor it needs to constrain, and applies the constraint to that factor.
+
+```julia
+struct GenericConstraintUpdate <: ConstraintUpdate
+    n::Integer
+    constraint::GenericConstraint
+end
+
+check(U::GenericConstraintUpdate, D::AbstractDecomposition) = check(U.constraint, factor(D, U.n))
+
+function (U::GenericConstraintUpdate)(x::T; kwargs...) where T
+    n = U.n
+    A = factor(x, n)
+    U.constraint(A)
+    check(U, A) ||
+        error("Something went wrong with GenericConstraintUpdate: $GenericConstraintUpdate")
+end
+```
+
+Projections follow this pattern closely.
+
+```julia
+struct Projection <: ConstraintUpdate
+    n::Integer
+    proj::Union{ProjectedNormalization, Entrywise}
+end
+
+check(P::Projection, D::AbstractDecomposition) = check(P.proj, factor(D, P.n))
+
+function (U::Projection)(x::T; kwargs...) where T
+    n = U.n
+    U.proj(factor(x, n))
+end
+```
+
+A more involved type of constraint update is a rescaled update. This uses a scaled normalization, but moves the weight of the factor to other factors.
+
+For example, if we have three matrix factors $A, B, C$ in a CP decomposition where we want the sum of the entries in $A$ to sum to $1$, we can divide $A$ by its sum, and multiple factor $B$ by this amount.
+That way the recombined tensor $\left\llbracket A, B, C \right\rrbracket$ remains unchanged, but now $A$ satisfies the desired constraint.
+We could instead multiply both $B$ and $C$ by the square root of the sum to achieve a similar outcome.
+When it is not specified what is rescaled (`whats_rescaled = missing`), we assume we should multiply every other factor by the geometric mean of the scaling.
+If we just want to scale the factor but skip any rescaling, we can use `whats_rescaled = nothing`.
+
+```julia
+struct Rescale{T<:Union{Nothing,Missing,Function}} <: ConstraintUpdate
+    n::Integer
+    scale::ScaledNormalization
+    whats_rescaled::T
+end
+
+check(S::Rescale, D::AbstractDecomposition) = check(S.scale, factor(D, S.n))
+
+function (U::Rescale{<:Function})(x; kwargs...)
+    Fn_scale = U.scale(factor(x, U.n))
+    to_scale = U.whats_rescaled(x)
+    to_scale .*= Fn_scale
+end
+
+(U::Rescale{Nothing})(x; kwargs...) =
+    U.scale(factor(x, U.n))
+
+function (U::Rescale{Missing})(x; skip_rescale=false, kwargs...)
+    Fn_scale = U.scale(factor(x, U.n))
+    x_factors = factors(x)
+    N = length(x_factors) - 1
+
+    # Nothing to rescale, so return here
+    if N == 0 || skip_rescale
+        return nothing
+    end
+
+    # Assume we want to evenly rescale all other factors by the Nth root of Fn_scale
+    scale = geomean(Fn_scale)^(1/N)
+    for (i, A) in zip(eachfactorindex(x), x_factors)
+        # skip over the factor we just updated
+        if i == U.n
+            continue
+        end
+        A .*= scale
+    end
+end
+```
+
+See @sec-ppr for a more in-depth discussion on when it may be beneficial to use this type of constraint update over a simple projection.
+
+We would like to combine all these updates to execute them in serial.
+We use a `BlockedUpdate` type to do this.
+This should be thought of as a list of `AbstractUpdates` that get applied one after another.
+
+```julia
+struct BlockedUpdate <: AbstractUpdate
+    updates::Vector{AbstractUpdate}
+end
+```
+
+::: {.callout-note}
+# Technical Julia Note
+
+We need the updates to be exactly something of the form `AbstractUpdate[]` since we want to push any type of `AbstractUpdate`s such as a `MomentumUpdate` or another `BlockedUpdate`, even if not already present.
+This means it cannot be `Vector{<:AbstractUpdate}` since a `BlockedUpdate` constructed with only `GradientDescent` would give a `GradientDescent[]` vector and we couldn't push a `MomentumUpdate`.
+And it cannot be `AbstractVector{AbstractUpdate}` since we may not be able to `insert!` or `push!` into other `AbstractVectors` like `Views`.
+:::
+
+We forward many standard methods so that `BlockedUpdate`s can behave like usual Julia vectors.
+
+TODO should I actually show all these details?
+
+```julia
+Base.getindex(U::BlockedUpdate, i::Int) = getindex(updates(U), i)
+Base.getindex(U::BlockedUpdate, I::Vararg{Int}) = getindex(updates(U), I...)
+Base.getindex(U::BlockedUpdate, I) = getindex(updates(U), I) # catch all
+Base.firstindex(U::BlockedUpdate) = firstindex(updates(U))
+Base.lastindex(U::BlockedUpdate) = lastindex(updates(U))
+Base.keys(U::BlockedUpdate) = keys(updates(U))
+Base.length(U::BlockedUpdate) = length(updates(U))
+Base.iterate(U::BlockedUpdate, state=1) = state > length(U) ? nothing : (U[state], state+1)
+Base.filter(f, U::BlockedUpdate) = BlockedUpdate(filter(f, updates(U)))
+```
+
+We define how `BlockedUpdates` get applied to a tensor with the following function.
+
+```julia
+function (U::BlockedUpdate)(x::T; recursive_random_order::Bool=false, random_order::Bool=recursive_random_order, kwargs...) where T
+    U_updates = updates(U)
+    if random_order
+        order = shuffle(eachindex(U_updates))
+        U_updates = U_updates[order]
+    end
+
+    for update! in U_updates
+        update!(x; recursive_random_order, kwargs...)
+        # note random_order does not get passed down
+    end
+end
+```
+
+The default order the blocks are updated is cyclically through each factor of the decomposition `D::AbstractDecomposition`, in the order of `factors(D)`.
+For `AbstractTucker` decompositions like Tucker, Tucker-1, and CP, this means starting with the core, followed by the matrix factor for the first dimension, second dimension, and so on.
+
+As an example, this would be the default order of updates for nonnegative CP decomposition on an order 3 tensor.
+
+```julia
+BlockedUpdate(
+    MomentumUpdate(1, lipschitz)
+    GradientStep(1, gradient, LipschitzStep)
+    Projection(1, Entrywise(ReLU, isnonnegative))
+    MomentumUpdate(2, lipschitz)
+    GradientStep(2, gradient, LipschitzStep)
+    Projection(2, Entrywise(ReLU, isnonnegative))
+    MomentumUpdate(3, lipschitz)
+    GradientStep(3, gradient, LipschitzStep)
+    Projection(3, Entrywise(ReLU, isnonnegative))
+)
+```
+
+The order of updates can be randomized with the `random_order` keyword.
+```julia
+X, stats, kwargs = factorize(Y; random_order=true)
+```
+
+By default, this will keep momentum steps, gradient steps, and constraint steps for each factor together as a block, in this order.
+
+A possible order of updates could be the following. Note that the updates for each factor are grouped together, but each factor is updated in a random order.
+
+```julia
+BlockedUpdate(
+    BlockedUpdate(
+        MomentumUpdate(2, lipschitz)
+        GradientStep(2, gradient, LipschitzStep)
+        Projection(2, Entrywise(ReLU, isnonnegative))
+    )
+    BlockedUpdate(
+        MomentumUpdate(1, lipschitz)
+        GradientStep(1, gradient, LipschitzStep)
+        Projection(1, Entrywise(ReLU, isnonnegative))
+    )
+    BlockedUpdate(
+        MomentumUpdate(3, lipschitz)
+        GradientStep(3, gradient, LipschitzStep)
+        Projection(3, Entrywise(ReLU, isnonnegative))
+    )
+)
+```
+
+For more randomization, use the `recursive_random_order` keyword which will also randomize the order in which the momentum steps, gradient steps, and constraint steps are performed.
+
+```julia
+X, stats, kwargs = factorize(Y; recursive_random_order=true)
+```
+
+A possible order of updates could now be the following. The updates for each factor are still grouped together, but the updates within each block appear in a random order.
+
+```julia
+BlockedUpdate(
+    BlockedUpdate(
+        Projection(2, Entrywise(ReLU, isnonnegative))
+        MomentumUpdate(2, lipschitz)
+        GradientStep(2, gradient, LipschitzStep)
+    )
+    BlockedUpdate(
+        MomentumUpdate(1, lipschitz)
+        Projection(1, Entrywise(ReLU, isnonnegative))
+        GradientStep(1, gradient, LipschitzStep)
+    )
+    BlockedUpdate(
+        GradientStep(3, gradient, LipschitzStep)
+        Projection(3, Entrywise(ReLU, isnonnegative))
+        MomentumUpdate(3, lipschitz)
+    )
+)
+```
+
+The opposite of this would be to keep the outer order of blocks as given, but randomize the order which the updates for each factor gets applied, use the following code.
+
+```julia
+X, stats, kwargs = factorize(Y; recursive_random_order=true, random_order=false, group_by_factor=true)
+```
+
+A possible order of updates could now be the following. Note the order of factors is preserved (1, 2, 3) but the inner `BlockedUpdate`s have a random order.
+
+```julia
+BlockedUpdate(
+    BlockedUpdate(
+        Projection(1, Entrywise(ReLU, isnonnegative))
+        MomentumUpdate(1, lipschitz)
+        GradientStep(1, gradient, LipschitzStep)
+    )
+    BlockedUpdate(
+        MomentumUpdate(2, lipschitz)
+        Projection(2, Entrywise(ReLU, isnonnegative))
+        GradientStep(2, gradient, LipschitzStep)
+    )
+    BlockedUpdate(
+        GradientStep(3, gradient, LipschitzStep)
+        MomentumUpdate(3, lipschitz)
+        Projection(3, Entrywise(ReLU, isnonnegative))
+    )
+)
+```
+
+Note all the previously mentioned options still keeps the various updates for each factor together. For full randomization, use the following code.
+
+```julia
+X, stats, kwargs = factorize(Y; recursive_random_order=true, group_by_factor=false)
+```
+
+A possible order of updates could now be the following. Note that every update can appear anywhere in the order.
+
+```julia
+BlockedUpdate(
+    Projection(3, Entrywise(ReLU, isnonnegative))
+    MomentumUpdate(2, lipschitz)
+    GradientStep(2, gradient, LipschitzStep)
+    MomentumUpdate(1, lipschitz)
+    GradientStep(1, gradient, LipschitzStep)
+    Projection(2, Entrywise(ReLU, isnonnegative))
+    MomentumUpdate(3, lipschitz)
+    MomentumUpdate(2, lipschitz)
+    Projection(1, Entrywise(ReLU, isnonnegative))
+    GradientStep(3, gradient, LipschitzStep)
+)
+```
+
+The complete behaviour is summarized in @tbl-blockupdate-randomization.
+
+We also use `BlockedUpdate` to handle a composition of constraints.
+
+<!-- ```julia
+function ConstraintUpdate(n, constraint::ComposedConstraint; kwargs...)
+    if constraint.inner == nonnegative! && typeof(constraint.outer) <: ScaledNormalization
+        norm = constraint.outer.norm
+        return BlockedUpdate(
+            SafeNNProjection(n, ProjectedNormalization(norm, makeNNprojection(norm); whats_normalized=constraint.outer.whats_normalized)),
+            ConstraintUpdate(n, constraint.outer; kwargs...))
+    else
+        return BlockedUpdate(ConstraintUpdate(n, constraint.inner; kwargs...), ConstraintUpdate(n, constraint.outer; kwargs...)) # note we apply inner constraint first
+    end
+end
+``` -->
+
+```julia
+ConstraintUpdate(n, constraint::ComposedConstraint; kwargs...) =
+    BlockedUpdate(ConstraintUpdate(n, constraint.inner; kwargs...),
+                  ConstraintUpdate(n, constraint.outer; kwargs...))
+end
+```
+
+The `BlockUpdate` language becomes especially helpful for automaically inserting additional updates as they are requested.
+For example, if we want to add momentum to a list of updates, we can call the following function.
+
+```julia
+function add_momentum!(U::BlockedUpdate)
+    # Find all the GradientDescent updates
+    U_updates = updates(U)
+    indexes = findall(u -> typeof(u) <: AbstractGradientDescent, U_updates)
+
+    # insert MomentumUpdates before each GradientDescent
+    # do this in reverse order so "i" correctly indexes a GradientDescent
+    # as we mutate updates
+    for i in reverse(indexes)
+        insert!(U_updates, i, MomentumUpdate(U_updates[i]))
+    end
+end
+```
+
+We can also interlace two lists of updates to put updates on the same factor next to each other, or group all updates by the factor they act on.
+
+```julia
+function smart_interlace!(U::BlockedUpdate, other_updates)
+    for V in other_updates
+        smart_insert!(U::BlockedUpdate, V::AbstractUpdate)
+    end
+end
+
+smart_interlace!(U::BlockedUpdate, V::BlockedUpdate) = smart_interlace!(U::BlockedUpdate, updates(V))
+
+function smart_insert!(U::BlockedUpdate, V::AbstractUpdate)
+    U_updates = updates(U)
+    i = findlast(u -> u.n == V.n, U_updates)
+
+    # insert the other update immediately after
+    # or if there is no update, push it to the end
+    isnothing(i) ? push!(U_updates, V) : insert!(U_updates, i+1, V)
+end
+```
+
+```julia
+function group_by_factor(blockedupdate::BlockedUpdate)
+    factor_labels = unique(getproperty(U, :n) for U in blockedupdate)
+    updates_by_factor = [filter(U -> U.n == n, blockedupdate) for n in factor_labels]
+    return BlockedUpdate(updates_by_factor)
+end
+```
+
+# Rescaling to Constrain Tensor Factorization {#sec-ppr}
 - for bounded linear constraints
@@
 
+Constraints on factors in a tensor decomposition can arise naturally when modeling physical problems.
+A common class of constraints is normalizations which restrict a factor or slices of a factor to have unit norm.
+These are sometimes intersected with interval constraints such as requiring entries to be nonnegative.
+
+For example, the demixing of $R$ probibility densities $b_1, \dots, b_R$ from $I$ mixtures $y_1,\dots,y_I$ for $I>R$ can be accomplished with a nonneagive matrix factorization (cite).
+We have the system of equations
+
+$$
+\begin{align}
+{\ y_1} &= a_{11} {\ b_1} + a_{12} {\ b_2} + \dots + a_{1R} {\ b_R} \\
+{\ y_2} &= a_{21} {\ b_1} + a_{22} {\ b_2} + \dots + a_{2R} {\ b_R} \\
+            &\vdots \\
+{\ y_I} &= a_{I1} {\ b_1} + a_{I2} {\ b_2} + \dots + a_{IR} {\ b_R} \\
+\end{align}
+$$
+
+with unknown mixing coefficients $a_{i,r}$ and densities $b_r$.
+If we can discritized the mixtures $y_i$, we can rewrite this system as a rank $R$ factorization of the matrix $Y$ where $Y[i,:]=y_i$ and
+
+$$
+\begin{align*}
+\begin{bmatrix}
+\leftarrow & \ y_{1}^\top & \rightarrow \\
+\leftarrow & \ y_{2}^\top & \rightarrow  \\
+\leftarrow & \ \vdots & \rightarrow \\
+\leftarrow & \ y_{I}^\top & \rightarrow
+\end{bmatrix}
+& = \begin{bmatrix}
+a_{11} & a_{12} & \dots & a_{1R} \\
+a_{21} & a_{22} & \dots & a_{2R} \\
+ \vdots &  &  & \vdots \\
+a_{I1} & a_{I2} & \dots & a_{IR} \\
+\end{bmatrix}
+\begin{bmatrix}
+ \leftarrow &{\ b_{1}}^\top & \rightarrow \\
+ \leftarrow &{\ b_2} ^\top & \rightarrow \\
+\ & \vdots & \ \\
+\leftarrow &{\ b_{R}}^\top& \rightarrow
+\end{bmatrix}
+\end{align*}
+$$
+
+$$
+Y = AB,
+$$
+
+for matricies $A$ and $B$.
+
+For the factorization to remain interpretable, we need to ensure each row $b_r$ of $B$ is a density.
+This means we would like to constrain each row $B[r,:]=b_r$ to the simplex^[This assumes the entries $b_r[j]$ in the discretization of the density $b_r$ represent probibilities or areas under some continuous 1D density function, not the sample values of the density $b_r(x_j)$.
+Assuming $J$ sample points $x_j$ of a grid on the interval $[x_0,x_J]$, the entries of the discretization vector can be defined as $b_r[j]=b_r(x_j) (x_j - x_{j-1})$.]
+
+$$
+b_r\in\Delta_J = \left\{ v\in\mathbb{R}^J \middle| \sum_{j=1}^J v[j] =1\quad \text{and}\quad  \forall j\in[J],\ \ v[j] \geq 0 \right\},
+$$
+
+which we can write as constraining the matrix $B$ to the simplex
+
+$$
+B\in\Delta_{R, J}=\left\{B\in\mathbb{R}^{R\times J} \middle| \forall r\in [R],\, \sum_{j=1}^J B[r,j] = 1  \quad \text{and}\quad \forall (r,j)\in [R]\times [J],\, B[r,j]\geq 0 \right\}.
+$$
+
+Given the rows of $B$ represent densities, to ensure the rows of the reconstructed matrix $\hat{Y}=AB$ are still densities, we need the mixing coefficients to be nonnegative $a_{ir}\geq 0$ and rows to sum to one $\sum_{r}a_{ir} = 1$.
+This constrains the matrix $A$ to the simplex
+
+$$
+A\in\Delta_{I, R}=\left\{A\in\mathbb{R}^{I\times R} \middle| \forall i\in [I],\, \sum_{r=1}^R A[i,r] = 1  \quad \text{and}\quad \forall (i,r)\in [I]\times [R],\, A[i,r]\geq 0 \right\}.
+$$
+
+<!--
+$\newcommand{\blue}[1]{\color{blue}{#1}}\newcommand{\green}[1]{\color{green}{#1}}\newcommand{\purple}[1]{\color{purple}{#1}}\newcommand{\orange}[1]{\color{orange}{#1}}$
+
+$$
+\begin{align}
+{\mathbf y_1} &= a_{11} \blue{\mathbf b_1} + a_{12} \green{\mathbf b_2} + \dots + a_{1R} \purple{\mathbf b_R} \\
+{\mathbf y_2} &= a_{21} \blue{\mathbf b_1} + a_{22} \green{\mathbf b_2} + \dots + a_{2R} \purple{\mathbf b_R} \\
+            &\vdots \\
+{\mathbf y_I} &= a_{I1} \blue{\mathbf b_1} + a_{I2} \green{\mathbf b_2} + \dots + a_{IR} \purple{\mathbf b_R} \\
+\end{align}
+$$
+
+$$
+\begin{align*}
+\begin{bmatrix}
+\leftarrow & \mathbf y_{1}^\top & \rightarrow \\
+\leftarrow & \mathbf y_{2}^\top & \rightarrow  \\
+\leftarrow & \vdots & \rightarrow \\
+\leftarrow & \mathbf y_{I}^\top & \rightarrow
+\end{bmatrix}
+& = \begin{bmatrix}
+a_{11} & a_{12} & \dots & a_{1R} \\
+a_{21} & a_{22} & \dots & a_{2R} \\
+ \vdots &  &  & \vdots \\
+a_{I1} & a_{I2} & \dots & a_{IR} \\
+\end{bmatrix}
+\begin{bmatrix}
+ \leftarrow &\blue{\mathbf b_{1}}^\top & \rightarrow \\
+ \leftarrow &\green{\mathbf b_2} ^\top & \rightarrow \\
+& \vdots & \\
+\leftarrow &\purple{\mathbf b_{R}}^\top& \rightarrow
+\end{bmatrix}
+\end{align*}
+$$
+
+-->
+
+## The two approaches for simplex constraints
+
+With constraints being defined in a flexible manner (see @sec-constraints), we decided to test conventional wisdom that Euclidean projections are the right kind of map to use when enforcing a constraint.
+A constraint that came up in applications was enforcing the $1$st mode slices of a tensor (e.g.
+rows in a matrix) or $3$rd mode fibres of a third order tensor to lie in their respective simplex.
+
+For example, to constrain a vector $v\in\mathbb{R}^J$ to the simplex
+
+$$
+\Delta_J = \left\{ v\in\mathbb{R}^J \middle| \sum_{j=1}^J v[j] =1,\quad \text{and}\quad  \forall j\in[J],\ \ v[j] \geq 0 \right\},
+$$
+
+we could apply a Euclidean projection
+
+$$
+v \leftarrow \arg\,\min_{u\in\Delta_J} \lVert u - v \rVert _2^2,
+$$
+
+or a generalized Kullback-Leibler divergence projection
+
+$$
+v \leftarrow \arg\,\min_{u\in\Delta_J} \sum_j u[j] \log{(\frac{u[j]}{v[j]})} - u[j] + v[j]
+$$ {#eq-kl-projection}
+
+amoung other reasonable maps onto $\Delta_J$.
+
+The Eucledian simplex projection can be done with the following implimentation of Chen and Ye's algorithm [@chen_projection_2011].
+The essence of the algorithm is to efficiently compute the special $t\in \mathbb{R}$ so that
+
+$$
+v \leftarrow \max(0, v - t\mathbb{1}) \in \Delta_J.
+$$ {#eq-simplex-projection}
+
+The $\max(0, x)$ function should be understood as operating entrywise on $x$. In BlockTensorDecomposition, we use the helper `ReLU(x) = max(0, x)` for this function to assist with broadcasting.
+
+```julia
+function projsplx(v)
+    J = length(v)
+
+    if J==1 # quick exit for trivial length-1 "vectors" (i.e. scalars)
+        return [one(eltype(v))]
+    end
+
+    v_sorted = sort(v[:]) # Vectorize/extract input and sort all entries
+    j = J - 1
+    t = 0 # need to ensure t has scope outside the while loop
+    while true
+        t = (sum(@view v_sorted[j+1:end]) - 1) / (J-j)
+        if t >= v_sorted[j]
+            break
+        else
+            j -= 1
+        end
+
+        if j >= 1
+            continue
+        else # j == 0
+            t = (sum(v_sorted) - 1) / J
+            break
+        end
+    end
+    return ReLU.(v .- t)
+end
+```
+
+This is turned into an operation that can mutated `v` with the following definition,
+
+```julia
+function projsplx!(y)
+    y .= projsplx(y)
+end
+```
+
+and can be turned into a `ProjectedNormalization` (see @sec-constraints) with the following code.
+
+```julia
+simplex! = ProjectedNormalization(isnonnegative_sumtoone, projsplx!)
+isnonnegative_sumtoone(x) = all(isnonnegative, x) && sum(x) ≈ 1
+```
+
+The generalized Kullback-Leibler divergence projection as stated in @eq-kl-projection is only well-defined when $v[j]>0$ for all $j\in[J]$. In this case the solution is given by
+
+$$
+v \leftarrow \frac{v}{\sum_{j} v[j]}
+$$
+
+which is well described by Ducellier et. al. [@ducellier_uncertainty_2024, Sec. 2.1].
+
+To extend the applicability of this map to any $v$ (when there is at least one positive entry $v[j]>0$), we can first (Euclidean) project onto the nonnegative orthant $\mathbb{R}_+^J$,
+
+$$
+v \leftarrow \arg\,\min_{u \in \mathbb{R}_+^J} \lVert u - v \rVert _2^2 = \max(0, v),
+$$
+
+and then apply the divergence projection.^[In the unfortunate case where every entry of $v$ is nonpositive, we can fallback to the Euclean simplex projection.] All together, this looks like
+$$
+v \leftarrow \frac{\max(0, v)}{\sum_{j} \max(0, v[j])}.
+$$ {#eq-nnpr}
+
+We will refer to @eq-nnpr as nonnegative projection and rescaling (NNPR). NNPR has the following implimentation in BlockTensorDecomposition.jl.
+
+```julia
+nnpr! = l1scale! ∘ nonnegative!
+```
+
+We define the two constraints as the following, using the constraint language from @sec-constraints.
+
+```julia
+nonnegative! = Entrywise(ReLU, isnonnegative)
+l1scale! = ScaledNormalization(l1norm)
+l1norm(x) = mapreduce(abs, +, x)
+```
+
+<!-- or a rescaling followed by a reflection (assuming $v\neq 0$)
+
+$$
+v[j] \leftarrow \frac{\lvert v[j]\rvert}{\sum_{j} \lvert v[j]\rvert}.
+$$
+ -->
+
+## Constraining Tensor Factorizations to Simplexes
+
 # Multi-scale {#sec-multi-scale}
@@
 $$
+
+## Randomizing the order of updates
+
+| `group_by_factor` | `random_order` | `recursive_random_order` | Description                                                                 |
+|-------------------|----------------|--------------------------|-----------------------------------------------------------------------------|
+| `false`           | `false`        | `false`                  | In the order given                                                          |
+| `false`           | `false`        | `true`                   | In order given, but randomize how existing blocks are ordered (recursively) |
+| `false`           | `true`         | `false`                  | Randomize updates, but keep existing blocks in order                        |
+| `false`           | `true`         | `true`                   | Fully random                                                                |
+| `true`            | `false`        | `false`                  | In the order given                                                          |
+| `true`            | `false`        | `true`                   | In order of factors, but updates for each factor a random order             |
+| `true`            | `true`         | `false`                  | Random order of factors, preserve order of updates within each factor       |
+| `true`            | `true`         | `true`                   | Almost fully random, but updates for each factor are done together          |
+
+: Full description of randomizing the order of updates within a `BlockUpdate`. {#tbl-blockupdate-randomization tbl-colwidths="[20,15,27,37]"}

### 2025-02-28 example/constrainedCP.jl
fix spelling of tolerance

@@ -17,7 +17,7 @@ X_absrandn = CPDecomposition(ntuple(_ -> N, D), R; init=abs_randn)
 X_randexp = CPDecomposition(ntuple(_ -> N, D), R; init=randexp)
 
 options = (
-    tolerence=.01,
+    tolerance=.01,
     maxiter=1000,
     converged=RelativeError,
     constraints=[l1scale_cols! ∘ nonnegative!, simplex_cols!],

### 2025-02-28 example/constrainedTucker.jl

@@ -13,7 +13,7 @@ decomposition, stats_data, kwargs = fact(Y;
     rank=(2,3,4),
     model=Tucker,
     momentum=true,
-    tolerence=(1, 0.045),
+    tolerance=(1, 0.045),
     converged=(GradientNNCone, RelativeError),
     constrain_init=true,
     constraints=nonnegative!,

### 2025-02-28 example/constrainedTucker1.jl

@@ -10,7 +10,7 @@ Y = array(Y)
 decomposition, stats_data, kwargs = fact(Y;
     rank=5,
     momentum=true,
-    tolerence=(1, 0.03),
+    tolerance=(1, 0.03),
     converged=(GradientNNCone, RelativeError),
     constrain_init=true,
     constraints=nonnegative!,

### 2025-02-28 src/Core/factorize.jl

@@ -106,8 +106,8 @@ Handles all keywords and options, and sets defaults if not provided.
 
 ## Stats
 - `stats`: `[Iteration, ObjectiveValue, GradientNorm]` or in the case of nonnegative `Y`, `GradientNNCone` in place of `GradientNorm`
-- `converged`: `GradientNorm` or in the case of nonnegative `Y`, `GradientNNCone`. What stat(s) to use for convergence. Will converge is any one of the provided stats is below their respective tolerence
-- `tolerence`: `1`. A list the same length as `converged`
+- `converged`: `GradientNorm` or in the case of nonnegative `Y`, `GradientNNCone`. What stat(s) to use for convergence. Will converge is any one of the provided stats is below their respective tolerance
+- `tolerance`: `1`. A list the same length as `converged`
 - `maxiter`: `1000`. Additional stopping criterion if the number of iterations exceeds this number
 """
 function default_kwargs(Y; kwargs...)
@@ -171,8 +171,8 @@ function default_kwargs(Y; kwargs...)
 	get!(kwargs, :converged, isnonnegative(Y) ? GradientNNCone : GradientNorm) # can be a single AbstractStat or a tuple/vector of them
 											 # and must be a subset of kwargs[:stats]
 	@assert all(s -> s <: AbstractStat, kwargs[:converged])
-	get!(kwargs, :tolerence, 1) # need one tolerence per stat
-	@assert length(kwargs[:tolerence]) == length(kwargs[:converged])
+	get!(kwargs, :tolerance, 1) # need one tolerance per stat
+	@assert length(kwargs[:tolerance]) == length(kwargs[:converged])
 	get!(kwargs, :maxiter, 1000) # maximum number of iterations
 
 	union!(kwargs[:stats], kwargs[:converged]) # add any missing converged criteria to stats
@@ -382,15 +382,15 @@ end
 
 update_t(t) = 0.5*(1 + sqrt(1 + 4*t^2))
 
-function make_converged(; converged, tolerence, maxiter, kwargs...)
-	converged_tol = zip(converged, tolerence)
+function make_converged(; converged, tolerance, maxiter, kwargs...)
+	converged_tol = zip(converged, tolerance)
 	function isconverged(stats_data; kwargs...)
 		if any(((c, t),) -> stats_data[end, Symbol(c)] < t, converged_tol) # note the use of ((c, t),) rather than (c, t) since each element of the zipped converged_tol is a tuple, not two values
 			if length(converged) == 1
[…]

### 2025-02-28 src/MatrixTensorFactor/nnmtf.jl

@@ -382,7 +382,7 @@ function _nnmtf_proxgrad(
         constraints=[constraintB, constraintA],
         converged=factorize_criterion,
         maxiter,
-        tolerence=tol,#(criterion == :ncone ? tol * sqrt(problem_size) : tol), # scale tolerance by √(R(I+JK)) if using distance ot NN cone criteria
+        tolerance=tol,#(criterion == :ncone ? tol * sqrt(problem_size) : tol), # scale tolerance by √(R(I+JK)) if using distance ot NN cone criteria
         momentum,
         δ=delta,
         constrain_output,

### 2025-02-28 test/factorize_initialization.jl

@@ -15,7 +15,7 @@ Y = array(Ydecomp)
 X = CPDecomposition(ntuple(_ -> N, D), R)
 
 options = (
-    tolerence=.01,
+    tolerance=.01,
     decomposition=X,
     maxiter=1000,
     converged=RelativeError,

### 2025-02-28 test/runtests.jl

@@ -421,26 +421,26 @@ end
     # check hitting maximum number of iterations
     decomposition, stats_data = fact(Y; rank=5, momentum=false, maxiter=2);
     # check convergence on first iteration
-    decomposition, stats_data = fact(Y; rank=5, momentum=false, tolerence=Inf);
+    decomposition, stats_data = fact(Y; rank=5, momentum=false, tolerance=Inf);
     # check momentum
-    decomposition, stats_data = fact(Y; rank=5, momentum=true, tolerence=Inf);
+    decomposition, stats_data = fact(Y; rank=5, momentum=true, tolerance=Inf);
     # check constraints
     ## a single constraint, to be applied on every block
-    decomposition, stats_data = fact(Y; rank=5, constraints=nonnegative!, tolerence=Inf);
+    decomposition, stats_data = fact(Y; rank=5, constraints=nonnegative!, tolerance=Inf);
     ## a collection of constraints
-    decomposition, stats_data = fact(Y; rank=5, tolerence=Inf,
+    decomposition, stats_data = fact(Y; rank=5, tolerance=Inf,
         constraints=[ConstraintUpdate(0, nonnegative!), ConstraintUpdate(0, l1scale_12slices!), ConstraintUpdate(1, nonnegative!)],
     );
     ## check if you can constrain the initialization
-    decomposition, stats_data = fact(Y; rank=5, tolerence=Inf, constrain_init=true,
+    decomposition, stats_data = fact(Y; rank=5, tolerance=Inf, constrain_init=true,
         constraints=[ConstraintUpdate(0, nonnegative!), ConstraintUpdate(0, l1scale_12slices!), ConstraintUpdate(1, nonnegative!)],
     );
 
-    fact(Y; rank=5, tolerence=Inf, constrain_init=true,
+    fact(Y; rank=5, tolerance=Inf, constrain_init=true,
         constraints=[l1scale_12slices! ∘ nonnegative!, nonnegative!],
     ); # constraints are ok since l1scale_12slices! can be applied to the core (the 0th factor)
 
-    @test_broken fact(Y; rank=5, tolerence=Inf, constrain_init=true,
+    @test_broken fact(Y; rank=5, tolerance=Inf, constrain_init=true,
         constraints=[nonnegative!, l1scale_12slices! ∘ nonnegative!],
     ); # the constraint l1scale_12slices! cannot be applied to a matrix (the 1st factor)
 
[…]

### 2025-02-28 test/subblock_momentum_benchmark.jl

@@ -15,7 +15,7 @@ end
 
 options = (
     :rank => (2,3,4),
-    :tolerence => (0.01),
+    :tolerance => (0.01),
     :converged => (RelativeError),
     :δ => 0.9999,
     :model => Tucker,

### 2025-03-06 paper/paper.qmd
fix spelling of tolerance
all but experiment subsection for rescaling section

@@
 - Implement new "tricks"
-  - a (Lipschitz) matrix step size for efficient sub-block updates
+  - a (Lipschitz) matrix stepsize for efficient sub-block updates
   - multi-scaled factorization when tensor entries are discretizations of a continuous function
@@
 
-We typically choose $L_n^t$ to be the Lipschitz constant of $\nabla f_n^t$, since it is a sufficient condition to guarantee $f_n^t(A_n^{t+1}) \leq f_n^t(A_n^{t})$, but other step sizes can be used in theory [@nesterov_NonlinearOptimization_2018, Sec.
+We typically choose $L_n^t$ to be the Lipschitz constant of $\nabla f_n^t$, since it is a sufficient condition to guarantee $f_n^t(A_n^{t+1}) \leq f_n^t(A_n^{t})$, but other stepsizes can be used in theory [@nesterov_NonlinearOptimization_2018, Sec.
 1.2.3].
@@
 $$
-\underbrace{ 2IJKR }_{ AB-Y }+\underbrace{ IJK(2I-1) }_{ A^\top (AB-Y) } \sim 2IJKR+2I^2JK
+\underbrace{ 2I JKR }_{ AB-Y }+\underbrace{ IJK(2I-1) }_{ A^\top (AB-Y) } \sim 2I JKR+2I^2JK
 $$
@@
 $$
-\underbrace{ \frac{R(R+1)}{2}(2I-1) }_{ A^\top A } + \underbrace{ RJK(2I-1) }_{ A^\top Y } + \underbrace{ 2R^2JK }_{ (A^{\top} A)B - (A^{\top}Y) } \sim 2IJKR+2R^2JK + IR^2
+\underbrace{ \frac{R(R+1)}{2}(2I-1) }_{ A^\top A } + \underbrace{ RJK(2I-1) }_{ A^\top Y } + \underbrace{ 2R^2JK }_{ (A^{\top} A)B - (A^{\top}Y) } \sim 2I JKR+2R^2JK + IR^2
 $$
@@
     :rank => 3,
-    :tolerence => (1, 0.03),
+    :tolerance => (1, 0.03),
     :converged => (GradientNNCone, RelativeError),
@@
 
-The `BlockUpdate` language becomes especially helpful for automaically inserting additional updates as they are requested.
+The `BlockUpdate` language becomes especially helpful for automatically inserting additional updates as they are requested.
 For example, if we want to add momentum to a list of updates, we can call the following function.
@@
 
-For example, the demixing of $R$ probibility densities $b_1, \dots, b_R$ from $I$ mixtures $y_1,\dots,y_I$ for $I>R$ can be accomplished with a nonneagive matrix factorization (cite).
+With constraints being defined in a flexible manner (see @sec-constraints), we decided to test conventional wisdom that Euclidean projections are the right kind of map to use when enforcing a constraint.
+A constraint that came up in recent applications of tensor decomposition to geology [@graham_tracing_2025] was enforcing the $1$st mode slices of a tensor (e.g.
+rows in a matrix) or $3$rd mode fibres of a third order tensor to lie in their respective simplex.
+
+For example, the demixing of $R$ probability densities $b_1, \dots, b_R$ from $I$ mixtures $y_1,\dots,y_I$ for $I>R$ can be accomplished with a nonnegative matrix factorization (cite).
 We have the system of equations
@@
 with unknown mixing coefficients $a_{i,r}$ and densities $b_r$.
-If we can discritized the mixtures $y_i$, we can rewrite this system as a rank $R$ factorization of the matrix $Y$ where $Y[i,:]=y_i$ and
+If we can discretized the mixtures $y_i$, we can rewrite this system as a rank $R$ factorization of the matrix $Y$ where $Y[i,:]=y_i$ and
 
@@
 
-for matricies $A$ and $B$.
+for matrices $A$ and $B$.
 
 For the factorization to remain interpretable, we need to ensure each row $b_r$ of $B$ is a density.
-This means we would like to constrain each row $B[r,:]=b_r$ to the simplex^[This assumes the entries $b_r[j]$ in the discretization of the density $b_r$ represent probibilities or areas under some continuous 1D density function, not the sample values of the density $b_r(x_j)$.
-Assuming $J$ sample points $x_j$ of a grid on the interval $[x_0,x_J]$, the entries of the discretization vector can be defined as $b_r[j]=b_r(x_j) (x_j - x_{j-1})$.]
-
-$$
-b_r\in\Delta_J = \left\{ v\in\mathbb{R}^J \middle| \sum_{j=1}^J v[j] =1\quad \text{and}\quad  \forall j\in[J],\ \ v[j] \geq 0 \right\},
+This means we would like to constrain each row $B[r,:]=b_r$ to the simplex^[This assumes the entries $b_r[j]$ in the discretization of the density $b_r$ represent probabilities or areas under some continuous 1D density function, and not the sample values of the density $b_r(x_j)$.
+One possible discretization is to take $J$ sample points $x_j$ of a grid on the interval $[x_0,x_J]$ where $b_r$ is supported, and define the entries of the discretization to be $B[r,j] = b_r[j] = b_r(x_j) (x_j - x_{j-1})$.
+For normalized densities $\int_{x_0}^{x_J}b_r(x)dx=1$, the sum of the entries $\sum_{j=1}^J b_r[j] \approx 1$ when large enough number of samples $J$ are taken.]
+
+$$
+b_r\in\Delta_J = \left\{ v\in\mathbb{R}_+^J \middle| \sum_{j=1}^J v[j] =1 \right\},
 $$
@@
 $$
-B\in\Delta_{R, J}=\left\{B\in\mathbb{R}^{R\times J} \middle| \forall r\in [R],\, \sum_{j=1}^J B[r,j] = 1  \quad \text{and}\quad \forall (r,j)\in [R]\times [J],\, B[r,j]\geq 0 \right\}.
+B\in\Delta_{J}^R=\left\{B\in\mathbb{R}_+^{R\times J} \middle| \forall r\in [R],\, \sum_{j=1}^J B[r,j] = 1 \right\}.
 $$
@@
 $$
-A\in\Delta_{I, R}=\left\{A\in\mathbb{R}^{I\times R} \middle| \forall i\in [I],\, \sum_{r=1}^R A[i,r] = 1  \quad \text{and}\quad \forall (i,r)\in [I]\times [R],\, A[i,r]\geq 0 \right\}.
+A\in\Delta_{R}^I=\left\{A\in\mathbb{R}_+^{I\times R} \middle| \forall i\in [I],\, \sum_{r=1}^R A[i,r] = 1 \right\}.
 $$
@@
 
+The question we investigate in this section is the following: how can we best constrain the factors $A$ and $B$ to their respective simplexes, while performing block gradient decent to minimize the least squared error $\frac12 \left \lVert AB - Y \right \rVert _F^2$?
+
 ## The two approaches for simplex constraints
 
-With constraints being defined in a flexible manner (see @sec-constraints), we decided to test conventional wisdom that Euclidean projections are the right kind of map to use when enforcing a constraint.
-A constraint that came up in applications was enforcing the $1$st mode slices of a tensor (e.g.
-rows in a matrix) or $3$rd mode fibres of a third order tensor to lie in their respective simplex.
-
-For example, to constrain a vector $v\in\mathbb{R}^J$ to the simplex
-
-$$
-\Delta_J = \left\{ v\in\mathbb{R}^J \middle| \sum_{j=1}^J v[j] =1,\quad \text{and}\quad  \forall j\in[J],\ \ v[j] \geq 0 \right\},
+To constrain a vector $v\in\mathbb{R}^J$ to the simplex
+
+$$
+\Delta_J = \left\{ v\in\mathbb{R}_+^J \middle| \sum_{j=1}^J v[j] =1\right\},
 $$
@@
 
-or a generalized Kullback-Leibler divergence projection
+or a generalized Kullback-Leibler (KL) divergence projection
 
@@
 
-amoung other reasonable maps onto $\Delta_J$.
-
-The Eucledian simplex projection can be done with the following implimentation of Chen and Ye's algorithm [@chen_projection_2011].
+among other reasonable maps onto $\Delta_J$.
+
+The Euclidean simplex projection can be done with the following implementation of Chen and Ye's algorithm [@chen_projection_2011].
 The essence of the algorithm is to efficiently compute the special $t\in \mathbb{R}$ so that
@@
 
-and then apply the divergence projection.^[In the unfortunate case where every entry of $v$ is nonpositive, we can fallback to the Euclean simplex projection.] All together, this looks like
+and then apply the divergence projection.^[In the unfortunate case where every entry of $v$ is nonpositive, we can fallback to the Euclidean simplex projection.] All together, this looks like
 $$
@@
 
-We will refer to @eq-nnpr as nonnegative projection and rescaling (NNPR). NNPR has the following implimentation in BlockTensorDecomposition.jl.
-
-```julia
-nnpr! = l1scale! ∘ nonnegative!
+We will refer to @eq-nnpr as nonnegative projection and rescaling (NNPR). NNPR has the following implementation in BlockTensorDecomposition.jl.
+
+```julia
+l1scale! ∘ nonnegative!
 ```
@@
 
-## Constraining Tensor Factorizations to Simplexes
+## The Rescaling Trick for Matrix Factorization {#sec-matrix-rescaling}
+
+- Explain that we can move the weight from one matrix to another
+
+Comparing the two methods of constraining a vector to the simplex 1) by Euclidean projection (@eq-simplex-projection) or 2) nonnegative projection and rescaling (NNPR, @eq-nnpr), the latter offers a few advantages.
+NNPR is cheaper and conceptually easier to compute.
+Another advantage of NNPR to tensor factorizations, is its ability to constrain a factor without loosing progress while performing gradient descent.
+
+For example, consider the low rank factorization problem of finding matrices $A,B$ such that $Y=AB$ where you would like the sum of entries in $B$ to be one
+
+$$
+\min_{A \in \mathbb{R}^{I\times R},  B \in \mathbb{R}^{R\times J}} \frac12\left\lVert AB - Y \right\rVert_{F}^2 \quad \text{s.t.} \quad B \in \Delta_{RJ} = \left\{ B \in \mathbb{R}_+^{R\times J} \middle| \sum_{r,j} B[r,j] = 1 \right\}.
+$$ {#eq-B-full-simplex-problem}
+
+The basic alternating projected gradient descent algorithm using a Euclidean projection would be
+
+$$
+\begin{align}
+A &\leftarrow A - \frac{1}{L_A} \nabla_A f(A,B)\\
+B &\leftarrow EP_{\Delta_{RJ}}\left(B - \frac{1}{L_B} \nabla_B f(A,B)\right)
+\end{align}
+$$
+
+where $f(A,B) = \frac12\left\lVert AB - Y \right\rVert_{F}^2$ and $EP_{\Delta_{RJ}}$ is the Euclidean projection onto the simplex $\Delta_{RJ}$. In the event the updated value for $B$,
+
+$$
+B - \frac{1}{L_B} \nabla_B f(A,B) :=\hat{B}\in \mathbb{R}_+^{R\times J}
+$$
+
+is already nonnegative, the objective $f$ at the new point $(A,EP_{\Delta_{RJ}}(\hat{B}))$ could be bigger or smaller than the objective before the Euclidean projection $f(A, \hat{B})$.
+
+If we use the nonnegative projection and rescaling $\text{NNPR}_{\Delta_{RJ}}$ instead of the Euclidean projection $EP_{\Delta_{RJ}}$ when $\hat{B}$ is already nonnegative, then
+$$
+\text{NNPR}_{\Delta_{RJ}}(\hat{B}) = \frac{1}{\sum_{rj}B[r,j]}\hat{B} := c^{-1}\hat{B}.
+$$
+
+This means the objective value $f$ at the point $(c^{-1}A, c\hat{B})$ will be the same as the objective value before the KL divergence projection
+
+$$
+f(cA, \text{NNPR}_{\Delta_{RJ}}(\hat{B})) = f(cA, c^{-1}\hat{B}) = \frac12\left\lVert Ac c^{-1}B - Y \right\rVert_{F}^2 = \frac12\left\lVert AB - Y \right\rVert_{F}^2 = f(A, \hat{B}).
+$$
+
+This suggests the following update may be a useful alternative to the standard projected gradient descent
+
+$$
+\begin{align}
+A &\leftarrow A - \frac{1}{L_A} \nabla_A f(A,B)\\
+B &\leftarrow B - \frac{1}{L_B} \nabla_B f(A,B)\\
+c &\leftarrow \sum_{r,j}B[r,j] \\
+A &\leftarrow cA \\
+B &\leftarrow c^{-1}B .
+\end{align}
+$$
+
+Of course, it is possible that $\hat{B}=B - \frac{1}{L_B} \nabla_B f(A,B)$ has negative entries.
+So we use both the nonnegative projection and rescaling part of NNPR in the alternating gradient descent with rescaling update
+
+$$
+\begin{align}
+A &\leftarrow A - \frac{1}{L_A} \nabla_A f(A,B)\\
+B &\leftarrow \max(0,B - \frac{1}{L_B} \nabla_B f(A,B))\\
+c &\leftarrow \sum_{r,j}B[r,j] \\
+A &\leftarrow cA \\
+B &\leftarrow c^{-1}B .
+\end{align}
+$$ {#eq-nnpr-gd}
+
+This algorithm can be called in BlockTensorDecomposition.jl with the following code.
+
+```julia
+options = (
+    model=Tucker1,
+    constraints=[l1scale! ∘ nonnegative!, noconstraint],
+)
+
+decomposition, stats, kwargs = factorize(Y; options...);
+```
+
+## Generalizing the Matrix Rescaling Trick
+
+The rescaling trick discussed in @sec-matrix-rescaling applies more generally to other tensor factorization, other simplex-type constraints, and other `ScaledNormalization`s.
+
+### Simplex-type constraints
+
+Instead of the matrix factorization problem where $B\in\Delta_{RJ}$ is constrained to the full simplex (@eq-B-full-simplex-problem), we can apply the rescaling trick the problem where the rows of $B$ are constrained to the simplex
+
+$$
+\min_\substack{A \in \mathbb{R}^{I\times R}\\  B \in \mathbb{R}^{R\times J}} \frac12\left\lVert AB - Y \right\rVert_{F}^2 \quad \text{s.t.} \quad B \in \Delta_{J}^R = \left\{ B \in \mathbb{R}_+^{R\times J} \middle| \forall r\in[R],\,\sum_{j=1}^J B[r,j] = 1 \right\}.
+$$ {#eq-B-row-simplex-problem}
+
+This could make sense in applications where rows of $B$ represent probability densities such as the demixing problem discussed at the start of this section (@sec-ppr).
+
+We can adjust the alternating gradient descent update with NNPR (@eq-nnpr-gd) to the following update
+$$
+\begin{align}
+A &\leftarrow A - \frac{1}{L_A} \nabla_A f(A,B)\\
+B &\leftarrow \max(0,B - \frac{1}{L_B} \nabla_B f(A,B))\\
+C[r,r] &\leftarrow \sum_{j}B[r,j]\quad \text{($C\in\mathbb{R}^{R\times R}$ is diagonal)} \\
+A &\leftarrow AC \\
+B &\leftarrow C^{-1}B .
+\end{align}
+$$ {#eq-nnpr-gd-B-rows}
+
+It is clear that the objective value would be maintained for any invertible matrix $C$
+
+$$
+f(AC,C^{-1}B) = \frac12\left\lVert ACC^{-1}B - Y \right\rVert_{F}^2 = \frac12\left\lVert AB - Y \right\rVert_{F}^2 = f(A,B).
+$$
+
+This algorithm can be called in BlockTensorDecomposition.jl with the following code.
+
+```julia
+options = (
+    model=Tucker1,
+    constraints=[l1scale_rows! ∘ nonnegative!, noconstraint],
+)
+
+decomposition, stats, kwargs = factorize(Y; options...);
+```
+
+::: {.callout-warning}
+This trick would also work if we wanted the columns of $A$ normalized to the simplex.
+But it does *not* work when we would like each column of $B$ to be constrained to the simplex.
+The normalizing matrix $C$ would have to be multiplied to the right of $B$ rather than between $A$ amd $B$.
+A similar story can be said with the rows of $A$.
+:::
+
+### Other `ScaledNormalization`'s
+
+This trick can also apply to other normalization constraints.
+For example, we may want the maximum magnitude of each row of $B$ to one.
+This could make sense in applications where each row represents a waveform audio file (WAV) which has an audio format that takes values between $-1$ and $1$.
+Instead of the Euclidean projection^[In audio processing, this is commonly called "clipping".]
+
+$$
+v \leftarrow \max(-1, \min(1, v))
+$$
+
+onto the infinity ball
+
+$$
+\mathcal{B}_J({\infty}) = \left\{v\in\mathbb{R}^J \middle| \max_{j\in[J]} \lvert v[j] \rvert \leq 1\right\},
+$$
+
+we can apply the rescaling^[In audio processing, this is commonly called "normalizing".
+Normalizing is often preferred to clipping since it maintains the perceived audio, but just at a different volume than the original signal.
+Clipping often introduces undesirable distortion (think of sound from a megaphone).]
+
+$$
+v \leftarrow \frac{v}{\max_{j\in[J]} \lvert v[j] \rvert}.
+$$
+
+Applying this principle to the factorization problem
+
+$$
+\min_\substack{A \in \mathbb{R}^{I\times R}\\  B \in \mathbb{R}^{R\times J}} \frac12\left\lVert AB - Y \right\rVert_{F}^2 \quad \text{s.t.} \quad B \in \mathcal{B}_J^R({\infty}) = \left\{ B \in \mathbb{R}^{R\times J} \middle| \forall r\in[R],\,\max_{j\in[J]} \lvert B[r,j] \rvert \leq 1 \right\}
+$$ {#eq-B-row-infinity-problem}
+
+gives us the update
+
+$$
+\begin{align}
+A &\leftarrow A - \frac{1}{L_A} \nabla_A f(A,B)\\
+B &\leftarrow B - \frac{1}{L_B} \nabla_B f(A,B)\\
+C[r,r] &\leftarrow \max_{j\in[J]} \lvert B[r,j] \rvert\quad \text{($C\in\mathbb{R}^{R\times R}$ is diagonal)} \\
+A &\leftarrow AC \\
+B &\leftarrow C^{-1}B .
+\end{align}
+$$ {#eq-nnpr-gd-B-rows-infinity}
+
+This algorithm can be called in BlockTensorDecomposition.jl with the following code.
+
+```julia
+options = (
+    model=Tucker1,
+    constraints=[linftyscale_rows! ∘ nonnegative!, noconstraint],
+)
+
+decomposition, stats, kwargs = factorize(Y; options...);
+```
+
+A similar story can be made about other $p$-norm balls $\mathcal{B}(p)$ or spheres.
+
+### Other Tensor Factorizations
+
+The rescaling trick is applicable to other tensor factorizations, but is dependent on the exact model and constraints.
+
+The simplest extension of matrix factorization is the Tucker-1 factorization of an order $N$ tensor $Y = B \times_1 A$.
+
+If we would like the first order slices of $B$ to be constrained to the simplex
+
+$$
+\begin{align}
+B[r, :, :] \in \Delta_{JK} &= \left\{ B[r, :, :] \in \mathbb{R}_+^{J\times K} \middle| \sum_{j,k} B[r,j,k] = 1\right\}\\
+B \in \Delta_{JK}^R &= \left\{ B \in \mathbb{R}_+^{R\times J\times K} \middle|\forall r\in[R],\, \sum_{j,k} B[r,j,k] = 1\right\},
+\end{align}
+$$
+
+we can solve the problem
+
+$$
+\min_\substack{A \in \mathbb{R}^{I\times R}\\  B \in \mathbb{R}^{R\times J\times K}} \frac12\left\lVert B\times_1 A - Y \right\rVert_{F}^2 \quad \text{s.t.} \quad B \in \Delta_{JK}^R.
+$$ {#eq-B-slice-simplex-problem}
+
+by iterating the update
+
+$$
+\begin{align}
+A &\leftarrow A - \frac{1}{L_A} \nabla_A f(A,B)\\
+B &\leftarrow B - \frac{1}{L_B} \nabla_B f(A,B)\\
+C[r,r] &\leftarrow \sum_{j\in[J],k\in[K]}  B[r, j, k] \quad \text{($C\in\mathbb{R}^{R\times R}$ is diagonal)} \\
+A &\leftarrow AC \\
+B &\leftarrow B \times_1 C^{-1} .
+\end{align}
+$$ {#eq-nnpr-gd-B-slices-simplex}
+
+This algorithm can be called in BlockTensorDecomposition.jl with the following code.
+
+```julia
+options = (
+    model=Tucker1,
+    constraints=[l1scale_1slices! ∘ nonnegative!, noconstraint],
+)
+
+decomposition, stats, kwargs = factorize(Y; options...);
+```
+
+A setting where this model applies would be if the first order slices of $B$ represent 2-dimensional probability densities.^[If sampling a 2-dimensional probability density $p_r(x,y)$ on a rectangular grid, entries of $B$ could be interpreted as $B[r, j, k]= p_r(x_j,y_k)(x_j - x_{j-1}) (y_k - y_{k-1})$.]
+
+For constraints that constrain an entire factor to some scale, the weight of that factor can be distributed to one or multiple other factors.
+For example, we may wish to find a CP-decomposition of an order 4-tensor $Y=\llbracket A,B,C,D\rrbracket$ where the Frobenius norm of $A$ is $1$.
+In the rescaling step, we could move the norm of $A$ to just the matrix $B$
+
+$$
+\begin{align}
+c \leftarrow \left\lVert A\right\rVert_F \\
+A \leftarrow c^{-1} A \\
+B \leftarrow cB
+\end{align}
+$$
+
+or all three other factors equally
+
+$$
+\begin{align}
+c \leftarrow \left\lVert A\right\rVert_F \\
+A \leftarrow c^{-1} A \\
+B \leftarrow c^{1/3}B \\
+C \leftarrow c^{1/3}C \\
+D \leftarrow c^{1/3}D.
+\end{align}
+$$
+
+In either case, the recombined tensor remains unchanged
+
+$$
+\llbracket c^{-1}A,cB,C,D\rrbracket = \llbracket c^{-1}A,c^{1/3}B,c^{1/3}C,c^{1/3}D\rrbracket = \llbracket A,B,C,D\rrbracket.
+$$
+
+The two algorithms can be called in BlockTensorDecomposition.jl with the following code.
+
+```julia
+options = (
+    model=CPDecomposition,
+    constraints=ConstraintUpdate(1, l2scaled!;
+        whats_rescaled=(Y -> factor(Y, 2))), # B is the second factor
+)
+
+decomposition, stats, kwargs = factorize(Y; options...);
+```
+
+```julia
+options = (
+    model=CPDecomposition,
+    constraints=ConstraintUpdate(1, l2scaled!),
+    # assumes all other factors are rescaled
+)
+
+decomposition, stats, kwargs = factorize(Y; options...);
+```
+
+## Constraining Multiple Factors
+
+When constraining multiple factors with the rescaling approach, there must be at least one factor that is not constrained with rescaling.
+
+Consider the following matrix factorization $Y=AB$ problem where we want both the columns of $A$ and rows of $B$ to be constrained to the simplex.^[We define the constraint on $A$ in terms of $A^{\top}$ to be consistent with the simplex constraint defined on $B$.]
+
+$$
+\begin{align}
+&\min_\substack{A \in \mathbb{R}^{I\times R}\\  B \in \mathbb{R}^{R\times J}} \frac12\left\lVert AB - Y \right\rVert_{F}^2 \\
+&\text{s.t.} \\
+\quad A^\top \in \Delta_{I}^R &= \left\{ A^\top \in \mathbb{R}_+^{R\times I} \middle|\forall r\in[R],\, \sum_{i} A^\top[r,i] = 1\right\} \\
+\quad B \in \Delta_{J}^R &= \left\{ B \in \mathbb{R}_+^{R\times J} \middle|\forall r\in[R],\, \sum_{j} B[r,j] = 1\right\}
+\end{align}
+$$ {#eq-AB-row-col-simplex-problem}
+
+If we try to rescale both $A$ and $B$ while moving the weights to the other factor, we often observe numerical instability or blow up.
+To be clear, the following update does not seem to work in practice.
+
+$$
+\begin{align}
+A &\leftarrow \max(0, A - \frac{1}{L_A} \nabla_A f(A,B))\\
+C_{A}[r,r] &\leftarrow \sum_{i\in[I]} A[i,r] \\
+A &\leftarrow AC_A^{-1} \\
+B &\leftarrow C_A B \\
+B &\leftarrow \max(0, B - \frac{1}{L_B} \nabla_B f(A,B))\\
+C_B[r,r] &\leftarrow \sum_{j\in[J]} B[r,j] \\
+A &\leftarrow A C_B \\
+B &\leftarrow C_B^{-1}B
+\end{align}
+$$
+
+The relevant call in BlockTensorDecomposition would be the following.
+
+```julia
+options = (
+    model=Tucker1,
+    constraints=[ # B is the 0th factor, A is the 1st factor
+        ConstraintUpdate(1, l1scale_cols! ∘ nonnegative!;
+            whats_rescaled=(x -> eachrow(factor(x, 0)))),
+        ConstraintUpdate(0, l1scale_rows! ∘ nonnegative!;
+            whats_rescaled=(x -> eachcol(factor(x, 1)))),
+    ],
+)
+
+decomposition, stats, kwargs = factorize(Y; options...);
+```
+
+Instead, one of the factors can be scaled without moving the weight to the other factor. For example, if we wanted to remove the 4th line $B\leftarrow C_A B$, we can call the following.
+
+```julia
+options = (
+    model=Tucker1,
+    constraints=[ # B is the 0th factor, A is the 1st factor
+        ConstraintUpdate(1, l1scale_cols! ∘ nonnegative!;
+            whats_rescaled=nothing),
+        ConstraintUpdate(0, l1scale_rows! ∘ nonnegative!;
+            whats_rescaled=(x -> eachcol(factor(x, 1)))),
+    ],
+)
+
+decomposition, stats, kwargs = factorize(Y; options...);
+```
+
+This approach seems to work better in practice. The same principle of relaxing how constraints are enforced can be applied to the very similar problem
+
+$$
+\begin{align}
+&\min_\substack{A \in \mathbb{R}^{I\times R}\\  B \in \mathbb{R}^{R\times J}} \frac12\left\lVert AB - Y \right\rVert_{F}^2 \\
+&\text{s.t.} \\
+\quad A \in \Delta_{R}^I &= \left\{ A \in \mathbb{R}_+^{I\times R} \middle|\forall i\in[I],\, \sum_{r} A[i,r] = 1\right\} \\
+\quad B \in \Delta_{J}^R &= \left\{ B \in \mathbb{R}_+^{R\times J} \middle|\forall r\in[R],\, \sum_{j} B[r,j] = 1\right\},
+\end{align}
+$$ {#eq-AB-row-simplex-problem}
+
+where we want both the rows of $A$ and $B$ to be constrained to the simplex.
+Here, we cannot move the weights from $A$ to $B$ since there are $I$ rows of $A$ but only $R$ rows of $B$.
+Instead, we relax the problem to
+
+$$
+\begin{align}
+&\min_\substack{A \in \mathbb{R}^{I\times R}\\  B \in \mathbb{R}^{R\times J}} \frac12\left\lVert AB - Y \right\rVert_{F}^2 \\
+&\text{s.t.} \\
+&\quad A \in \mathbb{R}_+^{I\times R}  \\
+\quad B \in \Delta_{J}^R &= \left\{ B \in \mathbb{R}_+^{R\times J} \middle|\forall r\in[R],\, \sum_{j} B[r,j] = 1\right\}.
+\end{align}
+$$ {#eq-AB-row-simplex-problem-relaxed}
+
+The relevant update for this relaxed problem would be
+
+$$
+\begin{align}
+A &\leftarrow \max(0, A - \frac{1}{L_A} \nabla_A f(A,B))\\
+B &\leftarrow \max(0, B - \frac{1}{L_B} \nabla_B f(A,B))\\
+C_B[r,r] &\leftarrow \sum_{j\in[J]} B[r,j] \\
+A &\leftarrow A C_B \\
+B &\leftarrow C_B^{-1}B
+\end{align}
+$$
+
+and is called in BlockTensorDecomposition.jl with the following code.
+
+```julia
+options = (
+    model=Tucker1,
+    constraints=[ # B is the 0th factor, A is the 1st factor
+        ConstraintUpdate(1, nonnegative!),
+        ConstraintUpdate(0, l1scale_rows! ∘ nonnegative!;
+            whats_rescaled=(x -> eachcol(factor(x, 1)))),
+    ],
+)
+
+decomposition, stats, kwargs = factorize(Y; options...);
+```
+
+We justify this relaxation with the following argument. When the rows of $Y$ are in the simplex, we can bound how close the rows of $A$ are to summing to one with @thm-closeness-to-simplex.
+
+::: {#thm-closeness-to-simplex}
+
+# Closeness of A's rows summing to one
+
+Let $Y\in\Delta_J^I$, $A\in\mathbb{R}_+^{I\times R}$, and $B\in\Delta_{J}^R$ where
+
+$$
+\left\lVert Y - AB \right\rVert _F \leq \epsilon.
+$$
+
+Then for any row $i\in[I]$, the sum of the row is $\epsilon\sqrt{J}$ close to 1
+
+$$
+\left\lvert 1 - \sum_{r\in[R]} A[i,r] \right\rvert \leq \epsilon\sqrt{J}.
+$$
+
+::: {.proof}
+
+We have the following inequalities.
+
+$$
+\begin{align}
+\left\lVert Y - AB \right \rVert_F &\leq \epsilon \\
+\frac{1}{\sqrt{J}}\left\lVert Y - AB \right \rVert_\infty &\leq \epsilon \quad (\frac{1}{\sqrt{J}} \left\lVert X \right\rVert _\infty  \leq \left\lVert X \right\rVert_2\leq \left\lVert X \right\rVert _F\ \text{for}\ X\in\mathbb{R}^{I\times J})\\
+\frac{1}{\sqrt{J}}\max_{i\in[I]} \left(\sum_{j\in[J]} \left\lvert(Y - AB)[i,j]\right\rvert \right) &\leq \epsilon\\
+\max_{i\in[I]} \left(\sum_{j\in[J]} \left\lvert(Y - AB)[i,j]\right\rvert \right) &\leq \epsilon\sqrt{J}
+\end{align}
+$$
+So for all rows $i\in[I]$,
+
+$$
+\begin{align}
+\epsilon \sqrt{J} &\geq \sum_{j\in[J]} \left\lvert(Y - AB)[i,j]\right\rvert \\
+&\geq  \left\lvert \sum_{j\in[J]} (Y - AB)[i,j]\right\rvert \\
+&=  \left\lvert \sum_{j\in[J]} Y[i,j] - \sum_{j\in[J]} (AB)[i,j]\right\rvert\\
+&=  \left\lvert 1 - \sum_{j\in[J]} \sum_{r\in[R]}A[i,r]B[r,j]\right\rvert \quad (\text{since}\ Y \in\Delta_J^I)\\
+&=  \left\lvert 1 - \sum_{r\in[R]}A[i,r] \left( \sum_{j\in[J]} B[r,j]\right)\right\rvert\\
+&=  \left\lvert 1 - \sum_{r\in[R]}A[i,r] \right\rvert  \quad (\text{since}\ B \in\Delta_J^R).
+\end{align}
+$$
+
+:::
+
+:::
+
+@thm-closeness-to-simplex implies that solutions to the relaxed problem (@eq-AB-row-simplex-problem-relaxed) are approximate solutions to the problem shown in @eq-AB-row-simplex-problem.
+Moreover, if there exists an exact factorization $Y=AB$ for the relaxed problem, then it is also a solution the original problem.
+
+## Experiment
+
+To illustrate the advantage of this rescaling trick over Euclidean projection, we will consider solving the problem shown in @eq-AB-row-simplex-problem where $Y\in\Delta_J^I$ and there exists an exact factorization $Y=AB$ with $A\in\Delta_{R}^I$ and $B\in\Delta_{J}^R$ using eight different algorithms.
 

### 2025-03-06 src/Core/decomposition.jl
add native matrix factorization TODO

@@ -355,3 +355,6 @@ Base.getindex(CPD::CPDecomposition, I::Vararg{Int}) = sum(reduce(.*, (@view f[i,
 # Additional CPDecomposition interface
 """The single rank for a CP Decomposition"""
 rankof(CPD::CPDecomposition) = size(matrix_factors(CPD)[begin], 2)
+
+# TODO Define a matrix factorization type Y=AB so that thinking about tensors can be avoided
+# Could even exploit Tucker1 or CPDecomposition to do this

### 2025-03-12 src/Core/blockupdates.jl
add a safe invert

@@ -33,8 +33,16 @@ end
 
 function (step::LipschitzStep)(x; kwargs...)
     L = step.lipschitz(x)
-    return L^(-1) # allow for Lipschitz to be a diagonal matrix
+    try
+        return L^(-1)  # allow for Lipschitz to be a diagonal matrix
+    catch
+        @warn "Could not invert the Lipschitz constant to get a stepsize. Ignoring zero coordinates."
+
+        return _safe_invert.(L)
+    end
 end
+
+_safe_invert(x) = iszero(x) ? x : x^(-1)
 
 function (step::LipschitzStep)(x::Tucker; kwargs...)
     L = step.lipschitz(x)

### 2025-03-12 test/rescaling_vs_projection.jl
add rescaling test file

@@ -1,0 +1,116 @@
+# This is a test of different approaches to solving the NMF problem
+# 1/2 ||AB-Y||_F^2
+# where the rows of A, B, & Y all sum to 1
+
+using BlockTensorDecomposition
+using Random
+using UnicodePlots
+using Plots
+
+Random.seed!(31415926535897)
+
+fact = BlockTensorDecomposition.factorize
+
+I, J = 25, 50
+R = 5
+
+A = abs_randn(I, R)
+l1scale_rows!(A)
+
+B = abs_randn(R, J)
+l1scale_rows!(B)
+
+Y = Tucker1((B, A))
+Y = array(Y)
+
+@assert check(l1scale_rows!, Y)
+
+A_init = abs_randn(I, R)
+l1scale_rows!(A_init)
+
+B_init = abs_randn(R, J)
+l1scale_rows!(B_init)
+
+n_iterations = 5000
+
+options = (
+    model=Tucker1,
+    tolerance=0.0, # Force the algorithm to halt at maxiter
+    maxiter=n_iterations,
+    converged=RelativeError,
+    stats=[
+        Iteration, ObjectiveValue, GradientNNCone, RelativeError, #PrintStats
+    ],
+    momentum=false,
+    do_subblock_updates=false,#false#true
+)
+
+# Possible updates
+# B is the 0th factor, A is the 1st factor
+
+scaleB_rows_rescaleA! = ConstraintUpdate(0, l1scale_rows! ∘ nonnegative!;
+    whats_rescaled=(x -> eachcol(factor(x, 1)))
+)
+
+scaleB_rows! = ConstraintUpdate(0, l1scale_rows! ∘ nonnegative!;
+    whats_rescaled=nothing
+)
+
+scaleA_rows! = ConstraintUpdate(1, l1scale_rows! ∘ nonnegative!;
+    whats_rescaled=nothing
+)
+
+nonnegativeA! = ConstraintUpdate(1, nonnegative!)
+
+nonnegativeB! = ConstraintUpdate(0, nonnegative!)
+
+euclideanProjA! = ConstraintUpdate(1, simplex_rows!)
+
+euclideanProjB! = ConstraintUpdate(0, simplex_rows!)
+
+algorithms = (
+    bothSimplex = [euclideanProjB!, euclideanProjA!],
+    simplexA = [nonnegativeB!, euclideanProjA!],
+    simplexB = [euclideanProjB!, nonnegativeA!],
+    bothScale = [scaleB_rows!, scaleA_rows!],
[…]

### 2025-03-12 paper/paper.qmd
start on multi scale section

@@
 
+In many applications [@saylor_CharacterizingSedimentSources_2019] [TODO add spatial transcriptomics], the data tensor $Y$ represents a discretization of continuous data.
+In these cases, we can decide how finely or coarsely to sample the data and effectively have control over the dimension of tensor needing to be factored.
+
+For high quality results, we would like as fine of a discretization as possible.
+Or if the data is already collected in a discretized form, we would like to incorporate all the data.
+But smaller tensors are faster to decompose because there are fewer parameters to fit and learn.
+Most matrix operations like addition, multiplication, and finding their norm are also faster for smaller tensors.
+
+We propose a multiresolution approach inspired by multi-grid [] and wavelets [] that factorizes the data at progressively finer scales, using the .
+The can greatly speed up how fast large scale tensors can be factorized.
+
+## Basic Approach {#sec-basic-multi-scale}
+
+We will start with same example from the beginning of @sec-ppr.
+Give a tensor $Y\in\mathbb{R}_+^{I\times J}$ representing mixtures of discretized densities, we would like to demix the densities according to the model
+
+$$
+\begin{align*}
+\begin{bmatrix}
+\leftarrow & \ y_{1}^\top & \rightarrow \\
+\leftarrow & \ y_{2}^\top & \rightarrow  \\
+\leftarrow & \ \vdots & \rightarrow \\
+\leftarrow & \ y_{I}^\top & \rightarrow
+\end{bmatrix}
+& = \begin{bmatrix}
+a_{11} & a_{12} & \dots & a_{1R} \\
+a_{21} & a_{22} & \dots & a_{2R} \\
+ \vdots &  &  & \vdots \\
+a_{I1} & a_{I2} & \dots & a_{IR} \\
+\end{bmatrix}
+\begin{bmatrix}
+ \leftarrow &{\ b_{1}}^\top & \rightarrow \\
+ \leftarrow &{\ b_2} ^\top & \rightarrow \\
+\ & \vdots & \ \\
+\leftarrow &{\ b_{R}}^\top& \rightarrow
+\end{bmatrix}
+\end{align*}
+$$
+
+or
+$$
+Y = AB.
+$$
+
+Notice that the rows of $Y$ and $B$ represent samples of continuous densities, so the size of their second dimension $J$ is arbitrary.
+Suppose each of the 1-dimensional densities are uniformly discretized on an interval $[a,b]$ with $J_s$ number of points.
+We use $s$ to represent the scale or spacing of the number of points.
+For example, $J_1=J$ would be the finest scale using every point in a discretization $x_1,x_2,x_3\dots,x_{J_1}$ with $x_1=a$ and $x_{J_1}=b$, $J_2=J_1/2$ would be coarser and use every other point $x_1,x_3,x_5,\dots x_{J_1 - 1}$.^[We assume $J$ is even here, but we could define $J_2=(J_1 -1)/2$ if $J_{1}=J$ is odd.]
+
+The basic approach is to factorize $Y_2\in\mathbb{R}^{I\times J_2}$ with entries $Y_2[i,j]=y_i(x_{2j-1})$ to obtain $A_2^{T_2}\in\mathbb{R}^{I\times R}$ and $B_^{T_2}\in\mathbb{R}^{I\times J_2}$ after $T_2$ many iterations.
+We use the factors $A_2^{T_2}$ and $B_2^{T_2}$ to initialize the factorization of $Y_1=Y\in\mathbb{R}^{I\times J_1}$.
+We can initialize $A_1^0 = A_2^{T_2}$ since the size of $A$ is the same at both scales, and repeat every entry of $B$ to initialize $B_1^0$ with entries $B_1^0[i,j] = B_2^{T_2}[i,\lceil j/2 \rceil]$.
+
+Factorizing $Y_2$ is faster than factorizing $Y_1$ since 1) there are fewer parameters to learn^[Only $(I+J_2)R$ at the coarse scale which is less than $(I+J_1)R$ for the fine scale.] and 2) most arithmetic like addition and multiplication, as well as calling other operators like `norm` are faster to compute.
+This gives us a better initialization for $A$ and $B$ in the factorization of $Y_1$ than some other random initialization so that fewer iterations are needed at the more expensive finer scale.
+
+## General Approach
+
+The basic approach can be generalized in two ways: the data could be continuous in multiple dimensions, and we can recursively apply this multi-scale approach to progressively refine a very coarse factorization.
+
+Suppose we are given a tensor $Y\in\mathbb{R}^{I_1 \times \dots \times I_N}$ where the dimensions $I_{n_1}, \dots, I_{n_M}$ represent a grided discretization of $M$-dimensional continuous data.
+
+A natural extension of the example shown in @sec-basic-multi-scale would be to higher dimensional distributions.
+For example, we could consider an order-$3$ tensor where the horizontal slices correspond to a $2$-dimensional discretization of a bivariate density.
+Entries of the input tensor $Y$ would be given by $Y[i,j,k] = f_i(x_j, y_k)$ for continuous probability density functions $f_i:\mathbb{R}^2\rightarrow \mathbb{R}_+$ for a 2D grid of points $(x_j, y_k)$.
+In this example, the second and third dimensions would be continuous.
+
+Factors that de
+
+, and $J_s$ would be a discretization that uses every $2^{s}$ points, or $J_s = 2^{S-s}+1$ points in total.
+
+The basic approach will be to factorize the coarsest version of $Y$, $Y_S\in\mathbb{R}^{I\times J_S}$, to obtain $A_S$ and $B_S$.
+
+## Convergence of a Multi-scale Method
+
+- show the multi-scale method has tighter bounds than regular gradient descent for lipschitz data
+- this assumes no constraints
+
 # Conclusion

### 2025-03-19 src/Core/multiscale.jl
add interpolate function

@@ -22,3 +22,65 @@ coarsen(Y::AbstractArray, scale::Integer; dims=1:ndims(Y), kwargs...) =
     Y[(d in dims ? axis[begin:scale:end] : axis for (d, axis) in enumerate(axes(Y)))...]
 
 # Using axis[begin:scale:end] rather than 1:scale:size(Y, d) for more flexible indexing
+
+"""
+    interpolate(Y, scale; dims=1:ndims(Y), kwargs...)
+
+Interpolates Y to a larger array with repeated values.
+
+Like the opposite of [`coarsen`](@ref).
+
+Example
+=======
+
+julia> Y = collect(reshape(1:6, 2, 3))
+2×3 Matrix{Int64}:
+ 1  3  5
+ 2  4  6
+
+julia> interpolate(Y, 2)
+3×5 Matrix{Int64}:
+ 1  1  3  3  5
+ 1  1  3  3  5
+ 2  2  4  4  6
+
+julia> interpolate(Y, 3; dims=2)
+2×7 Matrix{Int64}:
+ 1  1  1  3  3  3  5
+ 2  2  2  4  4  4  6
+
+julia> interpolate(Y, 1) == Y
+true
+"""
+function interpolate(Y, scale; dims=1:ndims(Y), kwargs...)
+    Y = repeat(Y; inner=(d in dims ? scale : 1 for d in 1:ndims(Y)))
+
+    # Chop the last slice of repeated dimensions since we only interpolate between
+    # the values
+    return Y[(d in dims ? axis[begin:end-scale+1] : axis for (d, axis) in enumerate(axes(Y)))...]
+end
+
+# TODO add more intelligent interpolation e.g. linear
+# TODO add tests for interpolate
+# TODO do we want linear_smooth(linear_smooth(Z)) == linear_smooth(Z) ?
+# right now this linear smooth only works on arrays that came from interpolate(Y, 2)
+
+function linear_smooth(Y; dims=1:ndims(Y), kwargs...)
+    return _linear_smooth!(1.0 * Y, dims)
+    # makes a copy of Y and ensures the type can hold float like elements
+end
+
+function _linear_smooth!(Y, dims)
+    all_dims = 1:ndims(Y)
+    for d in dims
+        axis = axes(Y, d)
+        Y1 = @view Y[(i==d ? axis[begin+1:end-1] : (:) for i in all_dims)...]
[…]

### 2025-03-19 paper/paper.qmd
multi scale section

@@
 
-Then for any row $i\in[I]$, the sum of the row is $\epsilon\sqrt{J}$ close to 1
+Then for any row $i\in[I]$, the sum of the row $A[i,:]$ is $\epsilon\sqrt{J}$ close to 1
 
@@
 
-We propose a multiresolution approach inspired by multi-grid [] and wavelets [] that factorizes the data at progressively finer scales, using the .
-The can greatly speed up how fast large scale tensors can be factorized.
+We propose a multiresolution approach inspired wavelets [@benedetto_wavelets_1993] and multigrid [@trottenberg_multigrid_2001] that factorizes the data at progressively finer scales.
+This can greatly speed up how fast large scale tensors can be factorized.
 
@@
 
-A natural extension of the example shown in @sec-basic-multi-scale would be to higher dimensional distributions.
-For example, we could consider an order-$3$ tensor where the horizontal slices correspond to a $2$-dimensional discretization of a bivariate density.
+An example of this setting would be an extension of the example shown in @sec-basic-multi-scale to higher dimensional distributions.
+We could consider an order-$3$ tensor where the horizontal slices correspond to a $2$-dimensional discretization of a bivariate density.
 Entries of the input tensor $Y$ would be given by $Y[i,j,k] = f_i(x_j, y_k)$ for continuous probability density functions $f_i:\mathbb{R}^2\rightarrow \mathbb{R}_+$ for a 2D grid of points $(x_j, y_k)$.
@@
 
-Factors that de
-
-, and $J_s$ would be a discretization that uses every $2^{s}$ points, or $J_s = 2^{S-s}+1$ points in total.
-
-The basic approach will be to factorize the coarsest version of $Y$, $Y_S\in\mathbb{R}^{I\times J_S}$, to obtain $A_S$ and $B_S$.
+If we want to perform a rank-$(R_1,\dots,R_N)$ Tucker decomposition of $Y=\left\llbracket A_0; A_1, \dots, A_N\right\rrbracket$, we can initialize the factors with a very corse factorization $A_{n_m}^{s}\in\mathbb{R}^{J_{n_m}^{s} \times R_{n_m}}$ where $J^{s}$ would be a discretization that uses every $2^{s}$ points, or more accurately
+
+$$
+J_{n_m}^{s} = 2^{\max(S_m-s, 0)}+1
+$$
+
+points in total.
+We select $S_m$ so that at the finest scale $s=0$, we have $J_{n_m}^{0}=I_{n_m}$.
+The dimensions that do not represent continuous values will use the full sized factors $A_n^{s}\in\mathbb{R}^{J_n \times R_n}$ with $J_n = I_n$, and the core will remain $A_0^{s}\in\mathbb{R}^{R_1\times\dots\times R_N}$ at all scales $s$.^[This assumes the dimensions of $Y$ where $Y$ is continuous have been discretized as one more than a power of two.
+The same idea holds with some other discretization plan, but becomes more complicated to express notationally and keep track of the number of points at each scale.]
+
+The aim is to fit the product of the factors $\left\llbracket A_0^{s}; A_1^{s}, \dots, A_N^{s}\right\rrbracket$ to a lower resolution version of $Y\in\mathbb{R}^{I_1 \times \dots \times I_N}$, namely $Y^s\in\mathbb{R}^{J_1^s \times \dots \times J_N^s}$, and use the result to initialize a finer version of the factors $A$.
+The code for this looks like the following.
+
+```julia
+function multiscale_factorize(Y; kwargs...)
+    scales, kwargs = initialize_scales(Y, kwargs)
+    coarsest_scale, finer_scales... = scales
+
+    # Factorize Y at the coarsest scale
+    Yₛ = coarsen(Y, coarsest_scale; kwargs...)
+    decomposition, stats, kwargs = factorize(Yₛ; kwargs...)
+
+    # Factorize Y at progressively finer scales
+    for scale in finer_scales
+        # Use an interpolated version of the coarse factorization
+        # as the initialization
+        decomposition = interpolate(decomposition, scale; kwargs...)
+        kwargs[:decomposition] = decomposition
+
+        Yₛ = coarsen(Y, scale; kwargs...)
+        decomposition, stats, kwargs = factorize(Yₛ; kwargs...)
+    end
+    return decomposition, stats, kwargs
+end
+```
+
+Straightforward subsampled coarsening and constant interpolating can be used for `coarsen` and `interpolate`, but more sophisticated methods can be used in principle.
+Since the final solve of `factorize` is on the original sized problem, the choice of coarsening and interpolating only influences the initialization used at this finest scale.
+Bellow are examples of the basic coarsening and interpolation methods.
+
+```julia
+coarsen(Y::AbstractArray, scale::Integer; dims=1:ndims(Y), kwargs...) =
+    Y[(d in dims ? axis[begin:scale:end] : axis for (d, axis) in enumerate(axes(Y)))...]
+
+function interpolate(Y, scale; dims=1:ndims(Y), kwargs...)
+    Y = repeat(Y; inner=(d in dims ? scale : 1 for d in 1:ndims(Y)))
+
+    # Chop the last slice of repeated dimensions
+    # since we only interpolate between the values
+    return Y[(d in dims ? axis[begin:end-scale+1] : axis for (d, axis) in enumerate(axes(Y)))...]
+end
+```
 

### 2025-04-23 src/BlockTensorDecomposition.jl
Finish constrained multiscale with bug fixes to regular factorize

@@ -12,7 +12,7 @@ export ×₁, nmp, nmode_product, mtt, slicewise_dot, tuckerproduct, cpproduct
 
 # Low level types and interface
 #include("./decomposition.jl")
-export array, contractions, core, eachfactorindex, factor, factors, frozen, isfrozen, matrix_factor, matrix_factors, rankof
+export array, contractions, core, eachfactorindex, factor, factors, frozen, isfrozen, matrix_factor, matrix_factors, nfactors, rankof
 export AbstractDecomposition
 export GenericDecomposition, SingletonDecomposition
 export AbstractTucker, Tucker, Tucker1, CPDecomposition
@@ -64,7 +64,7 @@ export Projection, NNProjection, SafeNNProjection, Rescale
 
 export BlockedUpdate
 export updates
-export smart_insert!, smart_interlase!, group_by_factor
+export getconstraint, smart_insert!, smart_interlase!, group_by_factor
 
 #export block_gradient_decent, nn_block_gradient_decent, scale_nn_block_gradient_decent, proj_nn_block_gradient_decent
 

### 2025-04-23 src/Core/Core.jl

@@ -22,7 +22,7 @@ export ×₁, nmp, nmode_product, mtt, slicewise_dot, tuckerproduct, cpproduct
 
 # Low level types and interface
 include("./decomposition.jl")
-export array, contractions, core, eachfactorindex, factor, factors, frozen, isfrozen, matrix_factor, matrix_factors, rankof
+export array, contractions, core, eachfactorindex, factor, factors, frozen, isfrozen, matrix_factor, matrix_factors, nfactors, rankof
 export AbstractDecomposition
 export GenericDecomposition, SingletonDecomposition
 export AbstractTucker, Tucker, Tucker1, CPDecomposition
@@ -74,7 +74,7 @@ export Projection, NNProjection, SafeNNProjection, Rescale
 
 export BlockedUpdate
 export updates
-export smart_insert!, smart_interlase!, group_by_factor
+export getconstraint, smart_insert!, smart_interlase!, group_by_factor
 
 #export block_gradient_decent, nn_block_gradient_decent, scale_nn_block_gradient_decent, proj_nn_block_gradient_decent
 

### 2025-04-23 src/Core/blockupdates.jl

@@ -433,6 +433,10 @@ end
 
 abstract type ConstraintUpdate <: AbstractUpdate end
 
+function getconstraint(C::AbstractConstraint)
+    error("Getter for constraints of type $(typeof(C)) is not implemented.")
+end
+
 """
     ConstraintUpdate(n, constraint)
 
@@ -472,6 +476,7 @@ struct GenericConstraintUpdate <: ConstraintUpdate
 end
 
 check(U::GenericConstraintUpdate, D::AbstractDecomposition) = check(U.constraint, factor(D, U.n))
+getconstraint(U::GenericConstraintUpdate) = U.constraint
 
 function (U::GenericConstraintUpdate)(x::T; kwargs...) where T
     n = U.n
@@ -490,6 +495,7 @@ struct Projection <: ConstraintUpdate
 end
 
 check(P::Projection, D::AbstractDecomposition) = check(P.proj, factor(D, P.n))
+getconstraint(U::Projection) = U.proj
 
 function (U::Projection)(x::T; kwargs...) where T
     n = U.n
@@ -499,12 +505,28 @@ function (U::Projection)(x::T; kwargs...) where T
     U.proj(factor(x, n))
 end
 
+function Base.getproperty(U::Projection, sym::Symbol)
+    if sym === :whats_normalized
+        proj = U.proj
+        whats_normalized = try
+            proj.whats_normalized
+        catch # in the event proj is just an entrywise for example
+            identityslice
+        end
+        return whats_normalized
+    else # fallback to other fields
+        return getfield(U, sym)
+    end
+end
+
 NNProjection(n) = Projection(n, nonnegative!)
 
 struct SafeNNProjection <: ConstraintUpdate
     n::Integer
     backup::ProjectedNormalization
 end
+
+getconstraint(_::SafeNNProjection) = nonnegative! #U.backup
 
 function (U::SafeNNProjection)(x::T; kwargs...) where T
     n = U.n
@@ -526,7 +548,15 @@ function (U::SafeNNProjection)(x::T; kwargs...) where T
     check(U, x) || error("Something went wrong with SafeNNProjection using the backup projection: $(U.backup)")
 end
 
-check(S::SafeNNProjection, D::AbstractDecomposition) = check(nonnegative!, factor(D, S.n)) && all(!iszero, S.backup.whats_normalized(factor(D, S.n)))
[…]

### 2025-04-23 src/Core/factorize.jl

@@ -124,6 +124,11 @@ function default_kwargs(Y; kwargs...)
 	get!(kwargs, :model) do
 		isnothing(kwargs[:decomposition]) ? Tucker1 : typeof(kwargs[:decomposition])
 	end
+
+	# Ensure the model is valid and decomposition type matches the model (or is unspecified)
+	@assert kwargs[:model] <: AbstractDecomposition
+	@assert typeof(kwargs[:decomposition]) <: Union{Nothing, kwargs[:model]}
+
 	get!(kwargs, :rank) do # Can also be a tuple. For example, Tucker rank could be (1, 2, 3) for an order 3 array Y
 		isnothing(kwargs[:decomposition]) ? 1 : rankof(kwargs[:decomposition])
 	end
@@ -295,7 +300,7 @@ function make_update!(decomposition, Y; momentum, constraints, constrain_init, g
 		if !all(C -> check(C, decomposition), expanded_constraints!)
 			indexes = findall(C -> !check(C, decomposition), expanded_constraints!)
 			if constrain_init
-				error("decomposition failed to be constrained. Check the constraint(s) $(expanded_constraints![indexes]) operation or the checking function")
+				error("decomposition failed to be constrained. Check the constraint(s) $(expanded_constraints![indexes]) operation or the checking function. The full list of constraints is $(expanded_constraints!).")
 			else
 				@warn "Initial decomposition does not satisfy the following constraints: $(expanded_constraints![indexes]). This may be ok if later iterations satisfy the constraints"
 			end

### 2025-04-23 src/Core/multiscale.jl

@@ -147,7 +147,13 @@ function multiscale_factorize(Y; kwargs...)
     # Factorize Y at the coarsest scale
     Yₛ = coarsen(Y, coarsest_scale; dims=continuous_dims, kwargs...)
 
-    constraints, kwargs = scaled_constraints(Yₛ, coarsest_scale; kwargs...)
+    get!(kwargs, :constraints, nothing)
+
+    original_constraints = kwargs[:constraints]
+
+    constraints, kwargs = scale_constraints(Yₛ, coarsest_scale; continuous_dims, kwargs...)
+
+    @info "Factorizing at scale $coarsest_scale..."
     decomposition, stats, _ = factorize(Yₛ; kwargs...)
 
     # Factorize Y at progressively finer scales
@@ -160,8 +166,11 @@ function multiscale_factorize(Y; kwargs...)
 
         Yₛ = coarsen(Y, scale; dims=continuous_dims, kwargs...)
 
-        constraints, kwargs = scale_constraints(Yₛ; kwargs...)
+        # Reset constraints so they can be properly scaled
+        kwargs[:constraints] = original_constraints
+        constraints, kwargs = scale_constraints(Yₛ, scale; continuous_dims, kwargs...)
 
+        @info "Factorizing at scale $scale..."
         decomposition, stats, _ = factorize(Yₛ; kwargs...)
     end
     return decomposition, stats, kwargs
@@ -174,20 +183,24 @@ const IMPLEMENTED_DECOMPOSITION_CONSTRAINT_SCALING = [
 ]
 
 """
-    scaled_constraints(Y, scale; kwargs...)
+    scale_constraints(Y, scale; kwargs...)
 
 Scales any constraints that need to be modified to use at a coarser scale.
 """
-function scaled_constraints(Y, scale; kwargs...)
-    continuous_dims = kwargs[:continuous_dims]
-    S = log2(scale) # TODO Don't assume the scale is a power of 2
+function scale_constraints(Y, scale; continuous_dims, kwargs...)
+
+    kwargs = Dict{Symbol,Any}(kwargs)
+
+    kwargs[:continuous_dims] = continuous_dims
+    S = log2(scale) + 1 # TODO Don't assume the scale is a power of 2
+    # we want to map scales -> S where 1 -> 1, 2 -> 2, 4 -> 3, 8 -> 4, 16 -> 5 etc.
 
     decomposition, constraints = expand_decomposition_constraints(Y, kwargs)
 
[…]

### 2025-04-23 paper/paper.qmd
finish first pass of paper

@@
 
+Basic linear constrains `AX=B` are also implemented in the following manner.
+This constraint projects a factor `X` onto the affine space defined by `AX=B` for a linear operator or matrix `A` and bias `B`.
+
+```julia
+struct LinearConstraint{T <: Union{Function, AbstractArray}} <: AbstractConstraint
+    linear_operator::T
+    bias::AbstractArray
+end
+
+check(C::LinearConstraint{Function}, X::AbstractArray) = C.linear_operator(X) ≈ C.bias
+check(C::LinearConstraint{<:AbstractArray}, X::AbstractArray) = C.linear_operator * X ≈ C.bias
+
+# TODO implement linear constraint given an operator
+function (C::LinearConstraint{Function})(X::AbstractArray)
+    error("Linear Constraints defined in terms of an operator are not implemented (YET!)")
+end
+
+function (C::LinearConstraint{<:AbstractArray})(X::AbstractArray)
+    error("Linear Constraints defined in terms of a general array are not implemented (YET!)")
+end
+
+function (C::LinearConstraint{<:AbstractMatrix})(X::AbstractArray)
+    A = C.linear_operator
+    b = C.bias
+    X .-= A' * ( (A*A') \ (A*X .- b) ) # Projects X onto the subspace AX=b
+end
+```
+
 Constraints can also be composed.
@@
 function multiscale_factorize(Y; kwargs...)
-    scales, kwargs = initialize_scales(Y, kwargs)
+    continuous_dims, kwargs = initialize_continuous_dims(Y; kwargs...)
+    scales, kwargs = initialize_scales(Y; kwargs...)
     coarsest_scale, finer_scales... = scales
@@
     # Factorize Y at the coarsest scale
-    Yₛ = coarsen(Y, coarsest_scale; kwargs...)
-    decomposition, stats, kwargs = factorize(Yₛ; kwargs...)
+    Yₛ = coarsen(Y, coarsest_scale; dims=continuous_dims, kwargs...)
+
+    constraints, kwargs = scale_constraints(Yₛ, coarsest_scale; kwargs...)
+    decomposition, stats, _ = factorize(Yₛ; kwargs...)
 
@@
         # Use an interpolated version of the coarse factorization
-        # as the initialization
-        decomposition = interpolate(decomposition, scale; kwargs...)
+        # as the initialization.
+        decomposition = interpolate(decomposition, 2; dims=continuous_dims, kwargs...)
         kwargs[:decomposition] = decomposition
 
-        Yₛ = coarsen(Y, scale; kwargs...)
-        decomposition, stats, kwargs = factorize(Yₛ; kwargs...)
+        Yₛ = coarsen(Y, scale; dims=continuous_dims, kwargs...)
+
+        constraints, kwargs = scale_constraints(Yₛ, scale; kwargs...)
+
+        decomposition, stats, _ = factorize(Yₛ; kwargs...)
     end
@@
 ```
+
+### Coarsening and Interpolating
 
@@
 
+To apply a linear interpolation, we can first perform the constant interpolation and smooth out the result.
+The following code averages neighbouring values along the continuous dimensions specified.
+Since this is applied after the constant interpolation, every even indexed value becomes the average of its neighbours, and every odd indexed value remains fixed.
+
+```julia
+function linear_smooth!(Y, dims)
+    all_dims = 1:ndims(Y)
+    for d in dims
+        axis = axes(Y, d)
+        Y1 = @view Y[(i==d ? axis[begin+1:end-1] : (:) for i in all_dims)...]
+        Y2 = @view Y[(i==d ? axis[begin+2:end] : (:) for i in all_dims)...]
+
+        @. Y1 = 0.5 * (Y1 + Y2)
+    end
+    return Y
+end
+```
+
+When interpolating an array that is an `AbstractDecomposition`, we can interpolate the factors directly instead of the combined array.
+
+```julia
+function interpolate(CPD::CPDecomposition, scale; dims=1:ndims(CPD), kwargs...)
+    interpolated_matrix_factors = (d in dims ? interpolate(A, scale; dims=1, kwargs...) : A for (d, A) in enumerate(matrix_factors(CPD)))
+    return CPDecomposition(Tuple(interpolated_matrix_factors))
+end
+
+function interpolate(T::Tucker1, scale; dims=1:ndims(T), kwargs...)
+    core_dims = setdiff(dims, 1) # Want all dimensions except possibly the first
+    interpolated_core = interpolate(core(T), scale; dims=core_dims, kwargs...)
+
+    matrix = matrix_factor(T, 1)
+
+    interpolated_matrix = 1 in dims ? interpolate(matrix, scale; dims=1, kwargs) : matrix
+    return Tucker1((interpolated_core, interpolated_matrix))
+end
+
+function interpolate(T::Tucker, scale; dims=1:ndims(T), kwargs...)
+    interpolated_matrix_factors = (d in dims ? interpolate(A, scale; dims=1, kwargs...) : A for (d, A) in enumerate(matrix_factors(T)))
+    # Core is not interpolated
+    return Tucker(Tuple(core(T), interpolated_matrix_factors...))
+end
+```
+
+## Constraints with Multi-scale
+
+Some constraints like `Entrywise` constrains can be used as-is at any scale, but other constraints like normalizations and linear constraints require some more thought.
+The main strategy is to modify constraints along continuous dimensions.
+Given a constraint on a factor, and the number of continuous dimensions it constrains, we can construct a modified constraint by scaling the full sized constraint appropriately.
+
+For a $p$-norm constraint where some part of a factor $X_i$ needs to be normalized to $\lVert X_i \rVert_p=C$, we require the coarsened parts of the factor $\overline{X}_i$ to have norm $\lVert X_i \rVert_p= (C/s)^{1/p}$ where $s$ is the scale of the coarsening.
+For example, a scale of $3$ would mean we only take every $3$ entries of $X_i$.
+
+For linear constraints `AX=B`, we construct a new constraint that coarsens $A$, and scales the bias $B$ by the scale.
+For example, if $x$ is a vector that represents a continuous function that is constrained to the affine space $Ax=b$, we treat the rows of the matrix $A$ as also being continuous functions that are inner product-ed with $x$ and can be coarsened in a similar manner to $x$.
+This means a `LinearConstraint(A, b)` gets scaled to the constraint `LinearConstraint(A[:, begin:scale:end], b ./ scale)`.
+
+In all the constraints, the scale will need to be applied in each continuous dimension, so we implement this as `scale^n_continuous_dims`. The full implementation of `scale_constraint` is shown below.
+
+```julia
+function scale_constraint(constraint::AbstractConstraint, scale, n_continuous_dims)
+    @warn "Unsure how to scale constraints of type $(typeof(constraint)). Leaving the constraint $constraint alone."
+    return constraint
+end
+
+# Constraints that are fixed like entrywise constraints do not need to be scaled
+function scale_constraint(constraint::Union{FIXED_CONSTRAINTS...}, scale, n_continuous_dims)
+    return constraint
+end
+
+# TODO handle LinearConstraint{<:AbstractArray} or LinearConstraint{Function}
+function scale_constraint(constraint::LinearConstraint{<:AbstractMatrix}, scale, n_continuous_dims)
+    A = constraint.linear_operator
+    b = constraint.bias
+    return LinearConstraint(A[:, begin:scale:end], b ./ scale^n_continuous_dims)
+end
+
+function scale_constraint(constraint::ScaledNormalization{<:Union{Real,AbstractArray{<:Real}}}, scale, n_continuous_dims)
+    norm = constraint.norm
+    F = constraint.whats_normalized
+    S = constraint.scale
+    return ScaledNormalization(norm, F, S ./ scale^n_continuous_dims)
+end
+
+function scale_constraint(constraint::ScaledNormalization{<:Function}, scale, n_continuous_dims)
+    norm = constraint.norm
+    F = constraint.whats_normalized
+    S = constraint.scale
+    return ScaledNormalization(norm, F, (x -> x ./ scale^n_continuous_dims) ∘ S)
+end
+
+# TODO scale a projected normalization
+function scale_constraint(constraint::ProjectedNormalization, scale, n_continuous_dims)
+    @warn "Scaling ProjectedNormalization constraints is not implemented (YET!) Leaving the constraint $constraint alone."
+    return constraint
+end
+```
+
+We justify scaling constraints in this way with the following propositions and observations.
+
+::: {#thm-rem-constraint-assumptions}
+
+### Constraint Proposition Assumptions
+
+In the following proposition, we assume $x\in\mathbb{R}^I$ represents an evenly-spaced discretization of an $L_f$-Lipschitz function $f:[l, u]\to\mathbb{R}$ on a finite interval $t\in[l, u]$ with entries
+
+$$
+x[i]=f(t[i]),
+$$
+
+where
+
+$$
+t[i] = l + (i-1)\Delta t= l + (i-1) \frac{u-l}{I-1}.
+$$
+
+We will use $\overline{x}\in\mathbb{R}^{\lfloor (I+1)/2 \rfloor}$ to represent the subvector made by removing every other entry of $x\in\mathbb{R}^I$ where
+
+$$
+\overline{x} = (x[1], x[3], x[5], \dots, x[\lfloor I \rfloor_{o}]).
+$$
+
+We use the shorthand
+
+$$
+\lfloor I \rfloor_{o} = 2\left\lfloor\frac{I+1}{2}\right\rfloor-1
+$$
+
+to round $I$ down to the nearest odd integer.
+
+:::
+
+::: {#prp-linear-constraint-scaling}
+
+### Linear Constraint Scaling
+
+Let the assumption in @thm-rem-constraint-assumptions hold. Let $a\in\mathbb{R}^I$ be a discretization of an $L_g$-Lipschitz function $g:[l,u]\to\mathbb{R}$.
+
+If $\langle a, x \rangle = b$, then for even $I$,
+
+$$
+\left\lvert \left\langle \overline{a}, \overline{x} \right\rangle - b/2 \right\rvert \leq \max(\lVert f \rVert_{\infty},\lVert g \rVert_{\infty}) \frac{I}{(I-1)} \frac{(L_{f} + L_{g})(u-l)}{4}
+$$
+
+and for odd $I$,
+
+TODO check formatting
+$$
+\begin{align}
+\left\lvert 2\left\lVert \overline{x} \right\rVert _1 - \frac{I+1}{I} \lVert x \rVert_{1} \right\rvert &= \left\lvert 2\left\lVert \overline{x} \right\rVert _1 - \lVert x \rVert_{1} - \frac{\lVert x \rVert_{1}}{I} \right\rvert \\
+&=\left\lvert 2\sum_{i\ \text{odd}} \left\lvert x_{i} \right\rvert - \sum_{i=1}^I \left\lvert x_{i} \right\rvert  - \frac{\lVert x \rVert_{1}}{I} \right\rvert  \\
+& = \left\lvert  \sum_{i\ \text{odd}}\left\lvert x_{i} \right\rvert+ \sum_{i\ \text{odd}}\left\lvert x_{i} \right\rvert - \left( \sum_{i\ \text{odd}}\left\lvert x_{i} \right\rvert + \sum_{i\ \text{even}} \left\lvert x_{i} \right\rvert  \right) - \frac{\lVert x \rVert_{1}}{I}  \right\rvert \\
+& = \left\lvert  \sum_{i\ \text{odd}}\left\lvert x_{i} \right\rvert - \sum_{i\ \text{even}} \left\lvert x_{i} \right\rvert  - \frac{\lVert x \rVert_{1}}{I}  \right\rvert \\
+&=  \left\lvert  \sum_{j=1}^{(I-1)/2}\left\lvert x_{2j-1} \right\rvert + \left\lvert x_{I} \right\rvert - \sum_{j=1}^{(I-1)/2} \left\lvert x_{2j} \right\rvert  - \frac{\lVert x \rVert_{1}}{I}  \right\rvert \\
+&\leq \sum_{j=1}^{(I-1)/2}\left\lvert \left\lvert x_{2j-1} \right\rvert-\left\lvert x_{2j} \right\rvert \right\rvert + \left\lvert \left\lvert x_{I} \right\rvert - \frac{\lVert x \rVert_{1}}{I} \right\rvert \\
+&\leq \sum_{j=1}^{(I-1)/2}\left\lvert  x_{2j-1} - x_{2j}  \right\rvert + \frac{1}{I} \left\lvert I \left\lvert x_{I} \right\rvert - \lVert x \rVert_{1} \right\rvert \\
+&\leq \sum_{j=1}^{(I-1)/2} C + \frac{1}{I} \left\lvert \sum_{i=1}^I \left( \left\lvert x_{I} \right\rvert - \lvert x_{i} \rvert  \right)\right\rvert\\
+&\leq C \frac{I-1}{2} + \frac{1}{I} \sum_{i=1}^I \left\lvert  \left\lvert x_{I} \right\rvert - \lvert x_{i} \rvert  \right\rvert\\
+&\leq C \frac{I-1}{2} + \frac{1}{I} \sum_{i=1}^I \left\lvert   x_{I}  -  x_{i}  \right\rvert\\
+&\leq C \frac{I-1}{2} + \frac{1}{I} \sum_{i=1}^I C (I-i)\ \text{(Apply Lipschitz recursively)} \\
+& = C \frac{I-1}{2} + \frac{1}{I}\frac{CI(I-1)}{2}\\
+&= \frac{L (u-l)}{2} + \frac{L (u-l)}{2} \\
+&= L (u-l).
+\end{align}
+$$
+
+::: {.proof}
+See @sec-linear-constraint-scaling-proof.
+:::
+
+:::
+
+@prp-linear-constraint-scaling can be extended to a bound on the $L_1$ distance between $\overline{A}\overline{x}$ and $b/2$ where $Ax=b$ and $\overline{A}=A[:,begin:2:end]$ is the submatrix with every other column removed.
+TODO add corollary.
+
+We also have a similar proposition for a $1$-norm constraint.
+
+::: {#prp-l1-norm-constraint-scaling}
+
+### $L_1$-norm Constraint Scaling
+
+Let the assumption in @thm-rem-constraint-assumptions hold.
+
+If $\lVert x \rVert_1 = b$, then for even $I$,
+
+$$
+\left\lvert \left\lVert \overline{x} \right\rVert_1 - b/2 \right\rvert \leq \frac{I}{(I-1)}\frac{L (u-l)}{4}
+$$
+
+and for odd $I$,
+
+$$
+\left\lvert \left\lVert \overline{x} \right\rVert_1 - \frac{I+1}{I} \frac{b}{2} \right\rvert \leq \frac{L(u-l)}{2}.
+$$
+
+::: {.proof}
+See @sec-l1-norm-constraint-scaling-proof.
+:::
+
+:::
+
+Importantly, for normalized vectors $z = x / \lVert x \rVert_1$, the error bound in @prp-l1-norm-constraint-scaling (for even $I$) becomes
+
+$$
+\left\lvert \left\lVert \overline{z} \right\rVert_1 - 1/2 \right\rvert \leq \frac{I}{(I-1)}\frac{L (u-l)}{4b}.
+$$
+
+As the number of points get large, $I\to\infty$, a finer discretization will have more points and have a $1$-norm becoming unbounded $b\to\infty$.
+This means the error bound goes to zero and $\left\lVert\overline{z} \right\rVert_1 \to 1/2$.^[This also holds for odd $I$.]
+
 ## Convergence of a Multi-scale Method
@@
 
+As implemented, both the regular and multi-scale approaches execute many iterations at the finest scale.
+The multi-scale method uses its coarsened iterations to warm start the fine iteration.
+So we should expect the multi-scale method to converge.
+We make this explicit with the following lemmas and theorems.
+The first subsection (@sec-analysis-lemmas) shows general functional and convex analysis facts about Lipschitz, smooth, and strongly convex functions, and what happens when you linearly interpolate Lipschitz functions.
+The next subsection (@sec-resolved-multiscale) shows convergence for the multi-scale method when we resolve the problem along the entire new discretization at each scale.
+And the final subsection (@sec-freezed-multiscale) looks at what happens when you freeze previously computed points so you only need to solve the problem at the interpolated points for each scale.
+
+### Definitions
+
+::: {#def-smooth-function}
+
+A differentiable function $f:\mathbb{R}^I\to\mathbb{R}$ is $S$-smooth when $\nabla f$ is $S$-Lipschitz,
+
+$$
+\lVert \nabla f(x) - \nabla f(y) \rVert_2 \leq S \lVert x - y \rVert_2.
+$$
+
+:::
+
+::: {#def-strongly-convex-function}
+
+A function $f:\mathbb{R}^I\to \mathbb{R}$ is $\mu$-strongly convex when $g(x) = f(x) - \frac{\mu}{2}\lVert x\rVert_2^2$ is convex.
+
+When $f$ is differentiable, $f$ is $\mu$-strongly convex when
+
+$$
+f(y)\geq f(x)+\left\langle \nabla f(x),y-x  \right\rangle + \frac{\mu}{2}\left\lVert x-y \right\rVert_2^2
+$$
+
+for all $x,y\in \mathbb{R}^I$.
+
+:::
+
+::: {#def-projected-gradient-decent}
+One step of projected gradient descent for an $S$-smooth and $\mu$-strongly convex function $\mathcal{L}:\mathbb{R}^I\to\mathbb{R}$, a constraint set $\mathcal{C}\subseteq V$, at an iterate $x^k$ is the update,
+
+$$
+x^{k+1} \leftarrow P_{\mathcal{C}}\left( x^k - \frac{1}{S}\nabla\mathcal{L}(x^k) \right)
+$$
+
+where $P_{\mathcal{C}}:\mathbb{R}^I\to\mathcal{C}$ is the (Euclidean) projection operator onto $\mathcal{C}$.
+
+:::
+
+::: {#prp-descent-lemma}
+
+### Descent Lemma
+
+In the projected gradient descent setting with an $S$-smooth and $\mu$-strongly convex function (@def-projected-gradient-decent), we can bound the distance to the unique minimizer $x^*\in\mathcal{C}$ in terms of the initial point $x^0\in\mathbb{R}^I$,
+
+$$
+\left\lVert x^{t}-x^* \right\rVert _2 \leq \left( 1-c \right)^t \left\lVert x^{0} -x^* \right\rVert _2
+$$
+
+with *condition number* $c=\mu/S$.
+
+:::
+
+TODO cite this.
+
+The vector space $\mathbb{R}^I$ in @def-strongly-convex-function, @def-smooth-function, and @def-projected-gradient-decent can be extended to tensor spaces $\mathbb{R}^{I_1 \times \dots \times I_N}$ with the Frobenius inner product $\langle X, Y\rangle _F$ and norm $\lVert \cdot \rVert _F$ more generally.
+
+### Functional Analysis Lemmas {#sec-analysis-lemmas}
+
+Before we can analyze the convergence of a multi-scaled method, we first need a handle on how linear interpolations play with discretized Lipschitz functions.
+To simplify the analysis of multi-scaled methods, we look at solving the general problem
+
+$$
+\min_{x \in \mathcal{C}} \mathcal{L} (x) := \sum_{i\in [I]} \ell_i(x[i])
+$$
+
+using projected gradient descent
+
+$$
+x \leftarrow P_\mathcal{C}\left(x - \frac{1}{L}\nabla \mathcal{L} (x)\right)
+$$
+
+where the samples $x[i]$ come from some $L_f$-Lipschitz function $f:\mathbb{R}\to\mathbb{R}$
+
+$$
+x[i] = f(t_i)
+$$
+
+on an interval $t\in [a,b]$ that has been evenly discretized $t_{i+1} - t_i=\Delta t$ for all $i\in[I]$.
+
+We assume the loss functions $\ell_i$ over each entry $x[i]$ are $S$-smooth and $\mu$-strongly convex. A natural example of such a function $\mathcal{L}$ would be a least-squares loss
+
+$$
+\mathcal{L}(x) = \frac{1}{2} \left\lVert x - y \right\rVert _2^2 = \sum_{i\in[I]} \frac{1}{2} (x[i] - y[i])^2 = \sum_{i\in [I]} \ell_i (x[i]).
+$$
+
+The following lemmas explain that the smoothness and strong convexity of the functions $\ell_i$ carry over to the full loss $\mathcal{L}$.
+
+::: {#lem-smooth-component-functions}
+
+If $\ell_i:\mathbb{R}\to \mathbb{R}$ are each $S$-smooth, then $\mathcal{L}:\mathbb{R}^I\to \mathbb{R}$ is $S$-smooth.
+
+::: {.proof}
+First note that $(\nabla \mathcal{L}(x))_{i}= \frac{\partial}{\partial x[i]}\mathcal{L}(x)=\frac{\partial}{\partial x[i]}\sum_{j=1}^I\ell_i(x[j])=\sum_{j=1}^I\frac{\partial}{\partial x[i]}\ell_i(x[j])=\frac{\partial}{\partial x[i]}\ell_i(x[i])=\ell_i^'(x[i])$ since $\mathcal{L}$ is separable in each coordinate.
+This gives us,
+$$
+\begin{align}
+\left\lVert \nabla \mathcal{L}(x)-\nabla \mathcal{L}(y) \right\rVert_{2}^2 &= \sum_{i=1}^I(\nabla \mathcal{L}(x)-\nabla \mathcal{L}(y) )[i]^2 \\
+&=\sum_{i=1}^I \left( \ell_i^'(x[i])-\ell_i^'(x[i]) \right)^2 \\
+&\leq \sum_{i=1}^I S^2\left( x[i]-y[i] \right)^2 \tag{$\ell$ is $S$ smooth} \\
+&=S^2\left\lVert x-y \right\rVert_{2}^2.
+\end{align}
+$$
+
+Taking square roots completes the proof.
+
+:::
+
+:::
+
+::: {#lem-strongly-convex-component-functions}
+
+If $\ell_i:\mathbb{R}\to \mathbb{R}$ are all $m$ strongly convex, then $\mathcal{L}:\mathbb{R}^I\to \mathbb{R}$ is $m$ strongly convex (under the $2$-norm in $\mathbb{R}^I$).
+
+::: {.proof}
+
+Consider
+$$
+\begin{align}
+\mathcal{L}(x)-\frac{m}{2}\lVert x \rVert_{2}^2 &= \sum_{i=1}^I\ell_i(x[i]) - \frac{m}{2}\sum_{i=1}^I x[i]^2 \\
+&= \sum_{i=1}^I\left( \ell_i(x[i])- \frac{m}{2}x[i]^2 \right).
+\end{align}
+$$
+
+The function $g(x[i])=\ell_i(x[i])- \frac{m}{2}x[i]^2$ is convex because each function $\ell_i(x[i])$ is $m$ strongly convex.
+The sum of convex functions is also convex, so $\mathcal{L}(x)-\frac{m}{2}\lVert x \rVert_{2}^2$ is convex.
+
+:::
+
+:::
+
+These conditions can be relaxed to include larger sets of function like quasi-strongly convex functions [@necoara_linear_2019], but the general proof approach remains the same.
+We do require the separability of $\mathcal{L}$ so it can make sense to solve the problem over coarser discretizations.
+
+Now we discuss how Lipschitz functions play with discretizations and linear interpolations.
+
+::: {#lem-lipschitz-interpolation}
+
+### Lipschitz Function Interpolation
+Let $f:\mathbb{R}^n \to \mathbb{R}$, $a,b\in \mathbb{R}^n$, $t\in[0,1]$.
+Suppose $f$ is $L$-Lipshtiz.
+Then the error between the function at a point on the line segment between $a$ and $b$, and the linear interpolation is
+$$
+\left\lvert (tf(a)+(1-t)f(b)) - f(ta+(1-t)b) \right\rvert\leq 2Lt(1-t)\left\lVert a-b \right\rVert _2.
+$$
+
+:::
+
+Using the bound on the linear interpolation of a Lipschitz function repeatedly, for centre point interpolation ($t=1/2$ in @lem-lipschitz-interpolation), we can bound the error between an exact discretization of a function at a fine scale ($Y[j] = f(X[j])$) and a linear interpolation ($\hat{Y}$) coming from a coarser discretization ($y[k]=f(x[k])$).
+
+::: {#lem-exact-interpolation}
+
+### Exact Interpolation
+
+Given $K$ uniformly space points $x[k]$ on $[a,b]$, (nearly) double them to get $J=2K-1$ uniformly spaced points $X[j]$ on $[a,b]$.
+Let $y[k]=f(x[k])$ for some $L$-Lipshitz function $f:\mathbb{R}\to \mathbb{R}$, and linearly interpolate the function values
+
+$$
+\hat{Y}[j]=\begin{cases}
+y[\frac{j+1}{2}]   & \text{if $j$ is odd}\\
+\frac{1}{2}\left( y[\frac{j}{2}] + y[\frac{j}{2}+1] \right) & \text{if $j$ is even}
+\end{cases}\ ,
+$$
+
+where the true values are given by $Y[j]=f(X[j])$. Then the difference between the interpolated $\hat{Y}$ and exact values $Y$ is bounded by
+$$
+\left\lVert \hat{Y}-Y \right\rVert_{2}\leq \frac{L}{2\sqrt{ K-1 }}\left\lVert a-b \right\rVert_{2}.
+$$
+
+:::
+
+We also have an inexact version when we interpolate not from an exact coarse discretization ($y[k]=f(x[k])$), but from an approximate coarse discretization ($\tilde{y}[k] = f(x[k]) + \delta_k$).
+
+::: {#lem-inexact-interpolation}
+
+### Inexact Interpolation
+
+Given a linear interpolation $\hat{\tilde{Y}}$ of an inexact discretization $\tilde{y}[k] = f(x[k]) + \delta_k$ of a function $f$ as described in @lem-exact-interpolation, where the interpolation is defined as
+
+$$
+\begin{align}
+  \hat{\tilde{Y}}[j]&=\begin{cases}
+  \tilde{y}[\frac{j+1}{2}]   & \text{if $j$ is odd}\\
+  \frac{1}{2}\left( \tilde{y}[\frac{j}{2}] + \tilde{y}[\frac{j}{2}+1] \right) & \text{if $j$ is even}
+  \end{cases},
+\end{align}
+$$
+
+we have the error bound between the interpolated inexact discretization $\hat{\tilde{Y}}$ and the exact discretization of the function $Y$,
+
+$$
+\begin{align}
+\left\lVert \hat{\tilde{Y}} - Y \right\rVert_2 &\leq \left\lVert \hat{\tilde{Y}} - \hat{Y} \right\rVert_2 + \left\lVert \hat{Y}-Y \right\rVert_2 \\
+&\leq \sqrt{ 2 } \left\lVert \tilde{y}-y \right\rVert + \frac{L}{2\sqrt{ K-1 }}\left\lVert a-b \right\rVert_{2}.
+\end{align}
+$$
+
+::: {.proof}
+
+See @sec-inexact-interpolation-proof.
+
+:::
+
+:::
+
+This makes sense in the following way: the error in our interpolation of approximate points, is bounded by two errors.
+The first comes from the fact that we interpolated using approximated values $\tilde{y}$ in place of the true values $y$, and the second comes from using a linear interpolation $\hat{Y}$ in place of the exact values $Y$.
+
+### Re-solved Multi-scale {#sec-resolved-multiscale}
+
+LEFT OFF HERE
+
+We use the following general approach to show convergence.
+
+Given some initial guess $x_S^0$ at the coarsest scale, we use the decent lemma (@prp-descent-lemma) to bound the error between our iterate $x_S^{K_S}$ after $K_S$ iterations, and the solution $x_S^*$ at the scale $S$.
+We can linearly interpolate our point $\hat{x}_S^{K_S}$ and use it to initialize another round of projected gradient descent $x_{S-1}^0=\hat{x}_S^{K_S}$ at the slightly finer scale.
+We can link the error between our iterate $x_S^{K_S}$ and the solution $x_S^*$ at  scale $S$ with the initial error between an iterate at the finer scale $x_{S-1}^0$ and the solution $x_{S-1}^*$ at this scale using the inexact interpolation lemma (@lem-inexact-interpolation).
+We perform $K_s$ iterations at each scale $s=S,S-1,\dots,2,1$ to get an error bound between our iterate at the finest scale $x_1^K_1$ the solution at this scale $x_1^*$ in terms of the initial error at the coarsest scale $x_S^0$.
+
+We present the main descent theorem here, and leave the rest of the details in the appendix.
+
+::: {#thm-resolved-multiscale-descent-error}
+
+### Re-solved Multi-scale Descent Error
+
+Let $c$ be the condition number of the loss function $\mathcal{L}$, $L_f$ be the Lipschitz constant for the underlying continuous function $f$, and $S$ be the coarsest scale.
+For the setting described in @sec-analysis-lemmas, we have the following error bound on the fixed multi-scaled method.
+$$
+\begin{align}
+&\left\lVert x^0_{1} - x^*_{1} \right\rVert _2 \\
+&\leq \sqrt{ 2 ^{S-1}}\left(1-c\right)^{\sum_{s=1}^{S} K_{s}}\left\lVert x_{S}^{0} -x_{S}^* \right\rVert _2 + \frac{L_f\left\lvert a-b \right\rvert}{2\sqrt{2^{S+1}}}\sum_{s=1}^{S-1}  2^{s} \left(1-c\right)^{\sum_{t=1}^sK_{t}}
+\end{align}
+$$
+
+::: {.proof}
+
+See @sec-resolved-multiscale-descent-error-proof
+
+:::
+
+:::
+
+### Freezed Multi-scale {#sec-freezed-multiscale}
+
+We take a nearly identical approach to the re-solved multi-scale method as described in @sec-resolved-multiscale, but we instead freeze the entries of our interpolation $\hat{x}_s^{K_s}$ that correspond to the same points as the prior iteration $x_s^{K_s}$.
+The projected gradient update will only act on the interpolated and unfrozen values which we will call $x_{(s)}^k$.
+
+TODO can I just introduce this notation in the proofs so that the main paper is cleaner?
+
+To be clear let us look at an example where the finest scale has $2^3+1=9$ points.
+After performing $K_3$ iterations at the coarsest scale $S=3$ with a total of three points, and slightly finer scale $S=2$ with five points (but only two unfrozen points), the full vector at the finest scale would be
+
+$$
+\begin{align}
+& x_{1}^k = \\
+& \left( {{x}_{(3)}^{K_{3}}}[1],{ x_{(1)}^{k}}[1],{ x_{(2)}^{K_{2}}}[1], { x_{(1)}^{k}}[2], {{x}_{(3)}^{K_{3}}}[2], { x_{(1)}^{k}}[3], { x_{(2)}^{K_{2}}}[2], { x_{(1)}^{k}}[4], {{x}_{(3)}^{K_{3}}}[3] \right)
+\end{align}
+$$
+
+where the vector of free variables is
+
+$$
+\begin{align}
+& x_{(1)}^k = \left( { x_{(1)}^{k}}[1], { x_{(1)}^{k}}[2], { x_{(1)}^{k}}[3], { x_{(1)}^{k}}[4]\right).
+\end{align}
+$$
+
+**Summary of notation**
+
+- $\hat{x}$ is some approximation of $x$
+- $x_{(s)}$ is a vector of just the free variables at scale $s$
+- $x_{s}$ is a vector of the free and fixed variables
+- $x^k$ is the $k$th iteration
+- $K_{s}$ is the number of iterations performed at scale $s$
+- $e=\hat{x}-x^*$ is the error
+- have $e_{(s)},e_{s},e^k$ similarly.
+	- Example, $e_{1}^2=x_{1}^2-x_{1}^*$ is the error between our iterates at the finest scale $s=1$ and the true values $x_{1}^*$ after $2$ iterations
+
+We present the main descent theorem here, and leave the rest of the details in the appendix.
+
+::: {#thm-freezed-multiscale-descent-error}
+
+### Freezed Multi-scale Descent Error
+
+Let $c$ be the condition number of the loss function $\mathcal{L}$. For the setting described in @sec-analysis-lemmas, we have the following error bound on the freezed multi-scaled method.
+
+$$
+\begin{align}
+\left\lVert e_{1}^{K_{1}} \right\rVert &\leq \left\lVert e_{S}^{0}  \right\rVert (1-c)^{K_{S}}\prod_{s=1}^{S-1}(1+(1-c)^{K_{s}}) \\
+&\qquad+\ \frac{L_f}{2}\left\lvert b-a \right\rvert\sum_{s=1}^{S-1} \frac{1}{\sqrt{ 2^{S-s} }}(1-c)^{K_{s}}\prod_{j=1}^{s-1}\left( 1+(1-c)^{K_{j}} \right).
+\end{align}
+$$
+
+If we use the same number of iterations $K_s=K$ at each scale, this reduces to the closed form upper bound
+$$
+\begin{align}
+\left\lVert e_{1}^{K_{1}} \right\rVert &\leq \left\lVert e_{S}^{0}  \right\rVert d(K)(1+d(K))^{S-1}  \\
+&\qquad+\ \frac{L_f}{2}\left\lvert b-a \right\rvert\frac{d(K) \left( \sqrt{2^S} \left( d(K) + 1 \right)^S - \sqrt{2} \left( d(K) + 1 \right) \right)}{\sqrt{2^S} \left( d(K) + 1 \right) \left( \sqrt{2} d(K) + \sqrt{2} - 1 \right)}
+\end{align}
+$$
+
+where $d(K) = (1-c)^K$ is a small number between $0 < d(K) < 1$.
+
+::: {.proof}
+
+See @sec-freezed-multiscale-descent-error-proof.
+
+:::
+
+:::
+
+We summarize the convergence with @cor-multiscale-convergence that sends the number of iterations to infinity.
+
+::: {#cor-multiscale-convergence}
+
+### Multiscale Convergence
+
+As the total number of iterations grows $\sum_{s=1}^S K_s \rightarrow \infty$ (in the case of re-solve multi-scale in @thm-resolved-multiscale-descent-error) or the number of iterations at each scale $K_s\to \infty$ (in the case of freezed multi-scale in @thm-freezed-multiscale-descent-error), the final error goes to $0$, $\lVert e_1^{K_1} \rVert_2 \rightarrow 0$, and we converge to the solution $x_1^{K_1} \rightarrow x_1^*$ at the finest scale.
+
+:::
+
+## Comparison With Projected Gradient Descent
+
+From @prp-descent-lemma, projected gradient descent with a generic initialization on the fine grid gives us the following iterate convergence
+
+$$
+\left\lVert x_{1}^{K}-x_{1}^* \right\rVert\leq \left( 1-c \right)^{K} \left\lVert x_{1}^{0} -x_{1}^* \right\rVert.
+$$
+
+We can use this in combination with an expected initial error to get an expected number of iterations needed until we have converged to a desired tolerance.
+This is make precise by @thm-expected-pgd-convergence.
+
+::: {#thm-expected-pgd-convergence}
+
+### Expected Projected Gradience Descent Convergence
+
+Assume the problem is either scaled or shifted so that the solution is normalized $\left\lVert x_{1}^* \right\rVert _2 = 1$ or centred $\left\lVert x_{1}^* \right\rVert _2 = 0$ where $x_{1}^*\in\mathbb{R}^{I}$.
+Choose an initialization $x_{1}^0 \in \mathbb{R}^I$ with i.i.d.
+standard normal entries $(x^0_{1})_{i}\sim \mathcal{N}(0,1)$.
+Then the expected initial error is
+$$
+\mathbb{E} \left\lVert x_{1}^{0} -x_{1}^* \right\rVert_2 =\sqrt{ I+1 }.
+$$
+
+And, as the number of points grows $I\to \infty$, if we iterate projected gradient descent $K$ times where
+$$
+ K \geq \frac{\log(1/\epsilon)+{\log(I-1)}/{2}}{-\log \left( 1-c \right)},
+$$
+
+the expected error is less than $\epsilon$,
+
+$$
+\mathbb{E} \left\lVert x_{1}^{0} -x_{1}^* \right\rVert_2 \leq \epsilon.
+$$
+
+::: {.proof}
+
+See @sec-expected-pgd-convergence-proof.
+
+:::
+
+:::
+
+This is contrasted with the expected initial and final error for re-solved multi-scaled descent (@thm-expected-resolved-convergence) and freezed multi-scaled descent (@thm-expected-freezed-convergence).
+
+::: {#thm-expected-resolved-convergence}
+
+### Expected Re-solved Multi-scale Convergence
+
+Assume the same setting as in @thm-expected-pgd-convergence, but instead using re-solved multi-scaled descent method starting with a scale $s_0$.
+Let the number of points at the finest scale be $I = 2^S + 1$.
+Then we have the expected initial error
+
+$$
+\mathbb{E} \left\lVert x_{s_{0}}^{0} -x_{s_{0}}^* \right\rVert_2 = \sqrt{ 2^{S-s_{0}+1} + 2 }.
+$$
+
+Performing $K_s$ iterations at each scale gives us the expected final error
+
+$$
+\begin{align}
+&\mathbb{E} \left\lVert x_{1}^{K_1} -x_{1}^* \right\rVert_2 \\
+&\leq (1-c)^{K_{1}}\sqrt{ 2 ^{S+1}}\left((1-c)^{\sum_{s=2}^S K_{s}}+ \frac{L\left\lvert a-b \right\rvert}{2\cdot{2^{S}}}+ \frac{L\left\lvert a-b \right\rvert}{2\cdot {2}}\sum_{s=2}^{S-1}  \frac{(1-c)^{\sum_{t=2}^sK_{t}}}{2^{S-s}} \right).
+\end{align}
+$$
+
+::: {.proof}
+
+See @sec-expected-resolved-convergence-proof.
+
+:::
+
+:::
+
+It is not as straightforward to get the required number of iterations to achieve a desired level of accuracy.
+Additionally, this would not be a fair comparison with projected gradient descent since we expect an iteration at a coarse scale $S$ to be cheaper (less time and fewer floating point operations) than an iteration at the finest scale $s=1$.
+For this reason, we need to cost an iteration of projected gradient descent at a scale $s$ in terms of the size of the problem at that scale.
+
+::: {#lem-cost-of-gd-vs-ms}
+
+### Cost of Projected Gradient Descent vs Multi-scale
+
+Suppose the total cost of regular descent is given by
+$$
+C_{\text{GD}}=C_{1}K
+$$
+
+where $C_1$ is the cost of performing one iteration at the finest scale $s=1$.
+
+The total cost of multi-scale descent is
+$$
+C_{\text{MS}} = \sum_{s=1}^S C_{s}K_{s}
+$$
+
+similarly.
+
+If we assume the cost of projected gradient scales at least in the size of the problem ($C_{1} = \Omega(I)$, i.e.
+$C_1 \geq CI$ for some $C\geq 0$), then the cost of projected gradient descent at scale $s$ is
+$$
+C_{s} \geq \frac{ 2^{S-s+1}+1 }{ 2^{S-s}+1}C_{s+1} \geq \frac{3}{2} C_{s+1},
+$$
+
+which gives the total cost of multi-scale descent at most
+
+$$
+C_{\text{MS}} \leq C_{1}\sum_{s=1}^S \left( \frac{2}{3} \right)^{s-1} K_{s}.
+$$
+
+::: {.proof}
+
+See @sec-cost-of-gd-vs-ms-proof.
+
+:::
+
+:::
+
+We can use this to select a plan for the number of iterations at each scale $K_s$ that will be cheaper than projected gradient descent, yet still give the same upper bound on the expected final error.
+
+::: {#cor-resolved-cost}
+
+### Sufficient Conditions for Re-solved Multi-scale to be Cheaper
+
+Assume the problem is well conditioned with a condition number at least $c \geq 0.3$, and the finest scale problem is discretized with at least $I \geq 33 = 2^{4+1} + 1$ points.
+Then performing re-solved multi-scale with one iteration $K_s=1$ at each scale except the finest scale where we iterate $K_1 = K - 3$ times, yields a tighter upper bound on the expected final error with a cheaper cost, than performing projected gradient descent with $K$ iterations.
+
+::: {.proof}
+
+See @sec-resolved-cost-proof.
+
+:::
+
+We can play a similar game for the freezed multi-scale approach.
+
+:::
+
+::: {#thm-expected-freezed-convergence}
+
+### Expected Freezed Multi-scale Convergence
+
+Assumed the same setting as @thm-expected-pgd-convergence, but instead using the freezed multi-scale approach starting at the coarsest scale $S$, where the finest scale has $I=2^S + 1$ many points.
+Then the expected initial error is two,
+
+$$
+\mathbb{E} \left\lVert x_{s_{0}}^{0} -x_{s_{0}}^* \right\rVert_2 = 2.
+$$
+
+Performing the same number of iterations $K_s=K$ at each scale gives us the expected final error
+
+$$
+\begin{align}
+&\mathbb{E} \left\lVert x_{1}^{K} -x_{1}^* \right\rVert_2  \\
+&\leq  d(K)\left( 2(1+d(K))^{S-1} + \frac{L_f}{2}\left\lvert b-a \right\rvert \frac{\left( \sqrt{2^S} \left( d(K) + 1 \right)^S - \sqrt{2} \left( d(K) + 1 \right) \right)}{\sqrt{2^S} \left( d(K) + 1 \right) \left( \sqrt{2} d(K) + \sqrt{2} - 1 \right)} \right)
+\end{align}
+$$
+
+where $d(K)=(1-c)^K$, $c$ is the condition number for the problem, and the underlying continuous function $f$ is $L_f$ Lipschitz on the interval $[a,b]$.
+
+::: {.proof}
+
+See @sec-expected-freezed-convergence-proof.
+
+:::
+
+:::
+
+To better analyze when the bound for freezed multi-scale descent (@thm-expected-freezed-convergence) is small than for regular projected gradient descent (@thm-expected-pgd-convergence), we will look at the case when we have many iterations.
+This lets us approximate $d(K)\approx 0$ since we know $d(K)\rightarrow 0$ as $K\rightarrow \infty$.
+
+::: {#cor-freezed-cost}
+
+### Sufficient Conditions for Freezed Multi-scale to be Cheaper
+
+Assume the finest scale has at least $I \geq 2^S + 1$ many points where $S$ is at least
+
+$$
+S > \log_{2}{\left(\left( \frac{\frac{L_f}{2}\left\lvert b-a \right\rvert + 1/2}{\left(\sqrt{2} - 1 \right)}+2 \right)^2 - 2 \right)}.
+$$
+
+Then performing freezed multi-scale with $K_s=\lceil K/3\rceil -1$ iterations at each scale starting with a scale $s=S$, yields a tighter upper bound on the expected final error with a cheaper cost, than performing projected gradient descent with $K$ iterations.
+
+::: {.proof}
+
+See @sec-freezed-cost-proof.
+
+:::
+
+:::
+
+## Benchmarks
+
+### Synthetic Data
+
+TODO see multiscalesynthetic3d.jl
+
+For a synthetic test, we generate 3 source distributions. Each distribution is a 3 dimensional product distribution of some standard continuous distributions.
+
+```julia
+using Distributions
+
+source1a = Normal(4, 1)
+source1b = Uniform(-7, 2)
+source1c = Uniform(-1, 1)
+
+source2a = Normal(0, 3)
+source2b = Uniform(-2, 2)
+source2c = Exponential(2)
+
+source3a = Exponential(1)
+source3b = Normal(0, 1)
+source3c = Normal(0, 3)
+
+source1 = product_distribution([source1a, source1b, source1c])
+source2 = product_distribution([source2a, source2b, source2c])
+source3 = product_distribution([source3a, source3b, source3c])
+
+sources = (source1, source2, source3)
+```
+
+We generate the following $5\times 3$ mixing matrix
+
+```julia
+p1 = [0, 0.4, 0.6]
+p2 = [0.3, 0.3, 0.4]
+p3 = [0.8, 0.2, 0]
+p4 = [0.2, 0.7, 0.1]
+p5 = [0.6, 0.1, 0.3]
+
+C_true = hcat(p1,p2,p3,p4,p5)'
+```
+
+and use it to construct $5$ mixture distributions.
+
+```julia
+distribution1 = MixtureModel([sources...], p1)
+distribution2 = MixtureModel([sources...], p2)
+distribution3 = MixtureModel([sources...], p3)
+distribution4 = MixtureModel([sources...], p4)
+distribution5 = MixtureModel([sources...], p5)
+distributions = [distribution1, distribution2, distribution3, distribution4, distribution5]
+```
+
+These are discretized into $65\times 65\times 65$ sample tensors, and stacked into a $5\times 65\times 65\times 65$. We normalize the $1$-slices so that they sum to one.
+
+```julia
+sinks = [pdf.((d,), xyz) for d in distributions]
+Y = cat(sinks...; dims=4)
+# reorder so the first dimension indexes mixtures rather than the final dimension
+Y = permutedims(Y, (4,1,2,3))
+Y_slices = eachslice(Y, dims=1)
+correction = sum.(Y_slices) # normalize slices to 1
+Y_slices ./= correction
+```
+
+Because these are large tensors, we only test a single shot decomposition (after they have been compiled). This gives us the following.
+
+```julia
+# Run once to compile functions
+factorize(Y; options...);
+multiscale_factorize(Y; continuous_dims=[2, 3, 4], options...);
+```
+```julia
+# Time the functions
+@time decomposition, stats_data, kwargs = factorize(Y; options...);
+```
+```shell
+11.828295 seconds (214.97 k allocations: 15.197 GiB, 32.42% gc time, 0.00% compilation time)
+```
+```julia
+@time decomposition, stats_data, kwargs = multiscale_factorize(Y; continuous_dims=[2, 3, 4], options...);
+```
+```shell
+2.335374 seconds (201.33 k allocations: 2.905 GiB, 25.26% gc time, 0.00% compilation time)
+```
+
+We can see that `multiscale_factorize` is roughly five times as fast and uses about a fifth of the memory in this example.
+
+### Real Data
+
+We use the same sedimentary data and Tucker-$1$ model as described in [@graham_tracing_2025] to factorize a tensor containing mixtures of estimated densities.
+We discretize the densities with $K=2^10 + 1 = 1025$ points to obtain an input tensor $Y\in\mathbb{R}_+^{20\times 7 \times 1025}$ and normalize the depth fibres so that $\sum_{k\in[K]} Y[i,j,k] = 1$ for all $i\in[20]$ and $j\in[7]$.
+
+We run the multi-scale factorization algorithm
+
+```julia
+multiscale_factorize(Y; continuous_dims=3, options...)
+```
+
+with the following options.
+
+```julia
+options = (
+    rank=3,
+    momentum=false,
+    do_subblock_updates=false,
+    model=Tucker1,
+    tolerance=(0.12),
+    converged=(RelativeError), # relative error ≤ 12%
+    constrain_init=true,
+    constraints=[l1scale_average12slices! ∘ nonnegative!, nnonnegative!],
+    stats=[Iteration, ObjectiveValue, GradientNNCone, RelativeError],
+    maxiter=200
+)
+```
+
+We use the same convergence criteria at each scale and iterate until the relative error between the input $Y$ and our model $X$ is at most $12\%$, or until $200$ iterations have passed.
+We use $12\%$ because this is roughly the error Graham et.
+al.
+observe in their final factorization [@graham_tracing_2025].
+The third dimension is specified as continuous since each depth fibre $Y[i, j, :]$ is a discretized continuous probability density function.
+
+This is compared to the regular factorization algorithm
+
+```julia
+factorize(Y; options...)
+```
+
+using the Julia package `BenchmarkingTools`.
+This runs the algorithm as many times as it can within a default time window.
+Note that a new random initialization is generated for each run.
+After running the following,
+
+```julia
+using BenchmarkingTools
+
+# Run each function once so they compile
+factorize(Y; options...);
+multiscale_factorize(Y; continuous_dims=3,options...);
+
+benchmark1 = @benchmark factorize(Y; options...)
+display(benchmark1)
+
+benchmark2 = @benchmark multiscale_factorize(Y; continuous_dims=3,options...)
+display(benchmark2)
+```
+
+we observe the two benchmarks. For `factorize`, we have
+
+```julia
+BenchmarkTools.Trial: 22 samples with 1 evaluation per sample.
+Range (min … max):  103.648 ms … 541.438 ms  ┊ GC (min … max): 13.87% … 14.20%
+Time  (median):     191.769 ms               ┊ GC (median):    15.00%
+Time  (mean ± σ):   227.315 ms ± 130.666 ms  ┊ GC (mean ± σ):  15.04% ±  1.79%
+
+▃▃     █    ▃    ▃
+██▇▁▇▁▁█▁▁▇▇█▇▁▁▁█▁▁▁▇▁▁▇▁▁▁▁▁▁▁▁▁▁▇▁▁▁▁▁▁▁▁▁▁▁▁▁▁▁▁▁▇▁▇▁▁▁▁▇ ▁
+104 ms           Histogram: frequency by time          541 ms <
+
+Memory estimate: 201.53 MiB, allocs estimate: 78879.
+```
+
+and for `multiscale_factorize`, we observe the following.
+
+```julia
+BenchmarkTools.Trial: 59 samples with 1 evaluation per sample.
+Range (min … max):  63.974 ms … 147.756 ms  ┊ GC (min … max):  0.00% … 22.20%
+Time  (median):     81.719 ms               ┊ GC (median):    13.13%
+Time  (mean ± σ):   85.094 ms ±  13.163 ms  ┊ GC (mean ± σ):  12.13% ±  5.42%
+
+            ▁█  ▁ ▁      ▁
+▆▁▁▁▁▁▁▁▆▁▄▆▇██▆▇█▇█▄▆▄▄▄▆█▁▁▁▄▁▄▄▁▄▁▁▄▄▁▁▁▁▁▁▄▁▁▁▁▁▁▁▁▁▁▁▁▄ ▁
+64 ms           Histogram: frequency by time          125 ms <
+
+Memory estimate: 84.93 MiB, allocs estimate: 634376.
+```
+
+By every metric, the multi-scale approach is faster and uses less memory than the regular factorize approach.
+Looking at the median time and memory estimate, it is roughly twice as fast, and uses about half as much memory in the case of this problem.
+These wall clock times include a nontrivial amount of garbage collection (GC) so it means there could be design improvements such as using more preallocated arrays to reduce memory usage.
+But these improvements would speed up both `factorize` and `multiscale_factorize`.
+Without these improvements, we can see that `multiscale_factorize` is less likely to use garbage collection since more computations are occurring on smaller arrays.
+
 # Conclusion
-- all-in-one package
-- provide a playground to invent new decompositions
-- like auto-diff for factorizations
+
+The BlockTensorDecomposition.jl Julia provides a new all-in-one package for performing constrained tensor factorizations, and a playground for designing new decompositions and custom constraints.
+Careful design elements were engineered to balance flexibility and efficiency.
+By creating this package, new advancements like enforcing constraints through scaling rather than projection, and performing optimization over multiple scales were mathematically examined and numerically tested.
+These novel ideas are worth investigating further to see their applicability to continuous optimization beyond tensor factorization.
 
@@
 : Full description of randomizing the order of updates within a `BlockUpdate`. {#tbl-blockupdate-randomization tbl-colwidths="[20,15,27,37]"}
+
+## Constraint Rescaling Proofs
+
+In the following proofs, we have an $L_f$-Lipschitz function $f:[l,u]\to\mathbb{R}$ that is discretized according to
+
+$$
+x[i]=f(t[i]),
+$$
+
+where
+
+$$
+t[i] = l + (i-1)\Delta t= l + (i-1) \frac{u-l}{I-1}.
+$$
+
+See @thm-rem-constraint-assumptions.
+
+This ensures neighboring entries of $x$ are close together. Specifically, we have
+
+$$
+\left\lvert x_{i}-x_{i+1} \right\rvert=\left\lvert f(t_{i})-f(t_{i+1}) \right\rvert\leq L_f\left\lvert t_{i}-t_{i+1} \right\rvert=L_f \frac{{u-l}}{I-1}:=C_x.
+$$
+
+### Linear Constraint Scaling {#sec-linear-constraint-scaling-proof}
+
+::: {.proof}
+First, assume an even number of points $I$.
+$$
+\begin{align}
+\left\lvert 2\left\langle \overline{a}, \overline{x} \right\rangle - b \right\rvert &= \left\lvert 2\sum_{i\ \text{odd}} a_{i}x_{i} -\sum_{i=1}^I a_{i}x_{i}\right\rvert \\
+& = \left\lvert  \sum_{i\ \text{odd}}a_{i}x_{i} - \sum_{i\ \text{even}} a_{i}x_{i}\right\rvert \\
+&=  \left\lvert  \sum_{i\ \text{odd}}a_{i}x_{i} - \sum_{i\ \text{odd}} a_{i+1}x_{i+1}\right\rvert \\
+&\leq \sum_{i\ \text{odd}}\left\lvert a_{i}x_{i}-a_{i+1}x_{i+1} \right\rvert\\
+&= \sum_{i\ \text{odd}}\left\lvert a_{i}x_{i}-a_{i+1}x_{i}+a_{i+1}x_{i}-a_{i+1}x_{i+1} \right\rvert\\
+&\leq \sum_{i\ \text{odd}}\left( \left\lvert a_{i}x_{i}-a_{i+1}x_{i}\right\rvert+\left\lvert a_{i+1}x_{i}-a_{i+1}x_{i+1} \right\rvert \right)\\
+&=\sum_{i\ \text{odd}}\left( \left\lvert x_{i} \right\rvert\left\lvert a_{i}-a_{i+1}\right\rvert+\left\lvert a_{i+1} \right\rvert\left\lvert x_{i}-x_{i+1} \right\rvert \right)\\
+&\leq \max(\lVert a \rVert_{\infty},\lVert x \rVert_{\infty})\sum_{i\ \text{odd}}\left( \left\lvert a_{i}-a_{i+1}\right\rvert+\left\lvert x_{i}-x_{i+1} \right\rvert \right)\\
+&\leq \max(\lVert a \rVert_{\infty},\lVert x \rVert_{\infty})\sum_{i\ \text{odd}}\left( C_{a}+C_{x} \right)\\
+&= \max(\lVert a \rVert_{\infty},\lVert x \rVert_{\infty})(C_{a}+C_{x}) \frac{I}{2}\\
+& = \max(\lVert a \rVert_{\infty},\lVert x \rVert_{\infty}) \frac{I(L_{g} + L_{f})(u-l)}{2(I-1)}\\
+& \leq \max(\lVert g \rVert_{\infty},\lVert f \rVert_{\infty}) \frac{I(L_{g} + L_{f})(u-l)}{2(I-1)}
+\end{align}
+$$
+:::
+
+### $L_1$ norm Constraint {#sec-l1-norm-constraint-scaling-proof}
+
+::: {.proof}
+
+For $I$ even,
+$$
+\begin{align}
+\left\lvert 2\left\lVert \overline{x} \right\rVert - \lVert x \rVert_{1} \right\rvert &= \left\lvert 2\sum_{i\ \text{odd}} \left\lvert x_{i} \right\rvert -\sum_{i=1}^I \left\lvert x_{i} \right\rvert\right\rvert \\
+& = \left\lvert  \sum_{i\ \text{odd}}\left\lvert x_{i} \right\rvert+ \sum_{i\ \text{odd}}\left\lvert x_{i} \right\rvert - \left( \sum_{i\ \text{odd}}\left\lvert x_{i} \right\rvert + \sum_{i\ \text{even}} \left\lvert x_{i} \right\rvert  \right)\right\rvert \\
+& = \left\lvert  \sum_{i\ \text{odd}}\left\lvert x_{i} \right\rvert - \sum_{i\ \text{even}} \left\lvert x_{i} \right\rvert\right\rvert \\
+&=  \left\lvert  \sum_{i\ \text{odd}}\left\lvert x_{i} \right\rvert - \sum_{i\ \text{odd}} \left\lvert x_{i+1} \right\rvert\right\rvert \\
+&\leq \sum_{i\ \text{odd}}\left\lvert \left\lvert x_{i} \right\rvert-\left\lvert x_{i+1} \right\rvert \right\rvert \\
+&\leq \sum_{i\ \text{odd}}\left\lvert  x_{i} - x_{i+1}  \right\rvert \\
+&\leq \sum_{i\ \text{odd}}C\\
+&= C \frac{I}{2}\\
+&= \frac{IL (u-l)}{2(I-1)}.
+\end{align}
+$$
+
+For $I$ odd, we have
+$$
+\begin{align}
+\left\lvert 2\left\lVert \overline{x} \right\rVert - \frac{I+1}{I} \lVert x \rVert_{1} \right\rvert &= \left\lvert 2\left\lVert \overline{x} \right\rVert - \lVert x \rVert_{1} - \frac{\lVert x \rVert_{1}}{I} \right\rvert \\
+&=\left\lvert 2\sum_{i\ \text{odd}} \left\lvert x_{i} \right\rvert - \sum_{i=1}^I \left\lvert x_{i} \right\rvert  - \frac{\lVert x \rVert_{1}}{I} \right\rvert  \\
+& = \left\lvert  \sum_{i\ \text{odd}}\left\lvert x_{i} \right\rvert+ \sum_{i\ \text{odd}}\left\lvert x_{i} \right\rvert - \left( \sum_{i\ \text{odd}}\left\lvert x_{i} \right\rvert + \sum_{i\ \text{even}} \left\lvert x_{i} \right\rvert  \right) - \frac{\lVert x \rVert_{1}}{I}  \right\rvert \\
+& = \left\lvert  \sum_{i\ \text{odd}}\left\lvert x_{i} \right\rvert - \sum_{i\ \text{even}} \left\lvert x_{i} \right\rvert  - \frac{\lVert x \rVert_{1}}{I}  \right\rvert \\
+&=  \left\lvert  \sum_{j=1}^{(I-1)/2}\left\lvert x_{2j-1} \right\rvert + \left\lvert x_{I} \right\rvert - \sum_{j=1}^{(I-1)/2} \left\lvert x_{2j} \right\rvert  - \frac{\lVert x \rVert_{1}}{I}  \right\rvert \\
+&\leq \sum_{j=1}^{(I-1)/2}\left\lvert \left\lvert x_{2j-1} \right\rvert-\left\lvert x_{2j} \right\rvert \right\rvert + \left\lvert \left\lvert x_{I} \right\rvert - \frac{\lVert x \rVert_{1}}{I} \right\rvert \\
+&\leq \sum_{j=1}^{(I-1)/2}\left\lvert  x_{2j-1} - x_{2j}  \right\rvert + \frac{1}{I} \left\lvert I \left\lvert x_{I} \right\rvert - \lVert x \rVert_{1} \right\rvert \\
+&\leq \sum_{j=1}^{(I-1)/2} C + \frac{1}{I} \left\lvert \sum_{i=1}^I \left( \left\lvert x_{I} \right\rvert - \lvert x_{i} \rvert  \right)\right\rvert\\
+&\leq C \frac{I-1}{2} + \frac{1}{I} \sum_{i=1}^I \left\lvert  \left\lvert x_{I} \right\rvert - \lvert x_{i} \rvert  \right\rvert\\
+&\leq C \frac{I-1}{2} + \frac{1}{I} \sum_{i=1}^I \left\lvert   x_{I}  -  x_{i}  \right\rvert\\
+&\leq C \frac{I-1}{2} + \frac{1}{I} \sum_{i=1}^I C (I-i)\ \text{(Apply Lipschitz recursively)} \\
+& = C \frac{I-1}{2} + \frac{1}{I}\frac{CI(I-1)}{2}\\
+&= \frac{L (u-l)}{2} + \frac{L (u-l)}{2} \\
+&= L (u-l).
+\end{align}
+$$
+
+:::
+
+## Multi-scale Convergence Proofs
+
+### Inexact Interpolation Error Proof{#sec-inexact-interpolation-proof}
+
+Proof of @lem-inexact-interpolation.
+
+*Proof*. We let the inexact values be $\tilde{y}[k]=y[k]+\delta_{k}$, and we interpolate the inexact values to get $\hat{\tilde{Y}}$:
+$$
+\begin{align}
+  \hat{\tilde{Y}}[j]&=\begin{cases}
+  \tilde{y}[\frac{j+1}{2}]   & \text{if $j$ is odd}\\
+  \frac{1}{2}\left( \tilde{y}[\frac{j}{2}] + \tilde{y}[\frac{j}{2}+1] \right) & \text{if $j$ is even}
+  \end{cases}\\
+  &=\begin{cases}
+  y[\frac{j+1}{2}] +\delta_{\frac{{j+1}}{2}}  & \text{if $j$ is odd}\\
+  \frac{1}{2}\left( y[\frac{j}{2}] + y[\frac{j}{2}+1] \right) + \frac{1}{2}\left( \delta_{\frac{j}{2}}+\delta_{{\frac{j}{2}}+1} \right) & \text{if $j$ is even}
+  \end{cases}.
+\end{align}
+$$
+
+So for odd $j$,
+$$
+\left\lvert \hat{\tilde{Y}}[j]-\hat{Y}[j] \right\rvert= \left\lvert \delta_{\frac{{j+1}}{2}} \right\rvert:=e[j],
+$$
+
+and even $j$,
+$$
+\left\lvert \hat{\tilde{Y}}[j]-\hat{Y}[j] \right\rvert = \frac{1}{2}\left\lvert \delta_{\frac{j}{2}}+\delta_{{\frac{j}{2}}+1} \right\rvert:=e[j].
+$$
+
+So we have
+$$
+\left\lVert \hat{\tilde{Y}} - \hat{Y} \right\rVert = \sqrt{ \sum_{j\in[J]} \left\lvert \hat{\tilde{Y}}[j]-\hat{Y}[j] \right\rvert^2 } = \sqrt{ \sum_{j\in[J]} \left\lvert e[j] \right\rvert^2 } = \lVert e \rVert
+$$
+
+We would like some bound on $\lVert e \rVert$ in terms of the closeness of $y$ and $\tilde{y}$ (that is, in terms of $\delta$).
+
+$$
+\begin{align}
+\lVert e \rVert^2 &= \sum_{j\ \text{odd}} \left\lvert e[j] \right\rvert^2 + \sum_{j\ \text{even}} \left\lvert e[j] \right\rvert^2 \\
+&= \sum_{j\ \text{odd}} \left\lvert \delta_{\frac{{j+1}}{2}} \right\rvert^2 + \sum_{j\ \text{even}} \left\lvert \frac{1}{2} \left( \delta_{\frac{j}{2}}+\delta_{{\frac{j}{2}}+1} \right)\right\rvert^2  \\
+&= \sum_{k=1}^K \left\lvert \delta_{k}\right\rvert^2 + \frac{1}{4}\sum_{k=1}^{K-1} \left( \delta_{k}+\delta_{k+1} \right)^2 \\
+&= \left\lVert \delta \right\rVert^2 + \frac{1}{4}\sum_{k=1}^{K-1}\left( \delta_{k}^2 + \delta_{k+1}^2 +2\delta_{k}\delta_{{k+1}} \right) \\
+&= \left\lVert \delta \right\rVert^2 + \frac{1}{4}\left(\sum_{k=1}^{K-1} \delta_{k}^2 + \sum_{k=1}^{K-1}\delta_{k+1}^2 +2\sum_{k=1}^{K-1}\delta_{k}\delta_{{k+1}} \right) \\
+&\leq \left\lVert \delta \right\rVert^2 + \frac{1}{4}\left(\sum_{k=1}^{K} \delta_{k}^2 + \sum_{k=0}^{K-1}\delta_{k+1}^2 +2\sum_{k=1}^{K-1}\delta_{k}\delta_{{k+1}} \right) \\
+&= \left\lVert \delta \right\rVert^2 + \frac{1}{4}\left(\left\lVert \delta \right\rVert^2 + \left\lVert \delta \right\rVert^2 +2\sum_{k=1}^{K-1}\delta_{k}\delta_{{k+1}} \right) \\
+&= \frac{3}{2}\left\lVert \delta \right\rVert^2 +\frac{1}{2}\sum_{k=1}^{K-1}\delta_{k}\delta_{{k+1}}\\
+&\leq \frac{3}{2}\left\lVert \delta \right\rVert^2 +\frac{1}{2}\sqrt{  \sum_{k=1}^{K-1}\delta_{k}^2}\sqrt{  \sum_{k=1}^{K-1}\delta_{k+1}^2}\quad \text{(Cauchy–Schwarz)}\\
+&\leq \frac{3}{2}\left\lVert \delta \right\rVert^2 +\frac{1}{2}\sqrt{  \sum_{k=1}^{K}\delta_{k}^2}\sqrt{  \sum_{k=0}^{K-1}\delta_{k+1}^2}\\
+&= \frac{3}{2}\left\lVert \delta \right\rVert^2 +\frac{1}{2}\left\lVert \delta \right\rVert\left\lVert \delta \right\rVert\\
+&= 2\left\lVert \delta \right\rVert^2
+\end{align}
+$$
+
+Therefore,
+$$
+\lVert e \rVert\leq\sqrt{ 2 }\left\lVert \delta \right\rVert
+$$
+
+or substituting our notation,
+$$
+\left\lVert \hat{\tilde{Y}} - \hat{Y} \right\rVert \leq \sqrt{ 2 } \left\lVert \tilde{y}-y \right\rVert.
+$$
+
+This is saying that the error in our fine grid $(Y)$ is bounded by a factor of $\sqrt{ 2 }$ of the error in the coarse grid $(y)$ when we double the number of points.
+
+Now we can use the triangle inequality to bound the difference between the interpolated approximate values $\hat{\tilde{Y}}$ and the true values on the finer grid $Y$:
+$$
+\begin{align}
+\left\lVert \hat{\tilde{Y}} - Y \right\rVert &\leq \left\lVert \hat{\tilde{Y}} - \hat{Y} \right\rVert + \left\lVert \hat{Y}-Y \right\rVert \\
+&\leq \sqrt{ 2 } \left\lVert \tilde{y}-y \right\rVert + \frac{L}{2\sqrt{ K-1 }}\left\lVert a-b \right\rVert_{2}.
+\end{align}
+$$
+
+END OF PROOF
+
+### Re-solved Multi-scale Descent Error Proof {#sec-resolved-multiscale-descent-error-proof}
+
+Proof of @thm-resolved-multiscale-descent-error.
+
+Using the lemmas in @sec-analysis-lemmas, we showed
+$$
+\left\lVert \hat{\tilde{Y}} - Y \right\rVert \leq \sqrt{ 2 } \left\lVert \tilde{y}-y \right\rVert + \frac{L\left\lvert a-b \right\rvert}{2\sqrt{ K-1 }}.
+$$
+Note the $L$ is the Lipchitz constant of the continuous function $f$, not the smoothness of the objective $\mathcal{L}$ above, and the $K$ is the number of points in the discretization, not the number of iterations.
+Translating this into our notation for the multi-scaled descent, we have
+$$
+\left\lVert x^0_{\frac{s}{2}} - x^*_{\frac{s}{2}} \right\rVert = \left\lVert \hat{x}^{K_{s}}_{s} - x^*_{\frac{s}{2}} \right\rVert \leq \sqrt{ 2 } \left\lVert {x}^{K_{s}}_{s}-x^*_{s} \right\rVert + \frac{L\left\lvert a-b \right\rvert}{2\sqrt{2^{S-s+1}}}.
+$$
+Note that at scale $s$, we have $2^{S-s+1} +1$ many points in our discretization, where we have $2^S +1$ many points in our finest scale, and the discretization is over the interval $a\leq t\leq b$ for an $L$ Lipschitz function $f$.
+
+Combining this with our convergence for gradient descent gives us the inequality
+$$
+\begin{align}
+\left\lVert x^0_{\frac{s}{2}} - x^*_{\frac{s}{2}} \right\rVert  &\leq \sqrt{ 2 } \left\lVert {x}^{K_{s}}_{s}-x^*_{s} \right\rVert + \frac{L\left\lvert a-b \right\rvert}{2\sqrt{2^{S-s+1}}} \\
+&\leq \sqrt{ 2 } \left( 1-c \right)^{K_{s}} \left\lVert x_{s}^{0} -x_{s}^* \right\rVert + \frac{L\left\lvert a-b \right\rvert}{2\sqrt{2^{S-s+1}}}
+\end{align}
+$$
+I noticed I am using $s$ in two different ways here.
+Originally, $s$ was supposed to be the number of points we skip, as in "only keep every $s$ points in our discretization".
+This is meant to start at some large power of $2$ and keep halving until we hit $1$.
+But the $s$ in $2^{S-s+1}$ counts what scale we are at and starts at $S$ and decreasing by $1$ until we hit $1$.
+The second definition makes more sense so we will go with that from now on.
+This means our descent lemma looks like
+$$
+\begin{align}
+\left\lVert x^0_{s-1} - x^*_{s-1} \right\rVert
+&\leq \sqrt{ 2 } \left( 1-c \right)^{K_{s}} \left\lVert x_{s}^{0} -x_{s}^* \right\rVert + \frac{L\left\lvert a-b \right\rvert}{2\sqrt{2^{S-s+1}}}.
+\end{align}
+$$
+Or writing it in terms of $s+1$:
+$$
+\begin{align}
+\left\lVert x^0_{s} - x^*_{s} \right\rVert
+&\leq \sqrt{ 2 } \left( 1-c \right)^{K_{s+1}} \left\lVert x_{s+1}^{0} -x_{s+1}^* \right\rVert + \sqrt{ 2^s } \frac{L\left\lvert a-b \right\rvert}{2\sqrt{2^{S}}}.
+\end{align}
+$$
+Let $e_{s}=\left\lVert x^0_{s} - x^*_{s} \right\rVert$ and $C=\frac{L\left\lvert a-b \right\rvert}{2\sqrt{2^{S+1}}}$ to clean up notation
+
+$$
+\begin{align}
+e_{s}
+&\leq \sqrt{ 2 } \left( 1-c \right)^{K_{s+1}} e_{s+1} + \sqrt{ 2^{s+1} } C.
+\end{align}
+$$
+Now we recurse! Starting from the last initial error at the finest scale $e_{1}$, we want to write this in terms of the first initial error at the largest scale $e_{S}$.
+$$
+\begin{align}
+e_{1}
+&\leq \sqrt{ 2 } \left( 1-c \right)^{K_{2}} e_{2} + \sqrt{ 2^{2} } C \\
+&\leq \sqrt{ 2 } \left( 1-c \right)^{K_{2}} \left( \sqrt{ 2 } \left( 1-c \right)^{K_{3}} e_{3} + \sqrt{ 2^{3} } C \right) + \sqrt{ 2^{2} } C \\
+&= \sqrt{ 2 } \left( 1-c \right)^{K_{2}}  \sqrt{ 2 } \left( 1-c \right)^{K_{3}} e_{3} + \sqrt{ 2 } \left( 1-c \right)^{K_{2}}\sqrt{ 2^{3} } C + \sqrt{ 2^{2} } C \\
+&=\sqrt{ 2 }^2 \left( 1-c \right)^{K_{2}+K_{3}} e_{3} + \sqrt{ 2 } \left( 1-c \right)^{K_{2}}\sqrt{ 2^{3} } C + \sqrt{ 2^{2} } C
+\end{align}
+$$
+Before going further, we actually want to run gradient descent with the starting point $x^0_{1}$ (the $x$ inside $e_{1}$), so we should really be writing
+$$
+\begin{align}
+&\left\lVert x_{1}^{K_{1}}-x_{1}^* \right\rVert \\
+&\leq \left( 1-c \right)^{K_{1}} \left\lVert x_{1}^{0} -x_{1}^* \right\rVert \\
+&= \left( 1-c \right)^{K_{1}} e_{1} \\
+&\leq \left( 1-c \right)^{K_{1}}\left( \sqrt{ 2 } \left( 1-c \right)^{K_{2}} e_{2} + \sqrt{ 2^{2} } C  \right) \\
+&= \sqrt{ 2 }\left( 1-c \right)^{K_{1}+K_{2}}   e_{2} + \left( 1-c \right)^{K_{1}}\sqrt{ 2^{2} } C \\
+&\leq \sqrt{ 2 }\left( 1-c \right)^{K_{1}+K_{2}}   \left( \sqrt{ 2 } \left( 1-c \right)^{K_{3}} e_{3} + \sqrt{ 2^{3} } C  \right)  + \left( 1-c \right)^{K_{1}}\sqrt{ 2^{2} } C \\
+&= \sqrt{ 2 }^2\left( 1-c \right)^{K_{1}+K_{2}+K_{3}} e_{3} + \sqrt{ 2 }\left( 1-c \right)^{K_{1}+K_{2}}\sqrt{ 2^{3} } C   + \left( 1-c \right)^{K_{1}}\sqrt{ 2^{2} } C  \\
+&\leq \sqrt{ 2 }^2\left( 1-c \right)^{K_{1}+K_{2}+K_{3}} \left( \sqrt{ 2 } \left( 1-c \right)^{K_{4}} e_{4} + \sqrt{ 2^{4} } C  \right) + \sqrt{ 2 }\left( 1-c \right)^{K_{1}+K_{2}}\sqrt{ 2^{3} } C   + \left( 1-c \right)^{K_{1}}\sqrt{ 2^{2} } C  \\
+&= \sqrt{ 2 }^3\left( 1-c \right)^{K_{1}+K_{2}+K_{3}+K_{4}}e_{4} + \sqrt{ 2 }^2\left( 1-c \right)^{K_{1}+K_{2}+K_{3}} \sqrt{ 2^{4} } C + \sqrt{ 2 }\left( 1-c \right)^{K_{1}+K_{2}}\sqrt{ 2^{3} } C   + \left( 1-c \right)^{K_{1}}\sqrt{ 2^{2} } C \\
+&\ \ \vdots \\
+&\leq \sqrt{ 2 }^{S-1}(1-c)^{\sum_{s=1}^S K_{s}}e_{S} +\sum_{s=1}^{S-1} \sqrt{ 2 }^{s-1} (1-c)^{\sum_{t=1}^sK_{t}} \sqrt{ 2^{s+1} } C \\
+&= \sqrt{ 2 ^{S-1}}(1-c)^{\sum_{s=1}^S K_{s}}e_{S} + C\sum_{s=1}^{S-1}  2 ^{s} (1-c)^{\sum_{t=1}^sK_{t}}.
+\end{align}
+$$
+
+### Freezed Multi-scale Descent Error Proof {#sec-freezed-multiscale-descent-error-proof}
+
+Proof of @thm-freezed-multiscale-descent-error.
+
+Using the work below, we have
+$$
+\left\lVert e_{(s)}^0 \right\rVert\leq \frac{L\left\lvert b-a \right\rvert}{2\sqrt{ 2^{S-s} }} + \left\lVert e_{s+1}^{K_{s+1}} \right\rVert.
+$$
+Putting this back into the GD inequality gets us
+$$
+\left\lVert e_{(s)}^{K_{s}} \right\rVert \leq (1-c)^{K_{s}}\left( \frac{L\left\lvert b-a \right\rvert}{2\sqrt{ 2^{S-s} }} + \left\lVert e_{s+1}^{K_{s+1}} \right\rVert \right).
+$$
+And finally, we get the recursion relation
+$$
+\begin{align}
+\left\lVert e_{s}^{K_{s}} \right\rVert^2&\leq (1-c)^{2K_{s}}\left( \frac{L\left\lvert b-a \right\rvert}{2\sqrt{ 2^{S-s} }} + \left\lVert e_{s+1}^{K_{s+1}} \right\rVert \right)^2 + \left\lVert e_{s+1}^{K_{s+1}} \right\rVert^2
+\end{align}
+$$
+We will use big $C=\frac{L}{2}\left\lvert b-a \right\rvert$ to clean up a bit. Note little $c=\frac{\mu}{L_{\nabla \ell}}$ is the ratio of strongly convex to smoothness of the loss function $\ell$.
+$$
+\begin{align}
+\left\lVert e_{s}^{K_{s}} \right\rVert^2&\leq (1-c)^{2K_{s}}\left( \frac{C}{\sqrt{ 2^{S-s} }} + \left\lVert e_{s+1}^{K_{s+1}} \right\rVert \right)^2 + \left\lVert e_{s+1}^{K_{s+1}} \right\rVert^2
+\end{align}
+$$
+Or the looser bound by removing the squares
+$$
+\begin{align}
+\left\lVert e_{s}^{K_{s}} \right\rVert &\leq (1-c)^{K_{s}}\left( \frac{C}{\sqrt{ 2^{S-s} }} + \left\lVert e_{s+1}^{K_{s+1}} \right\rVert \right) + \left\lVert e_{s+1}^{K_{s+1}} \right\rVert \\
+&= \left( 1+(1-c)^{K_{s}} \right)\left\lVert e_{s+1}^{K_{s+1}} \right\rVert + C(1-c)^{K_{s}} \frac{1}{\sqrt{ 2^{S-s} }}
+\end{align}
+$$
+
+- we have true points $x$ which are true sample values of an $L$ Lipschitz function
+- we have an approximation of $x$, given by $\hat{x}$
+- then we have a linear interpolation of $\hat{x}$ given by $\hat{X}$. These are ONLY the in-between points
+- the true values at the in-between points are $Y$
+- the linear interpolation of the true points is $X$
+- the respective errors are $e$ and $E$
+- want to bound $E$ by $e$
+- suppose we have $M=2^{S-s}+1$ points at the coarse scale for $x$, for some positive integer $s$ (this is the scale $s+1$, the previous scale)
+- we will have $N=2^{S-s}=M-1$ in-between points at the finer grid for $X$
+	- Note, this gives $2^{S-s+1}+1$ may points at the current scale $s$
+
+We have
+$$\hat{x}_{i}-x_{i}=e_{i}$$
+and the linear interpolation (only the in-between points) is
+$$
+\hat{X}_{j} = \frac{1}{2}(\hat{x}_{j}+\hat{x}_{j+1})
+$$
+Note there is one fewer point.
+Similarly, with the interpolation of the true values
+$$
+{X}_{j} = \frac{1}{2}({x}_{j}+{x}_{j+1})
+$$
+The error is defined as
+$$
+E_{j} = \hat{X}_{j}-Y_{j}.
+$$
+This is what we want to bound.
+We also have the difference between the interpolation of the approximate and interpolation of the true values
+$$
+\hat{X}_{j}-X_{j}= \frac{1}{2}(e_{j}+e_{j+1}):=\delta_{j}
+$$
+
+There is also the difference between the true values at the in-between points $Y$ and the interpolation of the true points $X$.
+$$
+X_{j}-Y_{j}
+$$
+We have
+$$
+\left\lVert E \right\rVert=\left\lVert \hat{X}-Y \right\rVert \leq \left\lVert X-Y \right\rVert+ \left\lVert \hat{X}-X \right\rVert = \left\lVert X-Y \right\rVert + \left\lVert \delta \right\rVert
+$$
+Now
+$$
+\begin{align}
+\left\lVert X-Y \right\rVert^2 &= \sum_{j=1}^N \left( X_{j}-Y_{j} \right)^2 \\
+&\leq \sum_{j=1}^N \left( \frac{L}{2} \Delta_{j} \right)^2 \\
+&\leq N \left( \frac{L}{2} \Delta_{j} \right)^2
+\end{align}
+$$
+where $\Delta_{i}$ is the (input) space between the $i$ and $i+1$ points. For $M$ equally spaced points on (and including the boundary) the interval $[a,b]$,
+$$
+\Delta_{j} = \frac{b-a}{M-1}=\frac{b-a}{N}
+$$
+So we have
+$$
+\left\lVert X-Y \right\rVert\leq \sqrt{ N }\frac{L}{2} \frac{b-a}{N} = \frac{L\left\lvert b-a \right\rvert}{2\sqrt{ N }}.
+$$
+Now we look at the second term $\delta$.
+$$
+\begin{align}
+\left\lVert \delta \right\rVert^2 &= \sum_{j=1}^N \delta_{j}^2 \\
+&= \sum_{j=1}^N \frac{1}{2^2}(e_{j}+e_{j+1})^2 \\
+&= \frac{1}{4}\sum_{j=1}^N (e_{j}^2+e_{j+1}^2 +2e_{j}e_{j+1}) \\
+&= \frac{1}{4}\left( \sum_{j=1}^N e_{j}^2+ \sum_{j=1}^N e_{j+1}^2 +2\sum_{j=1}^Ne_{j}e_{j+1}  \right)\\
+&\leq \frac{1}{4}\left( \sum_{j=1}^N e_{j}^2+ \sum_{j=1}^N e_{j+1}^2 +2\sqrt{ \sum_{j=1}^Ne_{j}^2}\sqrt{ \sum_{j=1}^Ne_{j+1}^2 } \right)\\
+&\leq \frac{1}{4}\left( \sum_{i=1}^M e_{i}^2+ \sum_{i=1}^M e_{j}^2 +2\sqrt{ \sum_{i=1}^Me_{i}^2}\sqrt{ \sum_{i=1}^Me_{i}^2 } \right)\\
+&= \frac{1}{4}\left( \left\lVert e \right\rVert^2+ \left\lVert e \right\rVert^2 +2\left\lVert e \right\rVert\left\lVert e \right\rVert\right)\\
+&= \frac{1}{4}\left( 4\left\lVert e \right\rVert^2\right)\\
+&= \lVert e \rVert^2\\
+\end{align}
+$$
+
+So finally, we have
+$$
+\begin{align}
+\left\lVert E \right\rVert&\leq \frac{L\left\lvert b-a \right\rvert}{2\sqrt{ N }} + \lVert e \rVert.
+\end{align}
+$$
+This is the same thing as we got in the previous attempt, but without the factor of $\sqrt{ 2 }$.
+
+We have the relation
+$$
+\begin{align}
+\left\lVert e_{s}^{K_{s}} \right\rVert &\leq \left( 1+(1-c)^{K_{s}} \right)\left\lVert e_{s+1}^{K_{s+1}} \right\rVert + C(1-c)^{K_{s}} \frac{1}{\sqrt{ 2^{S-s} }}
+\end{align}
+$$
+for every $s=1,\dots,S$. So let's expand this.
+$$
+\begin{align}
+\left\lVert e_{1}^{K_{1}} \right\rVert &\leq \left( 1+(1-c)^{K_{1}} \right)\left\lVert e_{2}^{K_{2}} \right\rVert + C(1-c)^{K_{1}} \frac{1}{\sqrt{ 2^{S-1} }} \\
+&\leq \left( 1+(1-c)^{K_{1}} \right)\left( \left( 1+(1-c)^{K_{2}} \right)\left\lVert e_{3}^{K_{3}} \right\rVert + C(1-c)^{K_{2}} \frac{1}{\sqrt{ 2^{S-2} }}\right) + C(1-c)^{K_{1}} \frac{1}{\sqrt{ 2^{S-1} }} \\
+&= \left( 1+(1-c)^{K_{1}} \right) \left( 1+(1-c)^{K_{2}} \right)\left\lVert e_{3}^{K_{3}} \right\rVert \\
+&\qquad +\ C\left( 1+(1-c)^{K_{1}} \right)(1-c)^{K_{2}} \frac{1}{\sqrt{ 2^{S-2} }} + C(1-c)^{K_{1}} \frac{1}{\sqrt{ 2^{S-1} }} \\
+&\leq \left( 1+(1-c)^{K_{1}} \right) \left( 1+(1-c)^{K_{2}} \right)\left( \left( 1+(1-c)^{K_{3}} \right)\left\lVert e_{4}^{K_{4}} \right\rVert + C(1-c)^{K_{3}} \frac{1}{\sqrt{ 2^{S-3} }} \right) \\
+&\qquad +\ C\left( 1+(1-c)^{K_{1}} \right)(1-c)^{K_{2}} \frac{1}{\sqrt{ 2^{S-2} }} + C(1-c)^{K_{1}} \frac{1}{\sqrt{ 2^{S-1} }} \\
+&= \left( 1+(1-c)^{K_{1}} \right) \left( 1+(1-c)^{K_{2}} \right)\left( 1+(1-c)^{K_{3}} \right)\left\lVert e_{4}^{K_{4}} \right\rVert \\
+&\qquad +\ C\left( 1+(1-c)^{K_{1}} \right) \left( 1+(1-c)^{K_{2}} \right)(1-c)^{K_{3}} \frac{1}{\sqrt{ 2^{S-3} }} \\
+&\qquad +\ C\left( 1+(1-c)^{K_{1}} \right)(1-c)^{K_{2}} \frac{1}{\sqrt{ 2^{S-2} }} + C(1-c)^{K_{1}} \frac{1}{\sqrt{ 2^{S-1} }}
+\end{align}
+$$
+So the general formula goes to
+$$
+\begin{align}
+\left\lVert e_{1}^{K_{1}} \right\rVert &\leq \left\lVert e_{S}^{K_{S}} \right\rVert\prod_{s=1}^{S-1}(1+(1-c)^{K_{s}}) + C\sum_{s=1}^{S-1} \frac{1}{\sqrt{ 2^{S-s} }}(1-c)^{K_{s}}\prod_{j=1}^{s-1}\left( 1+(1-c)^{K_{j}} \right).
+\end{align}
+$$
+After adding the GD on the coarsest scale, we get
+$$
+\begin{align}
+\left\lVert e_{1}^{K_{1}} \right\rVert &\leq \left\lVert e_{S}^{0}  \right\rVert (1-c)^{K_{S}}\prod_{s=1}^{S-1}(1+(1-c)^{K_{s}}) + C\sum_{s=1}^{S-1} \frac{1}{\sqrt{ 2^{S-s} }}(1-c)^{K_{s}}\prod_{j=1}^{s-1}\left( 1+(1-c)^{K_{j}} \right).
+\end{align}
+$$
+
+This gives us the first upper bound for any plan of iterations $K_s$. If we now assume each scale uses the name number of iterations $K_s=K$ we get the following.
+
+$$
+\begin{align}
+\left\lVert e_{1}^{K_{1}} \right\rVert &\leq \left\lVert e_{S}^{0}  \right\rVert (1-c)^{K}(1+(1-c)^{K})^{S-1} + C\sum_{s=1}^{S-1} \frac{1}{\sqrt{ 2^{S-s} }}(1-c)^{K}(1+(1-c)^{K})^{s-1}.
+\end{align}
+$$
+
+Via wolfram alpha, we have
+
+$$
+\begin{align}
+&\sum_{s=1}^{S - 1} ((1 - c)^K ((1 - c)^K + 1)^{s - 1}) \frac{1}{\sqrt{2^{S - s}}}  \\
+&= \frac{(1 - c)^K \left( \sqrt{2^S} \left( (1 - c)^K + 1 \right)^S - \sqrt{2} \left( (1 - c)^K + 1 \right) \right)}{\sqrt{2^S} \left( (1 - c)^K + 1 \right) \left( \sqrt{2} (1 - c)^K + \sqrt{2} - 1 \right)}
+\end{align}
+$$
+
+Setting $d(K)=(1-c)^K$ gives the final upper bound in the theorem.
+
+<!--
+Let's assume we use the same number of steps $K_{s}=K$ for every scale.
+$$
+\begin{align}
+\left\lVert e_{1}^{K_{1}} \right\rVert &\leq \left\lVert e_{S}^{0}  \right\rVert (1-c)^{K}(1+(1-c)^{K})^{S-1} + C\sum_{s=1}^{S-1} \frac{1}{\sqrt{ 2^{S-s} }}(1-c)^{K}(1+(1-c)^{K})^{s-1}.
+\end{align}
+$$
+Via wolfram alpha, we have
+$$
+\begin{align}
+&\sum_{s=1}^{S - 1} ((1 - c)^K ((1 - c)^K + 1)^{s - 1}) \frac{1}{\sqrt{2^{S - s}}}  \\
+&= \frac{(1 - c)^K \left( \sqrt{2^S} \left( (1 - c)^K + 1 \right)^S - \sqrt{2} \left( (1 - c)^K + 1 \right) \right)}{\sqrt{2^S} \left( (1 - c)^K + 1 \right) \left( \sqrt{2} (1 - c)^K + \sqrt{2} - 1 \right)} \\
+&= \frac{a \left( \sqrt{2^S} \left( a + 1 \right)^S - \sqrt{2} \left( a + 1 \right) \right)}{\sqrt{2^S} \left( a + 1 \right) \left( \sqrt{2} a + \sqrt{2} - 1 \right)}  \\
+&:= D(K,S)
+\end{align}
+$$
+
+Recall our regular descent is
+$$
+\left\lVert x_{1}^{K}-x_{1}^* \right\rVert\leq \left( 1-c \right)^{K} \sqrt{ 2^S + 2 }.
+$$
+But for starting out, we say we start at an error of $\sqrt{ 2+2 }=2$
+
+$$
+\begin{align}
+\left\lVert e_{1}^{K_{1}} \right\rVert &\leq  2 d(K)(1+d(K))^{S-1} + CD(K,S).
+\end{align}
+$$
+
+If we factor out an $a$, we get
+$$
+\begin{align}
+\left\lVert e_{1}^{K_{1}} \right\rVert &\leq  d(K)\left( 2(1+d(K))^{S-1} + C \frac{\left( \sqrt{2^S} \left( a + 1 \right)^S - \sqrt{2} \left( a + 1 \right) \right)}{\sqrt{2^S} \left( a + 1 \right) \left( \sqrt{2} a + \sqrt{2} - 1 \right)} \right).
+\end{align}
+$$
+So we need the thing in the bracket to be less than $\sqrt{ 2^S + 2 }$.
+In the limit as $K\to \infty$, we have that $a\to {0}$. So for a very small $a$, we get the bracket expression to be
+$$
+2 + C \frac{\left( \sqrt{2^S} - \sqrt{2} \right)}{\sqrt{2^S}  \left(\sqrt{2} - 1 \right)}.
+$$
+For a scale of $S=2$, we have $2+\frac{{C}}{\sqrt{ 2 }}$ which is bigger than gradient descent's $2$.
+After simplifying the factor for multiscale we get
+$$
+2 + C \frac{\left( \sqrt{2}^{S-1} - 1 \right)}{\sqrt{2}^{S-1}  \left(\sqrt{2} - 1 \right)}.
+$$
+
+For $S=3$, this is $2+\frac{C}{2(\sqrt{ 2 }+1)}$ which could be smaller than $\sqrt{ 2^3+2 }=\sqrt{ 10 }=2\sqrt{ 5 }$. And this scale shrinks with $S$, but grows for GD.
+
+What $S$ do we need for multi scale to be more accurate?
+$$
+\begin{align}
+2 + C \frac{\left( \sqrt{2}^{S-1} - 1 \right)}{\sqrt{2}^{S-1}  \left(\sqrt{2} - 1 \right)} &< \sqrt{ 2^S + 2 } \\
+C&<\left( \sqrt{ 2^S + 2 }-2 \right)\frac{\sqrt{2}^{S-1}  \left(\sqrt{2} - 1 \right)}{\left( \sqrt{2}^{S-1} - 1 \right)} \\
+C&<\left(\sqrt{2} - 1 \right)\left( \sqrt{ 2^S + 2 }-2 \right)\frac{\sqrt{2}^{S-1}  }{\left( \sqrt{2}^{S-1} - 1 \right)}
+\end{align}
+$$
+The last factor is basically $1$ when $S>6$ so let's use that
+$$
+\begin{align}
+C&<\left( \sqrt{ 2^S + 2 }-2 \right) \\
+\left( \frac{C}{\left(\sqrt{2} - 1 \right)}+2 \right)^2 - 2&< 2^S \\
+\log_{2}{\left(\left( \frac{C}{\left(\sqrt{2} - 1 \right)}+2 \right)^2 - 2 \right)}&< S
+\end{align}
+$$
+note because of our approximation, if the above is *not* satisfied, *then* GD is more accurate. This is not an iff condition.
+There are a number of ways to modify it so that it becomes "if this condition holds, then multiscale is better". You could replace $C$ with $C+0.5$ as a simple one.
+
+As a simple case, if $C=\frac{1}{2}$ (the function is 1-Lipschitz on $[0,1]$), then a scale of $S=2$ or higher is already better for multi scale. -->
+
+### Expected Projected Gradient Descent Convergence Proof {#sec-expected-pgd-convergence-proof}
+
+Proof of @thm-expected-pgd-convergence.
+
+**Case 1:** $\lVert x^*_{1} \rVert_2 = 1$.
+
+Without loss of generality, assume (by symmetry) that $(x^*_{1})_{I} = 1$ and $(x^*_{1})_{i}=0$ (fix a point on the sphere at the pole) so that $x^*_{1}=e_{I}$ (the unit vector!
+Not the error!).
+Note that this is not a realistic solution since we already assume the solution comes from a continuous function but we'll use this to illustrate how a warm start from these interpolations can do better than a random start.
+
+$$
+\begin{align}
+\mathbb{E}_{g_{i}\sim \mathcal{N}}\left\lVert g - e_{I} \right\rVert_{2}^2 &= \mathbb{E}_{g_{i}\sim \mathcal{N}}\left[ \sum_{i=1}^I (g_{i}-(e_{I})_{i})^2 \right] \\
+&=\mathbb{E}_{g_{i}\sim \mathcal{N}}\left[ \sum_{i=1}^{I-1} (g_{i}-0)^2 +(g_{i}-1)^2\right] \\
+&=\sum_{i=1}^{I-1} \mathbb{E}_{g_{i}\sim \mathcal{N}}\left[g_{i}^2\right] +\mathbb{E}_{g_{i}\sim \mathcal{N}}\left[(g_{i}-1)^2\right] \\
+&=\sum_{i=1}^{I-1} 1 +\mathbb{E}_{g_{i}\sim \mathcal{N}}\left[g_{i}^2-2g_{i}+1\right] \\
+&=I-1 + (1) - 2(0) +1 \\
+&=I+1.
+\end{align}
+$$
+
+With some Gaussian concentration, we can square root both sides.
+
+**Case 2:** $\lVert x^*_{1} \rVert_2 = 0$.
+
+Assume the solution is centred so that $x^*_{1} = 0\in\mathbb{R}^{I}$. Here, we have the well-known result
+$$
+\mathbb{E}\left[ \left\lVert x^0_1 - x^*_{1}\right\rVert \right] = \mathbb{E}\left[ \left\lVert g - 0\right\rVert \right] = \sqrt{ I }.
+$$
+
+So either way, our initial error for a scaled or centred problem goes like $\sim \sqrt{ I }$.
+Since the number of points we have is $I$ is one plus a power of two $I=2^S+1$, we *expect* (that is, with high probability) the following convergence
+$$
+\left\lVert x_{1}^{K}-x_{1}^* \right\rVert\leq \left( 1-c \right)^{K} \sqrt{ 2^S + 2 }.
+$$
+
+So to ensure $\left\lVert x_{1}^{K}-x_{1}^* \right\rVert \leq\epsilon$, we need
+$$
+\begin{align}
+\left( 1-c \right)^{K} \sqrt{ 2^S + 2 } &\leq \epsilon \\
+ \frac{\sqrt{ 2^S + 2 }}{\epsilon} &\leq  \left( 1-c \right)^{-K}\\
+ \log\left( \frac{\sqrt{ 2^S + 2 }}{\epsilon} \right) &\leq K \log\left( \left( 1-c \right)^{-1} \right)\\
+ \frac{\log\left( \frac{\sqrt{ 2^S + 2 }}{\epsilon} \right)}{-\log \left( 1-c \right)} &\leq K \\
+ \frac{\frac{1}{2}\log\left(  2^S + 2  \right) + \log(1/\epsilon)}{-\log \left( 1-c \right)} &\leq K.
+\end{align}
+$$
+
+For large $S$, this is approximately
+$$
+ \frac{\log(1/\epsilon)+\frac{S}{2}\log\left(  2\right)}{-\log \left( 1-c \right)} \leq K.
+$$
+
+### Expected Resolved Multi-scale Descent Convergence Proof {#sec-expected-resolved-convergence-proof}
+
+Proof of @thm-expected-resolved-convergence.
+
+First we need to have a handle on how close we are to our coarsest discretization at scale $s=s_{0}$.
+The coarsest we could go would be when $s_{0}=S$ which would result in $2^{S-s_{0}+1}+1=3$ points total at $t=a, \frac{1}{2}(a+b)$, and $b$.
+
+For the centred problem, $x_{s_{0}}^*=0\in\mathbb{R}^{2^{S-s_{0}+1}+1}$, so we would expect a Gaussian initialization to have an initial error of
+$$
+\mathbb{E}\left[ \left\lVert x_{s_{0}}^{0} -x_{s_{0}}^* \right\rVert \right] =\sqrt{ 2^{S-s_{0}+1}+1 }.
+$$
+If we used the scaled problem, it is a bit tricker to consider what happens when we only include $2^{S-s_{0}+1}+1$ many points, out of the total $2^{S}+1$ possible points.
+
+We will actually work out a bound of $\sqrt{ 2^{S-s_{0}+1}+2 }$ by considering the "smoothest" case where $(x_{1}^*)_{i}=\frac{1}{\sqrt{ 2^S+1 }}$ (normalized perfectly to a diagonal), and the "roughest" case where $x_{1}^*=e_{2^S+1}\in\mathbb{R}^{2^S+1}$ was aligned to the pole.
+After discretization, the smooth case still has all the same entries, but there are just less of them ($2^{S-s_{0}+1}+1$ many).
+In the rough case, we stay aligned to a pole $x_{s_{0}}^*=e_{2^{S-s_{0}+1}+1}\in\mathbb{R}^{2^{S-s_{0}+1}+1}$ since the last entry stays at a one, and the rest are still zero.
+We use a standard normal initialization of $g\in\mathbb{R}^{2^{S-s_{0}+1}+1}$ in the smaller space.
+This is identical to choosing the same initialization on the finer grid, and dropping our the coordinates that we skip.
+So this really is a fair comparison to the regular gradient descent!
+In the smooth case,
+$$
+\begin{align}
+\mathbb{E}\left[ \left\lVert x_{s_{0}}^{0} -x_{s_{0}}^* \right\rVert^2 \right] &=\sum_{i=1}^{2^{S-s_{0}+1}+1} \mathbb{E}\left( g_{i} - \frac{1}{\sqrt{ 2^S+1 }} \right)^2 \\
+&= \sum_{i=1}^{2^{S-s_{0}+1}+1}\left(  \mathbb{E}\left[ g_{i}^2 \right] -2\frac{1}{\sqrt{ 2^S+1 }} \mathbb{E}\left[ g_{i} \right] +\frac{1}{2^S+1} \mathbb{E}\left[ 1 \right]  \right) \\
+&= \sum_{i=1}^{2^{S-s_{0}+1}+1}\left(  1 - 0 +\frac{1}{2^S+1} \right) \\
+&= (2^{S-s_{0}+1}+1)\left(  1 +\frac{1}{2^S+1} \right) \\
+&= (2^{S-s_{0}+1}+1)\left( \frac{2^S+2}{2^S+1} \right).
+\end{align}
+$$
+In the rough case
+$$
+\begin{align}
+\mathbb{E}\left[ \left\lVert x_{s_{0}}^{0} -x_{s_{0}}^* \right\rVert^2 \right] &= \sum_{i=1}^{2^{S-s_{0}+1}+1-1} \mathbb{E}\left( g_{i} - 0 \right)^2 + \mathbb{E}\left( g_{2^{S-s_{0}+1}+1} - 1 \right)^2 \\
+&= 2^{S-s_{0}+1} + 2.
+\end{align}
+$$
+For $2^{S-s_{0}+1}\geq 3$ (which is the case), $2^{S-s_{0}+1} + 2 > (2^{S-s_{0}+1}+1)\left( \frac{2^S+2}{2^S+1} \right)$ (in fact, it is bigger by $1$ asymptotically as $S$ grows large).
+
+So we are justified in using $\sqrt{ 2^{S-s_{0}+1} + 2 }$ as our *expected* bound on $\left\lVert x_{s_{0}}^{0} -x_{s_{0}}^* \right\rVert$.
+
+This means our descent is
+$$
+\begin{align}
+&\left\lVert x_{1}^{K_{1}}-x_{1}^* \right\rVert\\
+&\leq \sqrt{ 2 ^{S-1}}(1-c)^{\sum_{s=1}^{S}K_{s}}e_{S} + C\sum_{s=1}^{S-1}  2 ^{s} (1-c)^{\sum_{t=1}^sK_{t}} \\
+&\leq \sqrt{ 2 ^{S-1}}(1-c)^{\sum_{s=1}^S K_{s}}\sqrt{ 2^{S-S+1} + 2 } + C\sum_{s=1}^{S-1}  2 ^{s} (1-c)^{\sum_{t=1}^sK_{t}} \\
+&= \sqrt{ 2 ^{S+1}}(1-c)^{\sum_{s=1}^S K_{s}}+ C\sum_{s=1}^{S-1}  2 ^{s} (1-c)^{\sum_{t=1}^sK_{t}} \\
+&= (1-c)^{K_{1}}\left(\sqrt{ 2 ^{S+1}}(1-c)^{\sum_{s=2}^S K_{s}}+ 2C+ C\sum_{s=2}^{S-1}  2 ^{s} (1-c)^{\sum_{t=2}^sK_{t}} \right) \\
+&= (1-c)^{K_{1}}\left(\sqrt{ 2 ^{S+1}}(1-c)^{\sum_{s=2}^S K_{s}}+ 2\frac{L\left\lvert a-b \right\rvert}{2\sqrt{2^{S+1}}}+ \frac{L\left\lvert a-b \right\rvert}{2\sqrt{2^{S+1}}}\sum_{s=2}^{S-1}  2 ^{s} (1-c)^{\sum_{t=2}^sK_{t}} \right) \\
+&= (1-c)^{K_{1}}\sqrt{ 2 ^{S+1}}\left((1-c)^{\sum_{s=2}^S K_{s}}+ \frac{L\left\lvert a-b \right\rvert}{2\cdot{2^{S}}}+ \frac{L\left\lvert a-b \right\rvert}{2\cdot {2}}\sum_{s=2}^{S-1}  \frac{(1-c)^{\sum_{t=2}^sK_{t}}}{2^{S-s}} \right).
+\end{align}
+$$
+
+### Cost of Projected Gradient Descent vs Multi-scale Proof{#sec-cost-of-gd-vs-ms-proof}
+
+Proof of @lem-cost-of-gd-vs-ms.
+
+The total cost of regular descent is given by
+$$C_{\text{GD}}=C_{1}K.$$
+For multiscale, it is
+$$
+C_{\text{MS}} = \sum_{s=1}^S C_{s}K_{s}.
+$$
+It is reasonable to assume that
+$$
+C_{s} \geq \frac{3}{2} C_{s+1}
+$$
+since one has to compute $2^{S-s+1}+1$ entries of $x_{s}$ at scale $s$. If it is a fixed cost $C$ to compute an entry of $x$, then we have
+$$
+C_{s} = C \left( 2^{S-s+1}+1 \right).
+$$
+So
+$$
+\frac{C_{s}}{C_{s+1}} = \frac{ 2^{S-s+1}+1 }{ 2^{S-(s+1)+1}+1}=\frac{ 2^{S-s+1}+1 }{ 2^{S-s}+1} \geq \frac{3}{2}
+$$
+since the function $\frac{ 2^{x+1}+1 }{ 2^{x}+1}$ (for nonnegative $x$) is minimized at $x=0$, where $x=S-s\geq {0}$.
+This gives us the chain,
+$$
+C_{1}\geq \frac{{3}}{2}C_{2}\geq \left( \frac{3}{2} \right)^2 C_{3}\geq \dots \geq \left( \frac{3}{2} \right)^{S-1} C_{S}
+$$
+or
+$$
+C_{S}\leq\left( \frac{2}{3} \right)C_{S-1}\leq\dots\leq\left( \frac{2}{3} \right)^{S-2}C_{2}\leq\left( \frac{2}{3} \right)^{S-1}C_{1}.
+$$
+
+This means
+$$
+C_{\text{MS}} = \sum_{s=1}^S C_{s}K_{s} \leq C_{1}\sum_{s=1}^S \left( \frac{2}{3} \right)^{s-1} K_{s}.
+$$
+
+### Re-solved Multi-scale Descent Cost Proof {#sec-resolved-cost-proof}
+
+Proof of @cor-resolved-cost.
+
+What we need to ensure, for multi-scale to be cheaper, with our mild assumption on cost at each scale, is
+$$
+C_{\text{MS}} = \sum_{s=1}^S C_{s}K_{s} \leq C_{1}\sum_{s=1}^S \left( \frac{2}{3} \right)^{s-1} K_{s} \leq C_{1}K.
+$$
+
+To leave the most budget left for $K_{1}$, we can make all other $K_{s}=1$ which gives
+$$
+C_{1}\sum_{s=1}^S\left( \frac{2}{3} \right)^{s-1}K_{s}=C_{1}\left( K_{1}+\sum_{s=2}^S \left( \frac{2}{3} \right)^{s-1} \right)= C_{1}\left( K_{1} + 3\left( 1-\left( \frac{2}{3} \right)^{S}\ \right)\right)
+$$
+
+And we can upper bound by considering what happens when we increase the number of scales $S\to \infty$ (as in the total number of points $I\to \infty$).
+$$
+C_{\text{MS}}< C_{1}(K_{1}+3)
+$$
+
+So interestingly, we can set $K_{1}=K-3$, and the rest of the scales $K_{s}=1$ and still be cheaper!
+
+TODO
+By hand I've shown that for this setup to give a better accuracy, it does not matter what $K$ is (as long at it is $4$ or bigger).
+There is just a minimum scale $S$ that is needed.
+And for well conditioned problems $0.3\leq c\leq 1$, a scale bigger than $4$ will do the trick.
+(Note it is not quite $0.3$, it can be made slightly lower)
+
+### Expected Freezed Multi-scale Descent Convergence Proof {#sec-expected-freezed-convergence-proof}
+
+Proof of @thm-expected-freezed-convergence.
+
+For multi-scale, we start out with only $I = 2^{S-s_0 +1} +1$ points. Taking the largest scale we can $s_0=S$, we have an initial error bound of $\sqrt{ 2+2 }=2$ giving us the expected error bound
+
+$$
+\begin{align}
+\left\lVert e_{1}^{K_{1}} \right\rVert &\leq  2 d(K)(1+d(K))^{S-1} + C \frac{d(K) \left( \sqrt{2^S} \left( d(K) + 1 \right)^S - \sqrt{2} \left( d(K) + 1 \right) \right)}{\sqrt{2^S} \left( d(K) + 1 \right) \left( \sqrt{2} d(K) + \sqrt{2} - 1 \right)}
+\end{align}
+$$
+
+for $C = \frac{L}{2}\left\lvert b-a \right\rvert$ (see @thm-expected-pgd-convergence).
+
+If we factor out a $d(K)$, we get
+$$
+\begin{align}
+\left\lVert e_{1}^{K_{1}} \right\rVert &\leq  d(K)\left( 2(1+d(K))^{S-1} + C \frac{\left( \sqrt{2^S} \left( d(K) + 1 \right)^S - \sqrt{2} \left( d(K) + 1 \right) \right)}{\sqrt{2^S} \left( d(K) + 1 \right) \left( \sqrt{2} d(K) + \sqrt{2} - 1 \right)} \right).
+\end{align}
+$$
+
+### Freezed Multi-scale Descent Cost Proof {#sec-freezed-cost-proof}
+
+Proof of @cor-freezed-cost.
+
+Recall our expected regular descent error have the bound
+$$
+\left\lVert x_{1}^{K}-x_{1}^* \right\rVert\leq \left( 1-c \right)^{K} \sqrt{ 2^S + 2 }.
+$$
+
+From @thm-expected-freezed-convergence, we have the expected error bound for freezed multi-scale
+
+$$
+\begin{align}
+\left\lVert e_{1}^{K_{1}} \right\rVert &\leq  d(K)\left( 2(1+d(K))^{S-1} + C \frac{\left( \sqrt{2^S} \left( d(K) + 1 \right)^S - \sqrt{2} \left( d(K) + 1 \right) \right)}{\sqrt{2^S} \left( d(K) + 1 \right) \left( \sqrt{2} d(K) + \sqrt{2} - 1 \right)} \right),
+\end{align}
+$$
+
+so need the thing in the bracket to be less than $\sqrt{ 2^S + 2 }$.
+
+In the limit as $K\to \infty$, we have that $d(K)\to {0}$. So for a very small $d(K)$, we get the bracket expression to be
+$$
+2 + C \frac{\left( \sqrt{2^S} - \sqrt{2} \right)}{\sqrt{2^S}  \left(\sqrt{2} - 1 \right)}.
+$$
+
+<!-- For a scale of $S=2$, we have $2+\frac{{C}}{\sqrt{ 2 }}$ which is bigger than gradient descent's $2$. -->
+
+After simplifying we get
+$$
+2 + C \frac{\left( \sqrt{2}^{S-1} - 1 \right)}{\sqrt{2}^{S-1}  \left(\sqrt{2} - 1 \right)}.
+$$
+
+<!-- For $S=3$, this is $2+\frac{C}{2(\sqrt{ 2 }+1)}$ which could be smaller than $\sqrt{ 2^3+2 }=\sqrt{ 10 }=2\sqrt{ 5 }$. And this scale shrinks with $S$, but grows for GD. -->
+
+What $S$ do we need for multi scale to be more accurate?
+$$
+\begin{align}
+2 + C \frac{\left( \sqrt{2}^{S-1} - 1 \right)}{\sqrt{2}^{S-1}  \left(\sqrt{2} - 1 \right)} &< \sqrt{ 2^S + 2 } \\
+C&<\left( \sqrt{ 2^S + 2 }-2 \right)\frac{\sqrt{2}^{S-1}  \left(\sqrt{2} - 1 \right)}{\left( \sqrt{2}^{S-1} - 1 \right)} \\
+C&<\left(\sqrt{2} - 1 \right)\left( \sqrt{ 2^S + 2 }-2 \right)\frac{\sqrt{2}^{S-1}  }{\left( \sqrt{2}^{S-1} - 1 \right)}
+\end{align}
+$$
+
+The first factor is less than one $\sqrt{2} - 1 < 1$, and the last factor is basically $1$ when $S>6$ so let's use that
+$$
+\begin{align}
+C&<\left( \sqrt{ 2^S + 2 }-2 \right) \\
+\left( \frac{C}{\left(\sqrt{2} - 1 \right)}+2 \right)^2 - 2&< 2^S \\
+\log_{2}{\left(\left( \frac{C}{\left(\sqrt{2} - 1 \right)}+2 \right)^2 - 2 \right)}&< S
+\end{align}
+$$
+
+note because of our approximation, if the above is *not* satisfied, *then* GD is more accurate. This is not an iff condition.
+
+There are a number of ways to modify it so that it becomes "if this condition holds, then multiscale is better". You could replace $C$ with $C+0.5$ as a simple one.
+
+As a simple case, if $C=\frac{1}{2}$ (the function is 1-Lipschitz on $[0,1]$), then a scale of $S=2$ or higher is already better for multi scale.
+
+The above work shows that multi-scale will give a tighter expected final error bound, but we need to make sure multi-scale is cheaper.
+
+Reusing the work shown in @sec-cost-of-gd-vs-ms-proof, if we assume each scale has the same number of iterations $K_s=K'$, then we get
+$$
+C_{\text{MS}} \leq C_{1}K' \sum_{s=1}^S \left( \frac{2}{3} \right)^{s-1} =C_{1}K'3\left( 1-\left( \frac{2}{3} \right)^S \right)\leq 3C_{1}K'.
+$$
+
+So if we make sure $K'< \frac{K}{3}$, where $K$ is the number of iterations we perform with regular projected gradient descent, then
+$$
+C_{\text{MS}}< C_{GD}.
+$$
+
+The largest integer less than $\frac{K}{3}$ would be $\lceil\frac{K}{3}\rceil -1$.
+
+TODO : Note I think the argument can be tightened to get something like $K'< \frac{K}{2+\delta}$ for a small delta, but this idea should be fine.
+
+The way the expected error upper bound works for the freezed multi-scale, the condition on the number of scales needed $S$ is enough to ensure that multi-scale gives a better error bound, assuming we take both methods to large enough iterations $K$.

### 2025-06-19 paper/paper.qmd
add hessian reference and more detailed low rank section

@@
 For a tensor input $A$ of order $N$, the Hessian tensor $\nabla^2 f(A)$ is of order $2N$.
+This contrasts the Hessian *matrix* constructed by vectorizing the function $f$'s input [@magnus_matrix_2019, Ch.
+10].
 
@@
 
-## Tensor rank
+## Tensor rank {#sec-tensor-rank}
 - tensor rank
@@
 :::
+
+By convention, we say the zero tensor $Y=0$ has rank $0$.
 
@@
 For applications where the rank is not known a priori, a common strategy is to attempt a decomposition for a variety of ranks, and select the model with smallest rank that still achieves good fit between the factorization and the original tensor.
+See section @sec-estimating-tensor-rank for an implementation of this strategy.
 
@@
     end
+end
+```
+
+### Estimating Tensor Rank {#sec-estimating-tensor-rank}
+
+In many applications, the rank of the input tensor $Y$ may not be known. In the case of CP and Tucker-1 decompositions, there are known bounds on the rank.
+
+::: {#lem-tensor-rank-bounds}
+
+Given a tensor $Y\in\mathbb{R}^{I_1\times \dots \times I_N}$, we have the following bounds on the Tucker-1 and CP rank [@kolda_TensorDecompositionsApplications_2009, Sec. 3.1].
+
+$$
+0 \leq \text{rank}_\text{CP} (Y) \leq \min_n\prod_{m\neq n} I_m = \min(I_2\cdots I_N, I_1 I_3 \cdots I_N, \dots, I_1 \cdots I_{N-1}),
+$$
+
+and
+$$
+0 \leq \text{rank}_\text{Tucker-1} (Y) \leq \min\left(I_1, \prod_{n=2}^N I_n\right) = \min(I_1, I_2\cdots I_N) .
+$$
+
+We have a rank of $0$ if and only if $Y=0$ is the zero tensor.
+:::
+
+Both of these bounds are natural extensions of the typical matrix rank bounds. For a matrix $Y\in\mathbb{R}^{I\times J}$, we know
+
+$$
+0 \leq \text{rank}(Y) \leq \min(I, J).
+$$
+
+In the case that we try to factorize a matrix $Y$ at a rank $R \geq \text{rank}(Y)$, we should still be able to achieve a final objective of zero $0=\frac{1}{2} \left\lVert A^*B^* - Y \right\rVert_F^2$, and at any rank $R < \text{rank}(Y)$, the objective is positive and decreasing.
+This fact extends to CP and Tucker-1 decompositions.
+So a naive strategy for finding the rank of a tensor $Y$, would be first check if it is rank zero, and then try factorizing $Y$ at incrementally larger ranks these two types of decompositions until the final objective hits zero.
+
+This approach has two problems.
+With real work data, the input tensor is often a noisy version of a low rank tensor $Y = Y_{\text{clean}} + Z$.
+Noise is often full rank and would cause the input tensor $Y$ to be the maximum rank [@vershynin_HighDimensionalProbability_2018].
+Secondly, even if we had a perfectly low rank tensor $Y$, our algorithm at best converges to the optimal solution in the limit.
+So any finite stopping point has a positive objective.
+
+A simple fix would be to extend the concept of numerical rank (also called $\epsilon$-rank) to tensors.[^numerical-rank-proofs] This serves as an approximation for the rank where taking $\epsilon \rightarrow 0^+$ would return the usual definition of the rank.
+
+[^numerical-rank-proofs]: See <https://mpf-optimization-laboratory.github.io/opt-blog/posts/epsilon-rank/> for an in-depth discussion on this topic. TODO move some proofs to this document.
+
+::: {#def-epsilon-rank}
+The $\epsilon$-rank [@golubRosetakDocumentRank1977] of a matrix $A$ is the smallest rank obtainable by an $\epsilon$-perturbation of the matrix $A+E$:
+$$
+\mathrm{rank}_{\epsilon}(A)=\min_{\left\lVert E \right\rVert\leq \epsilon}\mathrm{rank}(A+E).
+$$ {#eq-epsilon-rank}
+
+:::
+
+Here we use $\left\lVert E \right\rVert=\max_{\lVert v \rVert_2=1}\left\lVert Ev \right\rVert_{2}$ to denote the the operator norm.
+
+This definition is *stable* in the sense that adding a small amount of noise to a matrix does not effect the $\epsilon$-rank.
+This is unlike the traditional rank where adding noise to a low rank matrix can turn it into a full rank matrix, even for an "$\epsilon$" amount of noise.
+This is clear from @prp-singular-values which also offers a practical method for computing $\mathrm{rank}_{\epsilon}(A)$.
+
+::: {#prp-singular-values}
+# Singular Value Characterization
+
+The $\epsilon$-rank of $A$ is equal to the number of singular values of $A$ strictly bigger than $\epsilon$.
+:::
+
+The following new result, @thm-rank-recovery, shows that we can use the $\epsilon$-rank to compute the true rank of a matrix under some additive noise $B$ with the right choice of $\epsilon$.
+
+::: {#thm-rank-recovery}
+# Rank recovery
+
+Let $A,B\in\mathbb{R}^{m\times n}$ where $\left\lVert B \right\rVert_{2} < \sigma_r / 2$. Then,
+$$
+\mathrm{rank}_{\epsilon}(A+B)=\mathrm{rank}(A)=r
+$$
+
+for all $\epsilon$ between $\left\lVert B \right\rVert_{2} \leq \epsilon < \sigma_{r}/2.$ Here, we list the singular values of $A$ in non-increasing order: $\sigma_1\geq \dots \geq \sigma_r > \sigma_{r+1}=\dots=\sigma_{\min(m,n)}=0$.
+:::
+
+Note there are no assumptions on the noise $B$.
+It could be fixed, or come from any distribution.
+In the special case that you know the distribution $B$ comes from, you could use that information to estimate an appropriate choice of $\epsilon$.
+@cor-gaussian-rank-recovery gives the case where $B=Z$ is standard Gaussian noise.
+
+::: {#cor-gaussian-rank-recovery}
+# Rank Recovery with Gaussian Noise
+
+Let $Z\in\mathbb{R}^{m\times n}$ be a Gaussian matrix with standard normal entries $Z_{ij}\sim\mathcal{N}(0,1)$ and $\sigma_{r}$ be the largest nonzero singular value of $A\in\mathbb{R}^{m\times n}$.
+If $\sigma_{r}>2(\sqrt{ m }+\sqrt{ n })$, then for $(\sqrt{ m }+\sqrt{ n } +t) \leq \epsilon <\sigma_{r}/2$, $$\mathrm{rank}_{\epsilon}(A+Z)=\mathrm{rank}(A)$$ with high probability at least $1-2\exp(-ct^2)$ for some constant $0<c\in\mathbb{R}$.
+:::
+
+Extending these results to constrained tensors factorization should be possible, but there presents multiple ways we could approach the problem.
+In the case of identifying rank for a signal decomposition, we are interested the smallest sum of sources under constraints on the allowed sources.
+For nonnegative sources, the well studied nonnegative rank [@gillis_nonnegative_2020] is most appropriate:
+
+$$
+\mathrm{rank}_+(X)= \min \left\{R \in \mathbb{Z}_{+} \middle| \exists A,B\geq 0 \text{ s.t. } X[i,j] = \sum_{r=1}^R A[i,r]B[r,j]\right\}.
+$$ {#eq-nonnegative-rank}
+
+Unfortunately, computing this is NP hard in general [@gillis_nonnegative_2020; @vavasis_complexity_2010; @gillis_geometric_2012].
+This means extending the $\epsilon$-rank to something like an $\epsilon$-nonnegative-rank may not be practically computable.
+
+Even in the unconstrained case, we have only kicked the problem down the road and must estimate a suitable $\epsilon$ to extract the rank.
+
+There are a number of proposed solutions to this problem that all follow the general principle of Occam's razor of selecting the model with the smallest rank that still achieves a reasonable fit with the data.
+For automated rank selection, there are many criteria used to balance simplicity with accuracy.
+The following is a non exhaustive list of methods used in practice in roughly chronological order:
+
+- Akaike [@akaike_new_1974; @burnham_model_1998] and Bayesian [@schwarz_estimating_1978; @neath_bayesian_2012] Information Criterion
+- Numerical rank [@golubRosetakDocumentRank1977] and in combination with polynomial filtering [@ubaru_fast_2016]
+- Matrix perturbation theory [@ratsimalahelo_rank_2001]
+- Consensus clustering and cophenetic correlation coefficient [@monti_consensus_2003;@brunet_metagenes_2004]
+- Point of maximum curvature of the approximation error as a function of the rank [@satopaa_finding_2011]
+- Cross validation [@austin_tensor_2014]
+- "SCREE" plot and segmented linear regression [@saylor_CharacterizingSedimentSources_2019, Sec. 3.2]
+- Minimum description length [@fu_model_2019]
+- Concordance [@fogel_rank_2023]
+- And hypothesis testing [@cai_rank_2023].
+
+The number of ways to estimate rank---and lack of consensus on when to use each method---is best summarized by Fu et. al. [@fu_model_2019]:
+
+> "Even though the selection of rank is very important, in most studies on [non-negative tensor factorization] the value of R is simply determined by trial and error or specialists' insights, and there does not exist a good way to determine R automatically."
+
+We take an approach similar to Satopaa [@satopaa_finding_2011], Saylor [@saylor_CharacterizingSedimentSources_2019, Sec.
+3.2], and Graham [@graham_tracing_2025].
+We first factorize the tensor at every possible rank (as given by the bounds in @lem-tensor-rank-bounds), and then look at the function that takes as input a rank $r$, and output the final relative error between the reconstructed decomposition at that rank $X_r^*$ and the input tensor $Y$,
+
+$$
+f(r) = \left\lVert X_r^* - Y \right\rVert_F / \left\lVert Y \right\rVert_F.
+$$
+
+We select the "knee" or "elbow" of this function by finding the point of maximum curvature $\kappa_f(r)$ with finite differences,
+
+$$
+\hat{R} = \arg\,\max_r \kappa_f(r) := \frac{f''(r)}{(1+ (f'(x))^2)^{3/2}}.
+$$
+
+Since scaling the function $f$ would influence the curvature, we standardize the function by scaling it to the $[0, 1]^2$ box.
+The following is the specific implementation of the rank finding algorithm, and method for calculating the standard curvature.
+
+Since the maximum rank could be very large, we also include an option `online_rank_estimation` which stops looking are larger ranks if the standard curvature has noticeably dropped, implying a local maximum.
+
+```julia
+function rank_detect_factorize(Y; online_rank_estimation=false, rank=nothing, model=Tucker1, kwargs...)
+    if isnothing(rank)
+        # Initialize output and final error lists
+        all_outputs = []
+        final_rel_errors = Float64[]
+
+        # Make sure RelativeError is part of the stats keyword argument
+        kwargs = isempty(kwargs) ? Dict{Symbol,Any}() : Dict{Symbol,Any}(kwargs)
+        get!(kwargs, :stats) do # If stats is not given, populate stats with RelativeError
+            [Iteration, RelativeError, ObjectiveValue, isnonnegative(Y) ? GradientNNCone : GradientNorm]
+        end
+        if RelativeError ∉ kwargs[:stats] # If stats was given, make sure RelativeError is in the list stats
+            kwargs[:stats] = [RelativeError, kwargs[:stats]...] # not using pushfirst! since kwargs[:stats] could be a Tuple
+        end
+        kwargs[:model] = model # add the model back into kwargs
+
+        for rank in possible_ranks(Y, model)
+            @info "Trying rank=$rank..."
+
+            kwargs[:rank] = rank # add the rank into kwargs
+
+            output = factorize(Y; kwargs...) # safe to call factorize (rather than _factorize) since both factorize and rank_detect_factorize have checks to see if the keyword `rank` is provided
+            push!(all_outputs, output)
+            _, stats, _ = output
+
+            final_rel_error = stats[end, :RelativeError]
+            push!(final_rel_errors, final_rel_error)
+            @info "Final relative error = $final_rel_error"
+
+            if (online_rank_estimation == true) && length(final_rel_errors) >= 3 # Need at least 3 points to evaluate curvature
+                curvatures = standard_curvature(final_rel_errors)
+                if curvatures[end] ≈ maximum(curvatures) # want the last curvature to be significantly smaller than the max
+                    continue
+                else
+                    # we must have curvature[end] < maximum(curvature) so we can now return
+                    R = argmax(curvatures)
+                    @info "Optimal rank found: $R"
+                    return ((all_outputs[R])..., final_rel_errors)
+                end
+            end
+        end
+
+        # Return if online_rank_estimation == false, or a clear rank was not found
+        R = argmax(standard_curvature(final_rel_errors))
+        @info "Optimal rank found: $R"
+        return ((all_outputs[R])..., final_rel_errors)
+    else
+        return factorize(Y; rank, model, kwargs...)
+    end
+end
+```
+
+```julia
+"""
+Approximate first derivative with finite elements. Assumes y[i] = y(x_i) are samples with unit spaced inputs x_{i+1} - x_i = 1.
+"""
+function d_dx(y::AbstractVector{<:Real})
+    d = similar(y)
+    each_i = eachindex(y)
+
+    # centred estimate
+    for i in each_i[begin+1:end-1]
+        d[i] = (-y[i-1] + y[i+1])/2
+    end
+
+    # three point forward/backward estimate
+    i = each_i[begin+1]
+    d[begin] = (-3*y[i-1] + 4*y[i] - y[i+1])/2
+
+    i = each_i[end-1]
+    d[end] = (y[i-1] - 4*y[i] + 3*y[i+1])/2
+    return d
+end
+
+function d2_dx2(y::AbstractVector{<:Real})
+    d = similar(y)
+
+    for i in eachindex(y)[begin+1:end-1]
+        d[i] = y[i-1] - 2*y[i] + y[i+1]
+    end
+
+    # Assume the same second derivative at the end points
+    d[begin] = d[begin+1]
+    d[end] = d[end-1]
+    return d
+end
+
+function standard_curvature(y::AbstractVector{<:Real}; kwargs...)
+    # An interval 0:10 has length(0:10) = 11, but measure 10-0 = 10
+    # hence the length(y) - 1
+    Δx = 1 / (length(y) - 1)
+    y_max = maximum(y)
+    dy_dx = d_dx(y; kwargs...) / (Δx * y_max)
+    dy2_dx2 = d2_dx2(y; kwargs...) / (Δx^2 * y_max)
+    return @. dy2_dx2 / (1 + dy_dx^2)^1.5
 end

### 2025-06-19 test/geomultiscaletest/multiscaletest.jl
add more tests to multiscale and rational learning

@@ -1,0 +1,81 @@
+# Test of multiscale
+
+using BlockTensorDecomposition
+using HDF5
+using BenchmarkTools
+using Logging
+using Random
+
+Random.seed!(314)
+
+filename = "./test/geomultiscaletest/geodata.h5"
+
+myfile = h5open(filename, "r")
+
+#Y = myfile["Y_257_points"] |> read
+Y = myfile["Y_1025_points"] |> read
+
+close(myfile)
+
+size(Y)
+
+Y_backup = copy(Y)
+
+# # Add an extra slice so the continuous dimension is one plus a power of 2
+# Y = cat(Y, zeros(20, 7, 1) ;dims=3)
+
+scaleB_rescaleA! = ConstraintUpdate(0, l1scale_average12slices! ∘ nonnegative!;
+    whats_rescaled=(x -> eachcol(factor(x, 1)))
+)
+
+scaleA_rows! = ConstraintUpdate(1, l1scale_rows! ∘ nonnegative!;
+    whats_rescaled=nothing
+)
+
+nonnegativeA! = ConstraintUpdate(1, nonnegative!)
+
+nonnegativeB! = ConstraintUpdate(0, nonnegative!)
+# [l1scale_average12slices! ∘ nonnegative!, nonnegative!]
+# [scaleB_rescaleA!, scaleA_rows!]
+# [nonnegativeB!, scaleA_rows!]
+
+options = (
+    rank=3,
+    momentum=false,
+    model=Tucker1,
+    tolerance=(0.12),
+    converged=(RelativeError),
+    do_subblock_updates=false,
+    constrain_init=true,
+    constraints=[l1scale_average12slices! ∘ nonnegative!, nonnegative!],
+    stats=[Iteration, ObjectiveValue, GradientNNCone, RelativeError],
+    maxiter=200
+)
+
+#[l1scale_average12slices! ∘ nonnegative!,nonnegative!]
+
+println("First Compile Time")
+@time decomposition, stats_data, kwargs = factorize(Y; options...);
+
+@time decomposition, stats_data, kwargs = multiscale_factorize(Y;
+    continuous_dims=3,options...);
+println("")
+println("Regular Run")
+@time decomposition, stats_data, kwargs = factorize(Y; options...);
+
+@time decomposition, stats_data, kwargs = multiscale_factorize(Y;
+    continuous_dims=3,options...);
+
+println("Benchmark Run")
+
+
[…]

### 2025-06-19 test/multiscalesynthetic3d.jl

@@ -1,0 +1,259 @@
+# generate synthetic data for KDE and tensor factorization
+
+using Random
+using KernelDensity
+#using Pkg
+#Pkg.add("Distributions")
+#Pkg.add("Plots")
+#Pkg.add("PlotlyJS")
+
+using Distributions
+using BlockTensorDecomposition
+using Plots
+using PlotlyJS
+
+J = 65 # Number of samples in the x dimension
+K = 65 # Number of samples in the y dimension
+L = 65 # Number of samples in the z dimension
+
+Random.seed!(123)
+
+# Three sources, product distributions
+R = 3
+source1a = Normal(4, 1)
+source1b = Uniform(-7, 2)
+source1c = Uniform(-1, 1)
+
+source2a = Normal(0, 3)
+source2b = Uniform(-2, 2)
+source2c = Exponential(2)
+
+source3a = Exponential(1)
+source3b = Normal(0, 1)
+source3c = Normal(0, 3)
+
+source1 = product_distribution([source1a, source1b, source1c])
+source2 = product_distribution([source2a, source2b, source2c])
+source3 = product_distribution([source3a, source3b, source3c])
+
+sources = (source1, source2, source3)
+
+# x values to sample the densities at
+x = range(-10,10,length=J)
+Δx = x[2] - x[1]
+
+# y values to sample the densities at
+y = range(-10,10,length=K)
+Δy = y[2] - y[1]
+
+# z values to sample the densities at
+z = range(-10,10,length=K)
+Δz = z[2] - z[1]
+
+# Collect sample points into a matrix
+xyz = Array{Vector{Float64}, 3}(undef, length(y),length(x), length(z))
+for (j, x) in enumerate(x)
+    for (k, y) in enumerate(y)
+        for (l, z) in enumerate(z)
+            xyz[j, k, l] = [x, y, z]
+        end
+    end
+end
+
+# Plot the true sources
+XXX, YYY, ZZZ = mgrid(x, y, z)
+source1_density = pdf.((source1,), xyz)
+source2_density = pdf.((source2,), xyz)
+source3_density = pdf.((source3,), xyz);
+
+# function plot3d(pdf3d; isomin=1e-4)
+#     PlotlyJS.plot(isosurface(
+#         x=XXX[:], # isosurface wants all entries in a single list
+#         y=YYY[:],
+#         z=ZZZ[:],
[…]

### 2025-06-19 test/rationallearning.jl

@@ -1,0 +1,98 @@
+using BlockTensorDecomposition
+using Random
+using LinearAlgebra
+
+using Base: getindex, size
+using BlockTensorDecomposition: array
+
+######## SymPermTensor ###########
+
+struct SymPermTensor{T, N} <: AbstractArray{T,N}
+    nonzeroentry::T # TODO move the argument checking to inside the struct
+end
+
+"""
+    SymPermTensor(N::Int)
+
+Symmetric permutation tensor or order `n`.
+
+Entries T[i1, ..., iN] are 1/N! if (i1, ..., iN) are a permutation of (1, ..., N), and zero otherwise.
+"""
+function SymPermTensor(N)
+    isinteger(N) || throw(ArgumentError("n must be an integer, got $N"))
+    N ≥ 2 || throw(ArgumentError("n must be ≥ 2, got $N"))
+    nonzeroentry = 1/factorial(N)
+    return SymPermTensor{typeof(nonzeroentry), N}(nonzeroentry)
+end
+
+Base.size(_::SymPermTensor{T, N}) where {T, N} = Tuple(fill(N, N)) # 2×2, 3×3×3, 4×4×4×4 etc.
+Base.getindex(S::SymPermTensor, i::Int) = S[CartesianIndices(S)[i]]
+Base.getindex(S::SymPermTensor, I::CartesianIndex) = getindex(S, Tuple(I)...)
+function Base.getindex(S::SymPermTensor{T, N}, I::Vararg{Int}) where {T, N}
+    if allunique(I)
+        return S.nonzeroentry
+    else
+        return zero(T)
+    end
+end
+
+function BlockTensorDecomposition.array(S::SymPermTensor{T, N}) where {T, N}
+    nonzeroentry = S.nonzeroentry
+    output = zeros(T, size(S))
+    unique_indices = filter(I -> allunique(Tuple(I)),CartesianIndices(S))
+    output[unique_indices] .= nonzeroentry
+    return output
+end
+
+# @testset "SymPermTensor" begin
+#     S2 = SymPermTensor(2)
+#     P = [0.0 0.5; 0.5 0.0]
+#     @test S2[1] == S2[1,1] == 0
+#     @test S2[2] == S2[1,2] == 1/2
+#     @test S2[3] == S2[2,1] == 1/2
+#     @test S2[4] == S2[2,2] == 0
+#     @test S2 == P
+
+#     S3 = SymPermTensor(3)
[…]

### 2025-06-19 example/columnSimplexNMF.jl
add simplex constrained column NMF example

@@ -1,0 +1,69 @@
+using BlockTensorDecomposition
+using Random
+
+fact = BlockTensorDecomposition.factorize
+
+# Generate two matrices with columns constrained to the simplex
+I, J = 10, 20
+R = 3
+
+A = abs.(randn(I, R))
+B = abs.(randn(R, J))
+
+l1scale_cols!(A) # ensures columns sum to 1 since A is already nonnegative
+l1scale_cols!(B) # simplex_cols! can make the factors too sparse
+
+Ydecomp = Tucker1((B, A)) # This means Y = A*B
+
+@assert Ydecomp ≈ A*B
+@assert all(check.(simplex_cols!, factors(Ydecomp)))
+
+Y = array(Ydecomp) # Convert to a plain matrix type
+
+# Set options for factorization. Enter the following in the REPL for full list of options
+# julia>?BlockTensorDecomposition.Core.default_kwargs
+
+options = (
+    rank=R,
+    model=Tucker1,
+    momentum=false, # momentum causes slower iterations, but often fewer total iterations
+    do_subblock_updates=true, # uses independent stepsizes on each row, col, or fibre. Can speed up convergence but possibly unstable if an iterate is near the boundary of the constraints
+    tolerance=.01, # want less than 1% error
+    converged=RelativeError,
+    maxiter=2000,
+    stats=[
+        Iteration, ObjectiveValue, GradientNNCone, RelativeError,
+    ],
+
+    # Constrain columns of A and B to the simplex.
+    # If A has simplex constrained columns and B is only nonnegative,
+    # use [simplex_cols!, nonnegative!] or [l1scale_cols! ∘ nonnegative!, nonnegative!]
+    # Can also try [simplex_cols!, simplex_cols!] to constrain A via
+    # (Euclidean) simplex projections rather than nonnegative followed by a rescaling,
+    # but is usually slower than l1scale_cols! ∘ nonnegative!.
+
+    constraints=[l1scale_cols! ∘ nonnegative!, l1scale_cols! ∘ nonnegative!],
+    #constraints=[simplex_cols!, simplex_cols!],
+
+    # Ensure the initialization satisfies the constraints
+    constrain_init=true,
+
+    # Ensure the output satisfies the constraints given by `final_constraints`.
+    # If `final_constraints` is not given, fallback to using `constraints`.
+    # This will add one extra iteration in the stats
+    constrain_output=true,
[…]

### 2025-09-10 paper/paper.qmd
rename to BlockTensorFactorization.jl

@@
 ---
-title: "BlockTensorDecompositions.jl: A Unified Constrained Tensor Decomposition Julia Package"
+title: "BlockTensorFactorizations.jl: A Unified Constrained Tensor Decomposition Julia Package"
 subtitle: ""
@@
 Details like what rank to select, how the constraints should be enforced, and convergence criteria should be handled automatically, but customizable to the knowledgable user.
-These are the core specification for BlockTensorDecompositions.jl.
+These are the core specification for BlockTensorFactorizations.jl.
 
@@
 
-The main contribution is a description of a fast and flexible tensor decomposition package, along with a public implementation written in Julia: BlockTensorDecompositions.jl.
+The main contribution is a description of a fast and flexible tensor decomposition package, along with a public implementation written in Julia: BlockTensorFactorizations.jl.
 This package provides a framework for creating and performing custom tensor decompositions.
@@
 Natively in Julia, the 1-, 2-, and 3-fibres of a third order array `A` would be `eachslice(A, dims=(2,3))`, `eachslice(A, dims=(1,3))`, and `eachslice(A, dims=(1,2))`.
-BlockTensorDecomposition.jl defines the function `eachfibre(A; n)` to do exactly this.
+BlockTensorFactorization.jl defines the function `eachfibre(A; n)` to do exactly this.
 For example, the 1-fibres of an array `A` would be `eachfibre(A, n=1)`.
@@
 
-BlockTensorDecomposition.jl defines `identity_tensor(I, ndims)` to construct $\mathrm{id}_I$.
+BlockTensorFactorization.jl defines `identity_tensor(I, ndims)` to construct $\mathrm{id}_I$.
 
@@
 
-TODO Define in BlockTensorDecomposition.jl
+TODO Define in BlockTensorFactorization.jl
 
@@
 
-This product can also be thought of as taking the dot product $(A \cdot_n B)[i, j] = A_i \cdot B_j$ between all pairs of $n$th order slices of $A$ and $B$, which exactly how BlockTensorDecomposition.jl defines the operation.
+This product can also be thought of as taking the dot product $(A \cdot_n B)[i, j] = A_i \cdot B_j$ between all pairs of $n$th order slices of $A$ and $B$, which exactly how BlockTensorFactorization.jl defines the operation.
 
@@
 
-BlockTensorDecomposition.jl defines this operation with `slicewise_dot(A, B, n)`.
+BlockTensorFactorization.jl defines this operation with `slicewise_dot(A, B, n)`.
 In the special case where $A=B$, a more efficient method that only computes entries where $i \leq j$ is defined since $A \cdot_n A$ is a symmetric matrix.
@@
 
-BlockTensorDecomposition.jl defines this operation with `nmode_product(A, B, n)`.
+BlockTensorFactorization.jl defines this operation with `nmode_product(A, B, n)`.
 
@@
 Computationally, we can think of a generic decomposition as storing factors $(A,B,C,...)$ and operations $(\times_a,\times_b,...)$ for combining them.
-This is what we do in BlockTensorDecomposition.jl.
+This is what we do in BlockTensorFactorization.jl.
 
@@
 
-The *Tucker Product* defined by @eq-tucker-product is implemented in BlockTensorDecomposition.jl with `tuckerproduct(B, (A1, ..., AN))` and computes
+The *Tucker Product* defined by @eq-tucker-product is implemented in BlockTensorFactorization.jl with `tuckerproduct(B, (A1, ..., AN))` and computes
 $$
@@
 
-There are implemented in BlockTensorDecomposition.jl and can be called, for a third order tensor, with `Tucker((B, A₁, A₂, A₃))`, `Tucker1((B, A₁))`, and `CPDecomposition((A₁, A₂, A₃))`.
+There are implemented in BlockTensorFactorization.jl and can be called, for a third order tensor, with `Tucker((B, A₁, A₂, A₃))`, `Tucker1((B, A₁))`, and `CPDecomposition((A₁, A₂, A₃))`.
 These Julia `structs` store the tensor in its factored form.
@@
 
-To ensure the code stays flexible, the main algorithm of BlockTensorDecomposition.jl, `factorize`, is defined at a very high level.
+To ensure the code stays flexible, the main algorithm of BlockTensorFactorization.jl, `factorize`, is defined at a very high level.
 
@@
 
-The associated implementation with BlockTensorDecomposition.jl is shown below.
+The associated implementation with BlockTensorFactorization.jl is shown below.
 We define a `make_gradient` which takes the decomposition, factor index `n`, and data tensor `Y`, and creates a function that computes the gradient for the same type of decomposition.
@@
 
-Two more techniques are implemented in BlockTensorDecomposition.jl to improve convergence.
+Two more techniques are implemented in BlockTensorFactorization.jl to improve convergence.
 To the authors knowledge, these are new to tensor factorization, but may or may not be applicable depending on the exact factorization problem or data being studied.
@@
 using BenchmarkTools
-using BlockTensorDecomposition
-
-fact = BlockTensorDecomposition.factorize
+using BlockTensorFactorization
+
+fact = BlockTensorFactorization.factorize
 
@@
 
-There are a number of software engineering techniques used to ensure BlockTensorDecomposition.jl is flexible and applicable to a wide range of problems.
+There are a number of software engineering techniques used to ensure BlockTensorFactorization.jl is flexible and applicable to a wide range of problems.
 These enable key algorithmic choices to be hot-swapped and easily compared with each other.
@@
 Furthermore, depending on *why* we are decomposing a tensor, we may want different stats to be within a given a tolerance.
-BlockTensorDecomposition.jl attempts to solve this issue by defining some standard criteria that can be used to halt the algorithm.
+BlockTensorFactorization.jl attempts to solve this issue by defining some standard criteria that can be used to halt the algorithm.
 These are subtypes of the abstract type `AbstractStat` and are listed below.
@@
 
-One of the main motivations for developing BlockTensorDecomposition.jl is to solve constrained tensor problems.
+One of the main motivations for developing BlockTensorFactorization.jl is to solve constrained tensor problems.
 Other code did not have the expressivity to handle constraints beyond the most common constraints on the factors: nonnegativity and Euclidean normalized columns.
-To enable flexibility, BlockTensorDecomposition.jl defines
+To enable flexibility, BlockTensorFactorization.jl defines
 
@@
 
-These operations agree for the Frobenius norm (entry-wise $2$-norm), but are different operations in general. In BlockTensorDecomposition.jl, we define these classes as the following.
+These operations agree for the Frobenius norm (entry-wise $2$-norm), but are different operations in general. In BlockTensorFactorization.jl, we define these classes as the following.
 
@@
 
-The $\max(0, x)$ function should be understood as operating entrywise on $x$. In BlockTensorDecomposition, we use the helper `ReLU(x) = max(0, x)` for this function to assist with broadcasting.
+The $\max(0, x)$ function should be understood as operating entrywise on $x$. In BlockTensorFactorization, we use the helper `ReLU(x) = max(0, x)` for this function to assist with broadcasting.
 
@@
 
-We will refer to @eq-nnpr as nonnegative projection and rescaling (NNPR). NNPR has the following implementation in BlockTensorDecomposition.jl.
+We will refer to @eq-nnpr as nonnegative projection and rescaling (NNPR). NNPR has the following implementation in BlockTensorFactorization.jl.
 
@@
 
-This algorithm can be called in BlockTensorDecomposition.jl with the following code.
+This algorithm can be called in BlockTensorFactorization.jl with the following code.
 
@@
 
-The two algorithms can be called in BlockTensorDecomposition.jl with the following code.
+The two algorithms can be called in BlockTensorFactorization.jl with the following code.
 
@@
 
-The relevant call in BlockTensorDecomposition would be the following.
+The relevant call in BlockTensorFactorization would be the following.
 
@@
 
-and is called in BlockTensorDecomposition.jl with the following code.
+and is called in BlockTensorFactorization.jl with the following code.
 
@@
 
-The BlockTensorDecomposition.jl Julia provides a new all-in-one package for performing constrained tensor factorizations, and a playground for designing new decompositions and custom constraints.
+The BlockTensorFactorization.jl Julia provides a new all-in-one package for performing constrained tensor factorizations, and a playground for designing new decompositions and custom constraints.
 Careful design elements were engineered to balance flexibility and efficiency.

### 2025-09-10 test/opnorm/faster_opnorm.qmd
rename to BlockTensorFactorization.jl

@@
 Pkg.add(["LinearAlgebra","Random","Arpack","BenchmarkTools"])
-#Pkg.add(path="C:\\Users\\Nicholas\\OneDrive - UBC\\Research\\general-block-decomposition\\BlockTensorDecomposition.jl")
+#Pkg.add(path="C:\\Users\\Nicholas\\OneDrive - UBC\\Research\\general-block-decomposition\\BlockTensorFactorization.jl")
 ```
@@
 using Arpack: svds
-using BlockTensorDecomposition: slicewise_dot
+using BlockTensorFactorization: slicewise_dot
 using BenchmarkTools

### 2025-09-10 docs/make.jl

@@ -2,21 +2,21 @@ using Documenter
 
 #push!(LOAD_PATH,"../src/")
 
-using BlockTensorDecomposition
+using BlockTensorFactorization
 
 DocMeta.setdocmeta!(
-    BlockTensorDecomposition,
+    BlockTensorFactorization,
     :DocTestSetup,
-    :(using BlockTensorDecomposition;);
+    :(using BlockTensorFactorization;);
     recursive=true
 )
 
 makedocs(
     sitename="Block Tensor Decomposition",
-    modules = [BlockTensorDecomposition,],
+    modules = [BlockTensorFactorization,],
     checkdocs=:exports,
 )
 
 deploydocs(
-    repo = "github.com/MPF-Optimization-Laboratory/BlockTensorDecomposition.jl.git",
+    repo = "github.com/MPF-Optimization-Laboratory/BlockTensorFactorization.jl.git",
 )

### 2025-09-10 example/columnSimplexNMF.jl

@@ -1,7 +1,7 @@
-using BlockTensorDecomposition
+using BlockTensorFactorization
 using Random
 
-fact = BlockTensorDecomposition.factorize
+fact = BlockTensorFactorization.factorize
 
 # Generate two matrices with columns constrained to the simplex
 I, J = 10, 20
@@ -21,7 +21,7 @@ Ydecomp = Tucker1((B, A)) # This means Y = A*B
 Y = array(Ydecomp) # Convert to a plain matrix type
 
 # Set options for factorization. Enter the following in the REPL for full list of options
-# julia>?BlockTensorDecomposition.Core.default_kwargs
+# julia>?BlockTensorFactorization.Core.default_kwargs
 
 options = (
     rank=R,

### 2025-09-10 example/constrainedCP.jl

@@ -1,4 +1,4 @@
-using BlockTensorDecomposition
+using BlockTensorFactorization
 using Random
 using UnicodePlots
 
@@ -6,7 +6,7 @@ N = 100
 R = 5
 D = 2
 
-fact = BlockTensorDecomposition.factorize
+fact = BlockTensorFactorization.factorize
 matrices = [abs_randn(N, R) for _ in 1:D]
 l1scale_cols!.(matrices)
 Ydecomp = CPDecomposition(Tuple(matrices))#abs_randn

### 2025-09-10 example/constrainedTucker.jl

@@ -1,6 +1,6 @@
-using BlockTensorDecomposition
+using BlockTensorFactorization
 
-fact = BlockTensorDecomposition.factorize
+fact = BlockTensorFactorization.factorize
 
 G = abs_randn(2,3,4)
 A = abs_randn(10, 2)

### 2025-09-10 example/constrainedTucker1.jl

@@ -1,6 +1,6 @@
-using BlockTensorDecomposition
+using BlockTensorFactorization
 
-fact = BlockTensorDecomposition.factorize
+fact = BlockTensorFactorization.factorize
 
 C = abs_randn(5, 11, 12)
 A = abs_randn(10, 5)

### 2025-09-10 src/BlockTensorDecomposition.jl

@@ -1,4 +1,4 @@
-module BlockTensorDecomposition
+module BlockTensorFactorization
 
 include("./Core/Core.jl")
 
@@ -106,4 +106,4 @@ export DEFAULT_ALPHA, DEFAULT_N_SAMPLES
 export default_bandwidth, make_densities, make_densities2d, standardize_KDEs, standardize_2d_KDEs, filter_inner_percentile, filter_2d_inner_percentile # Functions
 export repeatcoord, kde2d, coordzip # 2d density estimation functions
 
-end # module BlockTensorDecomposition
+end # module BlockTensorFactorization

### 2025-09-10 test/factorize_initialization.jl

@@ -1,11 +1,11 @@
-using BlockTensorDecomposition
+using BlockTensorFactorization
 using Random
 
 N = 100
 R = 5
 D = 2
 
-fact = BlockTensorDecomposition.factorize
+fact = BlockTensorFactorization.factorize
 matrices = [abs_randn(N, R) for _ in 1:D]
 l1scale_cols!.(matrices)
 Ydecomp = CPDecomposition(Tuple(matrices))#abs_randn
@@ -29,9 +29,9 @@ options = (
 )
 
 # Test initialization functions of factorize
-kwargs = BlockTensorDecomposition.default_kwargs(Y; options...)
+kwargs = BlockTensorFactorization.default_kwargs(Y; options...)
 
-decomposition, kwargs = BlockTensorDecomposition.initialize_decomposition(Y; kwargs...)
-previous, updateprevious! = BlockTensorDecomposition.initialize_previous(decomposition, Y; kwargs...)
-parameters, updateparameters! = BlockTensorDecomposition.initialize_parameters(decomposition, Y, previous; kwargs...)
-update!, kwargs = BlockTensorDecomposition.make_update!(decomposition, Y; kwargs...)
+decomposition, kwargs = BlockTensorFactorization.initialize_decomposition(Y; kwargs...)
+previous, updateprevious! = BlockTensorFactorization.initialize_previous(decomposition, Y; kwargs...)
+parameters, updateparameters! = BlockTensorFactorization.initialize_parameters(decomposition, Y, previous; kwargs...)
+update!, kwargs = BlockTensorFactorization.make_update!(decomposition, Y; kwargs...)

### 2025-09-10 test/geomultiscaletest/multiscaletest.jl

@@ -1,6 +1,6 @@
 # Test of multiscale
 
-using BlockTensorDecomposition
+using BlockTensorFactorization
 using HDF5
 using BenchmarkTools
 using Logging

### 2025-09-10 test/multiscalesynthetic3d.jl

@@ -8,7 +8,7 @@ using KernelDensity
 #Pkg.add("PlotlyJS")
 
 using Distributions
-using BlockTensorDecomposition
+using BlockTensorFactorization
 using Plots
 using PlotlyJS
 
@@ -179,27 +179,27 @@ options = (
 # @show length(rel_errors)
 # @show mean_rel_error(C*F, Y; dims=1)
 
-# # BlockTensorDecomposition.Core.expand_decomposition_constraints(Y, options)[2]
-# # BlockedUpdate([BlockTensorDecomposition.Core.scale_decomposition_constraint([2,3,4], constraint, 2, Tucker1(size(Y),3)) for constraint in BlockTensorDecomposition.Core.expand_decomposition_constraints(Y, options)[2]])
+# # BlockTensorFactorization.Core.expand_decomposition_constraints(Y, options)[2]
+# # BlockedUpdate([BlockTensorFactorization.Core.scale_decomposition_constraint([2,3,4], constraint, 2, Tucker1(size(Y),3)) for constraint in BlockTensorFactorization.Core.expand_decomposition_constraints(Y, options)[2]])
 # constraints = nothing
 # kwargs=options
 # kwargs_copy = nothing
 # expanded_constraints = nothing
 # begin
-#     continuous_dims, kwargs = BlockTensorDecomposition.Core.initialize_continuous_dims(Y; kwargs...)
-#     scales, kwargs = BlockTensorDecomposition.Core.initialize_scales(Y; kwargs...)
+#     continuous_dims, kwargs = BlockTensorFactorization.Core.initialize_continuous_dims(Y; kwargs...)
+#     scales, kwargs = BlockTensorFactorization.Core.initialize_scales(Y; kwargs...)
 #     coarsest_scale, finer_scales... = scales
 
 #     # Factorize Y at the coarsest scale
-#     Yₛ = BlockTensorDecomposition.Core.coarsen(Y, coarsest_scale; dims=continuous_dims, kwargs...)
+#     Yₛ = BlockTensorFactorization.Core.coarsen(Y, coarsest_scale; dims=continuous_dims, kwargs...)
 
-#     #constraints, kwargs = BlockTensorDecomposition.Core.scale_constraints(Yₛ, coarsest_scale; continuous_dims, kwargs...)
+#     #constraints, kwargs = BlockTensorFactorization.Core.scale_constraints(Yₛ, coarsest_scale; continuous_dims, kwargs...)
 
[…]

### 2025-09-10 test/multiscalesynthetic3dconvergence.jl

@@ -8,11 +8,11 @@ using KernelDensity
 #Pkg.add("PlotlyJS")
 
 using Distributions
-using BlockTensorDecomposition
+using BlockTensorFactorization
 using Plots
 using PlotlyJS
 
-fact = BlockTensorDecomposition.factorize
+fact = BlockTensorFactorization.factorize
 
 show_plots = false
 
@@ -234,27 +234,27 @@ Plots.plot!(ts, regular_errors)
 # @show length(rel_errors)
 # @show mean_rel_error(C*F, Y; dims=1)
 
-# # BlockTensorDecomposition.Core.expand_decomposition_constraints(Y, options)[2]
-# # BlockedUpdate([BlockTensorDecomposition.Core.scale_decomposition_constraint([2,3,4], constraint, 2, Tucker1(size(Y),3)) for constraint in BlockTensorDecomposition.Core.expand_decomposition_constraints(Y, options)[2]])
+# # BlockTensorFactorization.Core.expand_decomposition_constraints(Y, options)[2]
+# # BlockedUpdate([BlockTensorFactorization.Core.scale_decomposition_constraint([2,3,4], constraint, 2, Tucker1(size(Y),3)) for constraint in BlockTensorFactorization.Core.expand_decomposition_constraints(Y, options)[2]])
 # constraints = nothing
 # kwargs=options
 # kwargs_copy = nothing
 # expanded_constraints = nothing
 # begin
-#     continuous_dims, kwargs = BlockTensorDecomposition.Core.initialize_continuous_dims(Y; kwargs...)
-#     scales, kwargs = BlockTensorDecomposition.Core.initialize_scales(Y; kwargs...)
+#     continuous_dims, kwargs = BlockTensorFactorization.Core.initialize_continuous_dims(Y; kwargs...)
+#     scales, kwargs = BlockTensorFactorization.Core.initialize_scales(Y; kwargs...)
 #     coarsest_scale, finer_scales... = scales
 
 #     # Factorize Y at the coarsest scale
-#     Yₛ = BlockTensorDecomposition.Core.coarsen(Y, coarsest_scale; dims=continuous_dims, kwargs...)
+#     Yₛ = BlockTensorFactorization.Core.coarsen(Y, coarsest_scale; dims=continuous_dims, kwargs...)
 
-#     #constraints, kwargs = BlockTensorDecomposition.Core.scale_constraints(Yₛ, coarsest_scale; continuous_dims, kwargs...)
[…]

### 2025-09-10 test/opnorm/faster_opnorm.jl

@@ -29,13 +29,13 @@
 # ```{julia}
 # using Pkg; Pkg.activate(".")
 # Pkg.add(["LinearAlgebra","Random","Arpack","BenchmarkTools"])
-# #Pkg.add(path="C:\\Users\\Nicholas\\OneDrive - UBC\\Research\\general-block-decomposition\\BlockTensorDecomposition.jl")
+# #Pkg.add(path="C:\\Users\\Nicholas\\OneDrive - UBC\\Research\\general-block-decomposition\\BlockTensorFactorization.jl")
 # ```
 
 using Random: rand, randn
 using LinearAlgebra: svdvals, opnorm, Symmetric
 using Arpack: svds
-using BlockTensorDecomposition: slicewise_dot
+using BlockTensorFactorization: slicewise_dot
 using BenchmarkTools
 
 

### 2025-09-10 test/rationallearning.jl

@@ -1,9 +1,9 @@
-using BlockTensorDecomposition
+using BlockTensorFactorization
 using Random
 using LinearAlgebra
 
 using Base: getindex, size
-using BlockTensorDecomposition: array
+using BlockTensorFactorization: array
 
 ######## SymPermTensor ###########
 
@@ -36,7 +36,7 @@ function Base.getindex(S::SymPermTensor{T, N}, I::Vararg{Int}) where {T, N}
     end
 end
 
-function BlockTensorDecomposition.array(S::SymPermTensor{T, N}) where {T, N}
+function BlockTensorFactorization.array(S::SymPermTensor{T, N}) where {T, N}
     nonzeroentry = S.nonzeroentry
     output = zeros(T, size(S))
     unique_indices = filter(I -> allunique(Tuple(I)),CartesianIndices(S))
@@ -77,7 +77,7 @@ W2 = randn(1,2)
 RD = RationalDecomp((W1, W2));
 
 # Example with a (2,2,1) network
-function BlockTensorDecomposition.array(RD::RationalDecomp)
+function BlockTensorFactorization.array(RD::RationalDecomp)
     coreRD = core(RD)
     top = tuckerproduct(coreRD, Tuple(fill(factor(RD, 1), ndims(RD))))
     bottom = tuckerproduct(coreRD, reverse(factors(RD)))

### 2025-09-10 test/rescaling_vs_projection.jl

@@ -2,14 +2,14 @@
 # 1/2 ||AB-Y||_F^2
 # where the rows of A, B, & Y all sum to 1
 
-using BlockTensorDecomposition
+using BlockTensorFactorization
 using Random
 using UnicodePlots
 using Plots
 
 Random.seed!(31415926535897)
 
-fact = BlockTensorDecomposition.factorize
+fact = BlockTensorFactorization.factorize
 
 I, J = 25, 50
 R = 5

### 2025-09-10 test/runtests.jl

@@ -7,11 +7,11 @@ using Test
 using Random
 using LinearAlgebra
 
-using BlockTensorDecomposition
+using BlockTensorFactorization
 
 const VERBOSE = true
 
-@testset verbose=VERBOSE "BlockTensorDecomposition" begin
+@testset verbose=VERBOSE "BlockTensorFactorization" begin
 
     @testset "Utils" begin
 
@@ -446,7 +446,7 @@ end
     Y = Tucker1((10,11,12), 5);
     Y = array(Y);
 
-    fact = BlockTensorDecomposition.factorize
+    fact = BlockTensorFactorization.factorize
 
     # check hitting maximum number of iterations
     decomposition, stats_data = fact(Y; rank=5, momentum=false, maxiter=2);
@@ -535,7 +535,7 @@ end
 
     core_constraint_update! = ConstraintUpdate(0, l1scale_average12slices! ∘ nonnegative!; whats_rescaled=(x -> eachcol(matrix_factor(x, 1))))
 
-    X, stats, kwargs = BlockTensorDecomposition.factorize(Y2;
+    X, stats, kwargs = BlockTensorFactorization.factorize(Y2;
         model=Tucker1, rank=R, tolerance=0.001, converged=RelativeError,
         constraints=[core_constraint_update!, ConstraintUpdate(1, nonnegative!)],
         )

### 2025-09-10 test/subblock_momentum_benchmark.jl

@@ -1,11 +1,11 @@
 using BenchmarkTools
 using Logging
-using BlockTensorDecomposition
+using BlockTensorFactorization
 
 
 global_logger(SimpleLogger(Warn))
 
-fact = BlockTensorDecomposition.factorize
+fact = BlockTensorFactorization.factorize
 
 function nnrankr_matrix((I, J), R)
     A = randn(I, R) .|> abs

### 2025-09-10 docs/make.jl
simplify to just the paper components left

@@ -1,22 +1,0 @@
-using Documenter
-
-#push!(LOAD_PATH,"../src/")
-
-using BlockTensorFactorization
-
-DocMeta.setdocmeta!(
-    BlockTensorFactorization,
-    :DocTestSetup,
-    :(using BlockTensorFactorization;);
-    recursive=true
-)
-
-makedocs(
-    sitename="Block Tensor Decomposition",
-    modules = [BlockTensorFactorization,],
-    checkdocs=:exports,
-)
-
-deploydocs(
-    repo = "github.com/MPF-Optimization-Laboratory/BlockTensorFactorization.jl.git",
-)

### 2025-09-10 example/columnSimplexNMF.jl

@@ -1,69 +1,0 @@
-using BlockTensorFactorization
-using Random
-
-fact = BlockTensorFactorization.factorize
-
-# Generate two matrices with columns constrained to the simplex
-I, J = 10, 20
-R = 3
-
-A = abs.(randn(I, R))
-B = abs.(randn(R, J))
-
-l1scale_cols!(A) # ensures columns sum to 1 since A is already nonnegative
-l1scale_cols!(B) # simplex_cols! can make the factors too sparse
-
-Ydecomp = Tucker1((B, A)) # This means Y = A*B
-
-@assert Ydecomp ≈ A*B
-@assert all(check.(simplex_cols!, factors(Ydecomp)))
-
-Y = array(Ydecomp) # Convert to a plain matrix type
-
-# Set options for factorization. Enter the following in the REPL for full list of options
-# julia>?BlockTensorFactorization.Core.default_kwargs
-
-options = (
-    rank=R,
-    model=Tucker1,
-    momentum=false, # momentum causes slower iterations, but often fewer total iterations
-    do_subblock_updates=true, # uses independent stepsizes on each row, col, or fibre. Can speed up convergence but possibly unstable if an iterate is near the boundary of the constraints
-    tolerance=.01, # want less than 1% error
-    converged=RelativeError,
-    maxiter=2000,
-    stats=[
-        Iteration, ObjectiveValue, GradientNNCone, RelativeError,
-    ],
-
-    # Constrain columns of A and B to the simplex.
-    # If A has simplex constrained columns and B is only nonnegative,
-    # use [simplex_cols!, nonnegative!] or [l1scale_cols! ∘ nonnegative!, nonnegative!]
-    # Can also try [simplex_cols!, simplex_cols!] to constrain A via
-    # (Euclidean) simplex projections rather than nonnegative followed by a rescaling,
-    # but is usually slower than l1scale_cols! ∘ nonnegative!.
-
-    constraints=[l1scale_cols! ∘ nonnegative!, l1scale_cols! ∘ nonnegative!],
-    #constraints=[simplex_cols!, simplex_cols!],
-
-    # Ensure the initialization satisfies the constraints
-    constrain_init=true,
-
-    # Ensure the output satisfies the constraints given by `final_constraints`.
-    # If `final_constraints` is not given, fallback to using `constraints`.
-    # This will add one extra iteration in the stats
-    constrain_output=true,
[…]

### 2025-09-10 example/constrainedCP.jl

@@ -1,45 +1,0 @@
-using BlockTensorFactorization
-using Random
-using UnicodePlots
-
-N = 100
-R = 5
-D = 2
-
-fact = BlockTensorFactorization.factorize
-matrices = [abs_randn(N, R) for _ in 1:D]
-l1scale_cols!.(matrices)
-Ydecomp = CPDecomposition(Tuple(matrices))#abs_randn
-@assert all(check.(l1scale_cols!, factors(Ydecomp)))
-Y = array(Ydecomp)
-
-X_absrandn = CPDecomposition(ntuple(_ -> N, D), R; init=abs_randn)
-X_randexp = CPDecomposition(ntuple(_ -> N, D), R; init=randexp)
-
-options = (
-    tolerance=.01,
-    maxiter=1000,
-    converged=RelativeError,
-    constraints=[l1scale_cols! ∘ nonnegative!, simplex_cols!],
-    constrain_init=true,
-    constrain_output=true,
-    momentum=true,
-    final_constraints = l1scale_cols!,
-    stats=[
-        Iteration, ObjectiveValue, GradientNNCone, RelativeError, FactorNorms, EuclidianLipschitz
-    ],
-)
-
-decomposition_randn, stats_data_randn, kwargs = fact(Y; decomposition=X_absrandn, options...);
-
-#display(stats_data_randn[:, :RelativeError])
-lineplot(log10.(stats_data_randn[:, :RelativeError])) |> display
-
-decomposition_randexp, stats_data_randexp, kwargs = fact(Y; decomposition=X_randexp, options...);
-#display(stats_data_randexp[:, :RelativeError])
-
-lineplot(log10.(stats_data_randexp[:, :RelativeError])) |> display
-
-@show stats_data_randn[end, :RelativeError], stats_data_randexp[end, :RelativeError]
-
-display(kwargs[:update])

### 2025-09-10 example/constrainedTucker.jl

@@ -1,30 +1,0 @@
-using BlockTensorFactorization
-
-fact = BlockTensorFactorization.factorize
-
-G = abs_randn(2,3,4)
-A = abs_randn(10, 2)
-B = abs_randn(10, 3)
-C = abs_randn(10, 4)
-Y = Tucker((G, A, B, C))
-Y = array(Y)
-
-decomposition, stats_data, kwargs = fact(Y;
-    rank=(2,3,4),
-    model=Tucker,
-    momentum=true,
-    tolerance=(1, 0.045),
-    converged=(GradientNNCone, RelativeError),
-    constrain_init=true,
-    constraints=nonnegative!,
-    stats=[Iteration, ObjectiveValue, GradientNNCone, RelativeError, EuclidianLipschitz, EuclidianStepSize]
-);
-
-display(stats_data)
-
-display(kwargs[:update])
-
-# using Pkg
-# Pkg.add("Plots")
-# using Plots
-# plot(stats_data[2:end, :EuclidianLipschitz], stats_data[2:end, :EuclidianStepSize])

### 2025-09-10 example/constrainedTucker1.jl

@@ -1,27 +1,0 @@
-using BlockTensorFactorization
-
-fact = BlockTensorFactorization.factorize
-
-C = abs_randn(5, 11, 12)
-A = abs_randn(10, 5)
-Y = Tucker1((C, A))
-Y = array(Y)
-
-decomposition, stats_data, kwargs = fact(Y;
-    rank=5,
-    momentum=true,
-    tolerance=(1, 0.03),
-    converged=(GradientNNCone, RelativeError),
-    constrain_init=true,
-    constraints=nonnegative!,
-    stats=[Iteration, ObjectiveValue, GradientNNCone, RelativeError, EuclidianLipschitz, EuclidianStepSize]
-);
-
-display(stats_data)
-
-display(kwargs[:update])
-
-# using Pkg
-# Pkg.add("Plots")
-# using Plots
-# plot(stats_data[2:end, :EuclidianLipschitz], stats_data[2:end, :EuclidianStepSize])

### 2025-09-10 src/BlockTensorDecomposition.jl

@@ -1,109 +1,0 @@
-module BlockTensorFactorization
-
-include("./Core/Core.jl")
-
-using .Core
-
-# Basic functionality
-#include("./utils.jl")
-export SuperDiagonal, abs_randn, all_recursive, eachfibre, getnotindex, geomean, identityslice, interlace, multifoldl, norm2, proj_one_hot, projsplx, proj_one_hot!, projsplx!, reshape_ndims
-
-#include("./curvaturetools.jl")
-export d_dx, d2_dx2, curvature, standard_curvature
-
-#include("./tensorproducts.jl")
-export nmp, nmode_product, mtt, slicewise_dot, tuckerproduct, cpproduct
-export ×₁, ×₂, ×₃, ×₄, ×₅, ×₆, ×₇, ×₈, ×₉
-export ⋅₁, ⋅₂, ⋅₃, ⋅₄, ⋅₅, ⋅₆, ⋅₇, ⋅₈, ⋅₉
-
-# Low level types and interface
-#include("./decomposition.jl")
-export array, contractions, core, eachfactorindex, eachrank1term, factor, factors, frozen, isfrozen, matrix_factor, matrix_factors, nfactors, rankof
-export AbstractDecomposition
-export GenericDecomposition, SingletonDecomposition
-export AbstractTucker, Tucker, Tucker1, CPDecomposition
-
-#include("./objective.jl")
-export AbstractObjective
-export L2
-
-#include("./factormatching.jl")
-export match_cols!, match_rows!, match_slices!
-export match_factors!
-
-#include("./constraint.jl")
-export AbstractConstraint
-export check
-
-export GenericConstraint
-export NoConstraint, noconstraint
-export ComposedConstraint
-
-export ProjectedNormalization
-export l2normalize!, l2normalize_rows!, l2normalize_cols!, l2normalize_1slices!, l2normalize_12slices!
-export l1normalize!, l1normalize_rows!, l1normalize_cols!, l1normalize_1slices!, l1normalize_12slices!
-export linftynormalize!, linftynormalize_rows!, linftynormalize_cols!, linftynormalize_1slices!, l1normalize_12slices!
-export simplex!, simplex_rows!, simplex_cols!, simplex_1slices!, simplex_12slices!
-
-export ScaledNormalization
-export l2scale!, l2scale_rows!, l2scale_cols!, l2scale_1slices!, l2scale_12slices!
-export l1scale!, l1scale_rows!, l1scale_cols!, l1scale_1slices!, l1scale_12slices!
[…]

### 2025-09-10 src/Core/Core.jl

@@ -1,103 +1,0 @@
-module Core
-
-# Dependencies
-using Random: randn, rand, seed!, shuffle
-using LinearAlgebra: ⋅, opnorm, Symmetric, mul!, Diagonal, I, norm, Tridiagonal
-using DataFrames: DataFrame, nrow
-
-const id = I # often use `I` for a dimension size, e.g. i ∈ 1:I, so here's a way to get the identity matrix
-
-# Method Extensions
-using Base: copy, deepcopy, eltype, filter, firstindex, getindex, getproperty, iterate, keys, lastindex, length, show, size, ndims
-using Base: +, -, *, /, \ # AbstractDecomposition methods
-using Base: ^, √, min # DiagonalTuple method
-using Base: ∘, convert # AbstractConstraint methods
-using LinearAlgebra: LinearAlgebra, diag
-
-# Basic functionality
-include("./utils.jl")
-export SuperDiagonal, abs_randn, all_recursive, eachfibre, getnotindex, geomean, identityslice, interlace, multifoldl, norm2, proj_one_hot, projsplx, proj_one_hot!, projsplx!, reshape_ndims
-
-include("./curvaturetools.jl")
-export d_dx, d2_dx2, curvature, standard_curvature
-
-include("./tensorproducts.jl")
-export nmp, nmode_product, mtt, slicewise_dot, tuckerproduct, cpproduct
-export ×₁, ×₂, ×₃, ×₄, ×₅, ×₆, ×₇, ×₈, ×₉
-export ⋅₁, ⋅₂, ⋅₃, ⋅₄, ⋅₅, ⋅₆, ⋅₇, ⋅₈, ⋅₉
-
-# Low level types and interface
-include("./decomposition.jl")
-export array, contractions, core, eachfactorindex, eachrank1term, factor, factors, frozen, isfrozen, matrix_factor, matrix_factors, nfactors, rankof
-export AbstractDecomposition
-export GenericDecomposition, SingletonDecomposition
-export AbstractTucker, Tucker, Tucker1, CPDecomposition
-
-include("./objective.jl")
-export AbstractObjective
-export L2
-
-include("./factormatching.jl")
-export match_cols!, match_rows!, match_slices!
-export match_factors!
-
-include("./constraint.jl")
-export AbstractConstraint
-export check
-
-export GenericConstraint
-export NoConstraint, noconstraint
-export ComposedConstraint
-
-export ProjectedNormalization
[…]

### 2025-09-10 src/Core/blockupdates.jl

@@ -1,831 +1,0 @@
-"""
-Mid level code that combines constraints with block updates to be used on an AbstractDecomposition
-"""
-
-
-"""
-Interface to make a step scheme is
-
-struct MyStep <: AbstractStep
-    ...
-end
-
-function (step::MyStep)(x::AbstractDecomposition; kwargs...)
-    ...
-    return step::Real
-end
-
-To use your scheme, construct an instance with any necessary parameters
-
-mystep = MyStep(...)
-
-and then you can call
-
-step = mystep(D; kwargs...)
-
-to compute the step size.
-"""
-abstract type AbstractStep <: Function end
-
-struct LipschitzStep <: AbstractStep
-    lipschitz::Function
-end
-
-function (step::LipschitzStep)(x; kwargs...)
-    L = step.lipschitz(x)
-    try
-        return L^(-1)  # allow for Lipschitz to be a diagonal matrix
-    catch
-        @warn "Could not invert the Lipschitz constant to get a stepsize. Ignoring zero coordinates."
-
-        return _safe_invert.(L)
-    end
-end
-
-_safe_invert(x) = iszero(x) ? x : x^(-1)
-
-function (step::LipschitzStep)(x::Tucker; kwargs...)
-    L = step.lipschitz(x)
-    if typeof(L) <: Tuple # Currently the only case is when we are updating the core of a Tucker factorization
-                          # Using this condition as a way to tell if it is the core we are calculating the constant for
-        return map(X -> X^(-1), L)
-    else
-        return L^(-1) # allow for Lipschitz to be a diagonal matrix
-    end
-end
-#LipschitzStep(L::Real) = 1/L
-
-# TODO have these be functions that act on decompositions more generally
-
-function make_lipschitz(T::Tucker1, n::Integer, Y::AbstractArray; objective::L2, kwargs...)
-    if n==0 # the core is the zeroth factor
-        function lipschitz0(T::Tucker1; kwargs...)
-            A = matrix_factor(T, 1)
-            return opnorm(A'A)
-        end
-        return lipschitz0
-
-    elseif n==1 # the matrix is the zeroth factor
-        function lipschitz1(T::Tucker1; kwargs...)
-            C = core(T)
-            return opnorm(slicewise_dot(C, C))
-        end
-        return lipschitz1
-
-    else
-        error("No $(n)th factor in Tucker1")
-    end
-end
-
[…]

### 2025-09-10 src/Core/constraint.jl

@@ -1,388 +1,0 @@
-"""
-Low level code for the types of constraints
-"""
-
-"""Abstract parent type for the various constraints"""
-abstract type AbstractConstraint <: Function end
-
-function Base.show(io::IO, X::AbstractConstraint)
-    print(io, typeof(X), "(", join((getfield(X, p) for p in propertynames(X)), ", "), ")")
-    return
-end
-
-function Base.show(io::IO, ::MIME"text/plain", X::AbstractConstraint)
-    println(io, typeof(X), "(")
-    for p in propertynames(X)
-        println(io, "  ", p, " = ", getfield(X, p))
-    end
-    print(io, ")")
-end
-
-"""
-    check(C::AbstractConstraint, A::AbstractArray)::Bool
-
-Returns `true` if `A` satisfies the constraint `C`.
-"""
-function check(C::AbstractConstraint, _::AbstractArray)
-    throw("checking constraints with type $(typeof(C)) is not implemented")
-end
-
-"""
-    GenericConstraint <: AbstractConstraint
-
-General constraint. Simply applies the function `apply` and checks it was successful with `check`.
-
-Calling a `GenericConstraint` on an `AbstractArray` will use the function in the
-field `apply`. Use `check(C::GenericConstraint, A)` to use the function in the field `check`.
-"""
-struct GenericConstraint <: AbstractConstraint
-    apply::Function # input a AbstractArray -> mutate it so that `check` would return true
-    check::Function
-end
-
-function (C::GenericConstraint)(D::AbstractArray)
-    (C.apply)(D)
-end
-
-check(C::GenericConstraint, A::AbstractArray) = (C.check)(A)
-
-"""
-    NoConstraint() <: AbstractConstraint
-
-The constraint that does nothing. Useful for giving a list of AbstractConstraint for each factor
-where you would like one factor to be unconstrained.
-"""
-struct NoConstraint <: AbstractConstraint end
-
-check(::NoConstraint, ::AbstractArray) = true
-(::NoConstraint)(A::AbstractArray) = A
-
-const noconstraint = NoConstraint()
-
-"""
-    ComposedConstraint{T<:AbstractConstraint, U<:AbstractConstraint}
-    outer_constraint ∘ inner_constraint
-
-Composing any two `AbstractConstraint`s with `∘` will return this type.
-
[…]

### 2025-09-10 src/Core/curvaturetools.jl

@@ -1,362 +1,0 @@
-"""Short helpers and operations related to finite differences and curavture"""
-
-# """
-#     d2_dx2(y::AbstractVector{<:Real}; order::Integer=length(y))
-
-# Approximates the 2nd derivative of a function using only given samples y of that function.
-
-# Assumes y came from f(x) where x was an evenly sampled, unit intervel grid.
-# Note the approximation uses centered three point finite differences for the
-# next-to-end points, and foward/backward three point differences for the begining/end points
-# respectively. The remaining interior points use five point differences.
-
-# Will use the largest order method possible by defult (currently 5 points), but can force
-# a specific order method with the keyword `order`.
-# See [`d_dx`](@ref).
-# """
-# function d2_dx2(y::AbstractVector{<:Real}; order::Integer=length(y))
-#     n = length(y)
-#     if n < 3
-#         throw(ArgumentError("input $y must have at least length 3"))
-#     elseif order <= 2
-#         throw(ArgumentError("order $order must be at least 3"))
-#     end
-
-#     if order == 3
-#         return _d2_dx2_3(y)
-#     elseif order == 4
-#         return _d2_dx2_4(y)
-#     elseif order >= 5
-#         return _d2_dx2_5(y)
-#     else
-#         throw(ErrorException("Something went wrong with the order $order."))
-#     end
-# end
-# # TODO is there a package that does this? The ones I've seen require the forward function.
-
-# function _d2_dx2_3(y::AbstractVector{<:Real})
-#     d = similar(y)
-#     for i in eachindex(y)[begin+1:end-1]
-#         d[i] = y[i-1] - 2*y[i] + y[i+1]
-#     end
-#     # Assume the same curvature at the end points
-#     d[begin] = d[begin+1]
-#     d[end] = d[end-1]
-#     return d
-# end
-
-# function _d2_dx2_4(y::AbstractVector{<:Real})
-#     d = similar(y)
-#     each_i = eachindex(y)
-
[…]

### 2025-09-10 src/Core/decomposition.jl

@@ -1,396 +1,0 @@
-"""
-Low level code defining the verious decomposition types like Tucker and CP
-"""
-
-"""
-Abstract type for the different decompositions.
-
-Main interface for subtypes are the following functions.
-Required:
-    `array(D)`: how to construct the full array from the decomposition representation
-    `factors(D)`: a tuple of arrays, the decomposed factors
-Optional:
-    `getindex(D, i::Int)` and `getindex(D, I::Vararg{Int, N})`: how to get the ith or Ith
-        element of the reconstructed array. Defaults to `getindex(array(D), x)`, but there
-        is often a more efficient way to get a specific element from large tensors.
-    `size(D)`: Defaults to `size(array(D))`.
-"""
-abstract type AbstractDecomposition{T, N} <: AbstractArray{T, N} end
-
-# Fallback interface for any AbstractDecomposition to behave like an AbstractArray
-Base.size(D::AbstractDecomposition) = size(array(D))
-Base.getindex(D::AbstractDecomposition, i::Int) = D[CartesianIndices(D)[i]] # usually more efficient to compute a single entry given the cartesian index rather than linear index
-Base.getindex(D::AbstractDecomposition, I::CartesianIndex) = getindex(D, Tuple(I)...)
-Base.getindex(D::AbstractDecomposition, I::Vararg{Int}) = getindex(array(D), I...)
-
-# copy/deepcopy all factors and any other properties before reconstructing
-Base.copy(D::AbstractDecomposition) = typeof(D)((copy.(getfield(D, p)) for p in propertynames(D))...)
-Base.deepcopy(D::AbstractDecomposition) = typeof(D)((deepcopy.(getfield(D, p)) for p in propertynames(D))...)
-
-#Mathematical Operations
-Base.:+(A::AbstractDecomposition, B::AbstractArray) = array(A) + B
-Base.:+(A::AbstractArray, B::AbstractDecomposition) = A + array(B)
-Base.:+(A::AbstractDecomposition, B::AbstractDecomposition) = array(A) + array(B)
-
-Base.:-(A::AbstractDecomposition, B::AbstractArray) = array(A) - B
-Base.:-(A::AbstractArray, B::AbstractDecomposition) = A - array(B)
-Base.:-(A::AbstractDecomposition, B::AbstractDecomposition) = array(A) - array(B)
[…]

### 2025-09-10 src/Core/factorize.jl

@@ -1,434 +1,0 @@
-"""
-	factorize(Y; rank=nothing, model=Tucker1, kwargs...)
-
-Factorizes `Y` according to the decomposition `model`.
-
-See [`default_kwargs`](@ref) for the default keywords.
-"""
-function factorize(Y; kwargs...)
-	kwargs = isempty(kwargs) ? Dict{Symbol,Any}() : Dict{Symbol,Any}(kwargs)
-
-	# Auto detect rank when an initial decomposition and rank are not given
-	# Do not mutate kwargs as to not confuse rank_detect_factorize
-	decomposition = get(kwargs, :decomposition, nothing)
-	rank = get(kwargs, :rank, nothing)
-	if isnothing(decomposition) && isnothing(rank)
-		return rank_detect_factorize(Y; kwargs...)
-	end
-
-	# Auto factorize at multiple scales when continuous_dims is given
-	# Do not mutate kwargs as to not confuse multiscale_factorize
-	# We differentiate
-	# continuous_dims==nothing: do not use multiscale_factorize
-	# continuous_dims==missing: guess if there are continuous dims, and use multiscale_factorize if any are found TODO implement this
-	continuous_dims = get(kwargs, :continuous_dims, missing)
-	if !ismissing(continuous_dims) && !isnothing(continuous_dims)
-		multiscale_factorize(Y; kwargs...)
-	end
-
-	return _factorize(Y; (default_kwargs(Y; kwargs...))...)
-end
-
-"""
-Inner level function once keyword arguments are set
-"""
-function _factorize(Y; kwargs...)
-    decomposition, previous, updateprevious!, parameters, updateparameters!,
-	update!, stats_data, getstats, converged, kwargs = initialize(Y, kwargs)
-
-    while !converged(stats_data; kwargs...)
-		# Usually one cycle of updates through each factor in the decomposition
-        update!(decomposition; parameters...)
-
-		# This could be the next stepsize or other info used by update!
-        updateparameters!(parameters, decomposition, previous)
-
-        push!(stats_data,
-            getstats(decomposition, Y, previous, parameters, stats_data))
-
-		# Update one or two previous iterates. For example, used for momentum
-        updateprevious!(previous, parameters, decomposition)
-    end
-
-    kwargs = postprocess!(decomposition, Y, previous, parameters, stats_data, updateparameters!, getstats, kwargs)
-
-    return decomposition, stats_data, kwargs
-end
-
-"""
[…]

### 2025-09-10 src/Core/factormatching.jl

@@ -1,118 +1,0 @@
-"""
-    match_slices!(X, Y; dims, dist=L2)
-
-Reorders the order-`dims` slices of X to best match the slices of Y.
-"""
-function match_slices!(X::AbstractArray, Y::AbstractArray; dims, kwargs...)
-    ordering = _find_ordering(X, Y; dims, kwargs...)
-
-    # Permute slice of X
-    reslice = Any[(:) for _ in 1:ndims(X)]
-    reslice[dims] = ordering
-    X .= @view X[reslice...]
-
-    return ordering
-end
-
-function _find_ordering(X, Y; dims, dist=L2())
-    size(X) == size(Y) || ArgumentError("Shape of X, $(size(X)) does not match shape of Y, $(size(Y))")
-
-    n_slices = size(X, dims)
-    ordering = zeros(Int, n_slices)
-
-    # for every slice of Y, find the best matching slice in X
-    for (j, Yj) ∈ enumerate(eachslice(Y; dims))
-        _, j_match = findmin(Xj -> dist(Xj, Yj), eachslice(X; dims))
-        ordering[j] = j_match
-    end
-
-    if !allunique(ordering)
-        # Global optimal order would involve checking factorial(n_slices) options
-        # Use a greedy approach instead
-        @warn "`match_slices!` may not have found the best ordering; using O(n²) greedy approach."
-
-        dist_matrix = zeros(n_slices, n_slices)
-        for (j, Yj) ∈ enumerate(eachslice(Y; dims))
-            for (i, Xi) ∈ enumerate(eachslice(X; dims))
-                dist_matrix[i, j] = dist(Xi, Yj)
-            end
-        end
-
-        # Find the closest slices (Xi, Yj), remove the from the options, and repeat.
-        # We don't saved a sliced dist_matrix since that would renumber the rows/cols.
-        # Instead fill the row/col with Inf so argmin can no longer pick that row/col.
-        pairings = zeros(CartesianIndex{2}, n_slices)
-        for n in 1:n_slices
-            pairing = argmin(dist_matrix) # CartesianIndex(i, j)
-            dist_matrix[pairing[1], :] .= Inf # remove row i
-            dist_matrix[:, pairing[2]] .= Inf # remove col j
-            pairings[n] = pairing
-        end
-
-        sort!(pairings) # Sorts by second index (Yj)
-
[…]

### 2025-09-10 src/Core/multiscale.jl

@@ -1,484 +1,0 @@
-# Functionality for performing factorize at multiple scales
-# This is suitable for tensors that are discretizations of continuous data
-# TODO Finish file
-
-"""
-    coarsen(Y::AbstractArray, scale::Integer; dims=1:ndims(Y))
-
-Coarsens or downsamples `Y` by `scale`. Only keeps every `scale` entries along the dimensions specified.
-
-Example
-=======
-
-Y = randn(12, 12, 12)
-
-coarsen(Y, 2) == Y[begin:2:end, begin:2:end, begin:2:end]
-
-coarsen(Y, 4; dims=(1, 3)) == Y[begin:4:end, :, begin:4:end]
-
-coarsen(Y, 3; dims=2) == Y[:, begin:3:end, :]
-"""
-coarsen(Y::AbstractArray, scale::Integer; dims=1:ndims(Y), kwargs...) =
-    Y[(d in dims ? axis[begin:scale:end] : axis for (d, axis) in enumerate(axes(Y)))...]
-
-# Using axis[begin:scale:end] rather than 1:scale:size(Y, d) for more flexible indexing
-
-"""
-    interpolate(Y, scale; dims=1:ndims(Y), degree=0, kwargs...)
-
-Interpolates Y to a larger array with repeated values.
-
-Keywords
-========
-`scale`. How much to scale up the size of `Y`.
-A dimension with size `k` will be scaled to `scale*k - (scale - 1) = scale*(k-1) + 1`
-
-`dims`:`1:ndims(Y)`. Which dimensions to interpolate.
-
-`degree`:`0`. What degree of interpolation to use. `0` is constant interpolation, `1` is linear.
-
-Like the opposite of [`coarsen`](@ref).
-
-Example
-=======
-
-julia> Y = collect(reshape(1:6, 2, 3))
-2×3 Matrix{Int64}:
- 1  3  5
- 2  4  6
-
-julia> interpolate(Y, 2)
-3×5 Matrix{Int64}:
- 1  1  3  3  5
- 1  1  3  3  5
- 2  2  4  4  6
-
-julia> interpolate(Y, 3; dims=2)
-2×7 Matrix{Int64}:
- 1  1  1  3  3  3  5
- 2  2  2  4  4  4  6
-
-julia> interpolate(Y, 1) == Y
-true
-"""
-function interpolate(Y::AbstractArray, scale; dims=1:ndims(Y), degree=0, kwargs...)
-    # Quick exit if no interpolation is needed
[…]

### 2025-09-10 src/Core/objective.jl

@@ -1,20 +1,0 @@
-"""
-AbstractObjective <: Function
-
-General interface is
-
-struct L2 <: AbstractObjective end
-
-after constructing
-
-myobjective = L2()
-
-you can call
-
-myobjective(X, Y)
-"""
-abstract type AbstractObjective <: Function end
-
-struct L2 <: AbstractObjective end
-
-(O::L2)(X, Y) = norm2(X - Y)

### 2025-09-10 src/Core/rankdetection.jl

@@ -1,86 +1,0 @@
-"""
-    rank_detect_factorize(Y; online_rank_estimation=false, rank=nothing, model=Tucker1, kwargs...)
-
-Wraps `factorize()` with rank detection.
-
-Selects the rank that maximizes the standard curvature of the Relative Error (as a function of rank).
-"""
-function rank_detect_factorize(Y; online_rank_estimation=false, rank=nothing, model=Tucker1, kwargs...)
-    if isnothing(rank)
-        # Initialize output and final error lists
-        all_outputs = []
-        final_rel_errors = Float64[]
-
-        # Make sure RelativeError is part of the stats keyword argument
-        kwargs = isempty(kwargs) ? Dict{Symbol,Any}() : Dict{Symbol,Any}(kwargs)
-        get!(kwargs, :stats) do # If stats is not given, populate stats with RelativeError
-            [Iteration, RelativeError, ObjectiveValue, isnonnegative(Y) ? GradientNNCone : GradientNorm]
-        end
-        if RelativeError ∉ kwargs[:stats] # If stats was given, make sure RelativeError is in the list stats
-            kwargs[:stats] = [RelativeError, kwargs[:stats]...] # not using pushfirst! since kwargs[:stats] could be a Tuple
-        end
-        kwargs[:model] = model # add the model back into kwargs
-
-        for rank in possible_ranks(Y, model)
-            @info "Trying rank=$rank..."
-
-            kwargs[:rank] = rank # add the rank into kwargs
-
-            output = factorize(Y; kwargs...) # safe to call factorize (rather than _factorize) since both factorize and rank_detect_factorize have checks to see if the keyword `rank` is provided
-            push!(all_outputs, output)
-            _, stats, _ = output
-
-            final_rel_error = stats[end, :RelativeError]
-            push!(final_rel_errors, final_rel_error)
-            @info "Final relative error = $final_rel_error"
-
-            if (online_rank_estimation == true) && length(final_rel_errors) >= 3 # Need at least 3 points to evaluate curvature
-                curvatures = standard_curvature(final_rel_errors; method=:finite_differences) # method=:splines
-                if curvatures[end] ≈ maximum(curvatures) # want the last curvature to be significantly smaller than the max
-                    continue
-                else
-                    # we must have curvature[end] < maximum(curvature) so we can now return
-                    R = argmax(curvatures)
[…]

### 2025-09-10 src/Core/stats.jl

@@ -1,160 +1,0 @@
-"""
-An AbstractStat is a type which, when created, can be applied to the four arguments
-(X::AbstractDecomposition, Y::AbstractArray, previous::Vector{<:AbstractDecomposition}, parameters::Dict)
-to (usually) return a number.
-"""
-abstract type AbstractStat <: Function end
-
-# These methods allow us to check length([GradientNNCone, ObjectiveValue]) or length(GradientNNCone)
-# or iterate over a list / single element of AbstractStat
-Base.length(_::Type{<:AbstractStat}) = 1
-Base.iterate(x::Type{<:AbstractStat}, state=1) = state > 1 ? nothing : (x, state+1)
-#Base.eltype(x::Type{<:AbstractStat}) = typeof(x)
-
-struct Iteration <: AbstractStat
-    function Iteration(; kwargs...) # must define it this way so the constructor can take (and ignore) kwargs
-        new()
-    end
-end
-# This does not work since the pattern (; kwargs...) counts as the same input as the empty ()
-# so Julia thinks you are redefining Iteration() = Iteration() in a circular manner
-# Iteration(; kwargs...) = Iteration()
-
-struct GradientNorm{T} <: AbstractStat
-    gradients::T
-    function GradientNorm{T}(gradients) where T
-        @assert eltype(gradients) <: Function
-        new{T}(gradients)
-    end
-end
-GradientNorm(; gradients, kwargs...) = GradientNorm{typeof(gradients)}(gradients)
-
-struct GradientNNCone{T} <: AbstractStat
-    gradients::T
-    function GradientNNCone{T}(gradients) where T
-        @assert eltype(gradients) <: Function
-        new{T}(gradients)
-    end
-end
-GradientNNCone(; gradients, kwargs...) = GradientNNCone{typeof(gradients)}(gradients)
-
-struct ObjectiveValue{T<:AbstractObjective} <: AbstractStat
-    objective::T
-end
-ObjectiveValue(; objective, kwargs...) = ObjectiveValue(objective)
-
-struct ObjectiveRatio{T<:AbstractObjective} <: AbstractStat
-    objective::T
-end
-ObjectiveRatio(; objective, kwargs...) = ObjectiveRatio(objective)
-
-struct RelativeError{T<:Function} <: AbstractStat
-    norm::T
-end
[…]

### 2025-09-10 src/Core/tensorproducts.jl

@@ -1,202 +1,0 @@
-"""
-Different tensor products. Note these do not use Einsum/Tullio or any entrywise notation
-since these are defined generaly for arbitrarily ordered tensors/arrays.
-"""
-
-"""
-    nmode_product(A::AbstractArray, B::AbstractMatrix, n::Integer)
-
-This is defined for arbitrary n, which cannot be done using Einsum/Tullio since they can
-only handle fixed ordered-tensors.
-"""
-function nmode_product(A::AbstractArray, B::AbstractMatrix, n::Integer)
-    Aperm = swapdims(A, n)
-    Cperm = Aperm ×₁ B # convert the problem to the mode-1 product
-    return swapdims(Cperm, n) # swap back
-    #Cmat = B * mat(A, n)
-    #sizeA = size(A)
-    #sizeC = (sizeA[begin:n-1]..., size(B, 2), sizeA[n+1:end]...)
-    #return imat(Cmat, n, sizeC)
-end
-
-function nmode_product(A::AbstractArray, b::AbstractVector, n::Integer)
-    Aperm = swapdims(A, n)
-    Cperm = Aperm ×₁ b # convert the problem to the mode-1 product
-    return Cperm # no need to swap since the first dimention is dropped
-end
-
-const nmp = nmode_product # Short-hand alias
-
-"""1-mode product between a tensor and a matrix"""
-×₁(A::AbstractArray, B::AbstractMatrix) = mtt(B, A)
-×₁(A::AbstractArray, b::AbstractVector) = dropdims(mtt(b', A); dims=1)
-
-"""
-Matrix times Tensor
-
-Looks like C[i1, i2, ..., iN] = sum_r A[i1, r] * B[r, i2, ..., iN] entry-wise.
-"""
-function mtt(A::AbstractMatrix, B::AbstractArray)
-    sizeB = size(B)
-    Bmat = reshape(B, sizeB[1], :)
-
-    #Cmat = A * Bmat
-    #C = reshape(Cmat, size(A, 1), sizeB[2:end]...)
-
-    # Slightly faster implimentation
-    C = zeros(size(A, 1), sizeB[2:end]...)
-    Cmat = reshape(C, size(A, 1), prod(sizeB[2:end]))
-    mul!(Cmat, A, Bmat)
-
-    return C
-end
-
-# Additional n-mode products, generated programmatically
-
-# products = (×₂, ×₃, ×₄, ×₅, ×₆, ×₇, ×₈, ×₉)
[…]

### 2025-09-10 src/Core/utils.jl

@@ -1,421 +1,0 @@
-"""
-Utility functions used throughout the package that don't fit anywhere else
-"""
-
-"""
-    SuperDiagonal{T, N} <: AbstractArray{T,N}
-
-Array of order N that is zero everywhere except possibly along the super diagonal.
-"""
-struct SuperDiagonal{T, N, V<:AbstractVector{T}} <: AbstractArray{T,N}
-    diag::V
-
-    function SuperDiagonal{T, N, V}(diag) where {T, N, V<:AbstractVector{T}}
-        Base.require_one_based_indexing(diag)
-        new{T, N, V}(diag)
-    end
-end
-
-# SuperDiagonal Interface
-SuperDiagonal(v::AbstractVector, ndims::Integer=2) = SuperDiagonal{eltype(v), ndims, typeof(v)}(v)
-SuperDiagonal{T}(v::AbstractVector, ndims::Integer=2) where {T} = SuperDiagonal(convert(AbstractVector{T}, v)::AbstractVector{T}, ndims)
-
-LinearAlgebra.diag(S::SuperDiagonal) = S.diag
-function array(S::SuperDiagonal)
-    A = zeros(eltype(S), size(S))
-    A[_diagonal_indexes(S)] .= diag(S)
-    return A
-end
-_diagonal_indexes(S::SuperDiagonal) = CartesianIndex.(fill(1:length(diag(S)), ndims(S))...)
-
-# AbstractArray interface
-#Base.ndims(_::SuperDiagonal{T,N}) where {T,N} = N
-
-function Base.size(S::SuperDiagonal{T,N}) where {T, N}
-    n = length(diag(S))
-    # Why Val(N) and not just N?
-    # See https://docs.julialang.org/en/v1/manual/performance-tips/#man-performance-value-type
-    return ntuple(i -> n, Val(N)) #Tuple(n for _ in 1:ndims(S))
-end
-
-Base.getindex(S::SuperDiagonal, i::Int) = getindex(array(S), i) # TODO work out this efficiently
-function Base.getindex(S::SuperDiagonal, I::Vararg{Int})
-    # Check if index is valid
-    min, max = extrema(I)
-    n = length(diag(S))
-
-    if min < 1 || n < max
-        throw(BoundsError("attempt to access $(size(S)) SuperDiagonal at index $I"))
-    elseif allequal(I)
-        return getindex(diag(S), I[begin])
-    else
-        return zero(eltype(S))
-    end
-end
-
-"""
[…]

### 2025-09-10 src/DensityEstimationTools/DensityEstimationTools.jl

@@ -1,13 +1,0 @@
-module DensityEstimationTools
-
-using Statistics: mean, median, quantile, std
-using KernelDensity
-
-export DEFAULT_ALPHA, DEFAULT_N_SAMPLES
-export default_bandwidth, make_densities, make_densities2d, standardize_KDEs, standardize_2d_KDEs, filter_inner_percentile, filter_2d_inner_percentile # Functions
-export repeatcoord, kde2d, coordzip # 2d density estimation functions
-
-include("densityestimation.jl")
-include("densityestimation2d.jl")
-
-end # module DensityEstimationTools

### 2025-09-10 src/MatrixTensorFactor/MatrixTensorFactor.jl

@@ -1,28 +1,0 @@
-"""
-Matrix-Tensor Factorization
-"""
-module MatrixTensorFactor
-
-using Random: randn
-
-# include("../Core/Core.jl")
-using ..Core # note the two dots .. since the module Core is not in the same folder as MatrixTensorFactor
-
-# include("../Core/utils.jl")
-# include("../Core/tensorproducts.jl")
-# include("../Core/decomposition.jl")
-# include("../Core/objective.jl")
-# include("../Core/constraint.jl")
-# include("../Core/stats.jl")
-# include("../Core/blockupdates.jl")
-# include("../Core/factorize.jl")
-
-export nnmtf, nnmtf_proxgrad_online # Functions
-
-export IMPLIMENTED_OPTIONS, IMPLIMENTED_NORMALIZATIONS, IMPLIMENTED_METRICS
-export IMPLIMENTED_PROJECTIONS, IMPLIMENTED_CRITERIA, IMPLIMENTED_STEPSIZES # implimented options
-
-#include("curvaturetools.jl")
-include("nnmtf.jl")
-
-end

### 2025-09-10 src/MatrixTensorFactor/nnmtf.jl

@@ -1,497 +1,0 @@
-# Code for using the the new factorize internals with the old interface
-#
-# Wraps the new `factorize` function with nnmtf so that old code doesn't break,
-# but can still benifit from the newer and more optimized internals.
-
-# nnmtf can still call _nnmtf_proxgrad as before
-# We just need to change the internals
-
-#---------------------------- data ------------------------------------#
-struct UnimplimentedError <: Exception end
-
-"""
-    IMPLIMENTED_NORMALIZATIONS::Set{Symbol}
-
-- `:fibres`: set ``\\sum_{k=1}^K B[r,j,k] = 1`` for all ``r, j``, or when `projection==:nnscale`,
-    set ``\\sum_{j=1}^J\\sum_{k=1}^K B[r,j,k] = J`` for all ``r``
-- `:slices`: set ``\\sum_{j=1}^J\\sum_{k=1}^K B[r,j,k] = 1`` for all ``r``
-- `:nothing`: does not enforce any normalization of `B`
-- `:rows`: set a metric on eachslice(X; dims=1) to 1, which is equivelent to eachrow(X) when X is a matrix
-- `:cols`: set a metric on eachslice(X; dims=2) to 1, which is equivelent to eachcol(X) when X is a matrix
-"""
-const IMPLIMENTED_NORMALIZATIONS = Set{Symbol}((:fibres, :slices, :rows, :cols, :nothing))
-
-"""
-    IMPLIMENTED_METRICS::Set{Symbol}
-
-- `:L1`: the default; ensure sums of entries in each fibre or slice (according to `normalize`)
-    are equal to 1
-- TODO `:L2`: ensure the sum of squares of entries are 1
-- `:Linfty`: ensures the maximum entry is 1
-- `:nothing`: do not enforce a metric to equal 1
-"""
-const IMPLIMENTED_METRICS = Set{Symbol}((:L1, :Linfinity, :nothing)) # TODO :L2
-
-"""
-    IMPLIMENTED_PROJECTIONS::Set{Symbol}
-
-- `:nnscale`: Two stage block coordinate decent; 1) projected gradient decent onto nonnegative
-    orthant, 2) shift any weight from `B` to `A` according to normalization. Equivilent to
-    :nonnegative when `normalization==:nothing`.
-- `:simplex`: Euclidian projection onto the simplex blocks accoring to `normalization`
[…]

### 2025-09-10 test/factorize_initialization.jl

@@ -1,37 +1,0 @@
-using BlockTensorFactorization
-using Random
-
-N = 100
-R = 5
-D = 2
-
-fact = BlockTensorFactorization.factorize
-matrices = [abs_randn(N, R) for _ in 1:D]
-l1scale_cols!.(matrices)
-Ydecomp = CPDecomposition(Tuple(matrices))#abs_randn
-@assert all(check.(l1scale_cols!, factors(Ydecomp)))
-Y = array(Ydecomp)
-
-X = CPDecomposition(ntuple(_ -> N, D), R)
-
-options = (
-    tolerance=.01,
-    decomposition=X,
-    maxiter=1000,
-    converged=RelativeError,
-    constraints=[l1scale_cols! ∘ nonnegative!, simplex_cols!],
-    constrain_init=true,
-    constrain_output=true,
-    final_constraints = l1scale_cols!,
-    stats=[
-        Iteration, ObjectiveValue, GradientNNCone, RelativeError, FactorNorms, EuclidianLipschitz
-    ],
-)
-
-# Test initialization functions of factorize
-kwargs = BlockTensorFactorization.default_kwargs(Y; options...)
-
-decomposition, kwargs = BlockTensorFactorization.initialize_decomposition(Y; kwargs...)
-previous, updateprevious! = BlockTensorFactorization.initialize_previous(decomposition, Y; kwargs...)
-parameters, updateparameters! = BlockTensorFactorization.initialize_parameters(decomposition, Y, previous; kwargs...)
-update!, kwargs = BlockTensorFactorization.make_update!(decomposition, Y; kwargs...)

### 2025-09-10 test/geomean.jl

@@ -1,44 +1,0 @@
-using BenchmarkTools
-using Random
-
-f1(x) = prod(x)^(1/length(x))
-f2(x) = exp(mean(log.(x)))
-
-absrandn(x) = abs.(randn(x))
-v = absrandn(10)
-
-f1(v) ≈ f2(v)
-
-@benchmark f1(x) setup=(x = absrandn(10000))
-@benchmark f1(x) setup=(x = 100000*absrandn(10000))
-@benchmark f2(x) setup=(x = absrandn(10000))
-@benchmark f2(x) setup=(x = 100000*absrandn(10000))
-
-function geomean1(v)
-    p = prod(v)
-    if isinf(p) | iszero(p)
-        return exp(mean(log.(v)))
-    else
-        # Faster, but more sensitive
-        return p^(1/length(v))
-    end
-end
-
-function geomean2(v)
-    p = prod(v)
-    if iszero(p) | isinf(p)
-        return exp(mean(log.(v)))
-    else
-        # Faster, but more sensitive
-        return p^(1/length(v))
-    end
-end
-
-@benchmark geomean1(x) setup=(x = absrandn(10000))
-@benchmark geomean2(x) setup=(x = absrandn(10000))
-@benchmark geomean1(x) setup=(x = 10*absrandn(10000))
-@benchmark geomean2(x) setup=(x = 10*absrandn(10000))
-
-@benchmark geomean1(x) setup=(x = 10*absrandn(10))
-@benchmark f1(x) setup=(x = 10*absrandn(10))
-@benchmark f2(x) setup=(x = 10*absrandn(10))

### 2025-09-10 test/geomultiscaletest/multiscaletest.jl

@@ -1,81 +1,0 @@
-# Test of multiscale
-
-using BlockTensorFactorization
-using HDF5
-using BenchmarkTools
-using Logging
-using Random
-
-Random.seed!(314)
-
-filename = "./test/geomultiscaletest/geodata.h5"
-
-myfile = h5open(filename, "r")
-
-#Y = myfile["Y_257_points"] |> read
-Y = myfile["Y_1025_points"] |> read
-
-close(myfile)
-
-size(Y)
-
-Y_backup = copy(Y)
-
-# # Add an extra slice so the continuous dimension is one plus a power of 2
-# Y = cat(Y, zeros(20, 7, 1) ;dims=3)
-
-scaleB_rescaleA! = ConstraintUpdate(0, l1scale_average12slices! ∘ nonnegative!;
-    whats_rescaled=(x -> eachcol(factor(x, 1)))
-)
-
-scaleA_rows! = ConstraintUpdate(1, l1scale_rows! ∘ nonnegative!;
-    whats_rescaled=nothing
-)
-
-nonnegativeA! = ConstraintUpdate(1, nonnegative!)
-
-nonnegativeB! = ConstraintUpdate(0, nonnegative!)
-# [l1scale_average12slices! ∘ nonnegative!, nonnegative!]
-# [scaleB_rescaleA!, scaleA_rows!]
-# [nonnegativeB!, scaleA_rows!]
-
-options = (
-    rank=3,
-    momentum=false,
-    model=Tucker1,
-    tolerance=(0.12),
-    converged=(RelativeError),
-    do_subblock_updates=false,
-    constrain_init=true,
-    constraints=[l1scale_average12slices! ∘ nonnegative!, nonnegative!],
-    stats=[Iteration, ObjectiveValue, GradientNNCone, RelativeError],
-    maxiter=200
-)
-
-#[l1scale_average12slices! ∘ nonnegative!,nonnegative!]
-
-println("First Compile Time")
-@time decomposition, stats_data, kwargs = factorize(Y; options...);
-
-@time decomposition, stats_data, kwargs = multiscale_factorize(Y;
-    continuous_dims=3,options...);
-println("")
-println("Regular Run")
-@time decomposition, stats_data, kwargs = factorize(Y; options...);
-
-@time decomposition, stats_data, kwargs = multiscale_factorize(Y;
-    continuous_dims=3,options...);
-
-println("Benchmark Run")
-
-
[…]

### 2025-09-10 test/multiscalesynthetic3d.jl

@@ -1,264 +1,0 @@
-# generate synthetic data for KDE and tensor factorization
-
-using Random
-using KernelDensity
-#using Pkg
-#Pkg.add("Distributions")
-#Pkg.add("Plots")
-#Pkg.add("PlotlyJS")
-
-using Distributions
-using BlockTensorFactorization
-using Plots
-using PlotlyJS
-
-J = 65 # Number of samples in the x dimension
-K = 65 # Number of samples in the y dimension
-L = 65 # Number of samples in the z dimension
-
-Random.seed!(123)
-
-# Three sources, product distributions
-R = 3
-source1a = Normal(4, 1)
-source1b = Uniform(-7, 2)
-source1c = Uniform(-1, 1)
-
-source2a = Normal(0, 3)
-source2b = Uniform(-2, 2)
-source2c = Exponential(2)
-
-source3a = Exponential(1)
-source3b = Normal(0, 1)
-source3c = Normal(0, 3)
-
-source1 = product_distribution([source1a, source1b, source1c])
-source2 = product_distribution([source2a, source2b, source2c])
-source3 = product_distribution([source3a, source3b, source3c])
-
-sources = (source1, source2, source3)
-
-# x values to sample the densities at
-x = range(-10,10,length=J)
-Δx = x[2] - x[1]
-
-# y values to sample the densities at
-y = range(-10,10,length=K)
-Δy = y[2] - y[1]
-
-# z values to sample the densities at
-z = range(-10,10,length=K)
-Δz = z[2] - z[1]
-
-# Collect sample points into a matrix
-xyz = Array{Vector{Float64}, 3}(undef, length(y),length(x), length(z))
-for (j, x) in enumerate(x)
-    for (k, y) in enumerate(y)
-        for (l, z) in enumerate(z)
-            xyz[j, k, l] = [x, y, z]
-        end
-    end
-end
-
-# Plot the true sources
-XXX, YYY, ZZZ = mgrid(x, y, z)
-source1_density = pdf.((source1,), xyz)
-source2_density = pdf.((source2,), xyz)
-source3_density = pdf.((source3,), xyz);
-
-function plot3d(pdf3d; isomin=1e-4)
-    PlotlyJS.plot(isosurface(
-        x=XXX[:], # isosurface wants all entries in a single list
-        y=YYY[:],
-        z=ZZZ[:],
[…]

### 2025-09-10 test/multiscalesynthetic3dconvergence.jl

@@ -1,321 +1,0 @@
-# generate synthetic data for KDE and tensor factorization
-
-using Random
-using KernelDensity
-#using Pkg
-#Pkg.add("Distributions")
-#Pkg.add("Plots")
-#Pkg.add("PlotlyJS")
-
-using Distributions
-using BlockTensorFactorization
-using Plots
-using PlotlyJS
-
-fact = BlockTensorFactorization.factorize
-
-show_plots = false
-
-J = 65 # Number of samples in the x dimension
-K = 65 # Number of samples in the y dimension
-L = 65 # Number of samples in the z dimension
-
-Random.seed!(123)
-
-# Three sources, product distributions
-R = 3
-source1a = Normal(4, 1)
-source1b = Uniform(-7, 2)
-source1c = Uniform(-1, 1)
-
-source2a = Normal(0, 3)
-source2b = Uniform(-2, 2)
-source2c = Exponential(2)
-
-source3a = Exponential(1)
-source3b = Normal(0, 1)
-source3c = Normal(0, 3)
-
-source1 = product_distribution([source1a, source1b, source1c])
-source2 = product_distribution([source2a, source2b, source2c])
-source3 = product_distribution([source3a, source3b, source3c])
-
-sources = (source1, source2, source3)
-
-# x values to sample the densities at
-x = range(-10,10,length=J)
-Δx = x[2] - x[1]
-
-# y values to sample the densities at
-y = range(-10,10,length=K)
-Δy = y[2] - y[1]
-
-# z values to sample the densities at
-z = range(-10,10,length=K)
-Δz = z[2] - z[1]
-
-# Collect sample points into a matrix
-xyz = Array{Vector{Float64}, 3}(undef, length(y),length(x), length(z))
-for (j, x) in enumerate(x)
-    for (k, y) in enumerate(y)
-        for (l, z) in enumerate(z)
-            xyz[j, k, l] = [x, y, z]
-        end
-    end
-end
-
-if show_plots
-
-# Plot the true sources
-XXX, YYY, ZZZ = mgrid(x, y, z)
-source1_density = pdf.((source1,), xyz)
-source2_density = pdf.((source2,), xyz)
-source3_density = pdf.((source3,), xyz);
-
-function plot3d(pdf3d; isomin=1e-4)
-    PlotlyJS.plot(isosurface(
[…]

### 2025-09-10 test/opnorm/faster_opnorm.jl

@@ -1,132 +1,0 @@
-# Context
-#This document tests different ways to calculate the operator norm of a matrix in Julia.
-
-#The operator norm is needed to calculate the step-size used in each gradient descent update, so it is important to ensure this operation fast.
-
-# Background
-
-# We define the operator norm of a matrix $A\in\mathbb{R}^{I\times J}$ to be
-# $$
-# \left\lVert A \right\rVert_{op} = \sup_{\lVert v \rVert_{2}} \left\lVert A v\right\rVert_{2}.
-# $$
-
-# It is true that
-# $$
-# \left\lVert A^\top \right\rVert_{op} = \left\lVert A \right\rVert_{op}
-# $$
-
-# and
-
-# $$
-# \left\lVert A^\top A \right\rVert_{op} = \left\lVert A A^\top  \right\rVert_{op} = \left\lVert A \right\rVert_{op}^2.
-# $$
-
-# # Tests
-
-# Let's look at a few ways we can find the norm of a matrix in Julia. To ensure a fair comparison, we will wrap all of these in a function, and test them out on three types of dense matrices: tall, wide, and square.
-
-## Possible routines
-# ```{julia}
-# using Pkg; Pkg.activate(".")
-# Pkg.add(["LinearAlgebra","Random","Arpack","BenchmarkTools"])
-# #Pkg.add(path="C:\\Users\\Nicholas\\OneDrive - UBC\\Research\\general-block-decomposition\\BlockTensorFactorization.jl")
-# ```
-
-using Random: rand, randn
-using LinearAlgebra: svdvals, opnorm, Symmetric
-using Arpack: svds
-using BlockTensorFactorization: slicewise_dot
-using BenchmarkTools
-
-
-# Now for the test set of functions
-
-fopnorm1(X) = opnorm(X)
-fopnorm2(X) = sqrt(opnorm(X'X))
-fopnorm3(X) = sqrt(opnorm(X*X'))
-fopnorm4(X) = sqrt(opnorm(Symmetric(X'X)))
-fopnorm5(X) = sqrt(opnorm(Symmetric(X*X')))
-
-fopnorms = [fopnorm1,fopnorm2,fopnorm3,fopnorm4,fopnorm5]
-
-svdvals_top(X) = svdvals(X)[1]
-fsvdvals1(X) = svdvals_top(X)
[…]

### 2025-09-10 test/rescaling_vs_projection.jl

@@ -1,117 +1,0 @@
-# This is a test of different approaches to solving the NMF problem
-# 1/2 ||AB-Y||_F^2
-# where the rows of A, B, & Y all sum to 1
-
-using BlockTensorFactorization
-using Random
-using UnicodePlots
-using Plots
-
-Random.seed!(31415926535897)
-
-fact = BlockTensorFactorization.factorize
-
-I, J = 25, 50
-R = 5
-
-A = abs_randn(I, R)
-l1scale_rows!(A)
-
-B = abs_randn(R, J)
-l1scale_rows!(B)
-
-Y = Tucker1((B, A))
-Y = array(Y)
-
-@assert check(l1scale_rows!, Y)
-
-A_init = abs_randn(I, R)
-l1scale_rows!(A_init)
-
-B_init = abs_randn(R, J)
-l1scale_rows!(B_init)
-
-n_iterations = 20000
-
-options = (
-    model=Tucker1,
-    tolerance=0.0, # Force the algorithm to halt at maxiter
-    maxiter=n_iterations,
-    converged=RelativeError,
-    stats=[
-        Iteration, ObjectiveValue, GradientNNCone, RelativeError, #PrintStats
-    ],
-    momentum=false,
-    do_subblock_updates=false,#false#true
-)
-
-# Possible updates
-# B is the 0th factor, A is the 1st factor
-
-scaleB_rows_rescaleA! = ConstraintUpdate(0, l1scale_rows! ∘ nonnegative!;
-    whats_rescaled=(x -> eachcol(factor(x, 1)))
-)
-
-scaleB_rows! = ConstraintUpdate(0, l1scale_rows! ∘ nonnegative!;
-    whats_rescaled=nothing
-)
-
-scaleA_rows! = ConstraintUpdate(1, l1scale_rows! ∘ nonnegative!;
-    whats_rescaled=nothing
-)
-
-nonnegativeA! = ConstraintUpdate(1, nonnegative!)
-
-nonnegativeB! = ConstraintUpdate(0, nonnegative!)
-
-euclideanProjA! = ConstraintUpdate(1, simplex_rows!)
-
-euclideanProjB! = ConstraintUpdate(0, simplex_rows!)
-
-algorithms = (
-    bothSimplex = [euclideanProjB!, euclideanProjA!],
-    simplexA = [nonnegativeB!, euclideanProjA!],
-    simplexB = [euclideanProjB!, nonnegativeA!],
-    bothScale = [scaleB_rows!, scaleA_rows!],
[…]

### 2025-09-10 test/runtests.jl

@@ -1,605 +1,0 @@
-"""
-Unit tests
-"""
-
-using Test
-
-using Random
-using LinearAlgebra
-
-using BlockTensorFactorization
-
-const VERBOSE = true
-
-@testset verbose=VERBOSE "BlockTensorFactorization" begin
-
-    @testset "Utils" begin
-
-        @testset "interlace" begin
-        @test interlace(1:10,10:15) == [1, 10, 2, 11, 3, 12, 4, 13, 5, 14, 6, 15, 7, 8, 9, 10]
-        @test interlace(1:5,10:20) == [1, 10, 2, 11, 3, 12, 4, 13, 5, 14, 15, 16, 17, 18, 19, 20]
-        @test interlace(1:3, ("this", "that", "other")) == Any[1, "this", 2, "that", 3, "other"]
-        end
-
-        @testset "multifoldl" begin
-            @test multifoldl((+,*,-), (2,3,4,5)) == 15 # ((2 + 3) * 4) - 5
-            @test multifoldl([+,*,-], [2,3,4,5]) == 15
-            @test multifoldl((*,), [2,3]) == 6
-        end
-
-        @testset "norm2" begin
-            @test norm2([3, 4]) == 25
-            @test norm2(1:10) == sum((1:10) .^ 2)
-        end
-
-        @testset "geomean" begin
-            @test geomean(5) == 5
-            @test geomean(2, 1/2) == 1
-            @test geomean((1,2,3)) ≈ 1.8171205928321397
-        end
-
-        @testset "getnotindex" begin
-            @test getnotindex(1:10, 5) == [1, 2, 3, 4, 6, 7, 8, 9, 10]
-            @test getnotindex(1:10, (5, 8)) == [1, 2, 3, 4, 6, 7, 9, 10]
-            @test getnotindex(1:10, [5, 8]) == [1, 2, 3, 4, 6, 7, 9, 10]
-            @test getnotindex(1:10, 4:6) == [1, 2, 3, 7, 8, 9, 10]
-        end
-
-        @testset "proj_one_hot" begin
-            @test proj_one_hot([-6,-4,0,2]) == [0, 0, 0, 1]
-            @test proj_one_hot([-6,-4,-1,-2]) == [0, 0, 1, 0]
-            @test proj_one_hot([-6.1,-4.2,-1.3,-2.4]) == [0., 0., 1., 0.]
-        end
-
-        @testset "all_recursive" begin
[…]

### 2025-09-10 test/subblock_momentum_benchmark.jl

@@ -1,93 +1,0 @@
-using BenchmarkTools
-using Logging
-using BlockTensorFactorization
-
-
-global_logger(SimpleLogger(Warn))
-
-fact = BlockTensorFactorization.factorize
-
-function nnrankr_matrix((I, J), R)
-    A = randn(I, R) .|> abs
-    B = randn(R, J) .|> abs
-    return A*B
-end
-
-options = (
-    :rank => (2,3,4),
-    :tolerance => (0.01),
-    :converged => (RelativeError),
-    :δ => 0.9999,
-    :model => Tucker,
-    #:constraints => nonnegative!,
-)
-
-n_subblock_n_momentum(Y) = fact(Y;
-    do_subblock_updates=false,
-    momentum=false,
-    options...
-)
-
-y_subblock_n_momentum(Y) = fact(Y;
-    do_subblock_updates=true,
-    momentum=false,
-    options...
-)
-
-n_subblock_y_momentum(Y) = fact(Y;
-    do_subblock_updates=false,
-    momentum=true,
-    options...
-)
-
-y_subblock_y_momentum(Y) = fact(Y;
-    do_subblock_updates=true,
-    momentum=true,
-    options...
-)
-
-I, J = 10, 10
-R = 3
-
-@benchmark n_subblock_n_momentum(Y) setup=(Y=array(Tucker1((I, J), R)))
-
-@benchmark y_subblock_n_momentum(Y) setup=(Y=nnrankr_matrix((I, J), R))
-
-@benchmark n_subblock_y_momentum(Y) setup=(Y=nnrankr_matrix((I, J), R))
-
-@benchmark y_subblock_y_momentum(Y) setup=(Y=nnrankr_matrix((I, J), R))
-
-performance_increase(old, new) = (old - new) / new * 100
-
-time_decrease(old, new) = (old - new) / old * 100
-
-
-###################
-
-
-I, J, K = 10, 10, 10
-R1, R2, R3 = 2, 3, 4
-
-b = @benchmark n_subblock_n_momentum(Y) setup=(Y=Tucker((I, J, K), (R1, R2, R3))|>array)
-display(b)
-b = @benchmark y_subblock_n_momentum(Y) setup=(Y=Tucker((I, J, K), (R1, R2, R3))|>array)
-display(b)
[…]

### 2025-09-11 paper.qmd
render updated paper

--- 2025-09-10 paper/paper.qmd
@@
 ---
-title: "BlockTensorFactorizations.jl: A Unified Constrained Tensor Decomposition Julia Package"
+title: "BlockTensorFactorization.jl: A Unified Constrained Tensor Decomposition Julia Package"
 subtitle: ""
@@
 Details like what rank to select, how the constraints should be enforced, and convergence criteria should be handled automatically, but customizable to the knowledgable user.
-These are the core specification for BlockTensorFactorizations.jl.
+These are the core specification for BlockTensorFactorization.jl.
 
@@
 
-The main contribution is a description of a fast and flexible tensor decomposition package, along with a public implementation written in Julia: BlockTensorFactorizations.jl.
+The main contribution is a description of a fast and flexible tensor decomposition package, along with a public implementation written in Julia: BlockTensorFactorization.jl.
 This package provides a framework for creating and performing custom tensor decompositions.
